### 2013-09-23 paper/paper.tex
Added paper to repository and began editing it

% Template for the submittion to:
%   The Annals of Probability           [aop]
%   The Annals of Applied Probability   [aap]
%   The Annals of Statistics            [aos]
%
%Author: In this template, the places where you need to add information
%        (or delete line) are indicated by {???}.  Mostly the information
%        required is obvious, but some explanations are given in lines starting
%Author:
%All other lines should be ignored.  After editing, there should be
%no instances of ??? after this line.

% use option [preprint] to remove info line at bottom
% journal options: aop,aap,aos
%\documentclass[aos]{imsart}
\documentclass{imsart}

\usepackage{amsthm,amsmath,natbib}
%\RequirePackage[dvips]{hyperref}

% use this package if hyperref and natbib is used:
%\RequirePackage{hypernat}

% provide arXiv number if available:
%\arxiv{math.PR/0000000}

% put your definitions there:
\startlocaldefs

\usepackage{graphics,amssymb,color}

\newcommand{\argmax}{\mathop{\mathrm{argmax}}}
\newcommand{\argmin}{\mathop{\mathrm{argmin}}}
\newcommand{\minimize}{\mathop{\mathrm{minimize}}}
\newcommand{\maximize}{\mathop{\mathrm{maximize}}}

\newcommand{\todo}{\textcolor{red}{\textbf{To do: }}}

\newcommand{\real}{\mathbb{R}}
\newcommand{\Pp}{{\mathbb P}}
\newcommand{\Ee}{{\mathbb E}}
\newcommand{\lips}{{\cal L}}
\newcommand{\tube}{\mathrm{Tube}}
\newcommand{\sphere}[1]{O^{#1-1}}
\newcommand{\dimens}{\text{dim}}
\newcommand{\cone}{\text{Cone}}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{cor}[theorem]{Corollary}
\newcommand{\XK}{K_X}
\def\qed{\hfill $\Box$\newline}

\endlocaldefs

\begin{document}

\begin{frontmatter}

% "Title of the paper"
\title{Significance testing and forward stepwise with the group lasso}

%%%
%%% \runtitle{} - I don't know what this is
%%%

% indicate corresponding author with \corref{}
\author{\fnms{Jonathan} \snm{Taylor}\corref{Jonathan E. Taylor}\ead[label=e1]{jonathan.taylor@stanford.edu}\thanksref{t1}
\and \fnms{Joshua} \snm{Loftus}
\and ???}
\thankstext{t1}{Supported in part by NSF grant DMS 1208857 and
AFOSR grant 113039.}

\affiliation{Stanford University}

\address{Department of Statistics\\  Stanford University\\ Sequoia
Hall \\390 Serra Mall\\ Stanford, CA 94305, U.S.A.\\ \printead{e1}}

%%%
%%% \runauthor{Taylor} - I don't know what this is
%%%

\begin{abstract}
% perhaps de-emphasize group lasso and emphasize forward stepwise?
  We extend the work of \cite{significance:lasso} on significance
  testing for the lasso to the group lasso.  We derive the corresponding
  test statistic and its exact distribution, and show that these can be
  used with (non-penalized) forward stepwise regression and model
  selection.  For group lasso the forward stepwise procedure adds a
  group of variables in each step.  These procedures are shown to
  perform well in simulations and in a real data example.

\end{abstract}

%%%
%%% - Gotta change these
%%%
\begin{keyword}[class=AMS]
\kwd[Primary ]{62M40}
%\kwd{}
\kwd[; secondary ]{62H35}
\end{keyword}

\begin{keyword}
\kwd{penalized regression}
\kwd{convex analysis}
\kwd{least squares}
\kwd{Gaussian processes}
%\kwd{extreme value theory}
\end{keyword}

\end{frontmatter}

\section{Introduction}\label{sec:intro}

Forward stepwise regression is a stochastic model selection procedure
that begins with an empty model and adds the best predictor variable
in each step.  Classical significance tests fail when a model has been
selected this way and tend to be anti-conservative.  Recently,
\cite{significance:lasso} found a novel test statistic with an
appropriate null distribution that behaves well when a model has been
selected using the lasso \citep{tibshirani:lasso}.
\cite{tests:adaptive} modified and extended those results to the
group lasso \citep{grouplasso} and other adaptive regression
problems.  The present work
explores the behavior of those test statistics for models selected by
forward stepwise procedures and works out some of the details
involved in applying these methods to models with grouped variables.
We find that our test statistics can be used for valid significance
testing, or for tempering the aggressive nature of forward stepwise
model selection.  The resulting methods may be less computationally
intensive than choosing a regularization parameter for the group lasso
by cross-validation.

To fix notation for the group lasso problem we consider an outcome $y
\in \real^n$.  Let an integer $G \geq 1$ be the number of groups (or
factors).  Lowercase $g$ will be used as an index for groups.  For
each $1 \leq g \leq G$ the design matrix encoding the $g$th group is
the $n \times p_g$ matrix denoted $X_g$.  Throughout most of this
paper we assume orthogonality within groups, that is $X_g^T X_g =
I_{p_g \times p_g}$ for all $g$. ( \todo Take this out if we don't
actually need it).  Let $p = \sum_{g=1}^G p_g$ be the total number of
individual variables, and let $X$ be the matrix constructed by
column-binding the $X_g$, that is  
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation*}
X = \begin{pmatrix} X_1 & X_2 & \cdots & X_G  \end{pmatrix}
\end{equation*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
With each group we associate the $p_g \times 1$ coefficient vector
$\beta_g$, and write $\beta$ for the $p \times 1$ vector constructed
by stacking all of the $\beta_g$ in order.  Finally, our model for the
data is
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
\begin{aligned}
\label{eq:gmodel}
y & = X \beta + \sigma \epsilon \\
   & = \sum_{g=1}^G X_g \beta_g + \sigma \epsilon
\end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
where $\epsilon$ is noise.  Unless otherwise specified we will assume
i.i.d. Gaussian noise $\epsilon \sim N(0, I_{n \times n})$ throughout.
So far we have only introduced some new notation to the usual linear
regression setup, but now we define the \emph{group lasso estimator}
as the following solution to a convex problem 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
\begin{aligned}
\label{eq:gsoln}
\displaystyle \hat \beta_\lambda = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
   \lambda \sum_{g=1}^G w_g \| \beta_g \|_2
\end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The parameter $\lambda \geq 0$ enforces sparsity in groups: for large
$\lambda$ most of the $\beta_g$ will be zero vectors. The weights
$w_g$ are usually taken to be $\sqrt p_g$ to normalize the penalty
across groups.  Note that this includes the usual lasso estimator as a
special case when all of the groups are of size 1, since then the
penalty term is the $L^1$-norm of $\beta$.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%% Related works %%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

The group lasso estimator is discussed in (ref that guy's thesis) and
(ref Yuan and Lin).  An important extension is the sparse group lasso
(ref ???) which enforces sparsity in groups as well as sparsity of the
coefficients within the groups.  For a survey on group lasso and
related factor models see (ref???). \todo review some more literature
and add a few more references here if they seem worthwhile.

In a recent work \cite{significance:lasso} defined a \emph{covariance
  test statistic} for testing the significance of a variable entering
the model along the lasso solution path.  They derived a simple
asymptotic null distribution, proved a type of convergence under broad
``minimum growth'' conditions, and demonstrated in simulations that
the test statistic closely matches its asymptotic distribution even in
finite samples.  That work marked an important advance in the problem
of combining inference with model selection.  The current paper
extends some of their results to the group lasso (ref Yuan and Lin?).
In the process we had to derive an exact finite sample null
distribution for the test statistic (ref TLTT?).  We also show that
the techniques used to get these results can be used to do a forward
stepwise procedure related to the group lasso that adds groups of
variables (or factors) in each step.

(\todo Change this paragraph as the sections are completed)
In Section 2 we briefly review the work of (ref LTTT) concerning
significance testing for the lasso, and discuss how this result is
extended to group lasso. Section 2 also introduces more of the
terminology we use to describe the present problem. Section 3
describes a framework for forward stepwise procedures using the test
statistic derived in Section 2. In this framework groups of variables
are added in each step until some stopping criterion involving the
test statistic is satisfied. In Section 4 we show the results of some
simulations for several forward stepwise procedures, and in Section 5
we apply our method to a real data set (\todo brief description of
data).

%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Significance testing: from lasso to group lasso}
\label{sec:testing}

\subsection{Background}
In the ordinary least squares setting, a significance test for a
single variable can be conducted by comparing the drop in residual
sums of squares (RSS) to a $\chi^2_1$ distribution. Similarly, when
adding a group of $k$ variables we can compare the drop in RSS to a
$\chi^2_k$ random variable. This generally does not work when the
variable to be added has been chosen by a method that uses the data,
and in particular it fails for forward stepwise procedures which add
the ``best'' (e.g. most highly correlated) predictor in each step. In
that case, the test statistic (drop in RSS) does not match the
theoretical null distribution even when the null hypothesis is
true. Lockhart et al. (ref LTTT) introduced a new test statistic based
on the knots in the lasso solution path. Writing $\hat
\beta(\lambda)$, the solution for a fixed value of $\lambda$, we need
the following facts (ref Tibs2012)

\begin{itemize}

  \item The vector valued function $\hat \beta(\lambda)$ is a
    continuous function of $\lambda$. For the lasso path, the
    coordinates of $\hat \beta(\lambda)$ are piecewise linear with
    changes in slope occurring at a finite number of $\lambda$ values
    referred to as \emph{knots}. The knots depend on the data and are
    usually written in order $\lambda_1 \geq \lambda_2 \geq \cdots
    \geq \lambda_r \geq 0$. We follow this convention.

  \item The \emph{active set} $A_k$ is a set of indices of variables
    for which the corresponding coordinates of $\hat \beta(\lambda_k)$
    are potentially nonzero. Any variable with index not in $A_k$ has
    a zero coefficient in $\hat \beta(\lambda_k)$, but the converse is
    not true.

  \item Path algorithms for computing lasso solutions proceed by
    fitting models at a grid of $\lambda$ values. The active set
    changes whenever $\lambda$ crosses a knot, and predictor variables
    can both enter and leave the active set. However, at the first two
    knots $\lambda_1$ and $\lambda_2$ no variable can leave the active
    set. So the first two knots correspond to the first two variables
    entering the model.

\end{itemize}

Lockhart et al. prove that, under the null hypothesis that $A_k$
contains all the strong predictor variables, the distribution of a
test statistic $T_k \propto \lambda_k(\lambda_k - \lambda_{k+1})$ is
asymptotically Exp(1). In the lasso case we know a lot about the knots
and active set, but the group lasso picture is slightly more
complicated. For the group lasso, $\hat \beta(\lambda)$ does not have
piecewise linear components. To overcome this difficulty we will
restrict our attention to the first group of variables to enter the
active set since the analysis then follows almost exactly as for the
lasso. (\todo Change this if I can find $\lambda_k$).

Before considering the group lasso, we review some ingredients of the
proof for the lasso. Let $J = \{ 1, 2, \ldots, p \}$ index variables
and consider a stochastic process $f_{j,s} = sX_j^Ty$ defined on $T =
J \times [ -1, 1 ]$. This stochastic process is simply a collection of
linear combinations of $y$, hence it is Gaussian under the assumption
of Gaussian errors. The \emph{Karush-Kuhn-Tuckher (KKT) conditions}
(ref???) imply that $\lambda_1 = \max_j |X_j^Ty|$. By introducing the
sign variable $s$, we can remove the absolute value and write
$\lambda_1$ as the maximum of our Gaussian process

\begin{equation}
\lambda_1 = \max_{(j,s)} f_{j,s}
\end{equation}

We have exhibited the first knot as the maximum of a Gaussian
process. We can do this for the second knot by introducing a new
process. Let $(j_1, s_1)$ be the maximizer so that $\lambda_1 =
s_1X_{j_1}^Ty$, and define 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
\begin{aligned}
  f^{(j_1,s_1)}_{(j,s)} & = \frac{ sX_j^T y - s X_j^T X_{j_1} X_{j_1}^Ty  } { 1 -  ss_1 X_j^TX_{j_1}} \\
  & = \frac{ sX_j^T(I-P_{j_1}) y }{ 1 -  s_1 X_{j_1}^T X_js}
\end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
where $P_j$ is the projection onto the subspace spanned by $X_j$. We
can think of this as a ``residual process'' after regressing out the
maximum. Write $M = \max_{j \neq j_1, s} f^{(j_1,s_1)}_{(j,s)}$, the
maximum of this residual process. It can be shown from the KKT
conditions that $M  = \lambda_2$. To summarize, we have represented
the first two knots of the lasso solution path as the maxima of some
natural Gaussian processes. Distributional facts about Gaussian
processes now allow us to make conclusions about the distribution of
functions of the knots.

\subsection{Group lasso}
To extend this argument to the group lasso we need to define Gaussian
processes that characterize the knots of the group lasso solution
path.

\todo Either use the simplified argument in the case of equal weights
(equal group sizes), or finish adjusting the proof of $M = \lambda_2$
to include the weights and use that version. The proof can go in an
appendix.

\emph{(For this section I have a lot of stuff written that can just be pasted in once I'm done wrangling the KKT conditions for the unequal weight case)}

\subsection{Better p-values}

\begin{itemize}

\item \todo Add the figures to this section
\item \todo Convert to exposition instead of bulleted list
  \item As in LTTT, $T = \lambda_1 (\lambda_1 - M)$ and $M = \lambda_2$
  \item Convergence to the limiting Exp(1) distribution is too slow
    \[
      \frac{P(\chi_k / w_g \geq m + t/m)}{P(\chi_k / w_g \geq m)} \to e^{-t} \text{ as } m \to \infty
    \]
    (when the group achieving $\lambda_1$ is group $g$ and has rank $k$)
  \item The limiting distribution only depends on $T$, but we also observe $M$
  \item Let's just try the ratio (conditional $\chi_k$ tail probability) evaluated at $T$ and $M$ (it works better)
  \item Going one step further, instead of using the approximation (see LTTT Proof of Lemma 5)
    \[
      \frac{M + \sqrt{M^2+4t}}{2} \approx M + \frac{t}{M}
    \]
    we can just use the left hand side
  \item For $T = \lambda_1(\lambda_1 - M)$ the left hand side simplifies to $\lambda_1$
  \item Now our p-value is
    \[
      \frac{P(\chi_k / w_g \geq \lambda_1)}{P(\chi_k / w_g \geq \lambda_2)}
    \]   
\end{itemize}

\subsection{Can we do $\lambda_k$? Don't know yet}

\section{Forward stepwise procedures}
\todo This is potentially the second largest thing not done yet. However, if it turns out we can't think of a better stopping rule, then this section is done and just needs to be written (and would not take long).

\section{Simulations }
(I've done many of these and have a lot of code, so I can produce a lot of output for this section)
\todo Modify some of my previous simulations after deciding on a stopping rule, decide which simulations to put here.

\section{Real data example}
\todo This is definitely the single largest part I have not done yet.

%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
%% Check below for latex examples if necessary %
%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%
%
%
%In the classical setting $n > p$ the system eqref{eq:LS:KKT} often has a unique solution, the familiar
%$$
%\hat{\beta} = (X^TX)^{-1}(X^Ty).
%$$
%In many parametric models, the least squares model is of course too simple. In the exponential family setting  \citep{amari,efron}, the normal equations are similar, with  $(X^TX)$ replaced  by the observed Fisher information. We have focused on squared error-loss
%for its simplicity of exposition.
%
%In high-dimensional settings, $n$ is often less than $p$ and there is of course no unique solution 
%to eqref{eq:LS:KKT}. Many modifications are possible, for instance ridge or Tikhonov regularization which adds a strongly convex quadratic term to . The addition of such quadratic terms changes
%the quadratic part of the loss but does not fundamentally change much else until we begin
%to make assumptions about whether or not the model is correct, and how much bias might be incurred
%by such regularization.
%
%In modern high-dimensional settings the regularization term, or penalty, of choice is often a norm,
%with the LASSO \citep{tibshirani:lasso} being the most popular. The lasso problem is
%\begin{equation}
%\label{eq:lasso:0}
%\minimize_{\beta \in \real^p} \frac{1}{2} \|y-X\beta\|^2_2 + \lambda \|\beta\|_1.
%\end{equation}
%The duality between norms allows us to write $$
%\|\beta\|_1 = \sup_{u:\|u\|_{\infty} \leq 1} u^T\beta = h_{B_{\infty}}(\beta)
%$$
%with $B_{\infty}$ the $\ell_{\infty}$ ball of radius 1 in $\real^p$ and
%for any set $K$
%\begin{equation}
%\label{eq:support}
%h_K(\beta) = \sup_{u \in K} u^T\beta
%\end{equation}
%is the support function of the set $K$ which we assume to be closed and containing 0. In this notation,
%the LASSO problem can be expressed as
%\begin{equation}
%\label{eq:lasso}
%\minimize_{\beta \in \real^p} \frac{1}{2} \|y-X\beta\|^2_2 + \lambda h_{B_{\infty}}(\beta).
%\end{equation}
%
%
%Our canonical problem is therefore
%\begin{equation}
%\label{eq:canonical}
%\begin{aligned}
%\minimize_{\beta \in \real^p} \frac{1}{2} \|y-X\beta\|^2_2+ \lambda h_{K}(\beta)
%\end{aligned}
%\end{equation}
%with $K$ being a closed, convex set containing 0. For one of many possible infinite
%dimensional formulations of this canonical problem, see \citep{tsirelson} whose author is also associated
%to one of the most famous tools in the theory of Gaussian process, the Borell TIS inequality \citep{rfg}.
%
%
%The normal equations of the least squares problem are replaced with the KKT conditions \citep{boyd} for eqref{eq:canonical}. For our canonical problem,
%the KKT conditions are
%\begin{equation}
%\label{eq:KKT}
%X^T(X\hat{\beta}-y) + \hat{u}= 0, \qquad \hat{u} \in \lambda \cdot \partial h_K(\hat{\beta}),
%\end{equation}
%where the $\partial$ denotes the sub differential. In what follows, we denote
%a solution to this problem as $\hat{\beta}_{\lambda,\XK}$ to denote the dependence on
%the penalty $\XK$ and the penalty parameter $\lambda$.
%
%
%As we can encode linear or cone constraints in the support function,
%it is safe to say that a huge number of problems fit into this framework.
%Some examples include:
%\begin{itemize}
%\item LASSO \citep{tibshirani:lasso};
%\item compressed sensing \citep{candes:cs,donoho:cs};
%\item group LASSO  \citep{grouplasso,overlap:group:lasso};
%\item graphical LASSO \citep{graphical:lasso,buhlmann};
%\item matrix completion \citep{candes:recht:exact,mazumder:hastie:softsvd};
%\item sign restricted regression \citep{meinshausen:sign};
%\item hierarchically constrained models \citep{bien:hierarchical,bach:hierarchical};
%\item generalized LASSO \citep{tibshirani:taylor:path};
%\item total variation denoising \citep{nesta}.
%\end{itemize}
%
%The relation between eqref{eq:canonical} and the metric projection map is through a particular
%dual function, also developed by Legendre. The canonical dual problem is 
%\begin{equation}
%\label{eq:canonical:dual}
%\minimize_{u \in  \lambda \cdot \text{row}(X) \cap K} \frac{1}{2} (u-w)^T(X^TX)^{\dagger}(u-w)
%\end{equation}
%with $\text{row}(X)$ the row space of the matrix $X$. This dual problem can be derived by minimizing
%the following Lagrangian with respect to $\beta, \eta$
%\begin{equation}
%\label{eq:lagrangian}
%L(\beta,\eta;u) = \frac{1}{2} \|y-X\beta\|^2_2 + \lambda h_K(\eta) + u^T(\beta-\eta).
%\end{equation}
%After a sign change, the problem in eqref{eq:canonical:dual} is fairly easily seen to be equivalent to 
%\begin{equation}
%\label{eq:canonical:dual0}
%\maximize_{u}  \left[\inf_{\beta,\eta} L(\beta,\eta;u) \right]
%\end{equation}
%with the constraints in eqref{eq:canonical:dual} encoding the fact that
%$$
%\inf_{\beta,\eta} L(\beta,\eta;u) = -\infty
%$$
%whenever $u \not \in \lambda \cdot \text{row}(X) \cap K. $
%Any pair $\hat{\beta}, \hat{u}$ is related through
%\begin{equation}
%\label{eq:primal:dual}
%(X^TX)\hat{\beta} - w + \hat{u} = 0.
%\end{equation}
%
%Choosing an orthonormal basis for $\text{row}(X)$, we see that
%the dual problem can be phrased as the metric projection problem of $(X^TX)^{-\dagger/2}w$ onto
%$  \lambda \cdot  \text{row}(X) \cap K$. Alternatively, if we are interested only in 
%the fitted values $X\hat{\beta}_{\lambda, K}$, the original problem
%eqref{eq:canonical} can be expressed as the residual $\hat{\mu}=y-\hat{r}$ 
%from
%\begin{equation}
%\label{eq:resid0}
%\hat{r} = \argmin_{r \in \lambda \cdot \XK} \frac{1}{2} \|y-r\|^2_2,
%\end{equation}
%where
%\begin{equation}
%\label{eq:XK}
%\XK = (X^T)^{-1}K.
%\end{equation}
%Or, in another form,
%\begin{equation}
%\label{eq:resid}
%\hat{\mu}(y) = y - \pi_{\lambda \cdot \XK}(y).
%\end{equation}
%We see that our canonical regression problem is in fact a metric projection problem.
%
%Having posed the canonical high-dimensional regression problem as a metric projection problem, 
%we now try to describe how metric projection is related to some fundamental issues in the understanding
%of this problem both from an algorithmic view and an inferential view.
%
%\subsection{Algorithms to solve the canonical problem}
%\label{sec:algorithm}
%
%The general problem is phrased in terms of an arbitrary $K$. For many high-dimensional
%problems, this $K$ is chosen to emphasize expected structure in the data. For example,
%it is well-known that the LASSO yields sparse solutions, the group LASSO yields
%groups of non-zero coefficients, etc.
%
%This special structure is based on a particular structure encoded in $K$. Further, for many canonical
%choices of $K$ used in high dimensional statistics, such as those cited in \S \ref{sec:problem}, 
%the following metric projection is simple 
%$$
%\nu \mapsto \argmin_{ u \in \lambda \cdot K} \frac{1}{2} \|\nu-u\|^2_2 = \nu - \argmin_{\beta} \left[\frac{1}{2} \|\nu-\beta\|^2_2 + \lambda h_K(\beta) \right].
%$$
%
%Such optimization problems are referred to  as problems in composite form \citep{nesta,boyd,nesterov}. For such problems, 
%many modern first order solvers use a version of 
%generalized gradient descent. The steps in generalized gradient descent
%are essentially iterations of this metric projection map. Specifically, to
%solve the canonical problem eqref{eq:canonical} for a given step-size $\alpha$ a simple generalized
%gradient algorithm reads as
%\begin{equation}
%\begin{aligned}
%\hat{\beta}^{(k+1)} &= \argmin_{\beta} \frac{1}{2 \alpha} \|\nu^{(k)}_{\alpha}-\beta\|^2_2 + \lambda h_K(\beta) \\
%&= \nu^{(k)}_{\alpha} - \argmin_{\eta \in \lambda \alpha \cdot K} \frac{1}{2} \|\nu^{(k)}_{\alpha}-\eta\|^2_2, \\
%\nu^{(k)}_{\alpha} &= \beta^{(k)} - \alpha \cdot  X^T(X\beta^{(k)}-y).
%\end{aligned}
%\end{equation}
%
%The first line in the update above is the usual form of updates for generalized gradient descent, while the second line
%expresses this step as the residual after an application of the metric projection map. 
%Accelerated schemes can do much better with slightly different updates above, see \citep{nesta,nesterov,tseng}.
%With modern computing techniques, such simple algorithms can scale to huge
%problems, see \citep{admm,mazumder:hastie:softsvd}.
%
%\section{Inference for the canonical problem}
%\label{sec:inference}
%
%\subsection{Law of large numbers}
%
%Having solved the canonical problem, what can we say about its solution?
%As this special issue is devoted to the appearance of one of the first proofs of the law of large numbers, 
%we should at least hope to provide such an answer.
%
%In the classical setting, assuming independence, and the usual linear regression model
%\begin{equation}
%\label{eq:model}
%y = \mu + \epsilon
%\end{equation}
% with noise $\epsilon$ having scale $\sigma$,
%the central limit theorem
%can often be applied to yielding the usual result
%\begin{equation}
%\|X(\hat{\beta} - \beta_0)\|_2^2 \overset{D}{\approxeq} \sigma^2 \cdot \chi^2_p 
%\end{equation}
%under the null $H_0: \mu =X\beta_0 \in \text{col}(X)$. Of course, this
%forms the basis of much
%inference in modern (and not so modern) applied statistics in the fixed $p$, $n$ growing regime.
%In terms of the parameters themselves, this
%implies the weaker statement
%\begin{equation}
%\label{eq:LLN}
%\|\hat{\beta} - \beta\|_2 \leq n^{-1/2}\sigma \cdot \zeta_n
%\end{equation}
%for some random variable $\zeta_n=O_{\Pp}(1)$.
% 
%In the classical setting, assuming $X$ is full rank, the  bound eqref{eq:LLN} is a simple two line proof followed by
%some assertions. If
%we write ${\cal L}(\beta) = \frac{1}{2} \|y-X\beta\|^2_2$, then
%$$
%\begin{aligned}
%0 &\geq  {\cal L}(\hat{\beta}) - {\cal L}(\beta_0) \\
%& =   \nabla {\cal L}({\beta}_0)^T(\hat{\beta}-\beta_0) + \frac{1}{2}(\hat{\beta}-\beta_0)^T(X^TX)(\hat{\beta}-\beta_0) \\
%&= (X^T\epsilon)^T(\hat{\beta}-\beta_0) + \frac{1}{2}(\hat{\beta}-\beta_0)^T(X^TX)(\hat{\beta}-\beta_0) \\
%& \geq - \|X^T\epsilon\|_2 \|\hat{\beta}-\beta_0\|_2 + \frac{\lambda_{\min}(X^TX)}{2} \|\hat{\beta}-\beta_0\|^2_2
%\end{aligned}
%$$
%with $\lambda_{\min}(X^TX)$ denoting the smallest eigenvalue of $X^TX$. We see that we can take $$
%\zeta_n = \sigma^{-1/2} n^{1/2} \frac{\|X^T\epsilon\|_2}{\lambda_{\min}(X^TX)}.
%$$
%
%%Before proceeding, note that
%%$$
%%\|X^T\epsilon\|_2 = \sup_{u \in B_2(\real^p)} u^TX^T\epsilon
%%$$
%%which is the supremum of a (particularly simple) process. If the noise $\epsilon$ is
%%Gaussian, then, conditional on $X$, this is the supremum of a {\em Gaussian} process, the second
%%tool in the geometry of least squares toolbox.
%
%In the high-dimensional setting, of course this fails as $\lambda_{\min}(X^TX)=0$.
% What, then, can we say about
%our canonical estimator
%$$
%\hat{\beta}_{\lambda,\XK} = \argmin_{\beta \in \real^p} \frac{1}{2} \|y-X\beta\|^2_2+ \lambda h_K(\beta)?
%$$
%Is there even a weak law of large numbers?
%Without any assumptions on $K$, the answer is clearly no: if $K=\{0\}$, then this is the original ill-posed
%problem.
%
%Under an assumption of decomposability of $K$ 
%recent progress has been
%made in providing bounds on the estimation error in eqref{eq:canonical}, see  \citep{decomposable}. 
%The notion of decomposability in \citep{decomposable} has a precise definition which we will not dwell on here. However,
%a large set of examples of decomposable penalties are penalties of the form 
%\begin{equation}
%\label{eq:product}
%K = \prod_{i \in {\cal I}}K_i.
%\end{equation}
% That is,
%convex sets that can be expressed as products of convex sets lead to decomposable
%penalties.
% Every example of a decomposable norm in \citep{decomposable} has this form except the nuclear norm. For such penalties, the generalized gradient algorithms
%described in \S \ref{sec:algorithm} decompose into many smaller subproblems. Many efficient coordinate
%descent algorithms exploit this fact, see \citep{pathwise,glmnet}.
%
%A similar notion of decomposability we refer to as {\em additivity}
%is explored in \citep{tibs:taylor:additive} in which case $K$ can be expressed as $K=A+I$ with the
%sum being Minkowski addition of sets. In this case, the penalty has the form
%\begin{equation}
%\label{eq:additive}
%h_K(\beta) = h_A(\beta) + h_I(\beta).
%\end{equation}
%One concrete example of this is the $\ell_{\infty}$ ball in $\real^p$. For $A, I$ a partition of
%$\{1, \dots, p\}$ into {\em active} and {\em inactive} variables, we can write
%$$
%B_{\infty} = \left\{(u_A,0): \|u_A\|_{\infty} \leq 1 \right\} + \left\{(0,u_I): \|u_I\|_{\infty} \leq 1 \right\}.
%$$
%Any penalty of the form eqref{eq:product} can be expressed in the form eqref{eq:additive} in a similar fashion.
%If we are allowed to introduce a linear constraint to eqref{eq:canonical} then any problem with a  penalty of the form eqref{eq:additive} can be expressed as a problem with a penalty of the form eqref{eq:product} subject to an additional  set of linear constraints.
%
%The weak law of large numbers presented in the literature
%follow a similar path to the argument above. Of course, for precise results, the specific
%penalty as well as the data generating mechanism must be more precisely specified. 
%
%In the interest of space,
%we do not pursue such precise statements here.
%Rather, we will just attempt to paraphrase these results, of which there exist many in the literature (c.f. \citep{decomposable,bickel:ritov,support:recovery,highdim:stat}) with \citep{decomposable} being a
%particularly nice place to read in detail. Under various assumptions on the tails of 
%$\epsilon$, as well as the assumption $h_I(\beta_0)=0$ and $K^{\circ}$ is bounded\footnote{Recall the definition
%of the polar body of $K$
%\begin{equation}
%\label{eq:polar}
%K^{\circ} = \left\{\nu \in \real^p: u^T\nu \leq 1, \ \forall u \in K \right\}.
%\end{equation} For any $K$, the seminorm $h_{K^{\circ}}$ is the dual seminorm of $h_K$.},
% the canonical result has the form for $\lambda \geq C_1 \cdot \Ee\left(h_{K^{\circ}}(-X^T\epsilon) \right)$
%\begin{equation}
%\label{eq:error}
%\|\hat{\beta}_{\lambda, \XK} - \beta_0\|_2 \leq C_2 \cdot \sigma \frac{ \Ee\left(h_{K^{\circ}}(-X^T\epsilon) \right) \psi(A)}{\kappa(A, I, X)}
%\end{equation}
%with high probability
%for some universal $C_1, C_2$ 
%where $\psi(A)$ is referred to as a {\em compatibility} constant relating the $\ell_2$ norm and the 
%$h_A$ seminorm; the quantity 
%$\kappa(A,X)$ replaces $\lambda_{\min}(X^TX)$ and is referred to as a {\em restricted strong
%convexity} (RSC) constant. 
%
%The literature varies in their assumptions
%on the noise and the design matrix $X$. For instance, in the fixed design case one might
%consider the above probability only with respect to noise, while in a random design setting the dependence
%of the constants on $X$ are typically expressed with respect to the law that generates the design matrix $X$.
%
%% Above, we introduced
%%$K^{\circ}$, the polar body of $K$ defined as
%%\begin{equation}
%%\label{eq:polar}
%%K^{\circ} = \left\{\nu \in \real^p: u^T\nu \leq 1, \ \forall u \in K \right\}.
%%\end{equation}
%%
%
%Having established a bound such as eqref{eq:error}, 
%if one considers problems indexed by $n$, then one can obtain
%a law of large numbers for the problem eqref{eq:canonical} so long as the parameters
%are chosen so the right hand side decays to 0 and the bound holds with sufficiently high probability. 
%Again in the interest of space, we refer
%readers to the literature for more precise statements for specific versions of the problem such as the LASSO.
%We have tried to give a summary of some of the results related to the canonical problem, though we have
%barely exposed the tip of the iceberg. Under more specific assumptions much more can be said. For example, see \cite{donoho:montanari}.
%%
%%
%%If $\epsilon$ is Gaussian and $X$ is considered fixed, then the choice of $\lambda$ and 
%%the final constant in the bound above is
%%the canonical Gaussian
%%width of $\XK^{\circ}$, see \citep{tsirelson}. For other noise models,
%%this measures the noise under some other law. 
%%
%%The Gaussian width and analogous quantities arise in many contexts in machine learning. In some sense, made
%%proper in the machine learning and empirical processes literature,
%%this width, or its empirical process counterpart, is the fundamental complexity of the estimation problem eqref{eq:canonical}.
%%If we were to be so ambitious as to extend our problem to an infinite dimensional setting where
%%$\real^p$ is reproduced by some reproducing kernel space $H$, then the concept of 
%%Gaussian width is known to be related to the existence of maximum likelihood
%%estimators in such settings \citep{tsirelson}.
%
%\subsection{Gaussian width and metric projection: intrinsic volumes}
%
%For $\XK^{\circ} = X^TK^{\circ} \subset \real^n$, the error eqref{eq:error}
% depends on the
%quantity
%\begin{equation}
%\label{eq:width}
%\Ee(h_{K^{\circ}}(-X^T\epsilon)) = \Ee(h_{\XK}(-\epsilon)).
%\end{equation}
% For fixed $X$, this quantity is referred to as
% the Gaussian width of $\XK$ and it also intimately
%related to our first tool in the toolbox,  the metric projection. Specifically, 
%consider the tube of radius $r$ around $\XK^{\circ}$. That is,
%\begin{equation}
%\label{eq:tube}
%\tube(\XK^{\circ},r) = \left\{z \in \real^n: \|z - \pi_{\XK^{\circ}}(z)\|_2 \leq r \right\}.
%\end{equation}
%Then, a classical result of  Steiner in the case of convex bodies and Weyl in the case of manifolds  says that
%the Lebesgue measure of the tube, assuming $\XK^{\circ}$ is bounded, can be expressed as
%\begin{equation}
%\left|\tube(\XK^{\circ},r)\right|_{\real^n} = \sum_{j=0}^n r^j \omega_j \lips_{n-j}(\XK^{\circ}), \qquad r \leq r_c(\XK^{\circ})=\infty,
%\end{equation}
%where the $\lips_l(\XK^{\circ})$ are referred to as the intrinsic volumes of $\XK^{\circ}$
%and $\omega_l = |B_2(1)|_{\real^l}$ is the Lebesgue measure of the 
%unit $\ell_2$ ball in $\real^l$. See \citep{rfg,federer,schneider,weyl} for more details on such volume of tubes formulae.
%When $\XK^{\circ}$ is unbounded, Federer's curvature measures \citep{federer} can be used to define the volume of local tubular neighborhoods.
%Using  Gaussian process techniques \citep{vitale}, it can be shown that
%\begin{equation}
%\label{eq:gauss:width}
%\lips_1(\XK^{\circ}) =  (2\pi)^{-1/2}\Ee(h_{\XK^{\circ}}(\epsilon)|X) %= (2\pi)^{-1/2}\Ee(h_{K^{\circ}}(X^T\zeta)).
%\end{equation} 
%where $\epsilon |X \sim N(0, I_{n \times n})$. 
%
%For $D=\XK$ a smooth domain, that is convex set with non-empty interior bounded by a smooth hypersurface
%and for $j \leq n-1$
%\begin{equation}
%\label{eq:intrinsic:volumes}
%\lips_j(D)  \propto \int_{\partial D}P_{p-j-1}(\lambda_{1,x}, \dots, \lambda_{p-1,x}) \; \text{Vol}_{\partial D}(dx)
%\end{equation} 
%with $P_j$ the $j$-th elementary symmetric polynomial of the so-called
%principal curvatures of $\partial D$ at $x$ \citep{rfg,noniso}. These are just the eigenvalues of the second fundamental form in the unit inward normal direction. This formula can be derived
%by considering the inverse of the metric projection map eqref{eq:metric:proj}. The inverse takes $(x, \eta_x)$ defined on the extended outward normal bundle of
%$K^{\circ}$. The inverse of the map is simply the exponential map restricted to the outward normal bundle, or, more simply
%\begin{equation}
%\label{eq:exponential}
%x \mapsto x + \eta_x.
%\end{equation}
%A fairly straightforward calculation
%%\footnote{Weyl instructs in \citep{weyl} that this
%%is something any student of calculus should be able to do.} 
% yields the relation eqref{eq:intrinsic:volumes}.
%
%The main take away message above is that the functionals in the tube formula, such as the Gaussian width
%eqref{eq:width}, are related to properties of the metric projection map onto $\XK$.
%
%
%As written above, the intrinsic volumes are defined implicitly through a volume calculation, and it is not clear that they
%extend to the infinite dimensional setting. Under the right conditions of course, such extensions are indeed possible. See \citep{vitale} for a nice discussion of this problem. An alternative definition of intrinsic volumes specific to the Gaussian case
%was considered in \citep{taylor:vadlamani}, which were defined as coefficients in an expansion of the {\em Gaussian} measure of 
%$\tube(\XK^{\circ},r)$ as described in the Gaussian Kinematic Formula \citep{rfg,taylor:kinematic}.
%
%\subsection{Risk estimation}
%
%Another quantity of interest for our problem eqref{eq:canonical} is an estimate of
%how much ``fitting'' we are performing, as a function of $\lambda$. One quantitative
%measure of this is captured by Stein's estimate of risk  \citep{stein}, also known as
%SURE. Suppose now that $y \sim N(\mu, I)$ and we estimate $\mu$ by the estimator
%$\hat{\mu}$. The SURE estimate is an unbiased estimate of 
%$$\text{Risk}(\hat{\mu}) = \Ee(\|\mu - \hat{\mu}\|^2_2).$$ 
%The estimated degrees of freedom of this estimator is one part of the SURE estimate and is defined as
%\begin{equation}
%\widehat{\text{Cov}(y, \hat{\mu}}) = \text{div}(\nabla \hat{\mu}(y)).
%\end{equation}
%%As an aside, the proof that this is an unbiased estimate of risk is an application of integration by parts, something
%%that again, any student of calculus should be able to follow. This integration by parts formula is key to application
%%of Stein's method to prove central limit theorems. It is also at the heart of 
%%Malliavin calculus \citep{malliavin} via the representation of \^Ito integrals as divergence of vector fields. 
%%It seems, as in Weyl's proof
%%of the intrinsic nature of the intrinsic volumes, that it may have been asking too much to expect any student of calculus to see these deep connections.
%
%Suppose $y \sim N(\mu, I)$ and consider our residual form of the original estimation problem for $\mu=\Ee(y)$
%\begin{equation}
%\hat{\mu}(y) = y - \pi_{\lambda \XK}(y).
%\end{equation}
%If $K$ possesses a nice stratification, as do all the examples mentioned above, then,  for almost every $y$, $\pi_{\lambda \XK}(y)$ is in the relative interior
%of  some fixed stratum ${\cal S}$ of the normal bundle of $\XK$ over which the dimension
%of the tangent space is constant, and the normal bundle has a locally conic structure
%${\cal S} = {\cal T} \times {\cal N}$ of tangent and normal directions  \citep{schneider,rfg}. 
%Having fixed this stratum, we can write
%$y=x+\eta_x$ in  (ortho)normal coordinates centered at $(y-\hat{\mu}(y), \hat{\mu}(y))$. In these
%coordinates
%\begin{equation}
%\hat{\mu}(y(x,\eta_x)) = \eta_x.
%\end{equation}
%In order to relate the above to the problem eqref{eq:canonical}, one should
%invert the above chart to find $(x,\eta_x)$ in terms of $y$. 
%In the residual form  eqref{eq:resid}  it is easy to show that
%$$
%\begin{aligned}
%\left(x(y), \eta_x(y) \right) &= \left( \argmin_{r \in \lambda \cdot \XK} \frac{1}{2} \|y-r\|^2_2, y -  \argmin_{r \in \lambda \cdot \XK} \frac{1}{2} \|y-r\|^2_2 \right) \\
%&= (y - \hat{\mu}(y), \hat{\mu}(y)).
%\end{aligned}
%$$
%The derivatives along the normal directions yield 
%a purely dimensional term, while directions in the tangent directions yield
%curvature terms. This observation is enough to derive the following form of the degrees of freedom
%\begin{equation}
%\label{eq:SURE}
%\begin{aligned}
%\text{div}(\nabla \hat{\mu}(y))% &= n - \text{dim}({\cal T}) + \text{Tr}(S_{(x,\eta_x)})  \\
%&= 
% n - \text{dim}({\cal T}_y) + \text{Tr}(S_{(y - \hat{\mu}(y),\hat{\mu}(y))}).
% \end{aligned}
%\end{equation}
%Above, ${\cal T}_y$ is the tangential part of the stratum containing $x(y)$ and $S_{(x,\mu_x)}$ is the second fundamental form of 
%${\cal T}_y$ in $\lambda \cdot \XK$ as described \citep{rfg}. When $K$ is a polyhedral set, the second term disappears
%and the degrees of freedom can be computed by computing the rank of a certain matrix \citep{tibshirani_degrees_2012,bien:hierarchical}
%%
%%\begin{lemma}[SURE for the canonical problem]
%%Consider the
%%canonical problem in residual form eqref{eq:resid}. The 
%%SURE estimate of the degrees of freedom of this estimator is 
%%\begin{equation}
%%\label{eq:SURE}
%%\begin{aligned}
%%\text{div}(\nabla \hat{\mu}(y))% &= n - \text{dim}({\cal T}) + \text{Tr}(S_{(x,\eta_x)})  \\
%%&= 
%% n - \text{dim}({\cal T}_y) + \text{Tr}(S_{(y - \hat{\mu}(y),\hat{\mu}(y))}).
%% \end{aligned}
%%\end{equation}
%%where ${\cal T}_y$ is the tangent part of the stratum and $S_{(x,\mu_x)}$ is the second fundamental form of 
%%${\cal T}_y$ in $\lambda \cdot \XK$ as described \citep{rfg}.
%%\end{lemma}
%%{\bf Proof:}
%%Locally, as the estimate is in the fixed stratum with tangent component ${\cal T} = {\cal T}_y$ and
%%normal component ${\cal N} = {\cal N}_y$ the estimation procedure can be written as
%%$$
%%\begin{aligned}
%%\hat{\mu}(x+\eta_x) &= \eta_x \\
%%&= \sum_{j=1}^{n - \text{dim}({\cal T}_y)} a_j \eta_{j,x} \\
%%\end{aligned}
%%$$
%%where $a \in \real^{n - \text{dim}({\cal T}_x)}$ and $\eta_{j,x}$ form an orthonormal
%%frame for the space normal to ${\cal T}_x$. In fact, the $a_j$'s must be such that
%%$\eta_x \in N_x( \lambda \cdot \XK)$ but we can assume they are generic points in the relative interior.
%%Differentiating with respect to $a$ and summing yields $n - \text{dim}({\cal T}_x)$ while differentiating
%%with respect to an orthonormal basis for ${\cal T}_x$ yields $\text{Tr}(S_{x,\mu_x})$. Substituting
%%the parametrization $(y-\hat{\mu}(y), \hat{\mu}(y))=(x,\eta_x)$ concludes the proof. \qed
%
%\subsection{Hypothesis testing: weak convergence for high-dimensional inference?}
%
%Another fundamental tool in inference for least squares models is the ability to form hypothesis
%tests, as well as confidence intervals for the ``true'' mean. Such concepts
%clearly need a model, which we might take to be the usual model
%$$
%y \sim N(\mu,I).
%$$ Under the assumption that $\mu=X\beta_0$, classical inference in 
%linear models (assuming $X^TX$ is full rank) yields confidence intervals and hypothesis tests for any linear functional $\nu^T\mu$ based on the
%coefficients $\beta_0$. 
%
%What can we say about our canonical problem eqref{eq:canonical}? This is an area in which we still don't know all the answers. In some sense, we are in the situation analogous to Bernoulli having proved a weak 
%law of large numbers without the central limit theorem.
%Some progress  has been made for specific models of the design matrix $X$  for the  LASSO as well as group LASSO models, see \citep{meinshausen:buhlmann,buhlmann,minnier:cai,murphy:laber,wasserman:roeder,donoho:montanari}.
%
%Other recent work \citep{significance:lasso} gives some hints at what inferential tools may prove useful in this weak convergence theory. As described in the LARS algorithm \citep{lars,tibshirani:unique}, an entire path of solutions $\hat{\beta}_{\lambda,\XK}$ can be formed for the LASSO, i.e. when $K=B_{\infty}$. These paths are piecewise linear, with knots at points where the {\em active set} changes. The covariance statistic
%measures some change in the correlation of the fitted values between two knots $\lambda_k$ and $\lambda_{k+1}$ in the LASSO path. It has the form
%\begin{equation}
%\label{eq:covstat}
%T_k = C_k \lambda_k (\lambda_k - \lambda_{k+1})
%\end{equation}
%for some random scaling $C_k$ related to the active set and the variable added to the active set at $\lambda_{k+1}$. 
%The form for $k=1$ is particularly simple: suppose that $\hat{j}_1$ and $X$ is such the first variable in the LARS path, i.e.
%$$
%\hat{j}_1  \in \argmax_j |X_j^Ty|.
%$$
%Then,
%\begin{equation}
%\label{eq:covstat1}
%T_1 = \frac{1}{\sigma^2} y^TX\hat{\beta}_{\lambda_2} = y^T X_{\hat{j}_1} \hat{\beta}_{\lambda_2,\hat{j}_1}.
%\end{equation}
%For $k \geq 1$, the form of the test statistic is slightly more complicated, though it can be expressed in terms of 
%$A=A_k$, the active set at step $k$ as well as $s_{A_k}$, the signs of the active variables at step $k$ as well as 
%the active set $A_{k+1}$ and $s_{A_{k+1}}$, see \S 2.3 of \cite{significance:lasso} for the full expression.
%For a wide range of (sequences) covariance matrices, if the active set at $\lambda_k$ already contains all the strong active variables, then
%it is shown in \citep{significance:lasso} that
%\begin{equation}
%\label{eq:exp1}
%T_k \overset{D}{\rightarrow} \text{Exp}(1).
%\end{equation}
%In particular, under the global null $y \sim N(0,\sigma^2 I)$ as long as the design matrix satisfies some minimum growth condition, 
%$T_1 \overset{D}{\rightarrow} \text{Exp}(1)$.
% The main tools used in the above proof relate to the maxima of (discrete) Gaussian processes and the generalization of an
% argument previously applied to smooth Gaussian processes in \citet{taylor:validity} and \citet{rfg}.
% Ongoing work suggests that such a limiting distribution will hold for many (sequences) of $K$ and design matrices $X$.
%
%As for confidence intervals for the parameters related to $k$ strong variables, the relation to extreme values of Gaussian processes suggest that the bias-corrected or relaxed LASSO estimate of the active coefficients will have accurate coverage. This is ongoing work.
% 
%%However, there are some special cases for which some results are known. Specifically, 
%%suppose $K$ is a cone. Then eqref{eq:canonical} is a cone-restricted regression problem with the restriction being
%%$$
%%h_K(\beta) < \infty \iff \beta \in K^{\circ}
%%$$
%%where $K^{\circ}$ is the polar cone of $K$.
%%
%%Consider the test of $H_0:\beta=0$ versus the unrestricted alternative $H_a:\beta \in K^{\circ}$. 
%%The likelihood ratio test statistic \citep{takemura,kuriki,self,chernoff,perlman} for these hypotheses
%%is, perhaps not surprising to the readers at this point, related to the metric projection problem. Specifically, an inspection of the canonical problem in residual form eqref{eq:residual}
%%shows that the log-likelihood ratio test statistic  is given by
%%$$
%%\Lambda(H_0, H_a) = \|\pi_{\XK}(y)\|^2_2
%%$$
%%and we see the first tool in our toolbox  appearing again. 
%%
%%If $S(K,X) = \XK \cap S(\real^n)$ are the set of unit vectors in $\XK$, it is not
%%difficult to show that
%%\begin{equation}
%%\|\pi_{\XK}(y)\|_2 = \sup_{\nu \in S(K,X)} (\nu^Ty)^+
%%\end{equation}
%%and the log-likelihood ratio test statistic can be expressed as the supremum of a 
%% process indexed by $S(K,X)$. We see, then, that under $H_0$, it is the supremum of a unit variance
%% Gaussian process. Recalling our earlier discussion on Gaussian width, we know that there
%% is a metric projection hiding somewhere in the distribution of the supremum of this process.
%% We will see that this is indeed the case in the following section.
%% 
%
% 
%% Before continuing along the above path, we discuss the possibility of a central limit theorem
%% for functionals of $\mu$ in eqref{eq:canonical}. One of the fundamental questions that must be 
%% answered is what functionals of $\mu$ is it natural to consider? It is clear, and well known, that even if 
%% $y\sim N(X\beta_0,I)$ the estimator $\hat{\beta}_{\lambda,\XK}$ is providing
%% biased estimates of $\beta_0$. 
%% 
%% The natural question to ask is ``what functional of $\mu$ is $\hat{\beta}_{\lambda, \XK}$ estimating''? 
%% Writing $\hat{\beta}_{\lambda, \XK}(y,X)$ explicitly as a function of $y$ and $X$ we might suppose that
%% $\hat{\beta}_{\lambda, \XK}(y,X)$ is estimating 
%% \begin{equation}
%% \label{eq:estimating}
%%\hat{\beta}_{\lambda,\XK}(\mu,X) =\argmin_{\beta \in \real^p} \frac{1}{2} \|\mu-X\beta\|^2_2+ \lambda h_{K}(\beta).
%% \end{equation}
%% It remains to be seen whether this notion be used to establish a general weak convergence result for eqref{eq:canonical}.
%% 
% 
%\section{Smooth Gaussian processes: relaxing convexity}
%\label{sec:gaussian}
%
%In the special case that $K$ is a cone, we saw that the distribution of a particular 
%likelihood ratio test could be expressed in terms of the supremum of a Gaussian process indexed by a subset of the sphere. 
%Equivalently, this supremum could be expressed in terms of the metric projection onto the cone
%$\XK$. It is well-known that the distribution of this likelihood ratio test statistic is a mixture
%of $\chi^2$'s of varying degrees of freedom. This distribution is sometimes referred to as a  $\bar{\chi}_2$ distribution.
%
%The mixture weights can be expressed in terms of the geometry of the set
%$M = S(X,K) = \XK \cap S(\real^n)$. In particular, it is known \citep{rfg,takemura:kuriki1,takemura:kuriki2} that if 
%$\epsilon \sim N(0, I)$ for $u > 0$
%\begin{equation}
%\label{eq:sup}
%\Pp\left(\sup_{\nu \in M}( \nu^T \epsilon)^+ > u \right) =
%\sum_{j=0}^{\infty} \lips_j(M) \rho_j(u),
%\end{equation}
%where
%\begin{equation}
%\label{eq:rho}
%\rho_j(u) = \begin{cases}
%\int_u^{\infty} \frac{1}{\sqrt{2\pi}} e^{-t^2/2} \; dt  & j = 0 \\
%(2\pi)^{-(j+1)/2} H_{j-1}(u) e^{-u^2/2}& j  \geq 1,
%\end{cases}
%\end{equation}
%and$$
%H_j(u) =(-1)^j \frac{\partial ^j}{\partial u^j} e^{-u^2/2}
%$$
%are the standard Hermite polynomials. The functions eqref{eq:rho}
%are known as the EC or Euler characteristic densities for a
%Gaussian field, see \citep{rfg,takemura:kuriki2,worsley:boundary,worsley:unified1,worsley:unified2}.
%While the sum above is written as an infinite sum it terminates at $\text{dim}(M)$.
%
%Note that this implies
%$$
%\Ee \left(\sup_{\nu \in M} (\nu^T \epsilon)^+  \right) = \frac{1}{\sqrt{8\pi}} \lips_1(M)
%$$
%which is an analogous way to derive Gaussian width eqref{eq:gauss:width}.
%
%One of the derivations of the above formula, the so-called {\em volume of tubes} approach \citep{siegmund,sun,takemura:kuriki1} involves studying the Jacobian of the inverse of the
%spherical metric projection map eqref{eq:metric:proj:sphere}, i.e. the exponential
%map on $S(\real^n)$ which sends a pair $(x,\eta_x)$ to
%$$
%\cos(\|\eta_x\|) \cdot x + \sin (\|\eta_x\|) \cdot \eta_x.
%$$
%Another approach, the expected Euler characteristic approach
%\citep{rfg,worsley:boundary} involves counting critical points of 
%the Gaussian process above the level $u$ according to saddle type and applying
%Morse theory and the Rice-Kac formula \citep{azais,rfg} to count the expected number of such points.
%
%Neither of these approaches strictly require convexity of the cone generated by the parameter set $M$. Rather,
%they depend on a notion of local or infinitesimal convexity referred to as positive 
%reach \citep{federer}. Hence, the parameter sets may have finite critical radius. 
%They are both approaches used to form an approximation
%\begin{equation}
%\label{eq:sup:generic}
%\Pp\left(\sup_{\nu \in M} f(\nu) > u\right) \approxeq
%\sum_{j=0}^{\infty} \lips_j(M) \rho_j(u)
%\end{equation}
%for some centered, smooth, Gaussian process $f$ having constant variance 1 on a
%(possibly stratified) manifold $M$.
%In the volume of tubes approach, the critical radius appears in a natural way and enters into an estimate of the error of the volume of tubes approach. In both approaches, though it is clearer in  the expected Euler characteristic approach, the spherical critical radius is in fact the spherical critical radius of 
%$\Psi(M)$ where $\Psi:M \rightarrow S(H)$ where $S(H)$ is the unit sphere in $H$, the reproducing kernel Hilbert space of 
%$f$. Either approach yields roughly the same estimate of error: for $u$ large enough
%\begin{equation}
%\label{eq:sup:generic:tail}
%\liminf_{u \rightarrow \infty} -\frac{2}{u^2} \left|\Pp\left(\sup_{\nu \in M} f(\nu) > u\right) -
%\sum_{j=0}^{\infty} \lips_j(M) \rho_j(u) \right|= 1 + \tan^2 (r_c(M)).
%\end{equation}
%The above says that the {\em relative error} in the approximation is exponentially small whenever $r_c(M) > 0$.
%
%The critical radius of $M$ in the expected Euler characteristic approach arises in terms of 
%a functional of the original process $f$. Specifically, if we assume $M$ is a manifold without boundary, then define for each $x\neq y$ the process introduced in \citep{taylor:validity}
%\begin{equation}
%\label{eq:fxy}
%f^x(y)= \frac{f(y) - \Ee(f(y)|f(x), \nabla f(x))}{1 - \Ee(f(x) \cdot f(y))}.
%\end{equation}
%Then,
%\begin{equation}
%\label{eq:fxy2}
%\cot^2(r_c(M)) = \sup_{x \neq y} \Ee(f^x(y)^2).
%\end{equation}
%Hence, the critical radius depends in an explicit way on the covariance function of the 
%process $f$. 
%
%As mentioned previously, the argument related to derivation of eqref{eq:fxy2} in the smooth case led directly to the
%exponential limit in eqref{eq:exp1}. Such a connection suggests a relation between the
%distribution of the maxima of smooth Gaussian processes, specifically the spacings of the extreme values, can be 
%used to derive weak convergence results for high-dimensional inference. This is ongoing work.
%%
%%We might ask the question: does this critical radius manifest itself in any
%%sampling sense? That is, can we see this critical radius from observing
%%independent samples of $f$? The answer to this is affirmative.
%%In \citep{taylor:weinberger:adler} we consider the random embedding of $M$ given by
%%$\tilde{M}_k=f^k(M)$ where
%%$$
%%f^k(\nu) = (f_1(\nu), \dots, f_k(\nu)), \qquad f_j \overset{IID}{\sim} f
%%$$
%%and $k  \geq 2 \text{dim} + 1$ to ensure the image is generically an embedding. The set
%%$y$ itself has a metric projection map
%%$$
%%y \mapsto \pi_{\tilde{M}_k}(y)
%%$$
%%which is indeed a function on $\tube(\tilde{M}_k,r)$ for $r < r_c(\tilde{M}_k)$. 
%%
%%The size of $r_c(\tilde{M}_k)$ is related to the number of samples needed to recover the 
%%homology of $M$ from uniformly sampling points on $\tilde{M}_k$ \citep{NSW}. This is one of the 
%%more quantitative results in the nascent field of {\em topological data analysis}. But what are the
%%sampling properties of the random variable $\tilde{M}_k$? Not surprisingly, a
%%law of large numbers holds \citep{taylor:weinbergerg:adler} and $\tilde{M}_k \overset{\Pp}{\rightarrow} r_c(M)$ where $r_c(M)$ is the previously mentioned spherical critical radius in $H$.
%%Once again, we are at the point Bernoulli reached in {\em Ars  Conjectandi} and look to see if there
%%is a Gaussian, or CLT, analogue to this weak law.
%%
%
%\section{Conclusion}
%
%We have described what we call the two basic tools of the geometry of least squares
%that are just as relevant as when Gauss and Legendre disputed their original discovery
%over 200 years ago. While these tools are not the most technically sophisticated tools, 
%ceding that ground to exponential families for the canonical model eqref{eq:canonical} and 
%empirical processes for the fluctuation theory in \S \ref{sec:inference}, they nevertheless
%provide guiding principles for these more precise tools. We would argue that the Gaussian
% picture provided by the geometry of least squares, gets much of the picture correct under sufficient moment conditions. For heavier tailed results, of course much of \S \ref{sec:inference}
% would have to be reframed and \S \ref{sec:gaussian} paints quite a different picture
% \citep{samorodnitsky1,samorodnitsky2}.
% 
% As Bernoulli found himself with just a law of large  numbers, the field of statistics is roughly at this same stage in high-dimensional inference. We are hopeful that the geometry of least squares will eventually guide the field to weak convergence results in high-dimensional inference
% for the canonical problem eqref{eq:canonical}.

\bibliographystyle{ims}
%\bibliographystyle{acmtrans-ims}
\bibliography{paper}

\end{document}

### 2013-09-24 paper/paper.tex
Editing paper

@@
 % "Title of the paper"
-\title{Significance testing and forward stepwise with the group lasso}
+\title{Significance testing with forward stepwise and grouped forward
+  stepwise model selection}
 
@@
 group lasso \citep{grouplasso} and other adaptive regression
-problems.  The present work
-explores the behavior of those test statistics for models selected by
-forward stepwise procedures and works out some of the details
-involved in applying these methods to models with grouped variables.
-We find that our test statistics can be used for valid significance
-testing, or for tempering the aggressive nature of forward stepwise
-model selection.  The resulting methods may be less computationally
-intensive than choosing a regularization parameter for the group lasso
-by cross-validation.
+problems.  The present work explores the behavior of those test
+statistics for models selected by forward stepwise procedures and
+works out some of the detailsinvolved in applying these methods to
+models with grouped variables. We find that our test statistics can be
+used for valid significance testing, or for tempering the aggressive
+nature of forward stepwise model selection.  The resulting methods may
+be less computationally intensive than choosing a regularization
+parameter for the (group) lasso by cross-validation.
+
+\todo Write about more differences/advantages of this paper- maybe not
+in this section, but: using non-asymptotic p-values, forward step is
+simple(r than lasso), more?
 
@@
 
+\todo Fix references below
+
 The group lasso estimator is discussed in (ref that guy's thesis) and
@@
 \label{sec:testing}
+
+\todo Update this section to match/complement \citep{tests:adaptive}
 
@@
 
-\emph{(For this section I have a lot of stuff written that can just be pasted in once I'm done wrangling the KKT conditions for the unequal weight case)}
+\todo Modify write-up to match notation with \cite{tests:adaptive} and
+then paste it in here
 
 \subsection{Better p-values}
+
+Maybe just reference \cite{tests:adaptive}? It might be worthwhile to
+leave in the connection with \cite{significance:lasso} (getting our
+test statistic by not using the approximation)
 
@@
 \item \todo Convert to exposition instead of bulleted list
+\item \todo Update to reflect latest work
   \item As in LTTT, $T = \lambda_1 (\lambda_1 - M)$ and $M = \lambda_2$
@@
 
-\subsection{Can we do $\lambda_k$? Don't know yet}
-
 \section{Forward stepwise procedures}
-\todo This is potentially the second largest thing not done yet.
-However, if it turns out we can't think of a better stopping rule, then this section is done and just needs to be written (and would not take long).
-
-\section{Simulations }
-(I've done many of these and have a lot of code, so I can produce a lot of output for this section)
-\todo Modify some of my previous simulations after deciding on a stopping rule, decide which simulations to put here.
+
+\todo Expand this section
+
+\begin{itemize}
+  \item Brief background, cite some original papers
+  \item Describe exactly how we do each step (projection, residuals, etc) 
+  \item Stopping rules, cite new paper on lasso, perhaps we have
+    a computation advantage
+\end{itemize}
+
+\section{Simulations}
+
+\todo Expand, include simulation results (copy latex for
+includegraphics etc), do the simulation comparing stopping rules
+
+\begin{itemize}
+  \item Show null and non-null p-value-by-step plots for several examples
+  \item Discuss one of these carefully (perhaps with a noise variable
+    entering before a signal variable), for several fixed stopping points
+  \item Compare a few stopping rules, e.g. naive one(s),
+    TailStop/HybridStop, AIC/BIC/Cp
+  \item (the stopping comparisons are not done yet, but I am guessing
+    they will be comparable, perhaps AIC/BIC/Cp will be better, and in
+    that case look for other advantages of ours to mention,
+    e.g. interpretability for being based on p-values rather than a
+    complexity penalty)
+
+\end{itemize}
 
 \section{Real data example}
-\todo This is definitely the single largest part I have not done yet.
+\todo Simulation with non-Gaussian design matrix (from AIDS data probably)
+
+\section{Discussion}
+
+\todo After finising everything else, move some points of discussion
+here, re-read and see if there's anything else interesting to comment
+on here. Mention ongoing work
 

### 2013-09-25 paper/paper.tex
Added first small simulation figure

@@
 \usepackage{graphics,amssymb,color}
+\usepackage{graphicx,subfigure}
 
@@
 
+\begin{figure}
+\begin{center}
+\subfigure[Our test]{
+\label{fig:lassopvala}
+\includegraphics[width=0.5\textwidth]{../figs/small_sim_plot.pdf}}
+\hspace{-15pt}
+\subfigure[Covariance test]{
+\label{fig:lassopvalb}
+\includegraphics[width=0.5\textwidth]{../figs/small_sim_plot.pdf}}
+\caption{\small \it The left panel shows the empirical distribution
+  function of a sample of 20,000 p-values coming
+  from a variety of different lasso setups.  The agreement with
+  uniform here is excellent.  The right panel shows the empirical
+  distribution function of a sample of 10,000 covariance test
+  p-values, computed using an $\mathrm{Exp}(1)$ approximation,
+  using three different lasso setups.  The
+  $\mathrm{Exp}(1)$ approximation is generally conservative.}
+\end{center}
+\end{figure}
+
 \section{Real data example}

### 2013-09-25 fwd_step.R

@@ -22,7 +22,7 @@ add_group = function(X, Y, groups, weights, sigma, active.set = 0, eff.p = 0) {
   new.eff.p = eff.p + sum(gmax)
 
   if (new.eff.p >= n) {
-    stop("Too many variables added")
+    stop("Too many variables added. Try decreasing max.steps")
   }
   
   p.value = pvalue(results$L, results$Mplus, results$Mminus, sqrt(results$var), results$k, sigma=sigma)
@@ -44,7 +44,6 @@ add_group = function(X, Y, groups, weights, sigma, active.set = 0, eff.p = 0) {
 
   return(list(test.output = results, p.value = p.value, added = imax, active.set = new.active.set, eff.p = new.eff.p, Y.update = Y.resid, X.update = X.project))
 }
-
 
 
 
@@ -88,9 +87,9 @@ forward_group = function(X, Y, groups, weights = 0, sigma = 0, max.steps = 0) {
     # Some overfitting considerations
     if ((eff.p >= n - max(group.sizes)) & (i < max.steps)) {
       if (eff.p > n - mean(group.sizes)) {
-        warning("Overfitting likely to occur in next step")
+        warning("Overfitting imminent!")
       } else {
-        warning("Overfitting may occur in next step")
+        warning("Overfitting may occur soon")
       }
     }
     

### 2013-09-26 paper/paper.tex
Paper: began appendix with derivation

@@
 \newcommand{\XK}{K_X}
-\def\qed{\hfill $\Box$\newline}
+\newcommand{\norm}[1]{\lVert #1 \rVert}
+\newcommand{\innerp}[2]{\langle #1 , #2 \rangle}
+%\newcommand{\qed}{\hfill $\Box$\newline}
 
@@
 
+\appendix
+\section{Derivation of closed forms of statistic}
+
+Write $M^{\pm}_{\eta, h}$ and $M^0_{\eta, h}$ as the corresponding suprema and infima over the restricted parameter set $S_h$.
+Note that the characterization of the global maximizer (Lemma 1, general paper) implies $M^0_{\eta, h} \leq 0$ and $M^+_{\eta, h} \leq M^-_{\eta, h}$ for all $h \neq g$.
+This will be used below to eliminate some degenerate cases of the optimization sub-problem on each $S_h$.
+
+Now fix $h$, let $P_g = X_gX_g^T$ and define
+\begin{align*}
+a_h &= w_h^{-1} X_h^T (I-P_g) y \\
+b_h &= w_g w_h^{-1} X_h^T P_g y / \norm{X_g^T y} \\
+c_h &= a_h^T b_h / (\norm{a_h} \norm{b_h}) \\
+K_h   &= \{ x : \norm{x} = 1, b_h^Tx > w \}
+\end{align*}
+Note that $K_h = \emptyset$ if and only if $\norm{b_h} < w$.
+We will consider two cases.
+First, suppose $\norm{b_h} > w$.
+In this case we must rule out one sub-case, when $a_h$ is not in the polar cone of $K_h$.
+If this occurs, then on at least one side of the boundary of $K_h$ we have $a_h^Tx > 0$ there, so that the infimum inside $K$ is $-\infty$ but the supremum outside $K$ is $+\infty$.
+This violates the characterization of the global maximizer of the original process as described above, so this case cannot occur.
+
+Now, if $a_h$ is in the polar cone of $K_h$ (and ruling out the probability zero case where $a_h^Tx = 0$ on the boundary of $K_h$), then the numerator $a_h^Tx < 0$ on $K_h$ so there is a finite positive infimum in the interior $K_h$.
+Similarly there is a finite positive supremum on the interior of $K_h^c$ (the numerator is negative near the boundary of $K_h$ by continuity).
+
+To attain the infimum on $K_h$ requires making $x$ close to $b_h$ and simultaneously making the numerator closer to zero, so $x$ should be on the side of $b_h$ that is closer to $a_h$.
+To attain the supremum on $K_h^c$ requires making $x$ simultaneously close to $a_h$ (so that $a_h^Tx > 0$) and $b_h$ (to make the denominator small).
+In summary, both the infimum and supremum are attained with $x$ between $a_h$ and $b_h$, when the angle $\theta$ between $a_h$ and $b_h$ is larger than both the angle $\psi$ between $a_h$ and $x$ and the angle $\phi$ between $x$ and $b_h$.
+This allows the simplification $\phi = \theta - \psi$.
+This is easier to verify for the case when $\norm{b_h} < w$.
+
+We have verified that solving both cases requires finding the maxima of the trigonometric form $\norm{a_h} \cos (\psi) / (w - \norm{b_h} \cos(\theta - \psi))$ of the linear fraction on the interiors of $K_h$ (if it is nonempty) and $K_h^c$.
+The derivative is proportional to
+\[
+\frac{\norm{b_h} \sin(\theta) - w \sin(\psi)}{w - \norm{b_h} \cos(\theta - \psi)}
+\]
+Since we have ruled out the boundary of $K_h$, we can ignore the denominator and focus on critical points corresponding to the angles $\psi^{\pm}$ (symmetric about $\pi/2$) where $\sin(\psi^{\pm}) = (\norm{b_h} / w) \sin (\theta)$.
+That is,
+\[
+\psi^{\pm} \in \arcsin \frac{\norm{b_h}}{w} \sqrt{1-c_h^2}
+\]
+Note that these critical points only exist if $\sin(\theta) < w/\norm{b_h}$, and since we have argued that the infimum and supremum are attained it follows that this condition is met in our case.
+Let $\psi^+$ denote the smaller of the two angles.
+If $K_h$ is nonempty, then $\psi^-$ gives the infimum in $K_h$ and $\psi^+$ the supremum on $K_h^c$.
+This is necessary since if $x$ is in $K_h$ then its angle from $a_h$ is larger than $\pi/2$ because $a_h$ is in the polar cone of $K_h$.
+If $K_h$ is empty, then the supremum is still attained at $\psi^+$ since the numerator is negative at $\psi^- > \pi/2$ and the denominator is always positive.
+
+Finally we calculate the values of the linear fraction at $\psi^{\pm}$. Let us first record some facts:
+\begin{align*}
+0 < \psi^+ < \pi/2, \quad \psi^- = \pi - \psi^+ \\
+\cos(\theta) = c_h, \quad \sin(\theta) = \sqrt{1-c_h^2} \\
+\sin(\psi^+) = \sin(\psi^-) = \frac{\norm{b_h}}{w} \sin(\theta) \\
+\cos(\psi^+) = -\cos(\psi^-) = \sqrt{1 - (\norm{b_h}/w)^2(1-c_h^2)}
+\end{align*}
+Using the angle difference formula,
+\begin{align*}
+\cos(\theta - \psi^+) & = \frac{\norm{b_h}}{w} (1-c_h^2) +  c_h \cos(\psi^+) \\
+& = \frac{\norm{b_h}}{w} (1-c_h^2) +  c_h \cos(\psi^+)
+\end{align*}
+Hence
+\begin{align*}
+M^+_{\eta, h} & = \frac{ \norm{a_h} c_h }{ w - (\norm{b_h}^2/w) (1-c_h^2) - \norm{b_h} c_h \cos(\psi^+) } \\
+% & = 
+% & = \frac{ w \norm{a} a_h^Tb_h / \norm{b_h} }{ w^2 - \norm{a}\norm{b}^2 + (a^Tb)/\norm{a} - a^Tb \cos \psi^+ }
+\end{align*}
+And $M^-_{\eta, h}$ is $\infty$ when $\norm{b_h} < w$ and otherwise given by
+\begin{align*}
+M^+_{\eta, h} & = \frac{ \norm{a_h} c_h }{ w - (\norm{b_h}^2/w) (1-c_h^2) + \norm{b_h} c_h \cos(\psi^+) } \\
+\end{align*}
+
+I haven't been able to algebraically simplify these yet to a form that allows my $\lambda_2$ proof below to go through. The version below just assumes $\norm{b_h} < w = 1$.
+
+\newpage
+We can rewrite this by rationalizing the denominator:
+\[
+M =
+\frac
+{ a_h^T b_h +  \sqrt{ a_h^T a_h (1-b_h^Tb_h) + (a_h^T b_h)^2 }  }
+{ 1-b_h^Tb_h }
+\]
+Leaving $M$ in this form we now consider $\lambda_2$. The KKT conditions for the group lasso problem give (directly from Yuan and Lin)
+\begin{align*}
+\norm{ X_g^T(y - X \beta) } & \leq \lambda w_g \quad \forall \beta_g \neq 0 \\
+\beta_g & = \left( 1 - \frac{ \lambda w_g }{ \norm{S_g} } \right)_+ S_g
+\end{align*}
+where $S_g = X_g^T(y - X \beta_{-g})$. These expressions simplify for $\lambda_2 \leq \lambda < \lambda_1$. Letting $g$ be the index of the first group to enter we now have $S_g = X_g^Ty$ 
+\begin{align*}
+\beta_g & = \left( 1 - \frac{ \lambda w_g }{ \norm{X_g^Ty} } \right) X_g^Ty \\
+\lambda w_h & \geq \norm{ X_h^Ty - X_h^T P_g y (1-\lambda w_g / \norm{X_g^Ty})  } \quad \forall h \neq g
+\end{align*}
+Now let $\lambda = \lambda_2$, so the inequality above is strict for all other groups and equality is attained for the second group to enter. Hence $\lambda_2$ satisfies
+\begin{align*}
+\lambda_2 & = \max_{h \neq g} 
+\norm{ X_h^T(I - P_g) y - \lambda_2 ( w_g X_h^T P_g y / \norm{X_g^Ty} )  } / w_h\\
+& = \max_{h \neq g} \norm{ a_h - \lambda_2 b_h } 
+\end{align*}
+
+Write $\lambda_{2,h}$ as the positive root of the quadratic equation obtained by squaring the norm, so $\lambda_{2,h}^2 = \norm{ a_h - \lambda_2 b_h }^2$.
+Then $\lambda_2 = \max_{h \neq g} \lambda_{2,h}$.
+Solving this (note that $b_h^Tb_h < 1$), we find
+\[
+\lambda_{2,h} = \frac
+{ a_h^T b_h +  \sqrt{ a_h^T a_h (1-b_h^Tb_h) + (a_h^T b_h)^2 }  }
+{ 1-b_h^Tb_h }
+\]
+Hence when $h$ is the index of the second group to enter, we have $\lambda_2 = \lambda_{2,h} = M$.
+
 \bibliographystyle{ims}

### 2013-09-27 paper/paper.tex
Added code for categorical designs

@@
 
-\endlocaldefs
+\newcommand{\V}{\mathcal{V}}\endlocaldefs
+\newcommand{\K}{\mathcal{K}}
 
@@
 
+For the sake of completeness this appendix provides full derivations
+of closed forms for the quantities required to compute our
+statistic. We require two facts from \cite{tests:adaptive}
+
+A point $\eta \in \K$ maximizes $f_\eta$ over a convex
+set $\K$ if and only if the following conditions hold: 
+\begin{equation}
+\label{eq:maxcond}
+\grad f_{|T_{\eta}\K} = 0, \quad
+\tf^{\eta}_{\eta} \geq \V^+_{\eta}, \quad
+\tf^{\eta}_{\eta} \leq \V^-_{\eta}, \quad \text{and} \quad
+\V^0_{\eta} \leq 0.
+\end{equation}
+The same equivalence holds true even when $\K$ is only locally
+convex.
+
 Write $M^{\pm}_{\eta, h}$ and $M^0_{\eta, h}$ as the corresponding suprema and infima over the restricted parameter set $S_h$.

### 2013-09-28 paper/paper.tex
More model selection code, more paper editing

@@
 
+\newcommand{\grad}{\nabla}
 \newcommand{\V}{\mathcal{V}}\endlocaldefs
 \newcommand{\K}{\mathcal{K}}
+\newcommand\tf{\widetilde{f}}
 
@@
 \begin{center}
-\subfigure[Our test]{
+\subfigure[Gaussian design]{
 \label{fig:lassopvala}
-\includegraphics[width=0.5\textwidth]{../figs/small_sim_plot.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/small_sim_gaussian_design.pdf}}
 \hspace{-15pt}
-\subfigure[Covariance test]{
+\subfigure[Categorical design]{
 \label{fig:lassopvalb}
-\includegraphics[width=0.5\textwidth]{../figs/small_sim_plot.pdf}}
-\caption{\small \it The left panel shows the empirical distribution
-  function of a sample of 20,000 p-values coming
-  from a variety of different lasso setups.  The agreement with
-  uniform here is excellent.  The right panel shows the empirical
-  distribution function of a sample of 10,000 covariance test
-  p-values, computed using an $\mathrm{Exp}(1)$ approximation,
-  using three different lasso setups.  The
-  $\mathrm{Exp}(1)$ approximation is generally conservative.}
+\includegraphics[width=0.5\textwidth]{../figs/small_sim_categorical_design.pdf}}
+\caption{\small \it The left panel shows results of a simulation with
+  an independent Gaussian design matrix and a signal vector having
+  three groups with non-zero coefficients of sizes one, two, and
+  three, and seven groups of noise variables with sizes varying from
+  one to ten, for a total of 30 variables. For the right panel we
+  repeated this setup, but group sizes corresponded to levels of
+  categorical variables. The groups of size one were increased to two,
+so there are 10 categorical variables with a total of 32 levels.}
 \end{center}
 \end{figure}
+
+\label{fig:smallselection}
+
+\input{../small_sim_results.tex}
 
@@
 
-\newpage
+\hrule
+\hrule
 We can rewrite this by rationalizing the denominator:

### 2013-09-28 fwd_step.R

@@ -28,12 +28,17 @@ add_group = function(X, Y, groups, weights, sigma, active.set = 0, eff.p = 0) {
   p.value = pvalue(results$L, results$Mplus, results$Mminus, sqrt(results$var), results$k, sigma=sigma)
   new.active.set = update_active_set(active.set, imax)
 
-  Y.resid = lm(Y ~ X[, gmax] - 1)$residuals
-
   X.project = X
-  gmax = groups == results$i
   Xgmax = X[, gmax]
+  Xgmax.regress = Xgmax
   Pgmax = Xgmax %*% ginv(Xgmax)
+  # If X[, gmax] is categorical, leave one column out
+  if (sum(gmax) > 1) {
+    if (length(unique(rowSums(Xgmax))) == 1) {
+      Xgmax.regress = Xgmax[ , -1]
+    }
+  }
+  Y.resid = lm(Y ~ Xgmax.regress - 1)$residuals
   
   for (gind in 1:max(groups)) {
     if (gind != imax) {

### 2013-09-28 model_selection.R

@@ -6,7 +6,53 @@ source('stop_rules.R')
 # (those input p.list and alpha, output index of last inclusion)
 
 # from fwd_step_test.R results
-# names = null.p, signal.p, active.set, true.active
+
 
 # write a function taking  these inputs and calculating:
 # Rob/Alex/Max/Stefan show FDR, V, R, and average power (fraction of non-null rejected)
+
+selection_stats = function(p.list, active.set, true.step, m1, stop.rule, alpha = .1) {
+  rule_name = paste("stop_", stop.rule, sep = "")
+  if (exists(rule_name, mode = "function")) {
+    stop_fun = get(rule_name)
+    stop.ind = stop_fun(p.list, alpha)
+  } else {
+    stop(paste("Misspecified stopping rule:", stop.rule))
+  }
+#  if (stop.rule == "first") {
+#    stop.ind = stop_first(p.list, alpha)
+#  } else if (stop.rule == "forward") {
+#    stop.ind = stop_forward(p.list, alpha)
+#  } else if (stop.rule == "hybrid") {
+#    stop.ind = stop_hybrid(p.list, alpha)
+  if (stop.ind == 0) {
+    stopped.set = c()
+    S = 0
+  } else {
+    stopped.set = active.set[1:stop.ind]
+    S = sum(true.step[1:stop.ind])
+  }
+  R = stop.ind
+  V = R - S
+  Fdp = V/max(1, R)
+  Power = S/max(1, m1)
+  return(c(Fdp, R, S, V, Power))
+}
+
+# names = null.p, signal.p, active.set, true.set
+sim_select_stats = function(p.lists, active.sets, true.steps, m1, alpha = .1) {
+  first.mat = matrix(0, nrow = nrow(p.lists), ncol = 5)
+  forward.mat = first.mat
+  hybrid.mat = first.mat
+  for (i in 1:nrow(p.lists)) {
+    first.mat[i, ] = selection_stats(p.lists[i, ], active.sets[i, ], true.steps[i, ], m1, stop.rule = "first", alpha)
+    forward.mat[i, ] = selection_stats(p.lists[i, ], active.sets[i, ], true.steps[i, ], m1, stop.rule = "forward", alpha)
+    hybrid.mat[i, ] = selection_stats(p.lists[i, ], active.sets[i, ], true.steps[i, ], m1, stop.rule = "hybrid", alpha)
+  }
+  first = colMeans(first.mat)
[…]

### 2013-09-28 old/old_forward_step.R

@@ -1,0 +1,71 @@
+########################################
+# Functions for group-forward stepwise #
+########################################
+
+# TODO
+# * Add function to do forward stepwise
+# ** check FIX
+# *** Proceed until a fixed stopping point (don't interpolate)
+# *** Track order, pvals, gaps, etc, intelligently
+# *** Decide how to orthogonalize & residualize
+# * Check add1_group
+# * Triple check everything
+
+source("lambda_gap_stat.R")
+
+# Add the next group
+add1_group = function(X, Y, groups, weights, active.set=0, residualize=TRUE, alpha=0.1, origX=X, origY=Y, betahat=0, ...) {
+  V = test_statistic(X, Y, groups, weights, active.set)
+  V$X = X
+  V$Y = Y
+  V$betahat = betahat
+  V$active.set = active.set
+  imax = V$imax
+  ginds = groups[[imax]]
+  pval = pval_from_lambdas(V$lambda1, V$lambda2, V$weight, V$rank)
+  V$pval = pval
+  if ((length(betahat) > 1) & (length(active.set) < length(betahat) -1))  {
+    V$active.set = setdiff(union(active.set, imax), 0)
+    inds = unlist(groups[V$active.set])
+    lmnew = lm(origY ~ origX[,inds] - 1)
+    betanew = lmnew$coefficients
+    V$betahat[inds] = betanew
+    V$Y = lmnew$residuals
+  }
+  if (residualize == TRUE) {
+    Xnew = X
+    Xg = X[,ginds]
+    Hg = diag(1,length(Y)) - Xg %*% t(Xg)
+    for (g in setdiff(1:length(groups), imax)) {
+      group = groups[[g]]
+      Xnew[,group] = svd(Hg %*% X[,group])$u
+    }
+    V$X = Xnew
+  }
+  return(V)
+}
+
+# Forward stepwise function
+## Input design matrix, response vector, groups as list of lists
+## weights as list or "default" which weights by sqrt of group size
+### alpha? or separate function using alpha?
+## 
+forward_group <- function(X, Y, groups, weights="default", max.steps="default", alpha=0.1, ...) {
+  # In progress
+  if (weights == "default") {
+    # Default weights are sqrt of group size
+    weights <- sapply(groups, function(group) sqrt(length(group)))
+  }
+  if (max.steps == "default") {
+    # Default max steps is min of half of all groups and one less group than saturation
+    ## FIX
[…]

### 2013-09-28 sim_sgl_compare.R

@@ -1,12 +1,0 @@
-#############################################
-# Comparison of forward stepwise with cvSGL #
-#############################################
-
-
-# TODO
-# * Find a stopping rule first
-# * Keep it simple
-# *** Don't track everything
-# *** Don't vary many parameters
-
-# * source other files

### 2013-09-28 test.R

@@ -1,15 +1,0 @@
-###############################################
-# Run a battery of tests to check development #
-###############################################
-
-source("generate_data.R")
-source("forward_step.R")
-
-n <- 100
-g <- 20
-k <- 5
-
-data <- simulate_fixed(n, g, k)
-
-forward_group(data$X, data$Y, data$groups, weights="default")
-

### 2013-09-28 paper/paper.tex
Made a larger model selection simulation

@@
 \includegraphics[width=0.5\textwidth]{../figs/small_sim_gaussian_design.pdf}}
-\hspace{-15pt}
+\subfigure[Large Gaussian design]{
+\label{fig:lassopvala}
+\includegraphics[width=0.5\textwidth]{../figs/large_sim_gaussian_design.pdf}}
+%\hspace{-15pt}
 \subfigure[Categorical design]{
@@
 
-\label{fig:smallselection}
-
 \input{../small_sim_results.tex}
+
+\input{../selection_sim_results.tex}
+
+The large simulation has 50 groups, 25 of size 1, 10 of size 5, 10 of
+size 10, and 5 of size 15. Signal vectors were supported on random
+choices of 10 of these groups (changing in each realization), with
+non-zero signal magnitude around $\sqrt{2\log p}$ where $p = 250$
+(roughly 3.3). The design matrix had 200 rows and the simulation
+performed 100 realizations.
 

### 2013-09-28 selection_sim_results.tex
Added a few existing files

% latex table generated in R 3.0.1 by xtable 1.7-1 package
% Fri Sep 27 17:35:50 2013
\begin{table}[ht]
\centering
\begin{tabular}{rrrrrr}
  \hline
 & Fdp & R & S & V & Power \\ 
  \hline
first & 0.06 & 10.48 & 9.77 & 0.71 & 0.98 \\ 
  forward & 0.00 & 0.01 & 0.01 & 0.00 & 0.00 \\ 
  hybrid & 0.08 & 2.00 & 1.84 & 0.16 & 0.18 \\ 
   \hline
\end{tabular}
\caption{Evaluation of model selection using several stopping rules based on our p-values. The naive stopping rule performs well.} 
\end{table}

### 2013-10-01 fwd_step_test.R
Partial fix of categorical problem: signal zero group mean

@@ -78,7 +78,7 @@ fwd_group_simulation = function(n, sigma, groups, beta, nsim, max.steps, alpha =
     if (rand.beta == TRUE) {
       plot.main = paste(plot.main, "(randomized signal)")
     }
-    plot(xax, MSRS, type = "b", main = plot.main, xlab = "Step", ylab = "MSRS")
+    plot(xax, MSRS, type = "b", main = plot.main, xlab = "Step", ylab = "MSRS", ylim = c(0,1))
 
     points(nxax, null.Pvals, col="red")
     arrows(nxax, null.Pvals.CI[1, ], nxax, null.Pvals.CI[2, ],

### 2013-10-01 sim_small.R

@@ -5,7 +5,7 @@ source('fwd_step.R')
 source('generate_data.R')
 source('fwd_step_test.R')
 
-nsim = 500
+nsim = 20
 n = 50
 sigma = 1
 groups = c(1, 1, 2, 3, 3, 3, 4, 4, 4, 4, 5, 5, 6, 6, 7, 7, 7, 8, 9, 9, rep(10, 10))
@@ -15,18 +15,25 @@ lower = (1-.1)*sqrt(2*log(p))
 num.nonzero = 3
 max.steps = 8
 beta = beta_staircase(groups, num.nonzero, upper, lower)
+beta_old = beta
 
 pdf('figs/small_sim_gaussian_design.pdf')
 output.g <- fwd_group_simulation(n, sigma, groups, beta, nsim, max.steps, plot = TRUE)
 dev.off()
+warnings()
+
 
 groups = sort(c(groups, 2, 8))
-beta = beta_staircase(groups, num.nonzero, upper, lower)
+p = length(groups)
+upper = (1+.1)*sqrt(2*log(p))
+lower = (1-.1)*sqrt(2*log(p))
+beta = beta_staircase(groups, num.nonzero, upper, lower, rand.sign = TRUE, perturb = TRUE, cat.vars = 1:max(groups))
+warnings()
 
-#pdf('figs/small_sim_categorical_design.pdf')
-#output.c = fwd_group_simulation(n, sigma, groups, beta, nsim, max.steps, categorical = TRUE, plot = TRUE)
-#dev.off()
-#warnings()
+pdf('figs/small_sim_categorical_design.pdf')
+output.c = fwd_group_simulation(n, sigma, groups, beta, nsim, max.steps, categorical = TRUE, plot = TRUE)
+dev.off()
+warnings()
 
 caption = "Evaluation of model selection using several stopping rules based on our p-values. The naive stopping rule performs well."
 selection.results = with(output.g, xtable(sim_select_stats(signal.p, active.set, true.step, m1), caption = caption))
@@ -35,4 +42,9 @@ writeLines(print(selection.results), file)
 close(file)
 
 
-#with(output.c, print(sim_select_stats(signal.p, active.set, true.step, m1)))
+with(output.c, print(sim_select_stats(signal.p, active.set, true.step, m1)))
+
+print(c(max(abs(beta)), max(abs(beta_old)), upper, lower))
+
+print(beta_old)
+print(beta)

### 2013-10-02 sim_large.R
Almost broke group_lasso_knot trying to introduce Sigma

@@ -11,7 +11,7 @@ sigma = 1
 groups = sort(c(rep(1:10, 5), rep(11:20, 10), rep(21:25, 15), 26:50))
 p = length(groups)
 upper = (1+.1)*sqrt(2*log(p))
-lower = (1-.1)*sqrt(2*log(p))
+lower = (1-.3)*sqrt(2*log(p))
 num.nonzero = 10
 max.steps = 14
 beta = beta_staircase(groups, num.nonzero, upper, lower)

### 2013-10-02 sim_small.R

@@ -5,7 +5,7 @@ source('fwd_step.R')
 source('generate_data.R')
 source('fwd_step_test.R')
 
-nsim = 20
+nsim = 500
 n = 50
 sigma = 1
 groups = c(1, 1, 2, 3, 3, 3, 4, 4, 4, 4, 5, 5, 6, 6, 7, 7, 7, 8, 9, 9, rep(10, 10))

### 2013-10-04 fwd_step.R
Fixed categorical problem, rerunning simulations

@@ -18,7 +18,7 @@ update_active_set = function(active.set, group) {
 add_group = function(X, Y, groups, weights, sigma, active.set = 0, eff.p = 0) {
 
   n = length(Y)
-  results = group_lasso_knot(X, Y, groups, weights, active.set)
+  results = group_lasso_knot(X, Y, groups, weights, active.set = active.set)
   imax = results$i
   gmax = groups == imax
   new.eff.p = eff.p + sum(gmax)

### 2013-10-04 fwd_step_test.R

@@ -36,7 +36,7 @@ fwd_group_simulation = function(n, sigma, groups, beta, nsim, max.steps, alpha =
     }
 
     if (categorical == TRUE) {
-      X = categorical_design(n, groups, ortho.within = TRUE)
+      X = categorical_design(n, groups, ortho.within = FALSE)
     } else {
       X = gaussian_design(n, groups)
     }
@@ -57,12 +57,15 @@ fwd_group_simulation = function(n, sigma, groups, beta, nsim, max.steps, alpha =
                  is.element(x, true.active.groups))
   }
 
-  CI.probs <- c(.25, .75)
+  bar.quantiles <- c(.25, .75)
+  point.quantiles <- c(.05, .95)
   null.Pvals <- colMeans(P.mat)
-  null.Pvals.CI <- apply(P.mat, 2, function(col) quantile(col, probs = CI.probs))
+  null.Pvals.bar <- apply(P.mat, 2, function(col) quantile(col, probs = bar.quantiles))
+  null.Pvals.point <- apply(P.mat, 2, function(col) quantile(col, probs = point.quantiles))
 
   Pvals = colMeans(P.mat.b)
-  Pvals.CI <- apply(P.mat.b, 2, function(col) quantile(col, probs = CI.probs))
+  Pvals.bar <- apply(P.mat.b, 2, function(col) quantile(col, probs = bar.quantiles))
+  Pvals.point <- apply(P.mat.b, 2, function(col) quantile(col, probs = point.quantiles))
   
   #recover.mat = apply(AS.mat.b, 1:2, function(x) is.element(x, true.active.groups))
   MSRS = colMeans(recover.mat)
@@ -78,15 +81,19 @@ fwd_group_simulation = function(n, sigma, groups, beta, nsim, max.steps, alpha =
     if (rand.beta == TRUE) {
       plot.main = paste(plot.main, "(randomized signal)")
     }
-    plot(xax, MSRS, type = "b", main = plot.main, xlab = "Step", ylab = "MSRS", ylim = c(0,1))
+    plot(xax, MSRS, type = "b", main = plot.main, xlab = "Step", ylab = "MSRS", ylim = c(-.1,1.1))
 
     points(nxax, null.Pvals, col="red")
-    arrows(nxax, null.Pvals.CI[1, ], nxax, null.Pvals.CI[2, ],
+    arrows(nxax, null.Pvals.bar[1, ], nxax, null.Pvals.bar[2, ],
[…]

### 2013-10-04 generate_data.R

@@ -47,21 +47,27 @@ beta_staircase = function(groups, num.nonzero, upper, lower, rand.within=FALSE, 
   nz.inds = beta != 0
 
   if (perturb) {
-    beta = beta + rnorm(p) * sqrt(1/10) * nz.inds
+    beta[nz.inds] = beta[nz.inds] + rnorm(sum(nz.inds)) * sqrt(1/10)
+    maxmod = max(abs(beta))
+    beta[nz.inds] = beta[nz.inds] * upper / maxmod
   }
 
   # Ensure coefs for categorical variables sum to 0
   if (length(cat.vars) > 0) {
+    print("Oh, make a cat var")
     for (g in cat.vars) {
       gind = groups == g & nz.inds
       if (sum(gind) > 0) {
         gmod = max(abs(beta[gind]))
+        print(paste("Old modulus", gmod))
         beta[gind] = beta[gind] - mean(beta[gind])
         gnewmod = max(abs(beta[gind]))
+        print(paste("New modulus", gnewmod))
         if (gnewmod == 0) stop("Categorical variable with constant coeff (same for all levels)")
         beta[gind] = beta[gind] * gmod / gnewmod
       }
     }
+    print(paste("New upper", max(abs(beta))))
   }
 
   return(beta)
@@ -111,7 +117,7 @@ categorical_design = function(n, groups, ortho.within = FALSE) {
   return(X)
 }
 
-n = 6
-groups = c(1,1,1,2,2,3,3,3,3)
-X = categorical_design(n, groups)
+#n = 6
+#groups = c(1,1,1,2,2,3,3,3,3)
+#X = categorical_design(n, groups)
 

### 2013-10-04 model_selection.R

@@ -56,3 +56,5 @@ sim_select_stats = function(p.lists, active.sets, true.steps, m1, alpha = .1) {
   colnames(results) = c("Fdp", "R", "S", "V", "Power")
   return(results)
 }
+
+

### 2013-10-04 sim_small.R

@@ -1,17 +1,16 @@
 
-library(xtable)
-
 source('fwd_step.R')
 source('generate_data.R')
 source('fwd_step_test.R')
+source('tex_table.R')
 
-nsim = 500
+nsim = 1000
 n = 50
 sigma = 1
 groups = c(1, 1, 2, 3, 3, 3, 4, 4, 4, 4, 5, 5, 6, 6, 7, 7, 7, 8, 9, 9, rep(10, 10))
 p = length(groups)
-upper = (1+.1)*sqrt(2*log(p))
-lower = (1-.1)*sqrt(2*log(p))
+upper = 1 #sqrt(2*log(p))
+lower = .9
 num.nonzero = 3
 max.steps = 8
 beta = beta_staircase(groups, num.nonzero, upper, lower)
@@ -25,8 +24,8 @@ warnings()
 
 groups = sort(c(groups, 2, 8))
 p = length(groups)
-upper = (1+.1)*sqrt(2*log(p))
-lower = (1-.1)*sqrt(2*log(p))
+upper = 1 #(1+.1)*sqrt(2*log(p))
+lower = .9 #(1-.1)*sqrt(2*log(p))
 beta = beta_staircase(groups, num.nonzero, upper, lower, rand.sign = TRUE, perturb = TRUE, cat.vars = 1:max(groups))
 warnings()
 
@@ -36,13 +35,13 @@ dev.off()
 warnings()
 
 caption = "Evaluation of model selection using several stopping rules based on our p-values. The naive stopping rule performs well."
-selection.results = with(output.g, xtable(sim_select_stats(signal.p, active.set, true.step, m1), caption = caption))
-file = file("small_sim_results.tex")
-writeLines(print(selection.results), file)
-close(file)
+results.g = with(output.g, sim_select_stats(signal.p, active.set, true.step, m1))
+results.c = with(output.c, sim_select_stats(signal.p, active.set, true.step, m1))
+rownames(results.g) = paste("(1)", rownames(results.g))
+rownames(results.c) = paste("(2)", rownames(results.c))
+file = "small_sim_results.tex"
+tex_table(file, rbind(results.g, results.c), caption = caption)
 
-
-with(output.c, print(sim_select_stats(signal.p, active.set, true.step, m1)))
 
 print(c(max(abs(beta)), max(abs(beta_old)), upper, lower))
 

### 2013-10-06 paper/paper.tex
Paper: expanded forward stepwise description

@@
 \usepackage{graphicx,subfigure}
+\usepackage[boxed]{algorithm2e}
 
@@
 % "Title of the paper"
-\title{Significance testing with forward stepwise and grouped forward
-  stepwise model selection}
+\title{A significance test for forward stepwise model selection}
 
@@
 % perhaps de-emphasize group lasso and emphasize forward stepwise?
-  We extend the work of \cite{significance:lasso} on significance
-  testing for the lasso to the group lasso.  We derive the corresponding
-  test statistic and its exact distribution, and show that these can be
-  used with (non-penalized) forward stepwise regression and model
-  selection.  For group lasso the forward stepwise procedure adds a
-  group of variables in each step.  These procedures are shown to
-  perform well in simulations and in a real data example.
-
+  We apply the methods of \cite{tests:adaptive} and
+  \cite{significance:lasso} on significance tests for penalized
+  regression to forward stepwise model selection. For the $k$th
+  variable to enter the model, previous work relied on an asymptotic
+  null distribution that breaks down when grouped variables are
+  entering the model and is difficult to derive. We iteratively apply
+  an exact null distribution for the first (potentially grouped)
+  variable on the residual from previous steps. The resulting method
+  has the strengths of stepwise selection, for example parallel
+  computation, but also remedies the problem of inflated test
+  statistics and over-fitting.
 \end{abstract}
@@
 
-\section{Introduction}\label{sec:intro}
+\section{Introduction}
+\label{sec:intro}
 
@@
 statistics for models selected by forward stepwise procedures and
-works out some of the detailsinvolved in applying these methods to
-models with grouped variables. We find that our test statistics can be
-used for valid significance testing, or for tempering the aggressive
-nature of forward stepwise model selection.  The resulting methods may
-be less computationally intensive than choosing a regularization
-parameter for the (group) lasso by cross-validation.
-
-\todo Write about more differences/advantages of this paper- maybe not
-in this section, but: using non-asymptotic p-values, forward step is
-simple(r than lasso), more?
-
-To fix notation for the group lasso problem we consider an outcome $y
-\in \real^n$.  Let an integer $G \geq 1$ be the number of groups (or
-factors).  Lowercase $g$ will be used as an index for groups.  For
-each $1 \leq g \leq G$ the design matrix encoding the $g$th group is
-the $n \times p_g$ matrix denoted $X_g$.  Throughout most of this
-paper we assume orthogonality within groups, that is $X_g^T X_g =
-I_{p_g \times p_g}$ for all $g$. ( \todo Take this out if we don't
-actually need it).  Let $p = \sum_{g=1}^G p_g$ be the total number of
-individual variables, and let $X$ be the matrix constructed by
-column-binding the $X_g$, that is  
+works out some of the details involved in applying these methods to
+models with grouped variables.  Our test statistic can be
+used for valid significance testing when computed from the same
+data as the model selection.  The resulting method can
+be more statistically efficient than validation on held-out data, and
+also more computationally efficient than penalized methods with
+regularization parameters chosen by cross-validation.
+
+(\todo Change this paragraph as the sections are completed)
+In Section \ref{sec:stepwise} we establish notation and describe our
+forward stepwise procedure. Section \ref{sec:testing} briefly reviews
+the parts of \cite{significance:lasso} and \cite{tests:adaptive}
+relevant to our significance test, and describes the group lasso which
+we require for applying the test with grouped
+variables. Simulation results in Section \ref{sec:simulations} using
+various stopping rules for the forward stepwise procedure--including
+some from \cite{sequential:fdr}--appear promising.
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Forward stepwise model selection}
+\label{sec:stepwise}
+
+We allow forward stepwise selection to add groups of variables in each
+step, not only in the case of binary encoding for categorical
+variables but also for any grouping purpose. For example, groups of
+variables may be pre-designated factors such as expression
+measurements for all genes in a single regulatory pathway. An entire
+group is included or excluded together from the final model. For
+consistency we will use $g, h$ as indices rather than the usual $i, j$
+throughout. Since single variables can be considered groups of size 1,
+our general exposition includes non-grouped situations as a special case.
+
+Label the outcome variable of $n$ i.i.d. measurements $y \in \real^n$. Let an integer $G \geq 1$ 
+be the number of groups. For each $1 \leq g \leq G$ the design matrix encoding the
+$g$th group is the $n \times p_g$ matrix denoted $X_g$, where $p_g$ is
+the number of individual variables in group $g$. Define $p = \sum_{g=1}^G
+p_g$ be the total number of individual variables, so $p = G$ in the
+case where all groups have size 1. Let $X$ be the matrix constructed
+by column-binding the $X_g$, that is  
+%Throughout most of this paper we assume orthogonality within groups, that is
+%$X_g^T X_g = I_{p_g \times p_g}$ for all $g$. ( \todo Take this out if
+%we don't actually need it).  
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 by stacking all of the $\beta_g$ in order.  Finally, our model for the
-data is
+response is
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-where $\epsilon$ is noise.  Unless otherwise specified we will assume
-i.i.d. Gaussian noise $\epsilon \sim N(0, I_{n \times n})$ throughout.
-So far we have only introduced some new notation to the usual linear
-regression setup, but now we define the \emph{group lasso estimator}
-as the following solution to a convex problem 
+where $\epsilon$ is noise. Unless otherwise specified we assume
+i.i.d. Gaussian noise $\epsilon \sim N(0, I_{n \times n})$ and that
+$\sigma$ is known.
+
+Before we describe the forward stepwise procedure we require one last
+ingredient. To each group of variables we assign a weight $w_g$. These
+weights act like penalties or costs, so increasing $w_g$ makes it
+\textit{more difficult} for the group $X_g$ to enter the model. The
+modeler can choose weights arbitrarily, but we will only use one
+particular choice, based on $p_g$, that we discuss later. With this we
+are ready to describe the forward stepwise procedure Algorithm \ref{algo:fs}.
+
+\begin{algorithm}
+\DontPrintSemicolon
+\KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ grouped variables}
+\KwResult{Active set $A$ of variable groups included in the model}
+\SetKwFunction{lsfitResidual}{lsfitResidual}
+\caption{Forward stepwise procedure}
+\BlankLine
+$A \gets \emptyset$\;
+$A^c \gets \{ 1, \ldots, G \}$\;
+$r_0 \gets y$\;
+\For{$counter \leftarrow 1$ \KwTo $\min (n, G)$}{
+  $g^* \gets \argmax_{g \in A^c} \norm{X_g^T r_{g-1}}_2 / w_g$\;
+  $A \gets A \cup \{ g^* \}$\;
+  $A^c \gets A^c \backslash \{ g^* \}$\;
+  $r_g \gets $\lsfitResidual{$r_{g-1}, X_{g^*}$}\;
+}
+\KwRet{$A$}
+\label{algo:fs}
+\end{algorithm}
+
+\todo Expand this section
+
+\begin{itemize}
+  \item Brief background, cite some original papers
+  \item Stopping rules, cite new paper on lasso?
+\end{itemize}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%
+%%%%%%%%%%%%%%%%%%%%%%%%%
+
+\section{Significance testing: from lasso to group lasso}
+\label{sec:testing}
+
+\todo Update this section to match/complement \citep{tests:adaptive}
+
+In a recent work \cite{significance:lasso} defined a \emph{covariance
+  test statistic} for testing the significance of a variable entering
+the model along the lasso solution path.  They derived a simple
+asymptotic null distribution, proved a type of convergence under broad
+``minimum growth'' conditions, and demonstrated in simulations that
+the test statistic closely matches its asymptotic distribution even in
+finite samples.  That work marked an important advance in the problem
+of combining inference with model selection.  The current paper
+extends some of their results to the group lasso (ref Yuan and Lin?).
+In the process we had to derive an exact finite sample null
+distribution for the test statistic (ref TLTT?).  We also show that
+the techniques used to get these results can be used to do a forward
+stepwise procedure related to the group lasso that adds groups of
+variables (or factors) in each step.
+
+\subsection{Background}
+In the ordinary least squares setting, a significance test for a
+single variable can be conducted by comparing the drop in residual
+sums of squares (RSS) to a $\chi^2_1$ distribution. Similarly, when
+adding a group of $k$ variables we can compare the drop in RSS to a
+$\chi^2_k$ random variable. This generally does not work when the
+variable to be added has been chosen by a method that uses the data,
+and in particular it fails for forward stepwise procedures which add
+the ``best'' (e.g. most highly correlated) predictor in each step. In
+that case, the test statistic (drop in RSS) does not match the
+theoretical null distribution even when the null hypothesis is
+true. Lockhart et al. (ref LTTT) introduced a new test statistic based
+on the knots in the lasso solution path. Writing $\hat
+\beta(\lambda)$, the solution for a fixed value of $\lambda$, we need
+the following facts (ref Tibs2012)
+
+\begin{itemize}
+
+  \item The vector valued function $\hat \beta(\lambda)$ is a
+    continuous function of $\lambda$. For the lasso path, the
+    coordinates of $\hat \beta(\lambda)$ are piecewise linear with
+    changes in slope occurring at a finite number of $\lambda$ values
+    referred to as \emph{knots}. The knots depend on the data and are
+    usually written in order $\lambda_1 \geq \lambda_2 \geq \cdots
+    \geq \lambda_r \geq 0$. We follow this convention.
+
+  \item The \emph{active set} $A_k$ is a set of indices of variables
+    for which the corresponding coordinates of $\hat \beta(\lambda_k)$
+    are potentially nonzero. Any variable with index not in $A_k$ has
+    a zero coefficient in $\hat \beta(\lambda_k)$, but the converse is
+    not true.
+
+  \item Path algorithms for computing lasso solutions proceed by
+    fitting models at a grid of $\lambda$ values. The active set
+    changes whenever $\lambda$ crosses a knot, and predictor variables
+    can both enter and leave the active set. However, at the first two
+    knots $\lambda_1$ and $\lambda_2$ no variable can leave the active
+    set. So the first two knots correspond to the first two variables
+    entering the model.
+
+\end{itemize}
+
+Lockhart et al. prove that, under the null hypothesis that $A_k$
+contains all the strong predictor variables, the distribution of a
+test statistic $T_k \propto \lambda_k(\lambda_k - \lambda_{k+1})$ is
+asymptotically Exp(1). In the lasso case we know a lot about the knots
+and active set, but the group lasso picture is slightly more
+complicated. For the group lasso, $\hat \beta(\lambda)$ does not have
+piecewise linear components. To overcome this difficulty we will
+restrict our attention to the first group of variables to enter the
+active set since the analysis then follows almost exactly as for the
+lasso. (\todo Change this if I can find $\lambda_k$).
+
+The \emph{group lasso estimator} is the following solution to a convex
+problem
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 and add a few more references here if they seem worthwhile.
-
-In a recent work \cite{significance:lasso} defined a \emph{covariance
-  test statistic} for testing the significance of a variable entering
-the model along the lasso solution path.  They derived a simple
-asymptotic null distribution, proved a type of convergence under broad
-``minimum growth'' conditions, and demonstrated in simulations that
-the test statistic closely matches its asymptotic distribution even in
-finite samples.  That work marked an important advance in the problem
-of combining inference with model selection.  The current paper
-extends some of their results to the group lasso (ref Yuan and Lin?).
-In the process we had to derive an exact finite sample null
-distribution for the test statistic (ref TLTT?).  We also show that
-the techniques used to get these results can be used to do a forward
-stepwise procedure related to the group lasso that adds groups of
-variables (or factors) in each step.
-
-(\todo Change this paragraph as the sections are completed)
-In Section 2 we briefly review the work of (ref LTTT) concerning
-significance testing for the lasso, and discuss how this result is
-extended to group lasso. Section 2 also introduces more of the
-terminology we use to describe the present problem. Section 3
-describes a framework for forward stepwise procedures using the test
-statistic derived in Section 2. In this framework groups of variables
-are added in each step until some stopping criterion involving the
-test statistic is satisfied. In Section 4 we show the results of some
-simulations for several forward stepwise procedures, and in Section 5
-we apply our method to a real data set (\todo brief description of
-data).
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\section{Significance testing: from lasso to group lasso}
-\label{sec:testing}
-
-\todo Update this section to match/complement \citep{tests:adaptive}
-
-\subsection{Background}
-In the ordinary least squares setting, a significance test for a
-single variable can be conducted by comparing the drop in residual
-sums of squares (RSS) to a $\chi^2_1$ distribution. Similarly, when
-adding a group of $k$ variables we can compare the drop in RSS to a
-$\chi^2_k$ random variable. This generally does not work when the
-variable to be added has been chosen by a method that uses the data,
-and in particular it fails for forward stepwise procedures which add
-the ``best'' (e.g. most highly correlated) predictor in each step. In
-that case, the test statistic (drop in RSS) does not match the
-theoretical null distribution even when the null hypothesis is
-true. Lockhart et al. (ref LTTT) introduced a new test statistic based
-on the knots in the lasso solution path. Writing $\hat
-\beta(\lambda)$, the solution for a fixed value of $\lambda$, we need
-the following facts (ref Tibs2012)
-
-\begin{itemize}
-
-  \item The vector valued function $\hat \beta(\lambda)$ is a
-    continuous function of $\lambda$. For the lasso path, the
-    coordinates of $\hat \beta(\lambda)$ are piecewise linear with
-    changes in slope occurring at a finite number of $\lambda$ values
-    referred to as \emph{knots}. The knots depend on the data and are
-    usually written in order $\lambda_1 \geq \lambda_2 \geq \cdots
-    \geq \lambda_r \geq 0$. We follow this convention.
-
-  \item The \emph{active set} $A_k$ is a set of indices of variables
-    for which the corresponding coordinates of $\hat \beta(\lambda_k)$
-    are potentially nonzero. Any variable with index not in $A_k$ has
-    a zero coefficient in $\hat \beta(\lambda_k)$, but the converse is
-    not true.
-
-  \item Path algorithms for computing lasso solutions proceed by
-    fitting models at a grid of $\lambda$ values. The active set
-    changes whenever $\lambda$ crosses a knot, and predictor variables
-    can both enter and leave the active set. However, at the first two
-    knots $\lambda_1$ and $\lambda_2$ no variable can leave the active
-    set. So the first two knots correspond to the first two variables
-    entering the model.
-
-\end{itemize}
-
-Lockhart et al. prove that, under the null hypothesis that $A_k$
-contains all the strong predictor variables, the distribution of a
-test statistic $T_k \propto \lambda_k(\lambda_k - \lambda_{k+1})$ is
-asymptotically Exp(1). In the lasso case we know a lot about the knots
-and active set, but the group lasso picture is slightly more
-complicated. For the group lasso, $\hat \beta(\lambda)$ does not have
-piecewise linear components. To overcome this difficulty we will
-restrict our attention to the first group of variables to enter the
-active set since the analysis then follows almost exactly as for the
-lasso. (\todo Change this if I can find $\lambda_k$).
 
@@
 
-\section{Forward stepwise procedures}
-
-\todo Expand this section
-
-\begin{itemize}
-  \item Brief background, cite some original papers
-  \item Describe exactly how we do each step (projection, residuals, etc) 
-  \item Stopping rules, cite new paper on lasso, perhaps we have
-    a computation advantage
-\end{itemize}
-
 \section{Simulations}
+\label{sec:simulations}
 
@@
 \subfigure[Gaussian design]{
-\label{fig:lassopvala}
+\label{fig:smallgauss}
 \includegraphics[width=0.5\textwidth]{../figs/small_sim_gaussian_design.pdf}}
-\subfigure[Large Gaussian design]{
-\label{fig:lassopvala}
-\includegraphics[width=0.5\textwidth]{../figs/large_sim_gaussian_design.pdf}}
-%\hspace{-15pt}
+\hspace{-15pt}
 \subfigure[Categorical design]{
-\label{fig:lassopvalb}
+\label{fig:smallcat}
 \includegraphics[width=0.5\textwidth]{../figs/small_sim_categorical_design.pdf}}
@@
 
-\input{../selection_sim_results.tex}
+\begin{figure}
+\begin{center}
+\subfigure[Large Gaussian lasso]{
+\label{fig:largelasso}
+\includegraphics[width=0.5\textwidth]{../figs/large_sim_gaussian_lasso.pdf}}
+\hspace{-15pt}
+\subfigure[Large Gaussian group lasso]{
+\label{fig:largegrplasso}
+\includegraphics[width=0.5\textwidth]{../figs/large_sim_gaussian_design.pdf}}
+\caption{\small \it The left panel shows results of a simulation with
+  an independent Gaussian design matrix and a signal vector having
+  three groups with non-zero coefficients of sizes one, two, and
+  three, and seven groups of noise variables with sizes varying from
+  one to ten, for a total of 30 variables. For the right panel we
+  repeated this setup, but group sizes corresponded to levels of
+  categorical variables. The groups of size one were increased to two,
+so there are 10 categorical variables with a total of 32 levels.}
+\end{center}
+\end{figure}
+
+\input{../large_sim_results.tex}
 
@@
 \section{Discussion}
+\label{sec:discuss}
 

### 2013-10-06 fwd_step_test.R

@@ -81,8 +81,9 @@ fwd_group_simulation = function(n, sigma, groups, beta, nsim, max.steps, alpha =
     if (rand.beta == TRUE) {
       plot.main = paste(plot.main, "(randomized signal)")
     }
-    plot(xax, MSRS, type = "b", main = plot.main, xlab = "Step", ylab = "MSRS", ylim = c(-.1,1.1))
-
+    plot(xax, MSRS, type = "l", main = plot.main, xlab = "Step", ylab = "MSRS", ylim = c(-.1,1.1))
+    abline(v = num.nonzero, lty = "dotted")
+    
     points(nxax, null.Pvals, col="red")
     arrows(nxax, null.Pvals.bar[1, ], nxax, null.Pvals.bar[2, ],
            code = 3, angle = 90, length = 0, col = "red")

### 2013-10-06 sim_large.R

@@ -4,15 +4,31 @@ source('generate_data.R')
 source('fwd_step_test.R')
 source('tex_table.R')
 
-nsim = 500
+nsim = 200
 n = 150
 sigma = 1
+
+groups = 1:1050
+p = length(groups)
+mult = sqrt(2*log(p))
+upper = 1.1*mult
+lower = 0.5*mult
+num.nonzero = 20
+max.steps = 30
+beta = beta_staircase(groups, num.nonzero, upper, lower)
+
+pdf('figs/large_sim_gaussian_lasso.pdf')
+output.l <- fwd_group_simulation(n, sigma, groups, beta, nsim, max.steps, rand.beta = TRUE, plot = TRUE)
+dev.off()
+
+
 groups = sort(c(rep(1:100, 5), rep(101:150, 10), 151:200))
 p = length(groups)
-upper = 3
-lower = 2
-num.nonzero = 10
-max.steps = 14
+mult = sqrt(2*log(p))
+upper = 1.5*mult
+lower = 1.1*mult
+num.nonzero = 8
+max.steps = 12
 beta = beta_staircase(groups, num.nonzero, upper, lower)
 
 pdf('figs/large_sim_gaussian_design.pdf')
@@ -22,8 +38,9 @@ dev.off()
 
 groups = sort(c(groups, 151:200))
 p = length(groups)
-upper = 3 #(1+.1)*sqrt(2*log(p))
-lower = 2 #(1-.1)*sqrt(2*log(p))
+mult = sqrt(2*log(p))
+upper = 1.5*mult
+lower = 1.1*mult
 beta = beta_staircase(groups, num.nonzero, upper, lower, rand.sign = TRUE, perturb = TRUE, cat.vars = 1:max(groups))
 warnings()
 
@@ -33,9 +50,11 @@ dev.off()
 warnings()
 
 caption = "Evaluation of model selection using several stopping rules based on our p-values. The naive stopping rule performs well."
+results.l = with(output.l, sim_select_stats(signal.p, active.set, true.step, m1))
 results.g = with(output.g, sim_select_stats(signal.p, active.set, true.step, m1))
 results.c = with(output.c, sim_select_stats(signal.p, active.set, true.step, m1))
-rownames(results.g) = paste("(1)", rownames(results.g))
-rownames(results.c) = paste("(2)", rownames(results.c))
+rownames(results.l) = paste("(1)", rownames(results.l))
+rownames(results.g) = paste("(2)", rownames(results.g))
[…]

### 2013-10-09 fwd_step_test.R
Prediction error functioning, maybe done

@@ -2,7 +2,8 @@
 
 source('fwd_step.R')
 source('generate_data.R')
-source('model_selection.R')
+source('selection.R')
+source('pred_and_estim.R')
 
 active_groups = function(groups, beta) {
   beta.ind = aggregate(beta, by=list(groups), FUN = function(beta.coords) any(beta.coords != 0))
@@ -11,7 +12,7 @@ active_groups = function(groups, beta) {
 }
 
 
-fwd_group_simulation = function(n, sigma, groups, beta, nsim, max.steps, alpha = .1, categorical = FALSE, rand.beta = FALSE, plot = FALSE) {
+fwd_group_simulation = function(n, sigma, groups, beta, nsim, max.steps, alpha = .1, categorical = FALSE, predictions = FALSE, rand.beta = FALSE, plot = FALSE) {
 
   weights = sqrt(rle(groups)$lengths)
   p = length(groups)
@@ -26,9 +27,9 @@ fwd_group_simulation = function(n, sigma, groups, beta, nsim, max.steps, alpha =
   P.mat.b = P.mat
   AS.mat.b = P.mat
   recover.mat = P.mat
-  mses.train = c()
-  mses.test = c()
-  mse.beta = c()
+
+  pred.errs = matrix(0, nrow=3, ncol=3)
+  
   if (rand.beta == FALSE) {
     # track fitted values
     beta.mat = matrix(0, nrow=nsim, ncol=p)
@@ -69,19 +70,24 @@ fwd_group_simulation = function(n, sigma, groups, beta, nsim, max.steps, alpha =
                  is.element(x, true.active.groups))
 
     # Tracking prediction error and other things
-    #
-    #
-    # this is broken, need to use stopping rules!
-    #
-    #
-    
-    
-    
+    # stop.rules must agree with those in sim_pred_est_stats
+    if (predictions) {
+      if (categorical) {
+        stop("Tracking prediction error for categorical design is unsupported")
+      }
+      
+      stop.rules = c("first", "forward", "hybrid")
+      pred.errs = pred.errs + sim_pred_est_stats(results.b$p.vals,
+        results.b$active.set, X, Y.beta, groups, beta, X.test, Y.test, alpha)
+    }
+
     if (rand.beta == FALSE) {
-      beta.mat[i, groups.active] = fitted.beta
+      #beta.mat[i, groups.active] = fitted.beta
     }
[…]

### 2013-10-09 generate_data.R

@@ -54,20 +54,20 @@ beta_staircase = function(groups, num.nonzero, upper, lower, rand.within=FALSE, 
 
   # Ensure coefs for categorical variables sum to 0
   if (length(cat.vars) > 0) {
-    print("Oh, make a cat var")
+
     for (g in cat.vars) {
       gind = groups == g & nz.inds
       if (sum(gind) > 0) {
         gmod = max(abs(beta[gind]))
-        print(paste("Old modulus", gmod))
+
         beta[gind] = beta[gind] - mean(beta[gind])
         gnewmod = max(abs(beta[gind]))
-        print(paste("New modulus", gnewmod))
+
         if (gnewmod == 0) stop("Categorical variable with constant coeff (same for all levels)")
         beta[gind] = beta[gind] * gmod / gnewmod
       }
     }
-    print(paste("New upper", max(abs(beta))))
+
   }
 
   return(beta)

### 2013-10-09 sim_small.R

@@ -4,20 +4,20 @@ source('generate_data.R')
 source('fwd_step_test.R')
 source('tex_table.R')
 
-nsim = 500
+nsim = 50
 n = 50
 sigma = 1
 groups = c(1, 1, 2, 3, 3, 3, 4, 4, 4, 4, 5, 5, 6, 6, 7, 7, 7, 8, 9, 9, rep(10, 10))
 p = length(groups)
-upper = 1 #sqrt(2*log(p))
-lower = .9
+upper = 2 #sqrt(2*log(p))
+lower = 1.9
 num.nonzero = 3
 max.steps = 8
 beta = beta_staircase(groups, num.nonzero, upper, lower)
 beta_old = beta
 
 pdf('figs/small_sim_gaussian_design.pdf')
-output.g <- fwd_group_simulation(n, sigma, groups, beta, nsim, max.steps, plot = TRUE)
+output.g <- fwd_group_simulation(n, sigma, groups, beta, nsim, max.steps, predictions = TRUE, plot = TRUE)
 dev.off()
 warnings()
 
@@ -43,7 +43,12 @@ file = "small_sim_results.tex"
 tex_table(file, rbind(results.g, results.c), caption = caption)
 
 
-print(c(max(abs(beta)), max(abs(beta_old)), upper, lower))
+caption = "Prediction and estimation errors for a small simulation"
+p.results.g = output.g$pred.err
+file = "small_sim_pred_err.tex"
+tex_table(file, p.results.g, caption = caption)
 
-print(beta_old)
-print(beta)
+
+#print(c(max(abs(beta)), max(abs(beta_old)), upper, lower))
+#print(beta_old)
+#print(beta)

### 2013-10-09 paper/paper.tex
Moving files, small edits
Expanded forward stepwise section of paper

@@
 
+We allow for the possibility $p > n$ in which case \ref{eq:gmodel} is
+generically underdetermined. In such cases it still may be possible to
+estimate $\beta$ well if it is sparse--that is, if it has few nonzero
+entries \citep{donoho:pursuit}. In the rest of this paper we refer to
+variable groups $X_g$ as noise variables if $\beta_g$ is a zero vector
+and as true predictor variables if $\beta_g$ has nonzero entries.
+
 Before we describe the forward stepwise procedure we require one last
@@
 weights act like penalties or costs, so increasing $w_g$ makes it
-\textit{more difficult} for the group $X_g$ to enter the model. The
+more difficult for the group $X_g$ to enter the model. The
 modeler can choose weights arbitrarily, but we will only use one
 particular choice, based on $p_g$, that we discuss later. With this we
-are ready to describe the forward stepwise procedure Algorithm \ref{algo:fs}.
+are ready to describe the forward stepwise procedure, given in Algorithm
+\ref{algo:fs}. Note that to prevent saturation, and to enable our p-value
+computations, \textit{max.steps} should be at most $\min (n, G) - 1$.
 
@@
 \SetKwFunction{lsfitResidual}{lsfitResidual}
-\caption{Forward stepwise procedure}
+\caption{Forward stepwise procedure with groups and weights}
 \BlankLine
@@
 $r_0 \gets y$\;
-\For{$counter \leftarrow 1$ \KwTo $\min (n, G)$}{
+\For{$step \leftarrow 1$ \KwTo $max.steps$}{
   $g^* \gets \argmax_{g \in A^c} \norm{X_g^T r_{g-1}}_2 / w_g$\;
@@
 \end{itemize}
+
+Subset selection refers to the problem of choosing a subset of the $p$
+potential variables to include in a model. Since there are $2^p$ such
+choices, exhaustive search is computationally infeasible when $p$ is
+large. If exhaustive search were possible, performing a large
+search runs the risk of over-fitting unless model complexity is
+appropriately penalized. Algorithm \ref{algo:fs} produces a much
+smaller set of potential models, at most \textit{max.steps}. However,
+it is a greedy algorithm so the set of models it produces may not
+contain the best possible model. This is an inherent shortcoming which
+we do not address in this paper.
+
+So far we have left open the question of choosing among the models in
+the forward stepwise sequence, i.e. when to stop stepping
+forward. Some classical approaches can be posed as optimization
+criteria and stop at the step minimizing
+\begin{equation}
+\begin{aligned}
+\label{eq:subsetregress}
+\frac{1}{2} \| y - X \beta_s \|_2^2 + \lambda P(\beta_s)
+\end{aligned}
+\end{equation}
+where we have written $\{ \beta_s : s = 1, \ldots, max.steps \}$ as
+the sequence of models outputted by forward stepwise. The function
+$P(\beta)$ is a penalty on model complexity usually taken to be the
+number of nonzero entries of $\beta$. Proposals for $\lambda$ include
+2 ($C_p$, AIC \todo{cite}), $\log(n)$ (BIC of \cite{BIC}), and
+$2\log(p)$ (RIC of \cite{RIC}). In the current paper we provide
+accurate p-values when adding the first noise variable group in the forward
+stepwise sequence, and it is natural to consider using these p-values
+to choose a model. \cite{sequential:fdr} examined some stopping rules
+using the asymptotic p-values of \cite{significance:lasso} and showed
+their stopping rules control false discovery rate--the expected
+proportion of noise variables among variables declared significant
+\citep{fdr}. We explore this further in Section \ref{sec:simulations}.
 
@@
 
-\input{../small_sim_results.tex}
+\input{../small_sim_selection.tex}
+
+\input{../small_sim_estimation.tex}
 
@@
 
-\input{../large_sim_results.tex}
+\input{../large_sim_selection.tex}
 

### 2013-10-09 sim_large.R

@@ -5,7 +5,7 @@ source('fwd_step_test.R')
 source('tex_table.R')
 
 nsim = 200
-n = 150
+n = 100
 sigma = 1
 
 groups = 1:500
@@ -27,8 +27,8 @@ p = length(groups)
 mult = sqrt(2*log(p))
 upper = 1.1*mult
 lower = 0.9*mult
-num.nonzero = 8
-max.steps = 12
+num.nonzero = 6
+max.steps = 9
 beta = beta_staircase(groups, num.nonzero, upper, lower)
 
 pdf('figs/large_sim_gaussian_design.pdf')

### 2013-10-09 sim_small.R

@@ -4,7 +4,7 @@ source('generate_data.R')
 source('fwd_step_test.R')
 source('tex_table.R')
 
-nsim = 200
+nsim = 500
 n = 50
 sigma = 1
 groups = c(1, 1, 2, 3, 3, 3, 4, 4, 4, 4, 5, 5, 6, 6, 7, 7, 7, 8, 9, 9, rep(10, 10))

### 2013-10-09 paper/paper.tex
Paper editing

@@
 \newcommand\tf{\widetilde{f}}
+\newcommand{\pen}{\mathcal{P}}
 
@@
 
-\todo Expand this section
-
-\begin{itemize}
-  \item Brief background, cite some original papers
-  \item Stopping rules, cite new paper on lasso?
-\end{itemize}
-
 Subset selection refers to the problem of choosing a subset of the $p$
@@
 \label{eq:subsetregress}
-\frac{1}{2} \| y - X \beta_s \|_2^2 + \lambda P(\beta_s)
+\frac{1}{2} \| y - X \beta_s \|_2^2 + \lambda \pen(\beta_s)
 \end{aligned}
@@
 the sequence of models outputted by forward stepwise. The function
-$P(\beta)$ is a penalty on model complexity usually taken to be the
+$\pen(\beta)$ is a penalty on model complexity usually taken to be the
 number of nonzero entries of $\beta$. Proposals for $\lambda$ include
@@
 
-In a recent work \cite{significance:lasso} defined a \emph{covariance
-  test statistic} for testing the significance of a variable entering
-the model along the lasso solution path.  They derived a simple
-asymptotic null distribution, proved a type of convergence under broad
-``minimum growth'' conditions, and demonstrated in simulations that
-the test statistic closely matches its asymptotic distribution even in
-finite samples.  That work marked an important advance in the problem
-of combining inference with model selection.  The current paper
-extends some of their results to the group lasso (ref Yuan and Lin?).
-In the process we had to derive an exact finite sample null
-distribution for the test statistic (ref TLTT?).  We also show that
-the techniques used to get these results can be used to do a forward
-stepwise procedure related to the group lasso that adds groups of
-variables (or factors) in each step.
-
-\subsection{Background}
 In the ordinary least squares setting, a significance test for a
@@
 variable to be added has been chosen by a method that uses the data,
-and in particular it fails for forward stepwise procedures which add
+and in particular it fails for forward stepwise procedures adding
 the ``best'' (e.g. most highly correlated) predictor in each step. In
@@
 theoretical null distribution even when the null hypothesis is
-true. Lockhart et al. (ref LTTT) introduced a new test statistic based
-on the knots in the lasso solution path. Writing $\hat
+true. \cite{significance:lasso} introduced a new test statistic based
+on the knots in the lasso solution path. They derived a simple
+asymptotic null distribution, proved a type of convergence under broad
+``minimum growth'' conditions, and demonstrated in simulations that
+the test statistic closely matches its asymptotic distribution even in
+finite samples.  That work marked an important advance in the problem
+of combining inference with model selection.  The current paper
+extends some of their results to the group lasso \citep{grouplasso}.
+In the process we had to derive an exact finite sample null
+distribution for the test statistic, this appears in
+\cite{tests:adaptive}.
+
+Writing $\hat
 \beta(\lambda)$, the solution for a fixed value of $\lambda$, we need

### 2013-10-11 paper/paper.tex
Normalize columns, don't project variables

@@
 % "Title of the paper"
-\title{A significance test for forward stepwise model selection}
+\title{A significance test in forward stepwise model selection}
 
@@
 number of nonzero entries of $\beta$. Proposals for $\lambda$ include
-2 ($C_p$, AIC \todo{cite}), $\log(n)$ (BIC of \cite{BIC}), and
+2 ($C_p$ of \cite{CP}, AIC of \cite{AIC}), $\log(n)$ (BIC of \cite{BIC}), and
 $2\log(p)$ (RIC of \cite{RIC}). In the current paper we provide
@@
 
-Writing $\hat
-\beta(\lambda)$, the solution for a fixed value of $\lambda$, we need
-the following facts (ref Tibs2012)
+Writing $\hat \beta(\lambda)$ for the lasso solution for a fixed value of $\lambda$, we need
+the following facts summarized in \cite{significance:lasso} (ref Tibs2012?).
 
@@
     a zero coefficient in $\hat \beta(\lambda_k)$, but the converse is
-    not true.
+    not true. \todo{is this equicorrelation set?}
 
@@
 
-Lockhart et al. prove that, under the null hypothesis that $A_k$
+\cite{significance:lasso} prove that under the null hypothesis that $A_k$
 contains all the strong predictor variables, the distribution of a

### 2013-10-11 fwd_step.R

@@ -41,13 +41,14 @@ add_group = function(X, Y, groups, weights, sigma, active.set = 0, eff.p = 0) {
     }
   }
   Y.resid = lm(Y ~ Xgmax.regress - 1)$residuals
-  
-  for (gind in 1:max(groups)) {
-    if (gind != imax) {
-      group = groups == gind
-      X.project[, group] = (diag(rep(1, n)) - Pgmax) %*% X[, group] 
-    }
-  }
+
+####### is this necessary for p-value? ########  
+#  for (gind in 1:max(groups)) {
+#    if (gind != imax) {
+#      group = groups == gind
+#      X.project[, group] = (diag(rep(1, n)) - Pgmax) %*% X[, group] 
+#    }
+#  }
 
   return(list(test.output = results, p.value = p.value, added = imax, active.set = new.active.set, eff.p = new.eff.p, Y.update = Y.resid, X.update = X.project))
 }

### 2013-10-11 sim_large.R

@@ -4,17 +4,17 @@ source('generate_data.R')
 source('fwd_step_test.R')
 source('tex_table.R')
 
-nsim = 200
+nsim = 20
 n = 100
 sigma = 1
 
-groups = 1:500
+groups = 1:1000
 p = length(groups)
 mult = sqrt(2*log(p))
 upper = 1.1*mult
 lower = 0.9*mult
-num.nonzero = 20
-max.steps = 30
+num.nonzero = 8
+max.steps = 12
 beta = beta_staircase(groups, num.nonzero, upper, lower)
 
 pdf('figs/large_sim_gaussian_lasso.pdf')
@@ -25,10 +25,10 @@ dev.off()
 groups = sort(c(rep(1:50, 5), rep(51:75, 10)))
 p = length(groups)
 mult = sqrt(2*log(p))
-upper = 1.1*mult
-lower = 0.9*mult
-num.nonzero = 6
-max.steps = 9
+upper = 3.1*mult
+lower = 2.5*mult
+num.nonzero = 2
+max.steps = 4
 beta = beta_staircase(groups, num.nonzero, upper, lower)
 
 pdf('figs/large_sim_gaussian_design.pdf')

### 2013-10-11 sim_small.R

@@ -4,7 +4,7 @@ source('generate_data.R')
 source('fwd_step_test.R')
 source('tex_table.R')
 
-nsim = 500
+nsim = 40
 n = 50
 sigma = 1
 groups = c(1, 1, 2, 3, 3, 3, 4, 4, 4, 4, 5, 5, 6, 6, 7, 7, 7, 8, 9, 9, rep(10, 10))

### 2013-11-01 fwd_step/sim_fwdstep.R
Nothing to see here

@@ -1,0 +1,45 @@
+
+setwd('..')
+
+source('fwd_step.R')
+source('generate_data.R')
+source('fwd_step_sim.R')
+source('tex_table.R')
+
+nsim = 25
+sigma = 1
+n = 100
+p.list = c(25, 50, 200, 400)
+k.list = c(3, 4, 5, 6, 7, 8, 9, 10)
+max.steps = max(k.list) + 8
+mu.mat = matrix(Inf, nrow = length(k.list), ncol = length(p.list))
+mu.mat.sd = matrix(Inf, nrow = length(k.list), ncol = length(p.list))
+pow.mat = matrix(0, nrow = length(k.list), ncol = length(p.list))
+pow.mat.sd = matrix(0, nrow = length(k.list), ncol = length(p.list))
+
+for (p in p.list) {
+  j = which(p.list == p)
+  groups = 1:p
+  mult = 1*sqrt(2*log(p))
+  upper = 1.2*mult
+  lower = 1.0*mult  
+  for (k in k.list) {
+    i = which(k.list == k)
+    beta = beta_staircase(groups, num.nonzero = k, upper, lower)
+    output <- fwd_group_simulation(n, sigma, groups, beta, nsim, max.steps, rand.beta = TRUE, coherence = TRUE, plot = FALSE)
+    mus = output$coherence
+    pows = rowSums(output$true.step[, 1:k])/k
+    mu.mat[i, j] = mean(mus)
+    mu.mat.sd[i, j] = sd(mus)
+    pow.mat[i, j] = mean(pows)
+    pow.mat.sd[i, j] = sd(pows)
+    cat(paste("Finished k =", k, "\n"))
+  }
+  cat(paste("Finished p =", p, "\n"))
+}
+
+
+#caption = "Evaluation of model selection using several stopping rules based on our p-values. The naive stopping rule performs well."
+#results = with(output, sim_select_stats(signal.p, active.set, true.step, m1))
+#file = "large_sim_selection____________what.tex"
+#tex_table(file, results, caption = caption)

### 2013-12-01 paper/paper.tex
Nothing to see here
Adding to fwd stepwise section of paper

@@
 variables may be pre-designated factors such as expression
-measurements for all genes in a single regulatory pathway. An entire
+measurements for all genes in a single functional pathway. An entire
 group is included or excluded together from the final model. For
@@
 the number of individual variables in group $g$. Define $p = \sum_{g=1}^G
-p_g$ be the total number of individual variables, so $p = G$ in the
+p_g$ the total number of individual variables, so $p = G$ in the
 case where all groups have size 1. Let $X$ be the matrix constructed
@@
 are ready to describe the forward stepwise procedure, given in Algorithm
-\ref{algo:fs}. Note that to prevent saturation, and to enable our p-value
-computations, \textit{max.steps} should be at most $\min (n, G) - 1$.
+\ref{algo:fs}. Note that to enable our p-value computations,
+\textit{max.steps} should be at most $\min (n, G) - 1$. In our
+implementation we treat the active set $A$ as an ordered set to easily
+track the order of variables entering the model.
 
@@
   $A^c \gets A^c \backslash \{ g^* \}$\;
+  \For{$g \in A^c$}{
+    $X_g \gets (I - X_{g^*}X_{g^*}^\dagger) X_g$;
+  }
   $r_g \gets $\lsfitResidual{$r_{g-1}, X_{g^*}$}\;
@@
 forward. Some classical approaches can be posed as optimization
-criteria and stop at the step minimizing
+criteria which stop at the step minimizing
 \begin{equation}
@@
 \citep{fdr}. We explore this further in Section \ref{sec:simulations}.
+
+Although forward stepwise is a greedy algorithm producing a
+potentially sub-optimal sequence of models, under favorable conditions
+it can still perform well. There is a parallel literature to the
+compressed sensing literature \textit{(cite some here)} with theorems
+stating that forward stepwise can exactly select the true model under
+some stringent conditions involving quantities like the sparsity of
+the true model and the coherence of the design matrix. We do not
+pursue this any further because we find empirically that forward
+stepwise can work well in much less restrictive situations than those
+guaranteed by theorems.
+
+To explore the performance of forward stepwise empirically we
+conducted simulations with results shown in ~\ref{fig:fwdstepsims}
+
+\begin{figure}
+\begin{center}
+\subfigure[Gaussian design]{
+%\label{fig:smallgauss}
+\includegraphics[width=0.5\textwidth]{../figs/fwdstep/gaussian_p2n_n100_lower1p4_upper1p6.pdf}}
+\hspace{-15pt}
+\subfigure[Categorical design]{
+%\label{fig:smallcat}
+\includegraphics[width=0.5\textwidth]{../figs/fwdstep/gaussian_p2n_n100_lower1p4_upper1p6_corr0p2.pdf}}
+\caption{\small \it The left panel shows results of a simulation with
+  an independent Gaussian design matrix and a signal vector having
+  three groups with non-zero coefficients of sizes one, two, and
+  three, and seven groups of noise variables with sizes varying from
+  one to ten, for a total of 30 variables. For the right panel we
+  repeated this setup, but group sizes corresponded to levels of
+  categorical variables. The groups of size one were increased to two,
+so there are 10 categorical variables with a total of 32 levels.}
+\label{fig:fwdstepsim}
+\end{center}
+\end{figure}
 
@@
 \label{fig:smallgauss}
-\includegraphics[width=0.5\textwidth]{../figs/small_sim_gaussian_design.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/gaussian_size1_n100_p200_g200_k10_lower1p8_upper1p9.pdf}}
 \hspace{-15pt}
@@
 \label{fig:smallcat}
-\includegraphics[width=0.5\textwidth]{../figs/small_sim_categorical_design.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/gaussian_size1_n100_p200_g200_k10_lower1p8_upper1p9_corr0p1.pdf}}
 \caption{\small \it The left panel shows results of a simulation with
@@
 \label{fig:largelasso}
-\includegraphics[width=0.5\textwidth]{../figs/large_sim_gaussian_lasso.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/gaussian_size5-10_n100_p200_g35_k10_lower1p8_upper1p9.pdf}}
 \hspace{-15pt}
@@
 \label{fig:largegrplasso}
-\includegraphics[width=0.5\textwidth]{../figs/large_sim_gaussian_design.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/gaussian_size5-10_n100_p200_g35_k10_lower1p8_upper1p9_corr0p1.pdf}}
 \caption{\small \it The left panel shows results of a simulation with

### 2013-12-01 fwd_step/fstep_sim_gaussian.R

@@ -15,8 +15,10 @@ source('coherence.R')
 fname = "p4n"          ##
 ns = c(100) #, 200)    ##
 ps = ceiling(ns*4)     ##
-nsim = 50              ##
-kplus = 8              ##
+nsim = 75              ##
+kplus = 10             ##
+##########################
+corr = .1
 ##########################
 mus = n_to_mu(ns)
 ks = mu_to_k(mus)
@@ -28,8 +30,8 @@ for (i in 1:length(ns)) {
   n = ns[i]
   p = ps[i]
   mult = sqrt(2*log(p))
-  upper.coeff = 1.4
-  lower.coeff = 1.4
+  upper.coeff = 1.9
+  lower.coeff = 1.7
   upper = upper.coeff*mult
   lower = lower.coeff*mult
   mu = mus[i]
@@ -48,7 +50,7 @@ for (i in 1:length(ns)) {
       # Generate data
       beta = beta_staircase(groups, k, upper, lower, rand.sign=TRUE, permute=TRUE)
       true.groups = true_active_groups(groups, beta)
-      X = gaussian_design(n, groups)
+      X = gaussian_design(n, groups, corr = corr)
       Y = X %*% beta + rnorm(n)
       
       # Calculate coherence
@@ -68,8 +70,18 @@ for (i in 1:length(ns)) {
 
   # Save plot
   plot.main = paste0("n = ", n, ", p = ", p, ", signal strength ", lower.coeff, "/", upper.coeff)
-  pdf(paste0('../figs/fwdstep/gaussian_', fname, "_n", n, "_lower", lower.coeff, "_upper", upper.coeff, ".pdf"))
+  filename = paste0('../figs/fwdstep/gaussian_', fname, "_n", n, "_lower", lower.coeff, "_upper", upper.coeff)
+  if (corr != 0) {
+    filename = paste0(filename, "_corr", corr)
+    plot.main = paste0(plot.main, ", corr = ", corr)
+  }
+  filename = paste0(filename, ".pdf")
+  
+  pdf(filename)
   plot(klist, power.MCavg, type = "l", main = plot.main, xlab = "Sparsity", ylab = "Average power", ylim = c(-0.1, 1.1), lwd = 2)
+  abline(h = c(.9, .7, .5, .3, .1), lty = 2, col = "gray")
+  abline(h = c(1, .8, .6, .4, .2, 0), lty = 3, col = "gray")
+  
[…]

### 2013-12-01 sim_large.R

@@ -6,13 +6,15 @@ source('tex_table.R')
 
 
 design = 'gaussian'
-corr = .1 # nonzero only supported for gaussian design
-nsim = 500
+corr = 0 # nonzero only supported for gaussian design
+
+nsim = 400
 n = 100
 num.nonzero = 10
-max.steps = 12
-upper.coeff = 1.8
-lower.coeff = 1.8
+k = num.nonzero
+max.steps = 13
+upper.coeff = 1.9
+lower.coeff = 1.1
 
 sigma = 1
 groups = 1:200
@@ -57,8 +59,13 @@ caption = "Evaluation of model selection using several stopping rules based on o
 results.l = with(output.l, sim_select_stats(signal.p, active.set, true.step, m1))
 results.g = with(output.g, sim_select_stats(signal.p, active.set, true.step, m1))
 
-rownames(results.l) = paste("(1)", rownames(results.l))
-rownames(results.g) = paste("(2)", rownames(results.g))
+#rownames(results.l) = paste("(1)", rownames(results.l))
+rownames(results.g) = paste("(g)", rownames(results.g))
 
-file = paste0(design, "_large_selection.tex")
+file = paste0("tables/", design, "_n", n, "_p", p, "_k", k, "_lower", lower.coeff, "_upper", upper.coeff)
+if (corr != 0) {
+  file = paste0(file, "_corr", corr)
+}
+file = paste0(file, ".tex")
+
 tex_table(file, rbind(results.l, results.g), caption = caption)

### 2013-12-05 paper/paper.tex
Chipping away

@@
 throughout. Since single variables can be considered groups of size 1,
-our general exposition includes non-grouped situations as a special case.
+our general exposition includes non-grouped situations as a special
+case. Our variant of forward stepwise uses a different method than
+usual for choosing which group to add because we are using the
+same objective in the forward stepwise procedure that was used to
+derive our test statistic.
 
@@
 
-We allow for the possibility $p > n$ in which case \ref{eq:gmodel} is
+We allow for the possibility $p > n$ in which case ~\ref{eq:gmodel} is
 generically underdetermined. In such cases it still may be possible to
@@
 \ref{algo:fs}. Note that to enable our p-value computations,
-\textit{max.steps} should be at most $\min (n, G) - 1$. In our
-implementation we treat the active set $A$ as an ordered set to easily
-track the order of variables entering the model.
+\textit{max.steps} should be at most $\min (n, G) - 1$. If the noise
+variance is known--as we assume unless specified otherwise--and
+$\beta$ is expected to be quite sparse, then \textit{max.steps} can be
+much smaller as long as it is still above the expected sparsity of
+$\beta$. This last constraint is suggested so that forward stepwise
+still has a chance of recovering all the nonzero coefficients of
+$\beta$ before reaching \textit{max.steps}. In our implementation we
+treat the active set $A$ as an ordered set to easily track the order
+of variables entering the model.
 
@@
 $r_0 \gets y$\;
-\For{$step \leftarrow 1$ \KwTo $max.steps$}{
-  $g^* \gets \argmax_{g \in A^c} \norm{X_g^T r_{g-1}}_2 / w_g$\;
+\For{$s \gets 1$ \KwTo $max.steps$}{
+  $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
+  $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
   $A \gets A \cup \{ g^* \}$\;
@@
   \For{$g \in A^c$}{
-    $X_g \gets (I - X_{g^*}X_{g^*}^\dagger) X_g$;
+    $X_g \gets P X_g$\;
   }
-  $r_g \gets $\lsfitResidual{$r_{g-1}, X_{g^*}$}\;
+  $r_s \gets P r_{s-1}$\;
 }
@@
 
-Subset selection refers to the problem of choosing a subset of the $p$
-potential variables to include in a model. Since there are $2^p$ such
-choices, exhaustive search is computationally infeasible when $p$ is
+Subset selection refers to the problem of choosing a subset of the $G$
+potential groups to include in a model. Since there are $2^G$ such
+choices, exhaustive search is computationally infeasible when $G$ is
 large. If exhaustive search were possible, performing a large
 search runs the risk of over-fitting unless model complexity is
-appropriately penalized. Algorithm \ref{algo:fs} produces a much
-smaller set of potential models, at most \textit{max.steps}. However,
+appropriately penalized. Forward stepwise produces a much
+smaller set of potential models, with cardinality at most \textit{max.steps}. However,
 it is a greedy algorithm so the set of models it produces may not
-contain the best possible model. This is an inherent shortcoming which
-we do not address in this paper.
+contain the best possible model. This is an inherent shortcoming of
+forward stepwise procedures and should be kept in mind when choosing
+between model selection methods.
 
@@
 some stringent conditions involving quantities like the sparsity of
-the true model and the coherence of the design matrix. We do not
-pursue this any further because we find empirically that forward
-stepwise can work well in much less restrictive situations than those
-guaranteed by theorems.
-
-To explore the performance of forward stepwise empirically we
-conducted simulations with results shown in ~\ref{fig:fwdstepsims}
+the true model and the coherence of the design matrix. Recall that the
+coherence $\mu(X)$ of a matrix $X$ with columns scaled to have unit
+2-norm is defined as
+
+\begin{equation}
+  \mu(X) = \max_{i \neq j} \{ | \innerp{ X_i }{ X_j } | \}
+\end{equation}
+
+Writing $k$ for the sparsity of $\beta$, typical results in this
+literature say that if $k < (1/\mu + 1)/2$ and the nonzero
+coefficients of $\beta$ are sufficiently large then forward stepwise
+recovers $\beta$ with high probability. We refer the
+reader to the literature for details. For our purposes the
+conditions required to guarantee exact recovery are usually far too
+stringent. We conduct simulations to show empirically that forward
+stepwise can work well even when it is not working perfectly, and that
+it does so under a much wider range of conditions.
 
@@
 \begin{center}
-\subfigure[Gaussian design]{
-%\label{fig:smallgauss}
-\includegraphics[width=0.5\textwidth]{../figs/fwdstep/gaussian_p2n_n100_lower1p4_upper1p6.pdf}}
+\subfigure[Ternary design]{
+\label{fig:fsternary}
+\includegraphics[width=0.5\textwidth]{../figs/fwdstep/ternary_p2n_n100_lower1p4_upper1p4.pdf}}
 \hspace{-15pt}
-\subfigure[Categorical design]{
-%\label{fig:smallcat}
+\subfigure[Gaussian design with correlation]{
+\label{fig:fsgausscorr}
 \includegraphics[width=0.5\textwidth]{../figs/fwdstep/gaussian_p2n_n100_lower1p4_upper1p6_corr0p2.pdf}}
-\caption{\small \it The left panel shows results of a simulation with
-  an independent Gaussian design matrix and a signal vector having
-  three groups with non-zero coefficients of sizes one, two, and
-  three, and seven groups of noise variables with sizes varying from
-  one to ten, for a total of 30 variables. For the right panel we
-  repeated this setup, but group sizes corresponded to levels of
-  categorical variables. The groups of size one were increased to two,
-so there are 10 categorical variables with a total of 32 levels.}
+\caption{\small \it The left panel shows the results of the simulation
+  using design matrices with i.i.d. ternary entries taking values $0,
+  \pm 1$ with equal probability. The left panel shows the results
+  with design matrices of standard Gaussian entries with
+  equi-correlation of 0.2. }
 \label{fig:fwdstepsim}
@@
 \end{figure}
+
+For various sparsity levels $k$, Algorithm ~\ref{algo:fs} was run on
+various data sets with max.steps equal to $k$ and the ``average
+power'' was calculated as the number of true variables in the active
+set after $k$ steps divided by $k$. Nonzero coefficients had magnitude
+in a range of multiples of $\sqrt{2\log(p)}$, e.g. in $[1.4
+\sqrt{2\log(p)}, 1.6 \sqrt{2\log(p)}]$. Results are shown in
+Figure~\ref{fig:fwdstepsim}. Some calculations show the required
+sparsity level to guarantee exact recovery in these situations is
+about 2 or smaller, and the required nonzero coefficient magnitude is
+likely in the range of 10 to 100 times $\sqrt{2\log(p)}$.
 
@@
 active set since the analysis then follows almost exactly as for the
-lasso. (\todo Change this if I can find $\lambda_k$).
-
-The \emph{group lasso estimator} is the following solution to a convex
+lasso.
+
+The \emph{group lasso estimator} is a solution to the following convex
 problem

### 2013-12-05 fwd_step/fstep_minimal.R

@@ -1,4 +1,3 @@
-
 # A minimal version of forward stepwise for phase diagram simulations
 
 library(MASS)
@@ -12,7 +11,6 @@ update_active_set = function(active.set, group) {
   }
   return(new.active.set)
 }
-
 
 # Determine which group to add next
 add_group = function(X, R, groups, weights, active.set) {
@@ -34,7 +32,6 @@ add_group = function(X, R, groups, weights, active.set) {
   imax = which.max(terms)
   return(imax)
 }
-
 
 
 # Input data, output sequence of added variables
@@ -63,6 +60,7 @@ fstep_fit = function(X, Y, groups, max.steps = 0) {
     Xgmax.regress = Xgmax
     Pgmax = Xgmax %*% ginv(Xgmax)
     Hgmax = diag(rep(1, n)) - Pgmax
+    ### I think this part is never needed
     # If X[, gmax] is categorical, leave one column out
     #if (sum(gmax) > 1) {
     #  if (length(unique(rowSums(Xgmax))) == 1) {
@@ -71,7 +69,6 @@ fstep_fit = function(X, Y, groups, max.steps = 0) {
     #}
     R = R - Pgmax %*% R
 
-    ####### This is necessary for p-value? ########
     # Project all other groups orthogonal to the one being added
     for (gind in 1:max(groups)) {
       if (gind != next.step) {
@@ -79,6 +76,7 @@ fstep_fit = function(X, Y, groups, max.steps = 0) {
         X.project[, group] = Hgmax %*% X[, group]
       }
     }
+    # Re-normalize?
     X.project = X.project %*% diag(1/sqrt(colSums(X.project^2)))
     X = X.project
   }

### 2013-12-05 fwd_step/fstep_sim_ortho.R

@@ -6,19 +6,21 @@ source('coherence.R')
 
 ##########################
 ### Modify these lines ###
-#fname = "p.5n"         ##
-#ns = c(100) #c(100, 200, 400)  ##
-#ps = floor(ns/2)       ##
-#nsim = 80              ##
-#kplus = 10             ##
-##########################
+#fname = "pp5n"           ##
+#ns = c(100) #, 200, 400)   ##
+#ps = floor(ns/2)        ##
+#nsim = 75               ##
+#kplus = 20              ##
+#########################
 fname = "p2n"          ##
-ns = c(100) #, 200)       ##
+ns = c(100) #, 200)    ##
 ps = ceiling(ns*2)     ##
-nsim = 80              ##
-kplus = 10              ##
+nsim = 75              ##
+kplus = 10             ##
 ##########################
-#mus = n_to_mu(ns)
+corr = 0
+##########################
+mus = n_to_mu(ns)
 ks = mu_to_k(mus)
 
 for (i in 1:length(ns)) {
@@ -28,15 +30,15 @@ for (i in 1:length(ns)) {
   n = ns[i]
   p = ps[i]
   mult = sqrt(2*log(p))
-  upper.coeff = 1.4
-  lower.coeff = 1.4
+  upper.coeff = 1.1
+  lower.coeff = 1.1
   upper = upper.coeff*mult
   lower = lower.coeff*mult
-#  mu = mus[i]
+  mu = mus[i]
   kmin = 1 # ks[i]
   kmax = kmin + kplus
   # Something
-  klist = kmin:kmax
+  klist = 0:(kmax-1)
   klist = 2*klist + 1
   groups = 1:p
 
@@ -48,11 +50,10 @@ for (i in 1:length(ns)) {
       # Generate data
       beta = beta_staircase(groups, k, upper, lower, rand.sign=TRUE, permute=TRUE)
       true.groups = true_active_groups(groups, beta)
-      X = ternary_design(n, groups)
+      X = orthogonal_design(n, groups)
       Y = X %*% beta + rnorm(n)
       
       # Calculate coherence
-# slow slow slow slow slow slow slow slow slow slow slow slow
 #      mu.mat[j, iter] = coherence(X)
       
       # Fit forward stepwise
@@ -62,15 +63,25 @@ for (i in 1:length(ns)) {
   }
 
[…]

### 2013-12-05 generate_data.R

@@ -114,6 +114,27 @@ ternary_design = function(n, groups) {
   X = col_normalize(X)
 }
 
+orthogonal_design = function(n, groups) {
+  # If n > p, orthogonalize
+  # If p = k*n, orthogonalize the k-submatrices
+  # Else stop(error)
+  X = gaussian_design(n, groups)
+  if (n > p) {
+    X = svd(X)$u
+  } else {
+    if (p %% n == 0) {
+      k = p/n
+      for (j in 0:(k-1)) {
+        X[ , (j+1):(j+n)] = svd(X[ , (j+1):(j+n)])$u
+      }
+    } else {
+      stop("p not a multiple of n")
+    }
+  }
+  X = col_normalize(X)
+  return(X)
+}
+
 # Fixed group sizes, categorical design
 # Important: binary requires two indices in groups, e.g. c(1,1,...)
 categorical_design = function(n, groups, ortho.within = FALSE) {

### 2013-12-05 sim_small.R

@@ -4,53 +4,68 @@ source('generate_data.R')
 source('fwd_step_sim.R')
 source('tex_table.R')
 
-nsim = 50
+
+design = 'orthogonal'
+corr = 0 # nonzero only supported for gaussian design
+
+nsim = 500
 n = 100
+num.nonzero = 8
+k = num.nonzero
+max.steps = 12
+upper.coeff = 1.5
+lower.coeff = 1.1
+
 sigma = 1
-groups = c(1, 1, 2, 3, 3, 3, 4, 4, 4, 4, 5, 5, 6, 6, 7, 7, 7, 8, 9, 9, rep(10, 10))
+groups = 1:50
 p = length(groups)
-mult = 1*sqrt(2*log(p))
-upper = 1.5*mult
-lower = 1.1*mult
-num.nonzero = 3
-max.steps = 8
+mult = sqrt(2*log(p))
+upper = upper.coeff*mult
+lower = lower.coeff*mult
+
+if ((corr != 0) & (design != 'gaussian')) {
+  stop("nonzero only supported for gaussian design")
+}
+
 beta = beta_staircase(groups, num.nonzero, upper, lower)
-beta_old = beta
-
-pdf('figs/small_sim_gaussian_design.pdf')
-output.g <- fwd_group_simulation(n, sigma, groups, beta, nsim, max.steps, predictions = TRUE, plot = TRUE)
+filename = paste0('figs/', design, '_size1_n', n, '_p', p, '_g', p, '_k', num.nonzero, '_lower', lower.coeff, '_upper', upper.coeff)
+if (corr != 0) {
+  filename = paste0(filename, '_corr', corr)
+}
+filename = paste0(filename, '.pdf')
+pdf(filename)
+output.l <- fwd_group_simulation(n, sigma, groups, beta, nsim, max.steps, design = design, corr = corr, rand.beta = TRUE, plot = TRUE)
 dev.off()
-warnings()
 
 
-groups = sort(c(groups, 2, 8))
+groups = sort(c(rep(1:8, 5), 9:18))
 p = length(groups)
-mult = 1*sqrt(2*log(p))
-upper = 1.5*mult
-lower = 1.1*mult
-beta = beta_staircase(groups, num.nonzero, upper, lower, rand.sign = TRUE, perturb = TRUE, cat.vars = 1:max(groups))
-warnings()
-
-pdf('figs/small_sim_categorical_design.pdf')
[…]

### 2013-12-05 fwd_step.R
Possible fix of categorical variable problem

@@ -46,16 +46,15 @@ add_group = function(X.orig, X, Y, groups, weights, sigma, active.set = 0, eff.p
   X.project = X
   Xgmax = X[, gmax]
   Xgmax.regress = Xgmax
-  Pgmax = Xgmax %*% ginv(Xgmax)
   # If X[, gmax] is categorical, leave one column out
-
-  
   ########### Is this doing anything? Xgmax.regress isn't used ###########
   if (sum(gmax) > 1) {
-    if (length(unique(rowSums(Xgmax))) == 1) {
+    if (length(unique(rowSums(Xgmax == 0))) == 1) {
+#      print("categorical variable")
       Xgmax.regress = Xgmax[ , -1]
     }
   }
+  Pgmax = Xgmax.regress %*% ginv(Xgmax.regress)
   Y.resid = (diag(rep(1,n)) - Pgmax) %*% Y
       
 ####### This is necessary for p-value? ########
@@ -66,7 +65,8 @@ add_group = function(X.orig, X, Y, groups, weights, sigma, active.set = 0, eff.p
       X.project[, group] = (diag(rep(1, n)) - Pgmax) %*% X[, group]
     }
   }
-  X.project = X.project %*% diag(1/sqrt(colSums(X.project^2)))
+  # Renormalize
+#  X.project = X.project %*% diag(1/sqrt(colSums(X.project^2)))
   
   return(list(test.output = results, var = results$var, p.value = p.value, added = imax, active.set = new.active.set, eff.p = new.eff.p, Y.update = Y.resid, X.update = X.project))
 }

### 2013-12-05 fwd_step/fstep_sim_gaussian.R

@@ -12,13 +12,13 @@ source('coherence.R')
 #nsim = 60               ##
 #kplus = 20              ##
 #########################
-fname = "p4n"          ##
+fname = "p2n"          ##
 ns = c(100) #, 200)    ##
-ps = ceiling(ns*4)     ##
+ps = ceiling(ns*2)     ##
 nsim = 75              ##
 kplus = 10             ##
 ##########################
-corr = .1
+corr = 0
 ##########################
 mus = n_to_mu(ns)
 ks = mu_to_k(mus)
@@ -30,8 +30,8 @@ for (i in 1:length(ns)) {
   n = ns[i]
   p = ps[i]
   mult = sqrt(2*log(p))
-  upper.coeff = 1.9
-  lower.coeff = 1.7
+  upper.coeff = 1.5
+  lower.coeff = 1.5
   upper = upper.coeff*mult
   lower = lower.coeff*mult
   mu = mus[i]
@@ -75,7 +75,7 @@ for (i in 1:length(ns)) {
     filename = paste0(filename, "_corr", corr)
     plot.main = paste0(plot.main, ", corr = ", corr)
   }
-  filename = paste0(filename, ".pdf")
+  filename = paste0(filename, "_no_renormalize.pdf")
   
   pdf(filename)
   plot(klist, power.MCavg, type = "l", main = plot.main, xlab = "Sparsity", ylab = "Average power", ylim = c(-0.1, 1.1), lwd = 2)

### 2013-12-05 fwd_step/fstep_sim_ternary.R

@@ -67,7 +67,7 @@ for (i in 1:length(ns)) {
 
   # Save plot
   plot.main = paste0("n = ", n, ", p = ", p, ", signal strength ", lower.coeff, "/", upper.coeff)
-  pdf(paste0('../figs/fwdstep/ternary_', fname, "_n", n, "_lower", lower.coeff, "_upper", upper.coeff, ".pdf"))
+  pdf(paste0('../figs/fwdstep/ternary_', fname, "_n", n, "_lower", lower.coeff, "_upper", upper.coeff, "_no_renormalize.pdf"))
   plot(klist, power.MCavg, type = "l", main = plot.main, xlab = "Sparsity", ylab = "Average power", ylim = c(-0.1, 1.1), lwd = 2)
   abline(h = c(.9, .7, .5, .3, .1), lty = 2, col = "gray")
   abline(h = c(1, .8, .6, .4, .2, 0), lty = 3, col = "gray")

### 2014-03-07 paper/paper.tex
Possible fix of categorical variable problem
Fixed some issues with glinternet, prevent groups already counted from being added again

--- 2013-12-01 paper/paper.tex
@@
 % "Title of the paper"
-\title{A significance test in forward stepwise model selection}
+\title{A significance test in forward stepwise regression}
 
@@
 % indicate corresponding author with \corref{}
-\author{\fnms{Jonathan} \snm{Taylor}\corref{Jonathan E. Taylor}\ead[label=e1]{jonathan.taylor@stanford.edu}\thanksref{t1}
-\and \fnms{Joshua} \snm{Loftus}
-\and ???}
-\thankstext{t1}{Supported in part by NSF grant DMS 1208857 and
-AFOSR grant 113039.}
+%\author{\fnms{Jonathan} \snm{Taylor}\corref{Jonathan E. Taylor}\ead[label=e1]{jonathan.taylor@stanford.edu}\thanksref{t1}
+%\and \fnms{Joshua} \snm{Loftus}}
+%\thankstext{t1}{Supported in part by NSF grant DMS 1208857 and AFOSR grant 113039.}
 
@@
 \address{Department of Statistics\\  Stanford University\\ Sequoia
-Hall \\390 Serra Mall\\ Stanford, CA 94305, U.S.A.\\ \printead{e1}}
+Hall \\390 Serra Mall\\ Stanford, CA 94305, U.S.A.\\ }
+%\printead{e1}}
 
@@
 % perhaps de-emphasize group lasso and emphasize forward stepwise?
-  We apply the methods of \cite{tests:adaptive} and
-  \cite{significance:lasso} on significance tests for penalized
-  regression to forward stepwise model selection. For the $k$th
-  variable to enter the model, previous work relied on an asymptotic
-  null distribution that breaks down when grouped variables are
-  entering the model and is difficult to derive. We iteratively apply
-  an exact null distribution for the first (potentially grouped)
-  variable on the residual from previous steps. The resulting method
-  has the strengths of stepwise selection, for example parallel
-  computation, but also remedies the problem of inflated test
-  statistics and over-fitting.
+  We extend the methods developed by \cite{significance:lasso} and
+  \cite{tests:adaptive} on significance tests for penalized
+  regression to stepwise model selection by iteratively applying
+  the global null hypothesis test on the residual in a forward stepwise
+  procedure. While not fully theoretically justified, the resulting method
+  has the computational strengths of stepwise selection and partially
+  solves the problem of inflated test statistics due to model selection.
+  We illustrate the flexibility of this method by applying it to
+  novel specialized applications of forward stepwise to a hierarchically
+  constrained interactions model and generalized additive models.
 \end{abstract}
@@
 %%%
-%%% - Gotta change these
+%%% - Gotta change these??????????
 %%%
@@
 \begin{keyword}
-\kwd{penalized regression}
-\kwd{convex analysis}
-\kwd{least squares}
-\kwd{Gaussian processes}
-%\kwd{extreme value theory}
+\kwd{forward stepwise}
+\kwd{model selection}
+\kwd{significance test}
 \end{keyword}
@@
 
-Forward stepwise regression is a stochastic model selection procedure
-that begins with an empty model and adds the best predictor variable
-in each step.  Classical significance tests fail when a model has been
-selected this way and tend to be anti-conservative.  Recently,
-\cite{significance:lasso} found a novel test statistic with an
-appropriate null distribution that behaves well when a model has been
-selected using the lasso \citep{tibshirani:lasso}.
+Forward stepwise regression is a classical model selection procedure
+that begins with an empty model and sequentially adds the best predictor
+variable in each step. Classical significance tests fail
+when a model
+has been selected this way and tend to be anti-conservative.
+In the lasso setting, \cite{significance:lasso} derived a novel test
+statistic along with a correct asymptotic distribution, making possible
+valid inferences after model selection.
 \cite{tests:adaptive} modified and extended those results to the
-group lasso \citep{grouplasso} and other adaptive regression
-problems.  The present work explores the behavior of those test
-statistics for models selected by forward stepwise procedures and
-works out some of the details involved in applying these methods to
-models with grouped variables.  Our test statistic can be
-used for valid significance testing when computed from the same
-data as the model selection.  The resulting method can
+group lasso \citep{grouplasso} and other penalized regression
+problems, but only under the global null hypothesis.
+One of the strengths of these test statistics is that they can be
+used for valid significance testing when computed on the same
+data used for model selection, eliminating the need for data splitting.
+The present work iteratively applies the global null test
+of \cite{tests:adaptive} for each step in forward stepwise selection,
+and works out some of the details necessary for
+models with grouped variables. The resulting method can
 be more statistically efficient than validation on held-out data, and
-also more computationally efficient than penalized methods with
+more computationally efficient than penalized methods with
 regularization parameters chosen by cross-validation.
 
-(\todo Change this paragraph as the sections are completed)
-In Section \ref{sec:stepwise} we establish notation and describe our
-forward stepwise procedure. Section \ref{sec:testing} briefly reviews
+In Section~\ref{sec:stepwise} we establish notation and describe the
+forward stepwise algorithm used in most of the paper.
+Section~\ref{sec:testing} briefly reviews
 the parts of \cite{significance:lasso} and \cite{tests:adaptive}
 relevant to our significance test, and describes the group lasso which
-we require for applying the test with grouped
-variables. Simulation results in Section \ref{sec:simulations} using
-various stopping rules for the forward stepwise procedure--including
-some from \cite{sequential:fdr}--appear promising.
-
+we require for applying the test with grouped variables.
+Simulation results in Section~\ref{sec:simulations} show
+empirically that our method performs well in settings where forward
+stepwise itself performs well, and that
+various stopping rules using our test statistic--including
+some from \cite{sequential:fdr}--appear promising. In
+Section~\ref{sec:applications} we apply
+the method to several variants of forward
+stepwise tailored to models with interactions and generalized additive
+models, as well as to a real data example.
+(\todo Change this paragraph if necessary as the sections are completed)
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \label{sec:stepwise}
+
+\subsection{Background and notation}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.9\textwidth]{../figs/pubmed.pdf}
+\caption{\small \it Forward stepwise enjoys widespread use among
+practitioners}
+\label{fig:pubmed}
+\end{center}
+\end{figure}
+
+As a classical method dating back about half a century
+(see \cite{classical:selection} for a review),
+forward stepwise regression has not received much attention in
+recent years in the theoretical statistics community. But it continues
+to be widely used by practitioners.
+For example, search results on PubMed for forward stepwise
+(summarized in Figure~\ref{fig:pubmed}) show that many recent
+papers mention the method and there is an increasing trend over time.
+Its popularity among researchers continues despite the fact that it
+invalidates inferences using the standard $\chi^2$ or $F$-tests.
+
+Some classical attempts to address this issue include Monte Carlo
+estimation of tables of adjusted $F$-statistic values \citep{mc:ftoenter}
+and permutation statistics \citep{permutation:stop}. Aside from the works
+this paper is based on, there have been other recent attempts to do
+inference after model selection. Most of these make use of subsampling
+\citep{meinshausen:buhlmann} or data splitting \citep{wasserman:roeder}.
+Our approach allows use of the full data and does not require the
+extra computation involved in subsampling.
+Before describing the full approach we first introduce notation and
+specify our implementation of forward stepwise, which is slightly
+different from the most commonly used versions.
+(\todo Check R's \textit{step} for the group criterion)
 
@@
 variables may be pre-designated factors such as expression
-measurements for all genes in a single functional pathway. An entire
-group is included or excluded together from the final model. For
-consistency we will use $g, h$ as indices rather than the usual $i, j$
-throughout. Since single variables can be considered groups of size 1,
-our general exposition includes non-grouped situations as a special case.
-
-Label the outcome variable of $n$ i.i.d. measurements $y \in \real^n$. Let an integer $G \geq 1$ 
-be the number of groups. For each $1 \leq g \leq G$ the design matrix encoding the
+measurements for all genes in a single functional pathway.
+To emphasize this we will use $g, h$ as indices rather than the usual
+$i, j$ throughout. Since single variables can be considered groups of
+size 1, our general exposition includes non-grouped situations as a special
+case. Our variant of forward stepwise uses a different method than
+usual for choosing which group to add because we are using the
+same objective in the forward stepwise procedure that was used to
+derive the test statistics in penalized regression settings.
+
+Denote $y \in \real^n$ for $n$ i.i.d. measurements of the outcome variable.
+Let an integer $G \geq 2$ be the number of groups of explanatory variables.
+For each $1 \leq g \leq G$ the design matrix encoding the
 $g$th group is the $n \times p_g$ matrix denoted $X_g$, where $p_g$ is
-the number of individual variables in group $g$. Define $p = \sum_{g=1}^G
-p_g$ the total number of individual variables, so $p = G$ in the
+the number of individual variables or columns in group $g$. When a group
+encodes a categorical variable as binary indicators for the levels of
+that variable, we use the full encoding with a column for every level.
+Although this introduces collinearity, our method does not require the
+design matrix to have full rank.
+
+Define $p = \sum_{g=1}^Gp_g$,
+the total number of individual variables, so $p = G$ in the
 case where all groups have size 1. Let $X$ be the matrix constructed
 by column-binding the $X_g$, that is  
-%Throughout most of this paper we assume orthogonality within groups, that is
-%$X_g^T X_g = I_{p_g \times p_g}$ for all $g$. ( \todo Take this out if
-%we don't actually need it).  
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+We assume throughout that the individual columns of $X$ are scaled to
+have unit 2-norm (we do not scale groups, this is accomplished later
+by weighting).
 With each group we associate the $p_g \times 1$ coefficient vector
@@
 
-We allow for the possibility $p > n$ in which case \ref{eq:gmodel} is
+We allow for the possibility $p > n$ in which case \eqref{eq:gmodel} is
 generically underdetermined. In such cases it still may be possible to
 estimate $\beta$ well if it is sparse--that is, if it has few nonzero
-entries \citep{donoho:pursuit}. In the rest of this paper we refer to
-variable groups $X_g$ as noise variables if $\beta_g$ is a zero vector
-and as true predictor variables if $\beta_g$ has nonzero entries.
+entries. In the rest of this paper we refer to variable groups $X_g$
+as noise groups if $\beta_g$ is a zero vector and as true or signal
+groups if $\beta_g$ has any nonzero entries. We refer to the number
+of such nonzero groups as the \textit{sparsity} of the model, and
+denote this $k := \# \{ g : \beta_g \neq 0 \}$.
 
@@
 more difficult for the group $X_g$ to enter the model. The
-modeler can choose weights arbitrarily, but we will only use one
+modeler can choose weights arbitrarily, but we will mostly use one
 particular choice, based on $p_g$, that we discuss later. With this we
-are ready to describe the forward stepwise procedure, given in Algorithm
-\ref{algo:fs}. Note that to enable our p-value computations,
-\textit{max.steps} should be at most $\min (n, G) - 1$. In our
-implementation we treat the active set $A$ as an ordered set to easily
-track the order of variables entering the model.
+are ready to describe the forward stepwise procedure.
+
+\subsection{Description of the algorithm}
+
+First, the user must specify the maximum number of steps allowed,
+which we denote \textit{steps} (consistent with the \textit{R} function
+for stepwise). To enable our test statistic computations,
+\textit{steps} should be at most $\min (n, G) - 1$, but it is
+computationally desirable to set it as low as possible while still being
+larger than the (unknown) sparsity of $\beta$.
+This way forward stepwise has a chance of recovering all the nonzero
+coefficients of $\beta$ and then terminating without performing much
+additional computation.
+In our implementation we treat the active set $A$ as an ordered list to easily track the order of groups entering the model.
 
@@
 \DontPrintSemicolon
-\KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ grouped variables}
+\KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
 \KwResult{Active set $A$ of variable groups included in the model}
 \SetKwFunction{lsfitResidual}{lsfitResidual}
-\caption{Forward stepwise procedure with groups and weights}
+\caption{\small \it Forward stepwise procedure with groups and weights}
 \BlankLine
@@
 $r_0 \gets y$\;
-\For{$step \leftarrow 1$ \KwTo $max.steps$}{
-  $g^* \gets \argmax_{g \in A^c} \norm{X_g^T r_{g-1}}_2 / w_g$\;
+\For{$s \gets 1$ \KwTo $steps$}{
+  $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
+  $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
   $A \gets A \cup \{ g^* \}$\;
@@
 
-Subset selection refers to the problem of choosing a subset of the $p$
-potential variables to include in a model. Since there are $2^p$ such
-choices, exhaustive search is computationally infeasible when $p$ is
-large. If exhaustive search were possible, performing a large
-search runs the risk of over-fitting unless model complexity is
-appropriately penalized. Algorithm \ref{algo:fs} produces a much
-smaller set of potential models, at most \textit{max.steps}. However,
-it is a greedy algorithm so the set of models it produces may not
-contain the best possible model. This is an inherent shortcoming which
-we do not address in this paper.
+In the common case where all groups are single variables this version
+coincides with the standard implementations of forward stepwise.
+We will now use forward stepwise to refer specifically to
+Algorithm~\ref{algo:fs} unless otherwise specified.
+
+\subsection{Performance of forward stepwise}
+
+Among model selection procedures, forward stepwise is one which performs
+variable selection: from a potentially large set of variables it chooses
+a subset to include in the model. The most general form of variable
+selection is called subset selection, a procedure which searches among
+all $2^G$ subsets of $G$ variable groups and picks the best model.
+This exhaustive search is computationally infeasible when $G$ is
+large, and when possible it still
+runs the risk of over-fitting unless model complexity is
+appropriately penalized (as in \eqref{eq:subsetregress} below).
+Forward stepwise produces a much
+smaller set of potential models, with cardinality at most
+\textit{steps}. However
+it is a greedy algorithm, so the set of models it produces may not
+contain the best possible model. This is an inherent shortcoming of
+forward stepwise procedures and should be kept in mind when choosing
+between model selection methods.
 
@@
 the forward stepwise sequence, i.e. when to stop stepping
-forward. Some classical approaches can be posed as optimization
+forward. Some approaches for this problem can be posed as optimization
 criteria which stop at the step minimizing
@@
 \end{equation}
-where we have written $\{ \beta_s : s = 1, \ldots, max.steps \}$ as
+(\todo describe construction of $\beta$ after finding $A$)
+where we have written $\{ \beta_s : s = 1, \ldots, steps \}$ as
 the sequence of models outputted by forward stepwise. The function
@@
 2 ($C_p$ of \cite{CP}, AIC of \cite{AIC}), $\log(n)$ (BIC of \cite{BIC}), and
-$2\log(p)$ (RIC of \cite{RIC}). In the current paper we provide
-accurate p-values when adding the first noise variable group in the forward
-stepwise sequence, and it is natural to consider using these p-values
+$2\log(p)$ (RIC of \cite{RIC}). Stopping rules based on classical test
+statistics have also been used, so it is natural to consider using
+the new test statistics of \cite{significance:lasso} or
+\cite{tests:adaptive}
 to choose a model. \cite{sequential:fdr} examined some stopping rules
@@
 proportion of noise variables among variables declared significant
-\citep{fdr}. We explore this further in Section \ref{sec:simulations}.
+\citep{fdr}. We explore this further in Section~\ref{sec:simulations}.
 
@@
 potentially sub-optimal sequence of models, under favorable conditions
-it can still perform well. There is a parallel literature to the
-compressed sensing literature \textit{(cite some here)} with theorems
-stating that forward stepwise can exactly select the true model under
-some stringent conditions involving quantities like the sparsity of
-the true model and the coherence of the design matrix. We do not
-pursue this any further because we find empirically that forward
-stepwise can work well in much less restrictive situations than those
-guaranteed by theorems.
-
-To explore the performance of forward stepwise empirically we
-conducted simulations with results shown in ~\ref{fig:fwdstepsims}
+it can still perform well. There is a small segment of the
+compressed sensing literature \citep{donoho:pursuit, cai:wang:omp}
+with results stating that forward stepwise (often referred to in that
+literature as Orthogonal Matching Pursuit or OMP) can exactly select
+the true model
+under some stringent conditions involving quantities like the sparsity
+of the true model and the coherence of the design matrix. The
+coherence $\mu(X)$ of a matrix $X$ with columns scaled to have unit
+2-norm is defined as
+
+\begin{equation}
+  \mu := \mu(X) = \max_{i \neq j} \{ | \innerp{ X_i }{ X_j } | \}
+\end{equation}
+
+Recalling that $k$ is the sparsity of $\beta$, typical results in this
+literature say that if $k < (1/\mu + 1)/2$ and the nonzero
+coefficients of $\beta$ are sufficiently large then forward stepwise
+recovers $\beta$ with high probability. The coherence condition is
+necessary to guarantee exact recovery \citep{cai:wang:xu:sharp} in the
+sense that it is possible to construct counterexamples with
+$k = (1/\mu + 1)/2$. We refer the reader to the literature for details.
+For our purposes the
+conditions required to guarantee exact recovery are usually too
+stringent. Simulations show empirically that forward
+stepwise can work well even when it is not working perfectly, and that
+it does so under a wide range of conditions.
 
@@
 
+For various sparsity levels $k$, Algorithm~\ref{algo:fs} was run on
+various data sets with \textit{steps} equal to $k$ and the ``average
+power'' was calculated as the number of true variables in the active
+set after $k$ steps divided by $k$. Nonzero coefficients had magnitude
+in a range of multiples of $\gamma := \sqrt{2\log(G)}$, e.g. in $[1.4
+\gamma, 1.6 \gamma]$. Results are shown in
+Figure~\ref{fig:fwdstepsim}. After computing the
+coherence of these design matrices, some calculations show the required
+sparsity level to guarantee exact recovery in these situations is
+about 2 or smaller, and the required nonzero coefficient magnitude is
+likely in the range of 10 to 100 times $\gamma$. These simulations are far
+from the stringent conditions required by theory to guarantee perfect
+recovery, but the performance, while not perfect, may still be good
+enough for some purposes.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\todo Update this section to match/complement \citep{tests:adaptive}
+\todo Reorganize and update this section \\
 
@@
 $\chi^2_k$ random variable. This generally does not work when the
-variable to be added has been chosen by a method that uses the data,
+group to be added has been chosen by a method that uses the data,
 and in particular it fails for forward stepwise procedures adding
@@
 finite samples.  That work marked an important advance in the problem
-of combining inference with model selection.  The current paper
-extends some of their results to the group lasso \citep{grouplasso}.
-In the process we had to derive an exact finite sample null
-distribution for the test statistic, this appears in
-\cite{tests:adaptive}.
+of combining inference with model selection.  \cite{tests:adaptive}
+extended that work to the group lasso \citep{grouplasso} and other
+problems, and modified the test statistic to one with an exact finite
+sample distribution under the global null hypothesis.
 
@@
 $\lambda$ most of the $\beta_g$ will be zero vectors. The weights
-$w_g$ are usually taken to be $\sqrt p_g$ to normalize the penalty
+$w_g$ are usually taken to be $\sqrt {p_g}$ to normalize the penalty
 across groups.  Note that this includes the usual lasso estimator as a
@@
 
-\todo Expand, include simulation results (copy latex for
-includegraphics etc), do the simulation comparing stopping rules
-
-\begin{itemize}
-  \item Show null and non-null p-value-by-step plots for several examples
-  \item Discuss one of these carefully (perhaps with a noise variable
-    entering before a signal variable), for several fixed stopping points
-  \item Compare a few stopping rules, e.g. naive one(s),
-    TailStop/HybridStop, AIC/BIC/Cp
-  \item (the stopping comparisons are not done yet, but I am guessing
-    they will be comparable, perhaps AIC/BIC/Cp will be better, and in
-    that case look for other advantages of ours to mention,
-    e.g. interpretability for being based on p-values rather than a
-    complexity penalty)
-
-\end{itemize}
-
-\begin{figure}
+\todo Clean up and reorganize \\
+\todo Write a little more exposition \\
+\todo Compare to AIC and other classical stopping rules? \\
+
+\begin{figure}[h]
 \begin{center}
@@
 
-\begin{figure}
+\begin{figure}[h]
 \begin{center}
@@
 
-\section{Real data example}
-\todo Simulation with non-Gaussian design matrix (from AIDS data probably)
+\section{Applications}
+\label{sec:applications}
+We now turn to applying forward stepwise and examining the behavior of
+our test statistic in several unique settings, including a real data
+example.
+
+\subsection{Glinternet for hierarchical interactions}
+\label{sec:glint}
+With many variables, fitting models with pairwise interactions can drastically increase model complexity.
+\cite{glint} simplify this problem by imposing a hierarchical constraint on interactions (similar to that in \cite{bien:hierarchical}) where an interaction term can only be included if its main effects are also included.
+They accomplish this by creating a design matrix with main effects alone and also with groups including main effects and their first order interactions.
+Concretely, the expanded design matrix is
+\begin{equation}
+\label{eq:glintmat}
+\tilde X = \begin{pmatrix} X_1 & \cdots & X_G & X_{1:2} & \cdots & X_{1:G} & X_{2:3} & \cdots & X_{(G-1):G}  \end{pmatrix}
+\end{equation}
+where $X_{g:h}$ is the submatrix including $X_g$, $X_h$, and all
+$p_gp_h$  column multiples between columns in group $g$ and columns
+in group $h$. For example, if
+\[
+X_g = \begin{pmatrix} X_{g1} & X_{g2} \end{pmatrix}, \quad
+X_h = \begin{pmatrix} X_{h1} & X_{h2} \end{pmatrix}
+\]
+are two groups each containing two columns, then
+\[
+X_{g:h} = \begin{pmatrix} X_g & X_h & X_{g1} * X_{h1} & X_{g1} * X_{h2} & X_{g1} * X_{h1} & X_{g2} * X_{h2} \end{pmatrix}
+\]
+where * denotes the pointwise product (Hadamard product) of vectors (the $i$th entry of $X_{g1} * X_{h1}$ is the $i$th entry of $X_{g1}$ times the $i$th entry of $X_{h1}$).
+Note that this expanded matrix has $G(G+1)/2$ groups.
+For the rest of this section we drop the tilde
+and denote the expanded design matrix as $X$.
+After constructing the design matrix, they proceed by
+fitting a group lasso model to the data. Because interaction terms only
+appear in groups along with their main effects, the hierarchy condition
+holds for the fitted model.
+
+We can also fit a model by forward stepwise on the same expanded design
+matrix. The grouping still guarantees that our fitted model will
+satisfy the hierarchy condition. However, we must first modify our
+implementation of forward stepwise to prevent the possibility of
+including main effect groups after the main effects are already included
+in the model as part of an interaction group. To do this we compute the
+residual using the whole model in each step, rather than by subtracting
+the effect of the current group being included. For a set of group indices
+$A \subseteq \{ 1, \ldots G \}$ we write $X_A$ for the submatrix of the
+design matrix containing all the groups in $A$.
+
+\begin{algorithm}
+\DontPrintSemicolon
+\caption{\small \it Modification to compute residual from full model}
+\BlankLine
+\CommentSty{\# Everything the same as before, except:}\;
+%  $P_A \gets I - X_AX_A^\dagger$\;
+%  $r_s \gets P_A r_{s-1}$\;
+$r_s \gets y - \text{lsfit}(X_A, y)$\;
+\BlankLine
+\label{algo:fs:full}
+\end{algorithm}
+
+Testing this method by simulation, we constructed signals which have
+$k/3$ main effects nonzero but with no interactions and the remaining
+$2k/3$ nonzero main effects are matched to each other to form interactions.
+This way each nonzero main effect with any nonzero interactions has
+exactly one nonzero interaction.
+The design matrices are independent Gaussian with all groups of size 1.
+Results are shown in Figure~\ref{fig:glint}. The left
+panel shows average power of forward stepwise. Power is calculated
+using the group structure we impose, and not in terms of the original
+main effects and interactions. However, the dotted line shows a more
+forgiving definition of power where we are rewarded for discovering part
+of a nonzero group, i.e. for discovering only one main effect from a
+true interaction group. In the right panel we see that the global null
+result still holds (red vertical lines), and that with a signal the
+p-value is small until the signal has been recovered (blue). The
+interaction power is the proportion of nonzero interaction groups
+that were discovered and is labeled ``Int. Power.''
+
+\begin{figure}
+\begin{center}
+\subfigure[Forward stepwise on glinternet]{
+\label{fig:glint:fwd}
+\includegraphics[width=0.5\textwidth]{../figs/fwdstep/gauss_glint_p20_n200_lower2p9_upper3p6.pdf}}
+\hspace{-15pt}
+\subfigure[P-value with glinternet]{
+\label{fig:glint:pval}
+\includegraphics[width=0.5\textwidth]{../figs/gaussian_size1_n200_p40_g40_k6_lower2p5_upper3p9_glint.pdf}}
+\caption{\small \it The left panel shows the power of forward stepwise
+for various sparsity levels. The right panel shows our p-value marginally
+over each step, with red indiciating the global null case and blue
+indicating a signal with $k=6$ nonzero groups.}
+\label{fig:glint}
+\end{center}
+\end{figure}
+
+\subsection{Generalized additive model selection}
+\todo Finish code \\
+\todo Write section \\
+
+\subsection{Real data example}
+\todo Code simulation with HIVdb data \\
+\todo Write section \\
 
@@
 
+\todo Brief summary \\
 \todo After finising everything else, move some points of discussion
-here, re-read and see if there's anything else interesting to comment
-on here. Mention ongoing work
-
-%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Check below for latex examples if necessary %
-%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%
-%
-%
-%In the classical setting $n > p$ the system eqref{eq:LS:KKT} often has a unique solution, the familiar
-%$$
-%\hat{\beta} = (X^TX)^{-1}(X^Ty).
-%$$
-%In many parametric models, the least squares model is of course too simple.
-In the exponential family setting  \citep{amari,efron}, the normal equations are similar, with  $(X^TX)$ replaced  by the observed Fisher information.
-We have focused on squared error-loss
-%for its simplicity of exposition.
-%
-%In high-dimensional settings, $n$ is often less than $p$ and there is of course no unique solution 
-%to eqref{eq:LS:KKT}. Many modifications are possible, for instance ridge or Tikhonov regularization which adds a strongly convex quadratic term to . The addition of such quadratic terms changes
-%the quadratic part of the loss but does not fundamentally change much else until we begin
-%to make assumptions about whether or not the model is correct, and how much bias might be incurred
-%by such regularization.
-%
-%In modern high-dimensional settings the regularization term, or penalty, of choice is often a norm,
-%with the LASSO \citep{tibshirani:lasso} being the most popular. The lasso problem is
-%\begin{equation}
-%\label{eq:lasso:0}
-%\minimize_{\beta \in \real^p} \frac{1}{2} \|y-X\beta\|^2_2 + \lambda \|\beta\|_1.
-%\end{equation}
-%The duality between norms allows us to write $$
-%\|\beta\|_1 = \sup_{u:\|u\|_{\infty} \leq 1} u^T\beta = h_{B_{\infty}}(\beta)
-%$$
-%with $B_{\infty}$ the $\ell_{\infty}$ ball of radius 1 in $\real^p$ and
-%for any set $K$
-%\begin{equation}
-%\label{eq:support}
-%h_K(\beta) = \sup_{u \in K} u^T\beta
-%\end{equation}
-%is the support function of the set $K$ which we assume to be closed and containing 0. In this notation,
-%the LASSO problem can be expressed as
-%\begin{equation}
-%\label{eq:lasso}
-%\minimize_{\beta \in \real^p} \frac{1}{2} \|y-X\beta\|^2_2 + \lambda h_{B_{\infty}}(\beta).
-%\end{equation}
-%
-%
-%Our canonical problem is therefore
-%\begin{equation}
-%\label{eq:canonical}
-%\begin{aligned}
-%\minimize_{\beta \in \real^p} \frac{1}{2} \|y-X\beta\|^2_2+ \lambda h_{K}(\beta)
-%\end{aligned}
-%\end{equation}
-%with $K$ being a closed, convex set containing 0. For one of many possible infinite
-%dimensional formulations of this canonical problem, see \citep{tsirelson} whose author is also associated
-%to one of the most famous tools in the theory of Gaussian process, the Borell TIS inequality \citep{rfg}.
-%
-%
-%The normal equations of the least squares problem are replaced with the KKT conditions \citep{boyd} for eqref{eq:canonical}. For our canonical problem,
-%the KKT conditions are
-%\begin{equation}
-%\label{eq:KKT}
-%X^T(X\hat{\beta}-y) + \hat{u}= 0, \qquad \hat{u} \in \lambda \cdot \partial h_K(\hat{\beta}),
-%\end{equation}
-%where the $\partial$ denotes the sub differential. In what follows, we denote
-%a solution to this problem as $\hat{\beta}_{\lambda,\XK}$ to denote the dependence on
-%the penalty $\XK$ and the penalty parameter $\lambda$.
-%
-%
-%As we can encode linear or cone constraints in the support function,
-%it is safe to say that a huge number of problems fit into this framework.
-%Some examples include:
-%\begin{itemize}
-%\item LASSO \citep{tibshirani:lasso};
-%\item compressed sensing \citep{candes:cs,donoho:cs};
-%\item group LASSO  \citep{grouplasso,overlap:group:lasso};
-%\item graphical LASSO \citep{graphical:lasso,buhlmann};
-%\item matrix completion \citep{candes:recht:exact,mazumder:hastie:softsvd};
-%\item sign restricted regression \citep{meinshausen:sign};
-%\item hierarchically constrained models \citep{bien:hierarchical,bach:hierarchical};
-%\item generalized LASSO \citep{tibshirani:taylor:path};
-%\item total variation denoising \citep{nesta}.
-%\end{itemize}
-%
-%The relation between eqref{eq:canonical} and the metric projection map is through a particular
-%dual function, also developed by Legendre. The canonical dual problem is 
-%\begin{equation}
-%\label{eq:canonical:dual}
-%\minimize_{u \in  \lambda \cdot \text{row}(X) \cap K} \frac{1}{2} (u-w)^T(X^TX)^{\dagger}(u-w)
-%\end{equation}
-%with $\text{row}(X)$ the row space of the matrix $X$. This dual problem can be derived by minimizing
-%the following Lagrangian with respect to $\beta, \eta$
-%\begin{equation}
-%\label{eq:lagrangian}
-%L(\beta,\eta;u) = \frac{1}{2} \|y-X\beta\|^2_2 + \lambda h_K(\eta) + u^T(\beta-\eta).
-%\end{equation}
-%After a sign change, the problem in eqref{eq:canonical:dual} is fairly easily seen to be equivalent to 
-%\begin{equation}
-%\label{eq:canonical:dual0}
-%\maximize_{u}  \left[\inf_{\beta,\eta} L(\beta,\eta;u) \right]
-%\end{equation}
-%with the constraints in eqref{eq:canonical:dual} encoding the fact that
-%$$
-%\inf_{\beta,\eta} L(\beta,\eta;u) = -\infty
-%$$
-%whenever $u \not \in \lambda \cdot \text{row}(X) \cap K. $
-%Any pair $\hat{\beta}, \hat{u}$ is related through
-%\begin{equation}
-%\label{eq:primal:dual}
-%(X^TX)\hat{\beta} - w + \hat{u} = 0.
-%\end{equation}
-%
-%Choosing an orthonormal basis for $\text{row}(X)$, we see that
-%the dual problem can be phrased as the metric projection problem of $(X^TX)^{-\dagger/2}w$ onto
-%$  \lambda \cdot  \text{row}(X) \cap K$. Alternatively, if we are interested only in 
-%the fitted values $X\hat{\beta}_{\lambda, K}$, the original problem
-%eqref{eq:canonical} can be expressed as the residual $\hat{\mu}=y-\hat{r}$ 
-%from
-%\begin{equation}
-%\label{eq:resid0}
-%\hat{r} = \argmin_{r \in \lambda \cdot \XK} \frac{1}{2} \|y-r\|^2_2,
-%\end{equation}
-%where
-%\begin{equation}
-%\label{eq:XK}
-%\XK = (X^T)^{-1}K.
-%\end{equation}
-%Or, in another form,
-%\begin{equation}
-%\label{eq:resid}
-%\hat{\mu}(y) = y - \pi_{\lambda \cdot \XK}(y).
-%\end{equation}
-%We see that our canonical regression problem is in fact a metric projection problem.
-%
-%Having posed the canonical high-dimensional regression problem as a metric projection problem, 
-%we now try to describe how metric projection is related to some fundamental issues in the understanding
-%of this problem both from an algorithmic view and an inferential view.
-%
-%\subsection{Algorithms to solve the canonical problem}
-%\label{sec:algorithm}
-%
-%The general problem is phrased in terms of an arbitrary $K$. For many high-dimensional
-%problems, this $K$ is chosen to emphasize expected structure in the data. For example,
-%it is well-known that the LASSO yields sparse solutions, the group LASSO yields
-%groups of non-zero coefficients, etc.
-%
-%This special structure is based on a particular structure encoded in $K$. Further, for many canonical
-%choices of $K$ used in high dimensional statistics, such as those cited in \S \ref{sec:problem}, 
-%the following metric projection is simple 
-%$$
-%\nu \mapsto \argmin_{ u \in \lambda \cdot K} \frac{1}{2} \|\nu-u\|^2_2 = \nu - \argmin_{\beta} \left[\frac{1}{2} \|\nu-\beta\|^2_2 + \lambda h_K(\beta) \right].
-%$$
-%
-%Such optimization problems are referred to  as problems in composite form \citep{nesta,boyd,nesterov}. For such problems, 
-%many modern first order solvers use a version of 
-%generalized gradient descent. The steps in generalized gradient descent
-%are essentially iterations of this metric projection map. Specifically, to
-%solve the canonical problem eqref{eq:canonical} for a given step-size $\alpha$ a simple generalized
-%gradient algorithm reads as
-%\begin{equation}
-%\begin{aligned}
-%\hat{\beta}^{(k+1)} &= \argmin_{\beta} \frac{1}{2 \alpha} \|\nu^{(k)}_{\alpha}-\beta\|^2_2 + \lambda h_K(\beta) \\
-%&= \nu^{(k)}_{\alpha} - \argmin_{\eta \in \lambda \alpha \cdot K} \frac{1}{2} \|\nu^{(k)}_{\alpha}-\eta\|^2_2, \\
-%\nu^{(k)}_{\alpha} &= \beta^{(k)} - \alpha \cdot  X^T(X\beta^{(k)}-y).
-%\end{aligned}
-%\end{equation}
-%
-%The first line in the update above is the usual form of updates for generalized gradient descent, while the second line
-%expresses this step as the residual after an application of the metric projection map. 
-%Accelerated schemes can do much better with slightly different updates above, see \citep{nesta,nesterov,tseng}.
-%With modern computing techniques, such simple algorithms can scale to huge
-%problems, see \citep{admm,mazumder:hastie:softsvd}.
-%
-%\section{Inference for the canonical problem}
-%\label{sec:inference}
-%
-%\subsection{Law of large numbers}
-%
-%Having solved the canonical problem, what can we say about its solution?
-%As this special issue is devoted to the appearance of one of the first proofs of the law of large numbers, 
-%we should at least hope to provide such an answer.
-%
-%In the classical setting, assuming independence, and the usual linear regression model
-%\begin{equation}
-%\label{eq:model}
-%y = \mu + \epsilon
-%\end{equation}
-% with noise $\epsilon$ having scale $\sigma$,
-%the central limit theorem
-%can often be applied to yielding the usual result
-%\begin{equation}
-%\|X(\hat{\beta} - \beta_0)\|_2^2 \overset{D}{\approxeq} \sigma^2 \cdot \chi^2_p 
-%\end{equation}
-%under the null $H_0: \mu =X\beta_0 \in \text{col}(X)$. Of course, this
-%forms the basis of much
-%inference in modern (and not so modern) applied statistics in the fixed $p$, $n$ growing regime.
-%In terms of the parameters themselves, this
-%implies the weaker statement
-%\begin{equation}
-%\label{eq:LLN}
-%\|\hat{\beta} - \beta\|_2 \leq n^{-1/2}\sigma \cdot \zeta_n
-%\end{equation}
-%for some random variable $\zeta_n=O_{\Pp}(1)$.
-% 
-%In the classical setting, assuming $X$ is full rank, the  bound eqref{eq:LLN} is a simple two line proof followed by
-%some assertions. If
-%we write ${\cal L}(\beta) = \frac{1}{2} \|y-X\beta\|^2_2$, then
-%$$
-%\begin{aligned}
-%0 &\geq  {\cal L}(\hat{\beta}) - {\cal L}(\beta_0) \\
-%& =   \nabla {\cal L}({\beta}_0)^T(\hat{\beta}-\beta_0) + \frac{1}{2}(\hat{\beta}-\beta_0)^T(X^TX)(\hat{\beta}-\beta_0) \\
-%&= (X^T\epsilon)^T(\hat{\beta}-\beta_0) + \frac{1}{2}(\hat{\beta}-\beta_0)^T(X^TX)(\hat{\beta}-\beta_0) \\
-%& \geq - \|X^T\epsilon\|_2 \|\hat{\beta}-\beta_0\|_2 + \frac{\lambda_{\min}(X^TX)}{2} \|\hat{\beta}-\beta_0\|^2_2
-%\end{aligned}
-%$$
-%with $\lambda_{\min}(X^TX)$ denoting the smallest eigenvalue of $X^TX$. We see that we can take $$
-%\zeta_n = \sigma^{-1/2} n^{1/2} \frac{\|X^T\epsilon\|_2}{\lambda_{\min}(X^TX)}.
-%$$
-%
-%%Before proceeding, note that
-%%$$
-%%\|X^T\epsilon\|_2 = \sup_{u \in B_2(\real^p)} u^TX^T\epsilon
-%%$$
-%%which is the supremum of a (particularly simple) process. If the noise $\epsilon$ is
-%%Gaussian, then, conditional on $X$, this is the supremum of a {\em Gaussian} process, the second
-%%tool in the geometry of least squares toolbox.
-%
-%In the high-dimensional setting, of course this fails as $\lambda_{\min}(X^TX)=0$.
-% What, then, can we say about
-%our canonical estimator
-%$$
-%\hat{\beta}_{\lambda,\XK} = \argmin_{\beta \in \real^p} \frac{1}{2} \|y-X\beta\|^2_2+ \lambda h_K(\beta)?
-%$$
-%Is there even a weak law of large numbers?
-%Without any assumptions on $K$, the answer is clearly no: if $K=\{0\}$, then this is the original ill-posed
-%problem.
-%
-%Under an assumption of decomposability of $K$ 
-%recent progress has been
-%made in providing bounds on the estimation error in eqref{eq:canonical}, see  \citep{decomposable}. 
-%The notion of decomposability in \citep{decomposable} has a precise definition which we will not dwell on here. However,
-%a large set of examples of decomposable penalties are penalties of the form 
-%\begin{equation}
-%\label{eq:product}
-%K = \prod_{i \in {\cal I}}K_i.
-%\end{equation}
-% That is,
-%convex sets that can be expressed as products of convex sets lead to decomposable
-%penalties.
-% Every example of a decomposable norm in \citep{decomposable} has this form except the nuclear norm. For such penalties, the generalized gradient algorithms
-%described in \S \ref{sec:algorithm} decompose into many smaller subproblems. Many efficient coordinate
-%descent algorithms exploit this fact, see \citep{pathwise,glmnet}.
-%
-%A similar notion of decomposability we refer to as {\em additivity}
-%is explored in \citep{tibs:taylor:additive} in which case $K$ can be expressed as $K=A+I$ with the
-%sum being Minkowski addition of sets. In this case, the penalty has the form
-%\begin{equation}
-%\label{eq:additive}
-%h_K(\beta) = h_A(\beta) + h_I(\beta).
-%\end{equation}
-%One concrete example of this is the $\ell_{\infty}$ ball in $\real^p$. For $A, I$ a partition of
-%$\{1, \dots, p\}$ into {\em active} and {\em inactive} variables, we can write
-%$$
-%B_{\infty} = \left\{(u_A,0): \|u_A\|_{\infty} \leq 1 \right\} + \left\{(0,u_I): \|u_I\|_{\infty} \leq 1 \right\}.
-%$$
-%Any penalty of the form eqref{eq:product} can be expressed in the form eqref{eq:additive} in a similar fashion.
-%If we are allowed to introduce a linear constraint to eqref{eq:canonical} then any problem with a  penalty of the form eqref{eq:additive} can be expressed as a problem with a penalty of the form eqref{eq:product} subject to an additional  set of linear constraints.
-%
-%The weak law of large numbers presented in the literature
-%follow a similar path to the argument above. Of course, for precise results, the specific
-%penalty as well as the data generating mechanism must be more precisely specified. 
-%
-%In the interest of space,
-%we do not pursue such precise statements here.
-%Rather, we will just attempt to paraphrase these results, of which there exist many in the literature (c.f.
-\citep{decomposable,bickel:ritov,support:recovery,highdim:stat}) with \citep{decomposable} being a
-%particularly nice place to read in detail. Under various assumptions on the tails of 
-%$\epsilon$, as well as the assumption $h_I(\beta_0)=0$ and $K^{\circ}$ is bounded\footnote{Recall the definition
-%of the polar body of $K$
-%\begin{equation}
-%\label{eq:polar}
-%K^{\circ} = \left\{\nu \in \real^p: u^T\nu \leq 1, \ \forall u \in K \right\}.
-%\end{equation} For any $K$, the seminorm $h_{K^{\circ}}$ is the dual seminorm of $h_K$.},
-% the canonical result has the form for $\lambda \geq C_1 \cdot \Ee\left(h_{K^{\circ}}(-X^T\epsilon) \right)$
-%\begin{equation}
-%\label{eq:error}
-%\|\hat{\beta}_{\lambda, \XK} - \beta_0\|_2 \leq C_2 \cdot \sigma \frac{ \Ee\left(h_{K^{\circ}}(-X^T\epsilon) \right) \psi(A)}{\kappa(A, I, X)}
-%\end{equation}
-%with high probability
-%for some universal $C_1, C_2$ 
-%where $\psi(A)$ is referred to as a {\em compatibility} constant relating the $\ell_2$ norm and the 
-%$h_A$ seminorm; the quantity 
-%$\kappa(A,X)$ replaces $\lambda_{\min}(X^TX)$ and is referred to as a {\em restricted strong
-%convexity} (RSC) constant. 
-%
-%The literature varies in their assumptions
-%on the noise and the design matrix $X$. For instance, in the fixed design case one might
-%consider the above probability only with respect to noise, while in a random design setting the dependence
-%of the constants on $X$ are typically expressed with respect to the law that generates the design matrix $X$.
-%
-%% Above, we introduced
-%%$K^{\circ}$, the polar body of $K$ defined as
-%%\begin{equation}
-%%\label{eq:polar}
-%%K^{\circ} = \left\{\nu \in \real^p: u^T\nu \leq 1, \ \forall u \in K \right\}.
-%%\end{equation}
-%%
-%
-%Having established a bound such as eqref{eq:error}, 
-%if one considers problems indexed by $n$, then one can obtain
-%a law of large numbers for the problem eqref{eq:canonical} so long as the parameters
-%are chosen so the right hand side decays to 0 and the bound holds with sufficiently high probability. 
-%Again in the interest of space, we refer
-%readers to the literature for more precise statements for specific versions of the problem such as the LASSO.
-%We have tried to give a summary of some of the results related to the canonical problem, though we have
-%barely exposed the tip of the iceberg. Under more specific assumptions much more can be said. For example, see \cite{donoho:montanari}.
-%%
-%%
-%%If $\epsilon$ is Gaussian and $X$ is considered fixed, then the choice of $\lambda$ and 
-%%the final constant in the bound above is
-%%the canonical Gaussian
-%%width of $\XK^{\circ}$, see \citep{tsirelson}. For other noise models,
-%%this measures the noise under some other law. 
-%%
-%%The Gaussian width and analogous quantities arise in many contexts in machine learning. In some sense, made
-%%proper in the machine learning and empirical processes literature,
-%%this width, or its empirical process counterpart, is the fundamental complexity of the estimation problem eqref{eq:canonical}.
-%%If we were to be so ambitious as to extend our problem to an infinite dimensional setting where
-%%$\real^p$ is reproduced by some reproducing kernel space $H$, then the concept of 
-%%Gaussian width is known to be related to the existence of maximum likelihood
-%%estimators in such settings \citep{tsirelson}.
-%
-%\subsection{Gaussian width and metric projection: intrinsic volumes}
-%
-%For $\XK^{\circ} = X^TK^{\circ} \subset \real^n$, the error eqref{eq:error}
-% depends on the
-%quantity
-%\begin{equation}
-%\label{eq:width}
-%\Ee(h_{K^{\circ}}(-X^T\epsilon)) = \Ee(h_{\XK}(-\epsilon)).
-%\end{equation}
-% For fixed $X$, this quantity is referred to as
-% the Gaussian width of $\XK$ and it also intimately
-%related to our first tool in the toolbox,  the metric projection. Specifically, 
-%consider the tube of radius $r$ around $\XK^{\circ}$. That is,
-%\begin{equation}
-%\label{eq:tube}
-%\tube(\XK^{\circ},r) = \left\{z \in \real^n: \|z - \pi_{\XK^{\circ}}(z)\|_2 \leq r \right\}.
-%\end{equation}
-%Then, a classical result of  Steiner in the case of convex bodies and Weyl in the case of manifolds  says that
-%the Lebesgue measure of the tube, assuming $\XK^{\circ}$ is bounded, can be expressed as
-%\begin{equation}
-%\left|\tube(\XK^{\circ},r)\right|_{\real^n} = \sum_{j=0}^n r^j \omega_j \lips_{n-j}(\XK^{\circ}), \qquad r \leq r_c(\XK^{\circ})=\infty,
-%\end{equation}
-%where the $\lips_l(\XK^{\circ})$ are referred to as the intrinsic volumes of $\XK^{\circ}$
-%and $\omega_l = |B_2(1)|_{\real^l}$ is the Lebesgue measure of the 
-%unit $\ell_2$ ball in $\real^l$. See \citep{rfg,federer,schneider,weyl} for more details on such volume of tubes formulae.
-%When $\XK^{\circ}$ is unbounded, Federer's curvature measures \citep{federer} can be used to define the volume of local tubular neighborhoods.
-%Using  Gaussian process techniques \citep{vitale}, it can be shown that
-%\begin{equation}
-%\label{eq:gauss:width}
-%\lips_1(\XK^{\circ}) =  (2\pi)^{-1/2}\Ee(h_{\XK^{\circ}}(\epsilon)|X) %= (2\pi)^{-1/2}\Ee(h_{K^{\circ}}(X^T\zeta)).
-%\end{equation} 
-%where $\epsilon |X \sim N(0, I_{n \times n})$. 
-%
-%For $D=\XK$ a smooth domain, that is convex set with non-empty interior bounded by a smooth hypersurface
-%and for $j \leq n-1$
-%\begin{equation}
-%\label{eq:intrinsic:volumes}
-%\lips_j(D)  \propto \int_{\partial D}P_{p-j-1}(\lambda_{1,x}, \dots, \lambda_{p-1,x}) \; \text{Vol}_{\partial D}(dx)
-%\end{equation} 
-%with $P_j$ the $j$-th elementary symmetric polynomial of the so-called
-%principal curvatures of $\partial D$ at $x$ \citep{rfg,noniso}. These are just the eigenvalues of the second fundamental form in the unit inward normal direction. This formula can be derived
-%by considering the inverse of the metric projection map eqref{eq:metric:proj}. The inverse takes $(x, \eta_x)$ defined on the extended outward normal bundle of
-%$K^{\circ}$. The inverse of the map is simply the exponential map restricted to the outward normal bundle, or, more simply
-%\begin{equation}
-%\label{eq:exponential}
-%x \mapsto x + \eta_x.
-%\end{equation}
-%A fairly straightforward calculation
-%%\footnote{Weyl instructs in \citep{weyl} that this
-%%is something any student of calculus should be able to do.} 
-% yields the relation eqref{eq:intrinsic:volumes}.
-%
-%The main take away message above is that the functionals in the tube formula, such as the Gaussian width
-%eqref{eq:width}, are related to properties of the metric projection map onto $\XK$.
-%
-%
-%As written above, the intrinsic volumes are defined implicitly through a volume calculation, and it is not clear that they
-%extend to the infinite dimensional setting.
-Under the right conditions of course, such extensions are indeed possible.
-See \citep{vitale} for a nice discussion of this problem.
-An alternative definition of intrinsic volumes specific to the Gaussian case
-%was considered in \citep{taylor:vadlamani}, which were defined as coefficients in an expansion of the {\em Gaussian} measure of 
-%$\tube(\XK^{\circ},r)$ as described in the Gaussian Kinematic Formula \citep{rfg,taylor:kinematic}.
-%
-%\subsection{Risk estimation}
-%
-%Another quantity of interest for our problem eqref{eq:canonical} is an estimate of
-%how much ``fitting'' we are performing, as a function of $\lambda$. One quantitative
-%measure of this is captured by Stein's estimate of risk  \citep{stein}, also known as
-%SURE. Suppose now that $y \sim N(\mu, I)$ and we estimate $\mu$ by the estimator
-%$\hat{\mu}$. The SURE estimate is an unbiased estimate of 
-%$$\text{Risk}(\hat{\mu}) = \Ee(\|\mu - \hat{\mu}\|^2_2).$$ 
-%The estimated degrees of freedom of this estimator is one part of the SURE estimate and is defined as
-%\begin{equation}
-%\widehat{\text{Cov}(y, \hat{\mu}}) = \text{div}(\nabla \hat{\mu}(y)).
-%\end{equation}
-%%As an aside, the proof that this is an unbiased estimate of risk is an application of integration by parts, something
-%%that again, any student of calculus should be able to follow. This integration by parts formula is key to application
-%%of Stein's method to prove central limit theorems. It is also at the heart of 
-%%Malliavin calculus \citep{malliavin} via the representation of \^Ito integrals as divergence of vector fields. 
-%%It seems, as in Weyl's proof
-%%of the intrinsic nature of the intrinsic volumes, that it may have been asking too much to expect any student of calculus to see these deep connections.
-%
-%Suppose $y \sim N(\mu, I)$ and consider our residual form of the original estimation problem for $\mu=\Ee(y)$
-%\begin{equation}
-%\hat{\mu}(y) = y - \pi_{\lambda \XK}(y).
-%\end{equation}
-%If $K$ possesses a nice stratification, as do all the examples mentioned above, then,  for almost every $y$, $\pi_{\lambda \XK}(y)$ is in the relative interior
-%of  some fixed stratum ${\cal S}$ of the normal bundle of $\XK$ over which the dimension
-%of the tangent space is constant, and the normal bundle has a locally conic structure
-%${\cal S} = {\cal T} \times {\cal N}$ of tangent and normal directions  \citep{schneider,rfg}. 
-%Having fixed this stratum, we can write
-%$y=x+\eta_x$ in  (ortho)normal coordinates centered at $(y-\hat{\mu}(y), \hat{\mu}(y))$. In these
-%coordinates
-%\begin{equation}
-%\hat{\mu}(y(x,\eta_x)) = \eta_x.
-%\end{equation}
-%In order to relate the above to the problem eqref{eq:canonical}, one should
-%invert the above chart to find $(x,\eta_x)$ in terms of $y$. 
-%In the residual form  eqref{eq:resid}  it is easy to show that
-%$$
-%\begin{aligned}
-%\left(x(y), \eta_x(y) \right) &= \left( \argmin_{r \in \lambda \cdot \XK} \frac{1}{2} \|y-r\|^2_2, y -  \argmin_{r \in \lambda \cdot \XK} \frac{1}{2} \|y-r\|^2_2 \right) \\
-%&= (y - \hat{\mu}(y), \hat{\mu}(y)).
-%\end{aligned}
-%$$
-%The derivatives along the normal directions yield 
-%a purely dimensional term, while directions in the tangent directions yield
-%curvature terms. This observation is enough to derive the following form of the degrees of freedom
-%\begin{equation}
-%\label{eq:SURE}
-%\begin{aligned}
-%\text{div}(\nabla \hat{\mu}(y))% &= n - \text{dim}({\cal T}) + \text{Tr}(S_{(x,\eta_x)})  \\
-%&= 
-% n - \text{dim}({\cal T}_y) + \text{Tr}(S_{(y - \hat{\mu}(y),\hat{\mu}(y))}).
-% \end{aligned}
-%\end{equation}
-%Above, ${\cal T}_y$ is the tangential part of the stratum containing $x(y)$ and $S_{(x,\mu_x)}$ is the second fundamental form of 
-%${\cal T}_y$ in $\lambda \cdot \XK$ as described \citep{rfg}. When $K$ is a polyhedral set, the second term disappears
-%and the degrees of freedom can be computed by computing the rank of a certain matrix \citep{tibshirani_degrees_2012,bien:hierarchical}
-%%
-%%\begin{lemma}[SURE for the canonical problem]
-%%Consider the
-%%canonical problem in residual form eqref{eq:resid}. The 
-%%SURE estimate of the degrees of freedom of this estimator is 
-%%\begin{equation}
-%%\label{eq:SURE}
-%%\begin{aligned}
-%%\text{div}(\nabla \hat{\mu}(y))% &= n - \text{dim}({\cal T}) + \text{Tr}(S_{(x,\eta_x)})  \\
-%%&= 
-%% n - \text{dim}({\cal T}_y) + \text{Tr}(S_{(y - \hat{\mu}(y),\hat{\mu}(y))}).
-%% \end{aligned}
-%%\end{equation}
-%%where ${\cal T}_y$ is the tangent part of the stratum and $S_{(x,\mu_x)}$ is the second fundamental form of 
-%%${\cal T}_y$ in $\lambda \cdot \XK$ as described \citep{rfg}.
-%%\end{lemma}
-%%{\bf Proof:}
-%%Locally, as the estimate is in the fixed stratum with tangent component ${\cal T} = {\cal T}_y$ and
-%%normal component ${\cal N} = {\cal N}_y$ the estimation procedure can be written as
-%%$$
-%%\begin{aligned}
-%%\hat{\mu}(x+\eta_x) &= \eta_x \\
-%%&= \sum_{j=1}^{n - \text{dim}({\cal T}_y)} a_j \eta_{j,x} \\
-%%\end{aligned}
-%%$$
-%%where $a \in \real^{n - \text{dim}({\cal T}_x)}$ and $\eta_{j,x}$ form an orthonormal
-%%frame for the space normal to ${\cal T}_x$. In fact, the $a_j$'s must be such that
-%%$\eta_x \in N_x( \lambda \cdot \XK)$ but we can assume they are generic points in the relative interior.
-%%Differentiating with respect to $a$ and summing yields $n - \text{dim}({\cal T}_x)$ while differentiating
-%%with respect to an orthonormal basis for ${\cal T}_x$ yields $\text{Tr}(S_{x,\mu_x})$. Substituting
-%%the parametrization $(y-\hat{\mu}(y), \hat{\mu}(y))=(x,\eta_x)$ concludes the proof. \qed
-%
-%\subsection{Hypothesis testing: weak convergence for high-dimensional inference?}
-%
-%Another fundamental tool in inference for least squares models is the ability to form hypothesis
-%tests, as well as confidence intervals for the ``true'' mean. Such concepts
-%clearly need a model, which we might take to be the usual model
-%$$
-%y \sim N(\mu,I).
-%$$ Under the assumption that $\mu=X\beta_0$, classical inference in 
-%linear models (assuming $X^TX$ is full rank) yields confidence intervals and hypothesis tests for any linear functional $\nu^T\mu$ based on the
-%coefficients $\beta_0$. 
-%
-%What can we say about our canonical problem eqref{eq:canonical}?
-This is an area in which we still don't know all the answers.
-In some sense, we are in the situation analogous to Bernoulli having proved a weak
-%law of large numbers without the central limit theorem.
-%Some progress  has been made for specific models of the design matrix $X$  for the  LASSO as well as group LASSO models, see \citep{meinshausen:buhlmann,buhlmann,minnier:cai,murphy:laber,wasserman:roeder,donoho:montanari}.
-%
-%Other recent work \citep{significance:lasso} gives some hints at what inferential tools may prove useful in this weak convergence theory.
-As described in the LARS algorithm \citep{lars,tibshirani:unique}, an entire path of solutions $\hat{\beta}_{\lambda,\XK}$ can be formed for the LASSO, i.e.
-when $K=B_{\infty}$.
-These paths are piecewise linear, with knots at points where the {\em active set} changes.
-The covariance statistic
-%measures some change in the correlation of the fitted values between two knots $\lambda_k$ and $\lambda_{k+1}$ in the LASSO path. It has the form
-%\begin{equation}
-%\label{eq:covstat}
-%T_k = C_k \lambda_k (\lambda_k - \lambda_{k+1})
-%\end{equation}
-%for some random scaling $C_k$ related to the active set and the variable added to the active set at $\lambda_{k+1}$. 
-%The form for $k=1$ is particularly simple: suppose that $\hat{j}_1$ and $X$ is such the first variable in the LARS path, i.e.
-%$$
-%\hat{j}_1  \in \argmax_j |X_j^Ty|.
-%$$
-%Then,
-%\begin{equation}
-%\label{eq:covstat1}
-%T_1 = \frac{1}{\sigma^2} y^TX\hat{\beta}_{\lambda_2} = y^T X_{\hat{j}_1} \hat{\beta}_{\lambda_2,\hat{j}_1}.
-%\end{equation}
-%For $k \geq 1$, the form of the test statistic is slightly more complicated, though it can be expressed in terms of 
-%$A=A_k$, the active set at step $k$ as well as $s_{A_k}$, the signs of the active variables at step $k$ as well as 
-%the active set $A_{k+1}$ and $s_{A_{k+1}}$, see \S 2.3 of \cite{significance:lasso} for the full expression.
-%For a wide range of (sequences) covariance matrices, if the active set at $\lambda_k$ already contains all the strong active variables, then
-%it is shown in \citep{significance:lasso} that
-%\begin{equation}
-%\label{eq:exp1}
-%T_k \overset{D}{\rightarrow} \text{Exp}(1).
-%\end{equation}
-%In particular, under the global null $y \sim N(0,\sigma^2 I)$ as long as the design matrix satisfies some minimum growth condition, 
-%$T_1 \overset{D}{\rightarrow} \text{Exp}(1)$.
-% The main tools used in the above proof relate to the maxima of (discrete) Gaussian processes and the generalization of an
-% argument previously applied to smooth Gaussian processes in \citet{taylor:validity} and \citet{rfg}.
-% Ongoing work suggests that such a limiting distribution will hold for many (sequences) of $K$ and design matrices $X$.
-%
-%As for confidence intervals for the parameters related to $k$ strong variables, the relation to extreme values of Gaussian processes suggest that the bias-corrected or relaxed LASSO estimate of the active coefficients will have accurate coverage.
-This is ongoing work.
-% 
-%%However, there are some special cases for which some results are known. Specifically, 
-%%suppose $K$ is a cone. Then eqref{eq:canonical} is a cone-restricted regression problem with the restriction being
-%%$$
-%%h_K(\beta) < \infty \iff \beta \in K^{\circ}
-%%$$
-%%where $K^{\circ}$ is the polar cone of $K$.
-%%
-%%Consider the test of $H_0:\beta=0$ versus the unrestricted alternative $H_a:\beta \in K^{\circ}$. 
-%%The likelihood ratio test statistic \citep{takemura,kuriki,self,chernoff,perlman} for these hypotheses
-%%is, perhaps not surprising to the readers at this point, related to the metric projection problem. Specifically, an inspection of the canonical problem in residual form eqref{eq:residual}
-%%shows that the log-likelihood ratio test statistic  is given by
-%%$$
-%%\Lambda(H_0, H_a) = \|\pi_{\XK}(y)\|^2_2
-%%$$
-%%and we see the first tool in our toolbox  appearing again. 
-%%
-%%If $S(K,X) = \XK \cap S(\real^n)$ are the set of unit vectors in $\XK$, it is not
-%%difficult to show that
-%%\begin{equation}
-%%\|\pi_{\XK}(y)\|_2 = \sup_{\nu \in S(K,X)} (\nu^Ty)^+
-%%\end{equation}
-%%and the log-likelihood ratio test statistic can be expressed as the supremum of a 
-%% process indexed by $S(K,X)$. We see, then, that under $H_0$, it is the supremum of a unit variance
-%% Gaussian process. Recalling our earlier discussion on Gaussian width, we know that there
-%% is a metric projection hiding somewhere in the distribution of the supremum of this process.
-%% We will see that this is indeed the case in the following section.
-%% 
-%
-% 
-%% Before continuing along the above path, we discuss the possibility of a central limit theorem
-%% for functionals of $\mu$ in eqref{eq:canonical}. One of the fundamental questions that must be 
-%% answered is what functionals of $\mu$ is it natural to consider? It is clear, and well known, that even if 
-%% $y\sim N(X\beta_0,I)$ the estimator $\hat{\beta}_{\lambda,\XK}$ is providing
-%% biased estimates of $\beta_0$. 
-%% 
-%% The natural question to ask is ``what functional of $\mu$ is $\hat{\beta}_{\lambda, \XK}$ estimating''? 
-%% Writing $\hat{\beta}_{\lambda, \XK}(y,X)$ explicitly as a function of $y$ and $X$ we might suppose that
-%% $\hat{\beta}_{\lambda, \XK}(y,X)$ is estimating 
-%% \begin{equation}
-%% \label{eq:estimating}
-%%\hat{\beta}_{\lambda,\XK}(\mu,X) =\argmin_{\beta \in \real^p} \frac{1}{2} \|\mu-X\beta\|^2_2+ \lambda h_{K}(\beta).
-%% \end{equation}
-%% It remains to be seen whether this notion be used to establish a general weak convergence result for eqref{eq:canonical}.
-%% 
-% 
-%\section{Smooth Gaussian processes: relaxing convexity}
-%\label{sec:gaussian}
-%
-%In the special case that $K$ is a cone, we saw that the distribution of a particular 
-%likelihood ratio test could be expressed in terms of the supremum of a Gaussian process indexed by a subset of the sphere. 
-%Equivalently, this supremum could be expressed in terms of the metric projection onto the cone
-%$\XK$. It is well-known that the distribution of this likelihood ratio test statistic is a mixture
-%of $\chi^2$'s of varying degrees of freedom. This distribution is sometimes referred to as a  $\bar{\chi}_2$ distribution.
-%
-%The mixture weights can be expressed in terms of the geometry of the set
-%$M = S(X,K) = \XK \cap S(\real^n)$. In particular, it is known \citep{rfg,takemura:kuriki1,takemura:kuriki2} that if 
-%$\epsilon \sim N(0, I)$ for $u > 0$
-%\begin{equation}
-%\label{eq:sup}
-%\Pp\left(\sup_{\nu \in M}( \nu^T \epsilon)^+ > u \right) =
-%\sum_{j=0}^{\infty} \lips_j(M) \rho_j(u),
-%\end{equation}
-%where
-%\begin{equation}
-%\label{eq:rho}
-%\rho_j(u) = \begin{cases}
-%\int_u^{\infty} \frac{1}{\sqrt{2\pi}} e^{-t^2/2} \; dt  & j = 0 \\
-%(2\pi)^{-(j+1)/2} H_{j-1}(u) e^{-u^2/2}& j  \geq 1,
-%\end{cases}
-%\end{equation}
-%and$$
-%H_j(u) =(-1)^j \frac{\partial ^j}{\partial u^j} e^{-u^2/2}
-%$$
-%are the standard Hermite polynomials. The functions eqref{eq:rho}
-%are known as the EC or Euler characteristic densities for a
-%Gaussian field, see \citep{rfg,takemura:kuriki2,worsley:boundary,worsley:unified1,worsley:unified2}.
-%While the sum above is written as an infinite sum it terminates at $\text{dim}(M)$.
-%
-%Note that this implies
-%$$
-%\Ee \left(\sup_{\nu \in M} (\nu^T \epsilon)^+  \right) = \frac{1}{\sqrt{8\pi}} \lips_1(M)
-%$$
-%which is an analogous way to derive Gaussian width eqref{eq:gauss:width}.
-%
-%One of the derivations of the above formula, the so-called {\em volume of tubes} approach \citep{siegmund,sun,takemura:kuriki1} involves studying the Jacobian of the inverse of the
-%spherical metric projection map eqref{eq:metric:proj:sphere}, i.e. the exponential
-%map on $S(\real^n)$ which sends a pair $(x,\eta_x)$ to
-%$$
-%\cos(\|\eta_x\|) \cdot x + \sin (\|\eta_x\|) \cdot \eta_x.
-%$$
-%Another approach, the expected Euler characteristic approach
-%\citep{rfg,worsley:boundary} involves counting critical points of 
-%the Gaussian process above the level $u$ according to saddle type and applying
-%Morse theory and the Rice-Kac formula \citep{azais,rfg} to count the expected number of such points.
-%
-%Neither of these approaches strictly require convexity of the cone generated by the parameter set $M$. Rather,
-%they depend on a notion of local or infinitesimal convexity referred to as positive 
-%reach \citep{federer}. Hence, the parameter sets may have finite critical radius. 
-%They are both approaches used to form an approximation
-%\begin{equation}
-%\label{eq:sup:generic}
-%\Pp\left(\sup_{\nu \in M} f(\nu) > u\right) \approxeq
-%\sum_{j=0}^{\infty} \lips_j(M) \rho_j(u)
-%\end{equation}
-%for some centered, smooth, Gaussian process $f$ having constant variance 1 on a
-%(possibly stratified) manifold $M$.
-%In the volume of tubes approach, the critical radius appears in a natural way and enters into an estimate of the error of the volume of tubes approach.
-In both approaches, though it is clearer in  the expected Euler characteristic approach, the spherical critical radius is in fact the spherical critical radius of
-%$\Psi(M)$ where $\Psi:M \rightarrow S(H)$ where $S(H)$ is the unit sphere in $H$, the reproducing kernel Hilbert space of 
-%$f$. Either approach yields roughly the same estimate of error: for $u$ large enough
-%\begin{equation}
-%\label{eq:sup:generic:tail}
-%\liminf_{u \rightarrow \infty} -\frac{2}{u^2} \left|\Pp\left(\sup_{\nu \in M} f(\nu) > u\right) -
-%\sum_{j=0}^{\infty} \lips_j(M) \rho_j(u) \right|= 1 + \tan^2 (r_c(M)).
-%\end{equation}
-%The above says that the {\em relative error} in the approximation is exponentially small whenever $r_c(M) > 0$.
-%
-%The critical radius of $M$ in the expected Euler characteristic approach arises in terms of 
-%a functional of the original process $f$. Specifically, if we assume $M$ is a manifold without boundary, then define for each $x\neq y$ the process introduced in \citep{taylor:validity}
-%\begin{equation}
-%\label{eq:fxy}
-%f^x(y)= \frac{f(y) - \Ee(f(y)|f(x), \nabla f(x))}{1 - \Ee(f(x) \cdot f(y))}.
-%\end{equation}
-%Then,
-%\begin{equation}
-%\label{eq:fxy2}
-%\cot^2(r_c(M)) = \sup_{x \neq y} \Ee(f^x(y)^2).
-%\end{equation}
-%Hence, the critical radius depends in an explicit way on the covariance function of the 
-%process $f$. 
-%
-%As mentioned previously, the argument related to derivation of eqref{eq:fxy2} in the smooth case led directly to the
-%exponential limit in eqref{eq:exp1}. Such a connection suggests a relation between the
-%distribution of the maxima of smooth Gaussian processes, specifically the spacings of the extreme values, can be 
-%used to derive weak convergence results for high-dimensional inference. This is ongoing work.
-%%
-%%We might ask the question: does this critical radius manifest itself in any
-%%sampling sense? That is, can we see this critical radius from observing
-%%independent samples of $f$? The answer to this is affirmative.
-%%In \citep{taylor:weinberger:adler} we consider the random embedding of $M$ given by
-%%$\tilde{M}_k=f^k(M)$ where
-%%$$
-%%f^k(\nu) = (f_1(\nu), \dots, f_k(\nu)), \qquad f_j \overset{IID}{\sim} f
-%%$$
-%%and $k  \geq 2 \text{dim} + 1$ to ensure the image is generically an embedding. The set
-%%$y$ itself has a metric projection map
-%%$$
-%%y \mapsto \pi_{\tilde{M}_k}(y)
-%%$$
-%%which is indeed a function on $\tube(\tilde{M}_k,r)$ for $r < r_c(\tilde{M}_k)$. 
-%%
-%%The size of $r_c(\tilde{M}_k)$ is related to the number of samples needed to recover the 
-%%homology of $M$ from uniformly sampling points on $\tilde{M}_k$ \citep{NSW}. This is one of the 
-%%more quantitative results in the nascent field of {\em topological data analysis}. But what are the
-%%sampling properties of the random variable $\tilde{M}_k$? Not surprisingly, a
-%%law of large numbers holds \citep{taylor:weinbergerg:adler} and $\tilde{M}_k \overset{\Pp}{\rightarrow} r_c(M)$ where $r_c(M)$ is the previously mentioned spherical critical radius in $H$.
-%%Once again, we are at the point Bernoulli reached in {\em Ars  Conjectandi} and look to see if there
-%%is a Gaussian, or CLT, analogue to this weak law.
-%%
-%
-%\section{Conclusion}
-%
-%We have described what we call the two basic tools of the geometry of least squares
-%that are just as relevant as when Gauss and Legendre disputed their original discovery
-%over 200 years ago. While these tools are not the most technically sophisticated tools, 
-%ceding that ground to exponential families for the canonical model eqref{eq:canonical} and 
-%empirical processes for the fluctuation theory in \S \ref{sec:inference}, they nevertheless
-%provide guiding principles for these more precise tools. We would argue that the Gaussian
-% picture provided by the geometry of least squares, gets much of the picture correct under sufficient moment conditions. For heavier tailed results, of course much of \S \ref{sec:inference}
-% would have to be reframed and \S \ref{sec:gaussian} paints quite a different picture
-% \citep{samorodnitsky1,samorodnitsky2}.
-% 
-% As Bernoulli found himself with just a law of large  numbers, the field of statistics is roughly at this same stage in high-dimensional inference.
-We are hopeful that the geometry of least squares will eventually guide the field to weak convergence results in high-dimensional inference
-% for the canonical problem eqref{eq:canonical}.
+here \\
+\todo Mention ongoing work, like tracking all the constraints to
+get an exact p-value at every step. \\
+
+\newpage
+
+\bibliographystyle{ims}
+%\bibliographystyle{acmtrans-ims}
+\bibliography{paper}
+
+\newpage
 
@@
 \section{Derivation of closed forms of statistic}
+
+\todo Clean this section \\
+\todo Make $M=\lambda_2$ a special case corollary \\
 
@@
 
-\bibliographystyle{ims}
-%\bibliographystyle{acmtrans-ims}
-\bibliography{paper}
-
 \end{document}

### 2014-03-07 glint/fwd_glint_sim.R

@@ -83,13 +83,13 @@ fwd_glint_simulation = function(n, sigma, groups, num.nonzero, lower, upper, nsi
     Y.test = Y.noiseless.test + rnorm(n)*sigma
 
     # Null results
-    results = forward_group(X, Y, groups=all.groups, weights=weights, sigma, max.steps = max.steps)
+    results = forward_group(X, Y, groups=all.groups, int.groups=int.groups, weights=weights, sigma, max.steps = max.steps)
     P.mat[i, ] = results$p.vals
     AS.mat[i, ] = results$active.set
 
     # Non-null results
 
-    results.b = forward_group(X, Y.beta, groups=all.groups, weights, sigma, max.steps = max.steps)
+    results.b = forward_group(X, Y.beta, groups=all.groups, int.groups=int.groups, weights, sigma, max.steps = max.steps)
     P.mat.b[i, ] = results.b$p.vals
     rb.as = results.b$active.set
     recover.mat[i,] = sapply(rb.as, function(x) is.element(x, true.active.groups))
@@ -100,17 +100,16 @@ fwd_glint_simulation = function(n, sigma, groups, num.nonzero, lower, upper, nsi
     already.counted = c()
     cg = 1
     for (ag in rb.as) {
-      ps.recover.mat[i, cg] = true_step_glinternet(ag, p, int.groups, rb.as[1:cg], true.active.groups, already.counted)
       if (ag <= p) {
         already.counted = union(already.counted, ag)
       } else {
         already.counted = union(already.counted, c(main_effects_of(ag, int.groups), ag))
       }
+      ps.recover.mat[i, cg] = length(intersect(already.counted, all.active))/max(length(all.active),1)
       cg = cg + 1
     }
 
   }
-
 
 # HERE
   

### 2014-03-07 glint/group_lasso.R

@@ -51,7 +51,7 @@ trignometric_form = function(num, den, weight, tol=1.e-10) {
 }
 
 # fix Sigma
-group_lasso_knot <- function(X, Y, groups, weights, Sigma = NULL, active.set=0) {
+group_lasso_knot <- function(X, Y, groups, weights, Sigma = NULL, active.set=0, already.counted=c()) {
 
   U = t(X) %*% Y
   g = length(weights)
@@ -64,7 +64,7 @@ group_lasso_knot <- function(X, Y, groups, weights, Sigma = NULL, active.set=0) 
   }
   
   for (j in 1:g) {
-    if (is.element(j, active.set)) {
+    if (is.element(j, already.counted)) {
       terms[j] = 0
     } else {
       terms[j] = sqrt(terms[j]) / weights[j]

### 2014-03-07 glint/sim_glint.R

@@ -8,18 +8,18 @@ source('fwd_glint_sim.R')
 design = 'gaussian'
 corr = 0 # nonzero only supported for gaussian design
 
-nsim = 150
-n = 300
+nsim = 200
+n = 200
 num.nonzero = 6
 k = num.nonzero
 max.steps = 14
-upper.coeff = 9.9
-lower.coeff = 7.4
+upper.coeff = 3.91
+lower.coeff = 2.5
 
 sigma = 1
 groups = 1:20
 p = length(groups)
-mult = sqrt(2*log(p*(p+1)/2))
+mult = sqrt(2*log(p))
 upper = upper.coeff*mult
 lower = lower.coeff*mult
 
@@ -44,7 +44,8 @@ if (corr != 0) {
 }
 ps.fname = paste0(ps.fname, '_glint.pdf')
 m1 = output.l$m1
-psr = t(apply(output.l$psr.mat, 1, cumsum)/m1)
+#psr = t(apply(output.l$psr.mat, 1, cumsum)/m1)
+psr = output.l$psr.mat
 l = nrow(psr)
 m = ncol(psr)
 for (j in 1:nrow(psr)) {
@@ -68,12 +69,12 @@ abline(h = c(1, .8, .6, .4, .2, 0), lty = 3, col = "gray")
 abline(v = 1, col = "gray")
 dev.off()
 
-stop('not now')
 
-n = 400
+stop("Not this time!")
+
 num.nonzero = 6
 k = 6
-max.steps = 12
+max.steps = 10
 groups = sort(c(rep(1:15, 2), rep(16:20, 4)))
 p = length(groups)
 g = length(unique(groups))

### 2014-03-10 paper/paper.tex
Changed glint to normalize by frobenius norm, inflated interaction coeff

@@
 Some classical attempts to address this issue include Monte Carlo
-estimation of tables of adjusted $F$-statistic values \citep{mc:ftoenter}
+estimation of tables of adjusted $F$-statistic values \citep{mc:ftoenter},
 and permutation statistics \citep{permutation:stop}. Aside from the works
@@
 With many variables, fitting models with pairwise interactions can drastically increase model complexity.
-\cite{glint} simplify this problem by imposing a hierarchical constraint on interactions (similar to that in \cite{bien:hierarchical}) where an interaction term can only be included if its main effects are also included.
+\cite{glint} propose a method called \textsc{glinternet} to reduce the statistical complexity of this problem.
+The method imposes a strong hierarchical constraint on interactions (similar to that in \cite{bien:hierarchical}) where an interaction term can only be included if both its main effects are also included.
 They accomplish this by creating a design matrix with main effects alone and also with groups including main effects and their first order interactions.
@@
 where * denotes the pointwise product (Hadamard product) of vectors (the $i$th entry of $X_{g1} * X_{h1}$ is the $i$th entry of $X_{g1}$ times the $i$th entry of $X_{h1}$).
-Note that this expanded matrix has $G(G+1)/2$ groups.
-For the rest of this section we drop the tilde
-and denote the expanded design matrix as $X$.
+Note that this expanded matrix has $\tilde G = G(G+1)/2$ groups.
+%For the rest of this section we drop the tilde and denote the expanded design matrix as $X$.
 After constructing the design matrix, they proceed by
@@
 matrix. The grouping still guarantees that our fitted model will
-satisfy the hierarchy condition. However, we must first modify our
-implementation of forward stepwise to prevent the possibility of
-including main effect groups after the main effects are already included
-in the model as part of an interaction group. To do this we compute the
-residual using the whole model in each step, rather than by subtracting
-the effect of the current group being included. For a set of group indices
-$A \subseteq \{ 1, \ldots G \}$ we write $X_A$ for the submatrix of the
-design matrix containing all the groups in $A$.
-
+satisfy the hierarchy condition.
+To achieve a small computational speed-up, we first modify Algorithm~\ref{algo:fs} to keep track of which main effects have already been included.
+This reduces the search among potential variables.
+%However, we must first modify our implementation of forward stepwise to prevent the possibility of including main effect groups after the main effects are already included in the model as part of an interaction group.
+To do this we compute the residual using the whole model in each step, rather than by subtracting the effect of the current group being included.
+For a set of group indices $A \subseteq \{ 1, \ldots G \}$ we write $X_A$ for the submatrix of the design matrix containing all the groups in $A$.
+% \begin{algorithm}
+% \DontPrintSemicolon
+% \caption{\small \it Modification to compute residual from full model}
+% \BlankLine
+% \CommentSty{\# Everything the same as before, additionally:}\;
+% %  $P_A \gets I - X_AX_A^\dagger$\;
+% %  $r_s \gets P_A r_{s-1}$\;
+% \If{$g^*$ is an interaction group}{
+%   $g_1, g_2 \gets$ main effects groups of $g^*$\;
+%   A^c \gets A^c \slash \{ g_1, g_2 \}\;
+% }%\;
+% %$r_s \gets y - \text{lsfit}(X_A, y)$\;
+% \BlankLine
+% \label{algo:fs:full}
+% \end{algorithm}
 \begin{algorithm}
 \DontPrintSemicolon
-\caption{\small \it Modification to compute residual from full model}
+\KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
+\KwResult{Active set $A$ of variable groups included in the model}
+\SetKwFunction{lsfitResidual}{lsfitResidual}
+\caption{\small \it Forward stepwise for glinternet}
 \BlankLine
-\CommentSty{\# Everything the same as before, except:}\;
-%  $P_A \gets I - X_AX_A^\dagger$\;
-%  $r_s \gets P_A r_{s-1}$\;
-$r_s \gets y - \text{lsfit}(X_A, y)$\;
-\BlankLine
-\label{algo:fs:full}
+Compute design matrix~(\ref{eq:glintmat}) and expanded grouping $\tilde G$\;
+$A \gets \emptyset$\;
+$A^c \gets \{ 1, \ldots, G \}$\;
+$r_0 \gets y$\;
+\For{$s \gets 1$ \KwTo $steps$}{
+  $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
+  $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
+  $A \gets A \cup \{ g^* \}$\;
+  $A^c \gets A^c \backslash \{ g^* \}$\;
+  \For{$g \in A^c$}{
+    $X_g \gets P X_g$\;
+  }
+  $r_s \gets P r_{s-1}$\;
+}
+\KwRet{$A$}
+\label{algo:fsgl}
 \end{algorithm}

### 2014-03-10 fwd_step.R

@@ -93,6 +93,7 @@ forward_group = function(X, Y, groups, weights = 0, sigma = 0, max.steps = 0) {
   active.set = 0
   eff.p = 0
   p.vals = c()
+  chi.pvals = c()
   c.vars = c()
   Ls = c()
 
@@ -105,9 +106,12 @@ forward_group = function(X, Y, groups, weights = 0, sigma = 0, max.steps = 0) {
     output = add_group(X.orig, X.update, Y.update, groups, weights, sigma, active.set, eff.p)
     active.set = output$active.set
     eff.p = output$eff.p
+    RSS = sum(Y.update^2)
     Y.update = output$Y.update
+    RSS = RSS - sum(Y.update^2)
     X.update = output$X.update
     p.vals = c(p.vals, output$p.value)
+    chi.pvals = c(chi.pvals, pchisq(RSS, lower.tail=F, df=1))
     c.vars = c(c.vars, output$var)
     # tracking lambda_2
     Ls = c(Ls, output$test.output[1])
@@ -123,6 +127,6 @@ forward_group = function(X, Y, groups, weights = 0, sigma = 0, max.steps = 0) {
     
   }
 
-  return(list(active.set = active.set, p.vals = p.vals, Ls = Ls))
+  return(list(active.set = active.set, p.vals = p.vals, chi.pvals = chi.pvals, Ls = Ls))
 }
 

### 2014-03-10 fwd_step_sim.R

@@ -21,6 +21,7 @@ fwd_group_simulation = function(n, sigma, groups, beta, nsim,
   P.mat = matrix(nrow=nsim, ncol=max.steps)
   AS.mat = P.mat
   P.mat.b = P.mat
+  Chi.mat.b = P.mat
   AS.mat.b = P.mat
   recover.mat = P.mat
   pred.errs = matrix(0, nrow=3, ncol=3)
@@ -89,6 +90,7 @@ fwd_group_simulation = function(n, sigma, groups, beta, nsim,
 
     # Non-null results
     results.b = forward_group(X, Y.beta, groups, weights, sigma, max.steps = max.steps)
+    Chi.mat.b[i, ] = results.b$chi.pvals
     P.mat.b[i, ] = results.b$p.vals
     AS.mat.b[i, ] = results.b$active.set
     recover.mat[i, ] = sapply(results.b$active.set, function(x)
@@ -136,6 +138,11 @@ fwd_group_simulation = function(n, sigma, groups, beta, nsim,
     Pvals = colMeans(P.mat.b)
     Pvals.bar <- apply(P.mat.b, 2, function(col) quantile(col, probs = bar.quantiles))
     Pvals.point <- apply(P.mat.b, 2, function(col) quantile(col, probs = point.quantiles))
+
+    Chivals = colMeans(Chi.mat.b)
+    Chivals.bar <- apply(Chi.mat.b, 2, function(col) quantile(col, probs = bar.quantiles))
+    Chivals.point <- apply(Chi.mat.b, 2, function(col) quantile(col, probs = point.quantiles))
+
     
     xax <- 1:max.steps - 0.15
     nxax <- xax + 0.3
@@ -155,6 +162,12 @@ fwd_group_simulation = function(n, sigma, groups, beta, nsim,
            code = 3, angle = 90, length = 0, col = "blue")
     points(xax, Pvals.point[1, ], col = "blue", pch = 24, cex = .5)
     points(xax, Pvals.point[2, ], col = "blue", pch = 25, cex = .5)
+
+    points(xax, Chivals, col="green")
+    arrows(xax, Chivals.bar[1, ], xax, Chivals.bar[2, ],
+           code = 3, angle = 90, length = 0, col = "green")
+    points(xax, Chivals.point[1, ], col = "green", pch = 24, cex = .5)
+    points(xax, Chivals.point[2, ], col = "green", pch = 25, cex = .5)
 
   }
 

### 2014-03-10 generate_data.R

@@ -88,24 +88,38 @@ col_normalize = function(X) {
   return(t(t(X) / norms))
 }
 
+# Normalize groups by Frobenius norm
+frob_normalize = function(X, groups) {
+    X.out = X
+    for (g in 1:max(groups)) {
+        inds = groups == g
+        frob.norm = sqrt(sum(X[,inds]^2))
+        if (frob.norm > 0)
+            X.out[,inds] = X.out[,inds]/frob.norm
+    }
+    return(X.out)
+}
+
 # Fixed group sizes, gaussian design
-gaussian_design = function(n, groups, ortho.within = FALSE, corr = 0) {
+gaussian_design = function(n, groups, col.normalize = TRUE, corr = 0) {
   p = length(groups)
   X = matrix(rnorm(n*p), nrow=n)
 
-  if (ortho.within == TRUE) {
-    for (g in 1:max(groups)) {
-      group = groups == g
-      X[ , group] = svd(X[ , group])$u
-    }
-  }
+  ## if (ortho.within == TRUE) {
+  ##   for (g in 1:max(groups)) {
+  ##     group = groups == g
+  ##     X[ , group] = svd(X[ , group])$u
+  ##   }
+  ## }
 
   if (corr != 0) {
     Z = matrix(rep(t(rnorm(n)), p), nrow=n)
     X = sqrt(1-corr)*X + sqrt(corr)*Z
   }
 
-  X = col_normalize(X)
+  if (col.normalize) {
+    X = col_normalize(X)
+  }
   return(X)
 }
 
@@ -194,7 +208,6 @@ derangement = function(inds) {
 # beta nonzeros are for first k/3 main.groups
 
 
-
 # Input design matrix with groups
 # Output larger design matrix with all possible (grouped) interactions
 # main.groups: groups of main effects (original variables)
@@ -202,6 +215,7 @@ derangement = function(inds) {
 # int.groups: list, the [[g]]'th element contains all group indices that group g appears in
 generate_glinternet = function(X, groups) {
   X.out = X
+  n = dim(X)[1]
   g.out = groups
   main.out = groups
   gmax = max(groups)
@@ -213,16 +227,16 @@ generate_glinternet = function(X, groups) {
       hinds = which(groups == h)
       gs = length(ginds)
       hs = length(hinds)
-      X.out = cbind(X.out, X[, ginds]/sqrt(2))
[…]

### 2014-03-10 glint/fwd_glint_sim.R

@@ -28,6 +28,8 @@ fwd_glint_simulation = function(n, sigma, groups, num.nonzero, lower, upper, nsi
   start.time = as.numeric(Sys.time())
 
   ### Begin main loop ###
+main3 = c()  
+int37 = c()
   for (i in 1:nsim) {
 
     # Monitoring completion time
@@ -39,8 +41,8 @@ fwd_glint_simulation = function(n, sigma, groups, num.nonzero, lower, upper, nsi
     }
 
     if (design == 'gaussian') {
-      X = gaussian_design(n, groups, corr = corr)
-      X.test = gaussian_design(n, groups, corr = corr)
+      X = gaussian_design(n, groups, col.normalize = FALSE, corr = corr)
+      X.test = gaussian_design(n, groups, col.normalize = FALSE, corr = corr)
 
     } else {
       design_name = paste0(design, "_design")
@@ -55,11 +57,13 @@ fwd_glint_simulation = function(n, sigma, groups, num.nonzero, lower, upper, nsi
     # Glinternet specific part
     data = generate_glinternet(X, groups)
     data.test = generate_glinternet(X.test, groups)
-    X = data$X
-    X.test = data.test$X
     all.groups = data$all.groups
+    X = frob_normalize(data$X, all.groups)
+    X.test = frob_normalize(data.test$X, all.groups)
+
 #    weights = sqrt(sapply(rle(all.groups)$lengths - 1, function(x) max(1, x)))
     weights = rep(1, max(all.groups))
+#    weights = sqrt(rle(all.groups)$lengths)
     main.groups = data$main.groups
     int.groups = data$int.groups
     beta.data = beta_glinternet(all.groups=all.groups, int.groups=int.groups, num.nonzero=k, upper=upper, lower=lower)
@@ -81,6 +85,10 @@ fwd_glint_simulation = function(n, sigma, groups, num.nonzero, lower, upper, nsi
     Y.noiseless.test = X.test %*% beta
     Y.beta = Y.noiseless + Y
     Y.test = Y.noiseless.test + rnorm(n)*sigma
+
+main3 = c(main3, abs(t(X[,3]) %*% Y.beta))
+z=t(X[,all.groups==61])%*%Y.beta
+int37 = c(int37, sqrt(sum(z^2)))
 
     # Null results
[…]

### 2014-03-10 glint/fwd_step.R

@@ -1,7 +1,3 @@
-# Does this need to be modified for glinternet?
-# YES!
-# Form residuals using the full model after every step
-# (Do NOT only regress out the next variable being included)
 
 library(MASS) # for ginv
 source('group_lasso.R')
@@ -67,7 +63,7 @@ add_group = function(X, Y, groups, int.groups, weights, sigma, active.set = 0, a
 ##     }
 ##   }
   Pgmax = Xgmax.regress %*% ginv(Xgmax.regress)
-  Y.resid = lm(Y ~ Xgmax.regress)$residual
+  Y.resid = lm(Y ~ Xgmax.regress-1)$residual
 # full version?  
 #  Y.resid = Y
       

### 2014-03-10 glint/sim_glint.R

@@ -8,13 +8,13 @@ source('fwd_glint_sim.R')
 design = 'gaussian'
 corr = 0 # nonzero only supported for gaussian design
 
-nsim = 200
-n = 200
+nsim = 150
+n = 250
 num.nonzero = 6
 k = num.nonzero
-max.steps = 14
-upper.coeff = 3.91
-lower.coeff = 2.5
+max.steps = 10
+upper.coeff = 2.9
+lower.coeff = 1.4
 
 sigma = 1
 groups = 1:20
@@ -70,19 +70,20 @@ abline(v = 1, col = "gray")
 dev.off()
 
 
-stop("Not this time!")
+
+#stop("Not this time!")
 
 num.nonzero = 6
 k = 6
-max.steps = 10
-groups = sort(c(rep(1:15, 2), rep(16:20, 4)))
+max.steps = 8
+groups = sort(c(rep(1:15, 2), rep(16:20, 3)))
 p = length(groups)
 g = length(unique(groups))
 mult = sqrt(2*log(g))
 upper = upper.coeff*mult
 lower = lower.coeff*mult
 
-filename = paste0('../figs/', design, '_size2-4_n', n, '_p', p, '_g', g, '_k', num.nonzero, '_lower', lower.coeff, '_upper', upper.coeff)
+filename = paste0('../figs/', design, '_size2-3_n', n, '_p', p, '_g', g, '_k', num.nonzero, '_lower', lower.coeff, '_upper', upper.coeff)
 if (corr != 0) {
   filename = paste0(filename, '_corr', corr)
 }
@@ -92,6 +93,9 @@ output.g <- fwd_glint_simulation(n, sigma, groups, num.nonzero, lower, upper, ns
 dev.off()
 
 print(warnings())
+
+print(summary(output.l$main3))
+print(summary(output.l$int37))
 
 stop("No selection stats for this")
 

### 2014-03-10 sim_large.R

@@ -5,14 +5,15 @@ source('fwd_step_sim.R')
 source('tex_table.R')
 
 
-design = 'ternary'
+#design = 'ternary'
+design = 'gaussian'
 corr = 0 # nonzero only supported for gaussian design
 
 nsim = 500
 n = 100
-num.nonzero = 10
+num.nonzero = 6
 k = num.nonzero
-max.steps = 14
+max.steps = 12
 upper.coeff = 1.7
 lower.coeff = 1.5
 

### 2014-03-11 generate_data.R
Glint fixed... again

@@ -227,16 +227,18 @@ generate_glinternet = function(X, groups) {
       hinds = which(groups == h)
       gs = length(ginds)
       hs = length(hinds)
+      # Intercept term shrunk too much?
+#      X.out = cbind(X.out, rep(1/n,n)) 
       X.out = cbind(X.out, X[, ginds]) #/sqrt(2))
       X.out = cbind(X.out, X[, hinds]) #/sqrt(2))
       gnew = gnew + 1
       for (i in ginds) {
         for (j in hinds) {
           Xij = X[, i] * X[, j]
-          ## normij = sqrt(sum(Xij^2))
-          ## if (normij > 0) {
-          ##   Xij = Xij/normij
-          ## }
+          normij = sqrt(sum(Xij^2))
+          if (normij > 0) {
+            Xij = Xij/normij
+          }
           X.out = cbind(X.out, Xij)
         }
       }
@@ -301,15 +303,21 @@ beta_glinternet = function(all.groups, int.groups, num.nonzero, upper, lower, ra
       # i.e. they do not share their coeff mass with the main effects
       # otherwise it would be difficult to find them
       mains.of.g = main_effects_of(g, int.groups)
+      mgs = 0
+      for (h in mains.of.g) {
+        mgs = mgs + sum(h == all.groups)
+      }
+      igs = gs - mgs
       ginds = which(group)
       start.ind = 1
       for (h in mains.of.g) {
         hs = sum(h == all.groups)
-        beta[ginds[start.ind:hs]] = beta[ginds[start.ind:hs]]/sqrt(hs)
-        start.ind = start.ind + hs
+        end.ind = start.ind + hs - 1
+        beta[ginds[start.ind:end.ind]] = beta[ginds[start.ind:end.ind]]/sqrt(mgs)
+        start.ind = end.ind+1
       }
       # Inflate interaction size
-      beta[ginds[start.ind:gs]] = sqrt(gs)*beta[ginds[start.ind:gs]] #/sqrt(gs-start.ind+1)
+      beta[ginds[start.ind:gs]] = sqrt(2)*beta[ginds[start.ind:gs]]/sqrt(igs)
     }
   }
 
@@ -331,9 +339,9 @@ int_group_of = function(g, h, int.groups) {
 main_effects_of = function(gh, int.groups) {
   main.effs = which(int.groups == gh, arr.ind=TRUE)[1,]
   if (main.effs[1] < main.effs[2]) {
[…]

### 2014-03-11 glint/fwd_glint_sim.R

@@ -5,15 +5,26 @@ source('../generate_data.R')
 #source('pred_and_estim.R') not for glinternet
 #source('fwd_step/coherence.R')
 
+## n=250
+## sigma=1
+## groups=1:20
+## num.nonzero=6
+## max.steps=12
+## lower=7
+## upper=9
+## nsim=5
+## design="gaussian"
+## corr=0
+## k=6
 
 # Note: glint version passes k here instead of beta
 fwd_glint_simulation = function(n, sigma, groups, num.nonzero, lower, upper, nsim,
   max.steps, alpha = .1, design = 'gaussian', corr = 0, categorical = FALSE, predictions = FALSE, coherence = FALSE, plot = FALSE) {
 
   # Initialize
-  weights = sqrt(rle(groups)$lengths)
   p = length(groups)
   g = length(unique(groups))
+  G = g*(g+1)/2
   k=num.nonzero
   P.mat = matrix(nrow=nsim, ncol=max.steps)
   AS.mat = P.mat
@@ -28,10 +39,10 @@ fwd_glint_simulation = function(n, sigma, groups, num.nonzero, lower, upper, nsi
   start.time = as.numeric(Sys.time())
 
   ### Begin main loop ###
-main3 = c()  
-int37 = c()
+  main1 = c()  
+  int37 = c()
+  
   for (i in 1:nsim) {
-
     # Monitoring completion time
     if (i %% 10 == 0) {
       elapsed.time = as.numeric(Sys.time()) - start.time
@@ -41,8 +52,8 @@ int37 = c()
     }
 
     if (design == 'gaussian') {
-      X = gaussian_design(n, groups, col.normalize = FALSE, corr = corr)
-      X.test = gaussian_design(n, groups, col.normalize = FALSE, corr = corr)
+      X = gaussian_design(n, groups, col.normalize = TRUE, corr = corr)
+      X.test = gaussian_design(n, groups, col.normalize = TRUE, corr = corr)
 
     } else {
       design_name = paste0(design, "_design")
@@ -58,16 +69,19 @@ int37 = c()
     data = generate_glinternet(X, groups)
     data.test = generate_glinternet(X.test, groups)
     all.groups = data$all.groups
+    X = data$X
     X = frob_normalize(data$X, all.groups)
     X.test = frob_normalize(data.test$X, all.groups)
-
[…]

### 2014-03-11 glint/fwd_step.R

@@ -64,8 +64,6 @@ add_group = function(X, Y, groups, int.groups, weights, sigma, active.set = 0, a
 ##   }
   Pgmax = Xgmax.regress %*% ginv(Xgmax.regress)
   Y.resid = lm(Y ~ Xgmax.regress-1)$residual
-# full version?  
-#  Y.resid = Y
       
 ####### This is necessary for p-value? ########
   # Project all other groups orthogonal to the one being added
@@ -89,6 +87,7 @@ forward_group = function(X, Y, groups, int.groups, weights = 0, sigma = 0, max.s
   group.sizes = rle(groups)$lengths
 
   if ((length(weights) == 1) & (weights[1] == 0)) {
+    print("######## Making weights ############")
     weights = sqrt(rle(groups)$lengths)
   }
 

### 2014-03-11 glint/group_lasso.R

@@ -64,7 +64,7 @@ group_lasso_knot <- function(X, Y, groups, weights, Sigma = NULL, active.set=0, 
   }
   
   for (j in 1:g) {
-    if (is.element(j, already.counted)) {
+    if (is.element(j, active.set)) {
       terms[j] = 0
     } else {
       terms[j] = sqrt(terms[j]) / weights[j]

### 2014-03-13 paper/paper.tex
Glint fixed... again
Changed algorithm2e to algorithmicx

@@
 \startlocaldefs
-
+\usepackage{algorithm}
+\usepackage{algorithmic}
 \usepackage{graphics,amssymb,color}
 \usepackage{graphicx,subfigure}
-\usepackage[boxed]{algorithm2e}
+%\usepackage[boxed]{algorithm2e}
+
+\renewcommand{\algorithmicrequire}{\textbf{Input:}}
+\renewcommand{\algorithmicensure}{\textbf{Output:}}
 
@@
 \newcommand{\pen}{\mathcal{P}}
+\newcommand{\gstar}{g^*}
 
@@
 different from the most commonly used versions.
-(\todo Check R's \textit{step} for the group criterion)
 
@@
 
+% \begin{algorithm}
+% \DontPrintSemicolon
+% \KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
+% \KwResult{Active set $A$ of variable groups included in the model}
+% \SetKwFunction{lsfitResidual}{lsfitResidual}
+% \caption{\small \it Forward stepwise procedure with groups and weights}
+% \BlankLine
+% $A \gets \emptyset$\;
+% $A^c \gets \{ 1, \ldots, G \}$\;
+% $r_0 \gets y$\;
+% \For{$s \gets 1$ \KwTo $steps$}{
+%   $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
+%   $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
+%   $A \gets A \cup \{ g^* \}$\;
+%   $A^c \gets A^c \backslash \{ g^* \}$\;
+%   \For{$g \in A^c$}{
+%     $X_g \gets P X_g$\;
+%   }
+%   $r_s \gets P r_{s-1}$\;
+% }
+% \KwRet{$A$}
+% \label{algo:fs}
+% \end{algorithm}
+
 \begin{algorithm}
-\DontPrintSemicolon
-\KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
-\KwResult{Active set $A$ of variable groups included in the model}
-\SetKwFunction{lsfitResidual}{lsfitResidual}
-\caption{\small \it Forward stepwise procedure with groups and weights}
-\BlankLine
-$A \gets \emptyset$\;
-$A^c \gets \{ 1, \ldots, G \}$\;
-$r_0 \gets y$\;
-\For{$s \gets 1$ \KwTo $steps$}{
-  $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
-  $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
-  $A \gets A \cup \{ g^* \}$\;
-  $A^c \gets A^c \backslash \{ g^* \}$\;
-  \For{$g \in A^c$}{
-    $X_g \gets P X_g$\;
-  }
-  $r_s \gets P r_{s-1}$\;
-}
-\KwRet{$A$}
+  \caption{Forward stepwise variant with groups and weights}
   \label{algo:fs}
+  \begin{algorithmic}[1]
+    \REQUIRE An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ variable groups with weights $w_g$
+    \ENSURE Ordered active set $A$ of variable groups included in the model at each step
+    \STATE $A \gets \emptyset$
+    \STATE $A^c \gets \{ 1, \ldots, G\}$
+    \FOR{$s=1$ to $steps$}
+    \STATE $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$
+    \STATE $P \gets I - X_{g^*}X_{g^*}^\dagger$
+    \STATE $A \gets A \cup \{ g^* \}$
+    \STATE $A^c \gets A^c \backslash \{ g^* \}$
+    \FORALL{$g \in A^c$}
+      \STATE $X_g \gets P X_g$
+    \ENDFOR
+    \STATE $r_s \gets P r_{s-1}$
+    \ENDFOR
+    \RETURN $A$
+  \end{algorithmic}
 \end{algorithm}
@@
 We will now use forward stepwise to refer specifically to
-Algorithm~\ref{algo:fs} unless otherwise specified.
+Algorithm~\ref{algo:fs} unless otherwise specified. Note that other
+implementations of forward stepwise use different criteria for choosing
+the next variable to add, such as the correlation with the residual.
+Since we do not renormalize the columns after projecting the covariates
+(lines 8 to 10 above), and since we have weights, we are not computing
+correlations unless the design matrix is orthogonal and all weights are 1.
+There are advantages and disadvantages to both criteria which we do not
+discuss. Our choice was motivated by the group lasso result in
+\cite{tests:adaptive}, but we believe other criteria can be handled
+with appropriate modifications.
 
@@
 
-\subsection{Better p-values}
+\subsection{Relation to covariance test}
 
-Maybe just reference \cite{tests:adaptive}? It might be worthwhile to
-leave in the connection with \cite{significance:lasso} (getting our
-test statistic by not using the approximation)
+Consider the simple case for the first covariate to enter the model.
+In this case the covariance test statistic is $T = \lambda_1(\lambda_1 - M)$.
+\todo Does $M = \lambda_2$? Check proof, now that statistic has changed
+
+This is Lemma 5 from \cite{significance:lasso}.
 
@@
 
-\item \todo Add the figures to this section
-\item \todo Convert to exposition instead of bulleted list
-\item \todo Update to reflect latest work
   \item As in LTTT, $T = \lambda_1 (\lambda_1 - M)$ and $M = \lambda_2$
@@
 \end{itemize}
+
+\subsection{Exact p-value calculation}
+
+We now describe how to calculate the p-value, following the discussion of
+group lasso in \cite{tests:adaptive}. Let $g = \gstar$ be the index of
+the group attaining the maximum on line 4 of Algorithm~\ref{algo:fs},
+and let $r = \text{rank}(X_g)$ be the rank of the group attaining the maximum.
 
@@
 \label{sec:glint}
-With many variables, fitting models with pairwise interactions can drastically increase model complexity.
+In regression settings with many variables, considering models with pairwise interactions can drastically increase model complexity.
 \cite{glint} propose a method called \textsc{glinternet} to reduce the statistical complexity of this problem.
 The method imposes a strong hierarchical constraint on interactions (similar to that in \cite{bien:hierarchical}) where an interaction term can only be included if both its main effects are also included.
-They accomplish this by creating a design matrix with main effects alone and also with groups including main effects and their first order interactions.
-Concretely, the expanded design matrix is
+They accomplish this by creating a design matrix with both main effects alone and also with groups including main effects with their first order interactions.
+Then they fit a group lasso model with the expanded design matrix.
+Because interaction terms only appear in groups along with their respective main effects, the hierarchy condition holds for the fitted model.
+We now consider a related procedure as an example problem, but first modify their method to simplify some parts.
+Let the expanded design matrix be given by
 \begin{equation}
@@
 \]
-where * denotes the pointwise product (Hadamard product) of vectors (the $i$th entry of $X_{g1} * X_{h1}$ is the $i$th entry of $X_{g1}$ times the $i$th entry of $X_{h1}$).
-Note that this expanded matrix has $\tilde G = G(G+1)/2$ groups.
-%For the rest of this section we drop the tilde and denote the expanded design matrix as $X$.
-After constructing the design matrix, they proceed by
-fitting a group lasso model to the data. Because interaction terms only
-appear in groups along with their main effects, the hierarchy condition
-holds for the fitted model.
+where * denotes the pointwise product (Hadamard product) of vectors
+(the $i$th entry of $X_{g1} * X_{h1}$ is the $i$th entry of $X_{g1}$ times
+the $i$th entry of $X_{h1}$). Note that this expanded matrix has
+$\tilde G = G + \binom{G}{2} = O(G^2)$ groups. We refer to the first
+$G$ of these as main effects or main effect groups, and the remaining
+as interaction groups. Finally, instead of fitting a
+model by group lasso, we use forward stepwise on the expanded design
+matrix. The overlapping groups still guarantee that our fitted model will
+satisfy the strong hierarchy condition.
 
-We can also fit a model by forward stepwise on the same expanded design
-matrix. The grouping still guarantees that our fitted model will
-satisfy the hierarchy condition.
-To achieve a small computational speed-up, we first modify Algorithm~\ref{algo:fs} to keep track of which main effects have already been included.
-This reduces the search among potential variables.
 %However, we must first modify our implementation of forward stepwise to prevent the possibility of including main effect groups after the main effects are already included in the model as part of an interaction group.
@@
 % \end{algorithm}
-\begin{algorithm}
-\DontPrintSemicolon
-\KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
-\KwResult{Active set $A$ of variable groups included in the model}
-\SetKwFunction{lsfitResidual}{lsfitResidual}
-\caption{\small \it Forward stepwise for glinternet}
-\BlankLine
-Compute design matrix~(\ref{eq:glintmat}) and expanded grouping $\tilde G$\;
-$A \gets \emptyset$\;
-$A^c \gets \{ 1, \ldots, G \}$\;
-$r_0 \gets y$\;
-\For{$s \gets 1$ \KwTo $steps$}{
-  $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
-  $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
-  $A \gets A \cup \{ g^* \}$\;
-  $A^c \gets A^c \backslash \{ g^* \}$\;
-  \For{$g \in A^c$}{
-    $X_g \gets P X_g$\;
-  }
-  $r_s \gets P r_{s-1}$\;
-}
-\KwRet{$A$}
-\label{algo:fsgl}
-\end{algorithm}
+% \begin{algorithm}
+% \DontPrintSemicolon
+% \KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
+% \KwResult{Active set $A$ of variable groups included in the model}
+% \SetKwFunction{lsfitResidual}{lsfitResidual}
+% \caption{\small \it Forward stepwise for glinternet}
+% \BlankLine
+% Compute design matrix~(\ref{eq:glintmat}) and expanded grouping $\tilde G$\;
+% $A \gets \emptyset$\;
+% $A^c \gets \{ 1, \ldots, G \}$\;
+% $r_0 \gets y$\;
+% \For{$s \gets 1$ \KwTo $steps$}{
+%   $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
+%   $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
+%   $A \gets A \cup \{ g^* \}$\;
+%   $A^c \gets A^c \backslash \{ g^* \}$\;
+%   \For{$g \in A^c$}{
+%     $X_g \gets P X_g$\;
+%   }
+%   $r_s \gets P r_{s-1}$\;
+% }
+% \KwRet{$A$}
+% \label{algo:fsgl}
+% \end{algorithm}
 
-Testing this method by simulation, we constructed signals which have
-$k/3$ main effects nonzero but with no interactions and the remaining
-$2k/3$ nonzero main effects are matched to each other to form interactions.
+To demonstrate this method by simulation, we constructed signals which have
+the first $k/3$ main effects nonzero but with no interactions, and the
+remaining $2k/3$ nonzero main effects are matched to each other to form
+interactions.
 This way each nonzero main effect with any nonzero interactions has
-exactly one nonzero interaction.
-The design matrices are independent Gaussian with all groups of size 1.
+exactly one nonzero interaction. Furthermore, each nonzero interaction
+coefficient has been inflated to be larger than the corresponding main
+effect coefficients. This special case is favorable for our algorithm,
+but our purpose here is merely to demonstrate the flexibility of the
+hypothesis test and not to propose a general procedure for models with
+interactions.
+
 Results are shown in Figure~\ref{fig:glint}. The left
@@
 for various sparsity levels. The right panel shows our p-value marginally
-over each step, with red indiciating the global null case and blue
+over each step, with red indicating the global null case and blue
 indicating a signal with $k=6$ nonzero groups.}

### 2014-03-13 glint/fwd_glint_sim.R

@@ -29,6 +29,7 @@ fwd_glint_simulation = function(n, sigma, groups, num.nonzero, lower, upper, nsi
   P.mat = matrix(nrow=nsim, ncol=max.steps)
   AS.mat = P.mat
   P.mat.b = P.mat
+  Chi.mat.b = P.mat
   AS.mat.b = P.mat
   recover.mat = P.mat
   int.recover.mat = P.mat
@@ -111,6 +112,7 @@ fwd_glint_simulation = function(n, sigma, groups, num.nonzero, lower, upper, nsi
 
     # Non-null results
     results.b = forward_group(X, Y.beta, groups=all.groups, int.groups=int.groups, weights, sigma, max.steps = max.steps)
+    Chi.mat.b[i, ] = results.b$chi.pvals
     P.mat.b[i, ] = results.b$p.vals
     rb.as = results.b$active.set
     recover.mat[i,] = sapply(rb.as, function(x) is.element(x, true.active.groups))
@@ -155,6 +157,10 @@ fwd_glint_simulation = function(n, sigma, groups, num.nonzero, lower, upper, nsi
     Pvals = colMeans(P.mat.b)
     Pvals.bar <- apply(P.mat.b, 2, function(col) quantile(col, probs = bar.quantiles))
     Pvals.point <- apply(P.mat.b, 2, function(col) quantile(col, probs = point.quantiles))
+
+    Chivals = colMeans(Chi.mat.b)
+    Chivals.bar <- apply(Chi.mat.b, 2, function(col) quantile(col, probs = bar.quantiles))
+    Chivals.point <- apply(Chi.mat.b, 2, function(col) quantile(col, probs = point.quantiles))
     
     xax <- 1:max.steps - 0.15
     nxax <- xax + 0.3
@@ -175,6 +181,12 @@ fwd_glint_simulation = function(n, sigma, groups, num.nonzero, lower, upper, nsi
            code = 3, angle = 90, length = 0, col = "blue")
     points(xax, Pvals.point[1, ], col = "blue", pch = 24, cex = .5)
     points(xax, Pvals.point[2, ], col = "blue", pch = 25, cex = .5)
+
+    points(xax, Chivals, col="green")
+    arrows(xax, Chivals.bar[1, ], xax, Chivals.bar[2, ],
+           code = 3, angle = 90, length = 0, col = "green")
+    points(xax, Chivals.point[1, ], col = "green", pch = 24, cex = .5)
[…]

### 2014-03-13 glint/fwd_step.R

@@ -63,7 +63,8 @@ add_group = function(X, Y, groups, int.groups, weights, sigma, active.set = 0, a
 ##     }
 ##   }
   Pgmax = Xgmax.regress %*% ginv(Xgmax.regress)
-  Y.resid = lm(Y ~ Xgmax.regress-1)$residual
+  Y.resid = Y - Pgmax %*% Y
+    #lm(Y ~ Xgmax.regress-1)$residual
       
 ####### This is necessary for p-value? ########
   # Project all other groups orthogonal to the one being added
@@ -87,8 +88,7 @@ forward_group = function(X, Y, groups, int.groups, weights = 0, sigma = 0, max.s
   group.sizes = rle(groups)$lengths
 
   if ((length(weights) == 1) & (weights[1] == 0)) {
-    print("######## Making weights ############")
-    weights = sqrt(rle(groups)$lengths)
+    weights = sqrt(group.sizes)
   }
 
   # Estimate sigma instead?
@@ -104,6 +104,7 @@ forward_group = function(X, Y, groups, int.groups, weights = 0, sigma = 0, max.s
   already.counted = c()
   eff.p = 0
   p.vals = c()
+  chi.pvals = c()
   c.vars = c()
   Ls = c()
 
@@ -111,16 +112,18 @@ forward_group = function(X, Y, groups, int.groups, weights = 0, sigma = 0, max.s
   X.update = X
 
   for (i in 1:max.steps) {
-    
     output = add_group(X.update, Y.update, groups, int.groups, weights, sigma, active.set, already.counted, eff.p)
     active.set = output$active.set
+    imax = output$imax
+    grank = sum(groups == imax)
     already.counted = output$already.counted
     eff.p = output$eff.p
+    RSS = sum(Y.update^2)
     Y.update = output$Y.update
-    active.inds = sapply(groups, function(x) is.element(x, active.set))
-#    Y.update = lm(Y ~ X[, active.inds])$residual
+    RSS = RSS - sum(Y.update^2)
     X.update = output$X.update
     p.vals = c(p.vals, output$p.value)
+    chi.pvals = c(chi.pvals, pchisq(RSS, lower.tail=F, df=grank))
     c.vars = c(c.vars, output$var)
     # tracking lambda_2
     Ls = c(Ls, output$test.output[1])
@@ -134,6 +137,6 @@ forward_group = function(X, Y, groups, int.groups, weights = 0, sigma = 0, max.s
[…]

### 2014-03-13 glint/sim_glint.R

@@ -8,13 +8,13 @@ source('fwd_glint_sim.R')
 design = 'gaussian'
 corr = 0 # nonzero only supported for gaussian design
 
-nsim = 50
-n = 200
+nsim = 200
+n = 250
 num.nonzero = 6
 k = num.nonzero
 max.steps = 10
-upper.coeff = 5
-lower.coeff = 3
+upper.coeff = 8
+lower.coeff = 5
 
 sigma = 1
 groups = 1:20
@@ -73,7 +73,6 @@ dev.off()
 
 #stop("Not this time!")
 
-n = 250
 num.nonzero = 6
 k = 6
 max.steps = 9

### 2014-03-14 paper/paper.tex
Paper: editing pval calculation

@@
     \ENSURE Ordered active set $A$ of variable groups included in the model at each step
-    \STATE $A \gets \emptyset$
-    \STATE $A^c \gets \{ 1, \ldots, G\}$
+    \STATE $A \gets \emptyset$, $A^c \gets \{ 1, \ldots, G\}$
     \FOR{$s=1$ to $steps$}
@@
     \STATE $P \gets I - X_{g^*}X_{g^*}^\dagger$
-    \STATE $A \gets A \cup \{ g^* \}$
-    \STATE $A^c \gets A^c \backslash \{ g^* \}$
+    \STATE $A \gets A \cup \{ g^* \}$, $A^c \gets A^c \backslash \{ g^* \}$
     \FORALL{$g \in A^c$}
@@
 Since we do not renormalize the columns after projecting the covariates
-(lines 8 to 10 above), and since we have weights, we are not computing
+(lines 6 to 8 above), and since we have weights, we are not computing
 correlations unless the design matrix is orthogonal and all weights are 1.
@@
 
-We now describe how to calculate the p-value, following the discussion of
-group lasso in \cite{tests:adaptive}. Let $g = \gstar$ be the index of
-the group attaining the maximum on line 4 of Algorithm~\ref{algo:fs},
-and let $r = \text{rank}(X_g)$ be the rank of the group attaining the maximum.
+We now describe how to calculate our p-value, following the discussion of
+group lasso in \cite{tests:adaptive}. For the rest of this section let
+$g = \gstar$ be the index of
+the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
+%\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
+%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
+For a vector $u \in \real^p$, let $u_h$ denote the coordinates
+of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
+We can rearrange the columns of $X$ to group these adjacently, so that
+\[
+u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
+\]
+One step of the calculation will be to find an orthonormal basis for the
+linear space $V = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
+\text{ for all } h \neq g \}$ so that we can project orthogonally to
+this space. If $X_g$ is a single column and $X_g^Ty \neq 0$, then the
+space is trivial and the desired orthogonal projection is the identity.
+Otherwise, if $X_g$ has $p_g > 1$ columns,
+
+Define $U_g = X_g^Ty/(w_g\norm{X_g^Ty}_2)$.
+\begin{algorithm}
+  \caption{Computing p-value}
+  \label{algo:pval}
+  \begin{algorithmic}[1]
+    \REQUIRE Response $y$, design matrix $X$, index $g$ of next group to enter
+    \ENSURE P-value for testing if $g$ and all other remaining groups are noise
+    \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$, $r \gets \text{rank}(X_g)$
+    \IF{$X_g$ is a single column}
+    \STATE $P \gets 0$ is a $p \times p$ matrix
+    \STATE $\sigma^2_g \gets X_g^T \Sigma X_g$
+    \ELSE
+    \STATE Do that
+    \ENDIF
+    \RETURN $A$
+  \end{algorithmic}
+\end{algorithm}
 

### 2014-03-14 paper/paper.tex
Paper: editing projection calculation

@@
 \newcommand{\pen}{\mathcal{P}}
+\newcommand{\vecsp}{\mathcal{L}}
 \newcommand{\gstar}{g^*}
@@
 $g$th group is the $n \times p_g$ matrix denoted $X_g$, where $p_g$ is
-the number of individual variables or columns in group $g$. When a group
-encodes a categorical variable as binary indicators for the levels of
-that variable, we use the full encoding with a column for every level.
-Although this introduces collinearity, our method does not require the
-design matrix to have full rank.
+the number of individual variables or columns in group $g$.
+When a group encodes a categorical variable as binary indicators for the
+levels of that variable, we use the full encoding with a column for every
+level. Although this introduces collinearity, our method does not require
+the design matrix to have full rank.
 
@@
 One step of the calculation will be to find an orthonormal basis for the
-linear space $V = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
+linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
 \text{ for all } h \neq g \}$ so that we can project orthogonally to
-this space. If $X_g$ is a single column and $X_g^Ty \neq 0$, then the
+this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
+should be the case since $g$ maximizes the absolute value of this
+quantity), then the
 space is trivial and the desired orthogonal projection is the identity.
-Otherwise, if $X_g$ has $p_g > 1$ columns,
+Otherwise, if $X_g$ has $p_g > 1$ columns, the space $\vecsp_g$ will
+generally
+have dimension $r-1$ for $r = \text{rank}(X_g)$. This holds except in
+degenerate cases such as $X_g = 0$ or $r \ll p_g$, but these cases are not
+likely to occur since $\norm{X_g^T y}_2/w_g$ is supposed to be maximized.
+We proceed by forming the $n \times p_g$ matrix
+$X_g - X_g X_g^T y y^T X_g/\norm{X_g^Ty}_2^2$ and performing Gram-Schmidt
+on the first $r-1$ columns. For precise theorems showing that this will
+generally work, see \cite{outerp:indep}. Next we take the basis
+computed by Gram-Schmidt and form an $p \times r-1$ matrix, which we
+denote $V_g$, by appending 0's in the coordinates corresponding to all
+groups $h \neq g$ and to $p_g - r + 1$ columns in group $g$. With this
+we are ready to define the projection
+
+\begin{equation}
+  \begin{aligned}
+    P_g &= \Sigma XV_g (V_g^T X^T \Sigma X V_g)^\dagger V_g^TX^T \\
+  \end{aligned}
+\end{equation}
 

### 2014-03-14 group_lasso.R
Code: finished? general projection

@@ -3,6 +3,7 @@
 # This file contains core functions for computing the test statistic
 
 library(Matrix)
+library(MASS)
 
 trignometric_form = function(num, den, weight, tol=1.e-10) {
   
@@ -70,7 +71,6 @@ group_lasso_knot <- function(X, Y, groups, weights, Sigma = NULL, active.set=0) 
       terms[j] = sqrt(terms[j]) / weights[j]
     }
   }
-  
   imax = which.max(terms)
   L = terms[imax]
   if (L <= 0) {
@@ -81,74 +81,58 @@ group_lasso_knot <- function(X, Y, groups, weights, Sigma = NULL, active.set=0) 
   which = groups == imax
   Uwhich = U[which]
   Xmax = X[,which]
-  #
-  # assuming rank == num of variables in group...
   kmax = sum(which)
   kmaxrank = kmax
   if (length(dim(Xmax)) == 2) {
     kmaxrank = rankMatrix(Xmax)[1]
-    if (kmaxrank != kmax) {
-#      print(paste("Expected rank, computed rank =", kmax, kmaxrank))
-#      kmax = kmaxrank
-    }
   }
 
-  #
-  #
-  # fix this Sigma
-  Sigma = diag(rep(1, kmax))
-  # Uwhich = X_{g^star}^T y
-  
-  soln = (Uwhich / sqrt(sum(Uwhich^2))) / wmax
-  if (kmaxrank > 1) {
-#    print("kmaxrank > 1")
-    #soln = (Uwhich / sqrt(sum(Uwhich^2))) / wmax
-    Xeta = (Xmax %*% soln)[,1]
-    Xmax = Xmax - outer(Xeta, soln, '*') / sum(soln^2)
-    Wmax = Xmax[,1:(ncol(Xmax)-1)]
-    #Xeta = lsfit(Wmax, Xeta, intercept=FALSE)$residuals
-    Xeta = lm(Xeta ~ Wmax - 1)$residuals
-    conditional_variance = sum(Xeta^2)
-  } else if (kmax >= 2) {
-#    print("kmax >= 2")
-    # case where Xmax is a matrix but has rank 1
-    Xeta = Xmax / wmax * sign(U[which])
-    if (dim(Xmax)[2] == 1) {
-#      print("and 1 column")
-      conditional_variance = t(Xeta[,1]) %*% Sigma %*% Xeta[,1]
-    } else {
-#      print("and >1 columns")
[…]

### 2014-03-14 sim_large.R

@@ -9,13 +9,13 @@ source('tex_table.R')
 design = 'gaussian'
 corr = 0 # nonzero only supported for gaussian design
 
-nsim = 500
+nsim = 40
 n = 100
 num.nonzero = 6
 k = num.nonzero
-max.steps = 12
-upper.coeff = 1.7
-lower.coeff = 1.5
+max.steps = 10
+upper.coeff = 5.7
+lower.coeff = 3.9
 
 sigma = 1
 groups = 1:200
@@ -39,10 +39,11 @@ output.l <- fwd_group_simulation(n, sigma, groups, beta, nsim, max.steps, design
 dev.off()
 
 
+nsim = 80
 groups = sort(c(rep(1:30, 5), rep(31:35, 10)))
 p = length(groups)
 g = length(unique(groups))
-mult = sqrt(2*log(g))
+mult = sqrt(2*log(p)) # or log(g)?
 upper = upper.coeff*mult
 lower = lower.coeff*mult
 beta = beta_staircase(groups, num.nonzero, upper, lower)

### 2014-03-17 paper/paper.tex
Code: finished? general projection
Added HIV code and section

--- 2014-03-14 paper/paper.tex
@@
 $g$th group is the $n \times p_g$ matrix denoted $X_g$, where $p_g$ is
-the number of individual variables or columns in group $g$. When a group
-encodes a categorical variable as binary indicators for the levels of
-that variable, we use the full encoding with a column for every level.
-Although this introduces collinearity, our method does not require the
-design matrix to have full rank.
+the number of individual variables or columns in group $g$.
+% Do we need to assume $p_g \ll n$ or $\text{rank}(X_g) = p_g$?
+When a group encodes a categorical variable as binary indicators for the
+levels of that variable, we use the full encoding with a column for every
+level. Although this introduces collinearity, our method does not require
+the design matrix to have full rank.
 
@@
 One step of the calculation will be to find an orthonormal basis for the
-linear space $V = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
+linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
 \text{ for all } h \neq g \}$ so that we can project orthogonally to
-this space. If $X_g$ is a single column and $X_g^Ty \neq 0$, then the
+this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
+should be the case since $g$ maximizes the absolute value of this
+quantity), then the
 space is trivial and the desired orthogonal projection is the identity.
-Otherwise, if $X_g$ has $p_g > 1$ columns,
+Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$ will
+generally have dimension $p_g-1$.
+% have dimension $r-1$ for $r = \text{rank}(X_g)$. This holds except in
+% degenerate cases such as $X_g = 0$ or $r \ll p_g$, but these cases are not
+% likely to occur since $\norm{X_g^T y}_2/w_g$ is supposed to be maximized.
+% We proceed by forming the $n \times p_g$ matrix
+% $X_g - X_g X_g^T y y^T X_g/\norm{X_g^Ty}_2^2$ and performing Gram-Schmidt
+% on the first $r-1$ columns. For precise theorems showing that this will
+% generally work, see \cite{outerp:indep}.
+We compute an orthonormal basis by Gram-Schmidt and form a
+$p \times p_g$ matrix, which we denote $V_g$, by appending 0's in the
+coordinates corresponding to all groups $h \neq g$ and to one column
+in group $g$ (since Gram-Schmidt only produces $p_g-1$ columns).
+With this we are ready to define the projection
+
+\begin{equation}
+  \begin{aligned}
+    \label{eq:proj}
+    P_g &= \Sigma XV_g (V_g^T X^T \Sigma X V_g)^\dagger V_g^TX^T \\
+   \end{aligned}
+\end{equation}
+and the conditional variance
+
+\begin{equation}
+  \begin{aligned}
+    \label{eq:cvar}
+    \sigma^2_g &= y^TX_gX_g^T(I-P_g)\Sigma X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2) \\
+  \end{aligned}
+\end{equation}
+These simplify when $X_g$ is a single column, in which case $P_g = 0$
+and $\sigma^2_g = X_g^T\Sigma X_g^T$.
 
@@
     \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$, $r \gets \text{rank}(X_g)$
-    \IF{$X_g$ is a single column}
-    \STATE $P \gets 0$ is a $p \times p$ matrix
-    \STATE $\sigma^2_g \gets X_g^T \Sigma X_g$
-    \ELSE
-    \STATE Do that
-    \ENDIF
+%    \IF{$X_g$ is a single column}
+%    \STATE $P_g \gets 0$ is an $n \times n$ matrix
+%    \STATE $\sigma^2_g \gets X_g^T \Sigma X_g$
+%    \ELSE
+    \STATE Compute $P_g$, $\sigma^2_g$ from equations \ref{eq:proj} and
+      \ref{eq:cvar}
+%    \ENDIF
     \RETURN $A$
@@
 \subsection{Real data example}
-\todo Code simulation with HIVdb data \\
-\todo Write section \\
+\todo Sanity check my analysis \\
+\todo For rare mutations try binary wildtype/non-wt? \\
+\todo Finish writing section \\
+
+\cite{HIV} use genomic information to predict efficacy of antiretroviral
+drugs in treating HIV. Quantitative measurements of drug
+response/resistance were regressed on categorical covariates encoded
+the presence and type of certain genetic mutations. We attempt a similar
+analysis using forward stepwise, and report our P-value at each step.
+Categorical covariates are encoding as groups of binary variables using
+the full encoding, and weights are set to the square-root of the
+number of levels for that covariate. We perform forward stepwise once
+for each drug response, restricting to the subset of the data with
+no missing values and requiring all categorical variables to have 2 or
+greater observations in each level.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.9\textwidth]{../figs/HIV_PI.pdf}
+\caption{\small \it Results from PI dataset}
+\label{fig:HIVPI}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.9\textwidth]{../figs/HIV_NRTI.pdf}
+\caption{\small \it Results from NRTI dataset}
+\label{fig:HIVNRTI}
+\end{center}
+\end{figure}
+
+Results from two data sets are displayed in Figure~\ref{fig:HIVPI} and
+Figure~\ref{fig:HIVNRTI}. Each panel shows results for a different drug,
+with p-values plotted by step in forward stepwise. The solid black point
+is the last p-value to fall under the 0.1 threshold and points after it
+are colored red. If the \textit{last} stopping rule were used then the
+p-values for including variables are the black points, and the resulting
+models are shown in Tables 4 and 5. (\todo Fix this ref after re-running)
+
+%\begin{figure}
+%\begin{center}
+\input{../tables/HIV_PI.tex}
+%\label{tab:HIVPI}
+%\end{center}
+%\end{figure}
+
+%\begin{figure}
+%\begin{center}
+\input{../tables/HIV_NRTI.tex}
+%\label{tab:HIVNRTI}
+%\end{center}
+%\end{figure}
 

### 2014-03-17 fwd_step_sim.R

@@ -180,7 +180,8 @@ print(c(upper, lower, length(intersect(results.b$active.set[1:num.nonzero], true
 
     
     xax <- 1:max.steps - 0.15
-    nxax <- xax + 0.3
+    nxax <- xax + 0.2
+    cxax = xax + 0.4
     plot.main <- paste0("n: ", n, ", g: ", g, ", Signal: ", lower.coeff, "/", upper.coeff,  ", k-Oracle power: ", round(fwd.power, 3))
 
     plot(xax, TrueStep, type = "l", main = plot.main, xlab = "Step", ylab = "", ylim = c(-.05,1.05), xlim = c(min(xax) - .2, max(nxax) + .2), lwd=2)
@@ -198,11 +199,11 @@ print(c(upper, lower, length(intersect(results.b$active.set[1:num.nonzero], true
     points(xax, Pvals.point[1, ], col = "blue", pch = 24, cex = .5)
     points(xax, Pvals.point[2, ], col = "blue", pch = 25, cex = .5)
 
-    points(xax, Chivals, col="green")
-    arrows(xax, Chivals.bar[1, ], xax, Chivals.bar[2, ],
+    points(cxax, Chivals, col="green")
+    arrows(cxax, Chivals.bar[1, ], cxax, Chivals.bar[2, ],
            code = 3, angle = 90, length = 0, col = "green")
-    points(xax, Chivals.point[1, ], col = "green", pch = 24, cex = .5)
-    points(xax, Chivals.point[2, ], col = "green", pch = 25, cex = .5)
+    points(cxax, Chivals.point[1, ], col = "green", pch = 24, cex = .5)
+    points(cxax, Chivals.point[2, ], col = "green", pch = 25, cex = .5)
 
   }
 

### 2014-03-17 glint/fwd_glint_sim.R

@@ -5,18 +5,6 @@ source('../generate_data.R')
 #source('pred_and_estim.R') not for glinternet
 #source('fwd_step/coherence.R')
 
-## n=250
-## sigma=1
-## groups=1:20
-## num.nonzero=6
-## max.steps=12
-## lower=7
-## upper=9
-## nsim=5
-## design="gaussian"
-## corr=0
-## k=6
-
 # Note: glint version passes k here instead of beta
 fwd_glint_simulation = function(n, Sigma, groups, num.nonzero, lower, upper, nsim,
   max.steps, alpha = .1, design = 'gaussian', corr = 0, categorical = FALSE, predictions = FALSE, coherence = FALSE, plot = FALSE) {
@@ -121,13 +109,17 @@ fwd_glint_simulation = function(n, Sigma, groups, num.nonzero, lower, upper, nsi
       Y = unwhitener %*% rnorm(n)
       Y.t = unwhitener %*% rnorm(n)
     }
+
     Y.noiseless = X %*% beta
+    mu.max = max(abs(Y.noiseless))
+    mu.lc = sum(abs(Y.noiseless)>upper)
     Y.noiseless.test = X.test %*% beta
     Y.beta = Y.noiseless + Y
     Y.test = Y.noiseless.test + Y.t
+    X = col_normalize(X)
+    X.test = col_normalize(X.test)
     X = frob_normalize(X, all.groups)
-    X.test = frob_normalize(X.test, all.groups)
-    
+    X.test = frob_normalize(X.test, all.groups)    
     main1 = c(main1, abs(t(X[,1]) %*% Y.beta))
     z=t(X[,all.groups==61])%*%Y.beta
     int37 = c(int37, sqrt(sum(z^2)))
@@ -139,7 +131,7 @@ fwd_glint_simulation = function(n, Sigma, groups, num.nonzero, lower, upper, nsi
 
     # Non-null results
     results.b = forward_group(X, Y.beta, groups=all.groups, weights, Sigma, max.steps = max.steps)
-print(c(upper, lower, length(intersect(results.b$active.set[1:num.nonzero], true.active.groups))/num.nonzero))
+print(c(upper, lower, mu.max, mu.lc, length(intersect(results.b$active.set[1:num.nonzero], true.active.groups))/num.nonzero))
     
     Chi.mat.b[i, ] = results.b$chi.pvals
     P.mat.b[i, ] = results.b$p.vals

### 2014-03-17 glint/fwd_step.R

@@ -97,21 +97,23 @@ forward_group = function(X, Y, groups, weights = 0, Sigma = NULL, max.steps = 0)
   c.vars = c()
   Ls = c()
 
-  Y.update = Y
+  Y.update = Y - mean(Y)
   X.update = X
 
   for (i in 1:max.steps) {
     output = add_group(X.update, Y.update, groups, weights, Sigma, active.set, eff.p)
     active.set = output$active.set
-    imax = output$imax
+    imax = output$added
     grank = sum(groups == imax)
     eff.p = output$eff.p
     RSS = sum(Y.update^2)
     Y.update = output$Y.update
-    RSS = RSS - sum(Y.update^2)
+    RSSdrop = RSS - sum(Y.update^2)
+    chi.p = pchisq(RSSdrop, lower.tail=F, df=grank)
+    print(c(RSSdrop, round(grank), chi.p))
     X.update = output$X.update
     p.vals = c(p.vals, output$p.value)
-    chi.pvals = c(chi.pvals, pchisq(RSS, lower.tail=F, df=grank))
+    chi.pvals = c(chi.pvals, chi.p)
     c.vars = c(c.vars, output$var)
     # tracking lambda_2
     Ls = c(Ls, output$test.output[1])

### 2014-03-17 glint/sim_glint.R

@@ -6,16 +6,16 @@ source('fwd_glint_sim.R')
 
 
 design = 'gaussian'
-corr = 0 # nonzero only supported for gaussian design
-noisecorr = 0
+corr = .1 # nonzero only supported for gaussian design
+noisecorr = .1
 
-nsim = 100
-n = 250
+nsim = 200
+n = 400
 num.nonzero = 9
 k = num.nonzero
-max.steps = 12
-upper.coeff = 5
-lower.coeff = 3
+max.steps = 15
+upper.coeff = 2.5
+lower.coeff = 1.5
 Sigma = (1-noisecorr)*diag(rep(1,n)) + noisecorr
 groups = 1:20
 p = length(groups)
@@ -79,7 +79,7 @@ dev.off()
 
 num.nonzero = 6
 k = 6
-max.steps = 12
+max.steps = 10
 groups = sort(c(rep(1:15, 2), rep(16:20, 3)))
 p = length(groups)
 g = length(unique(groups))

### 2014-03-17 sim_small.R

@@ -7,17 +7,17 @@ source('tex_table.R')
 
 design = 'gaussian'
 corr = 0 # nonzero only supported for gaussian design
-noisecorr = .1
+noisecorr = 0
 
 nsim = 200
 n = 100
-num.nonzero = 5
+num.nonzero = 2
 k = num.nonzero
-max.steps = 10
-upper.coeff = 1.2
-lower.coeff = 1.1
+max.steps = 8
+upper.coeff = 0.8
+lower.coeff = 0.5
 Sigma = (1-noisecorr)*diag(rep(1,n)) + noisecorr
-groups = 1:50
+groups = 1:10
 p = length(groups)
 mult = sqrt(2*log(p))
 upper = upper.coeff*mult
@@ -42,9 +42,9 @@ output.l <- fwd_group_simulation(n, Sigma, groups, beta, nsim, max.steps, design
 dev.off()
 
 
-groups = sort(c(rep(1:8, 5), 9:18))
-num.nonzero = 3
-max.steps = 9
+groups = sort(c(rep(1:8, 5), 9:10))
+num.nonzero = 2
+max.steps = 8
 p = length(groups)
 g = length(unique(groups))
 mult = sqrt(2*log(p)) # or g?

### 2014-03-20 paper/paper.tex
Paper: simulation section exposition

@@
 \subsection{Exact p-value calculation}
+
+\todo Complete this section
 
@@
 
-\todo Clean up and reorganize \\
-\todo Write a little more exposition \\
 \todo Compare to AIC and other classical stopping rules? \\
+
+We performed simulations with several classes of design matrices including
+Gaussian, categorical, and categorical from a real data set. Gaussian
+design matrices are generated either independently or with some global
+correlation $\rho > 0$ between all pairs of columns. Categorical matrices
+were generated by first choosing a vector of probabilities for a given
+variable from a Dirichlet prior, and then sampling category levels with
+that vector of probabilities. Resampling was used to ensure the minimum
+number of observations in any level was at least five. Finally, categorical
+varables were encoded as groups of 0-1 vectors using the full encoding.
+The categorical data matrices from real data are described in
+Section~\ref{sec:hiv}
+
+Since we are interested in variable selection we generate signals with
+nonzero coefficients on the scale of $\gamma := \sqrt{2 \log(p)/n}$.
+(\todo Cite some paper on support recovery?). Coefficients within a
+nonzero group have roughly the same magnitude, and the magnitudes for
+each nonzero group range from a lower limit times $\gamma$ to an upper
+limit times $\gamma$. In plots, the limits are listed next to ``beta:''
+and the numbers in parentheses are the largest and smallest 2-norms of
+nonzero coefficient groups respectively. Each plot also displays the
+number of observations or rows, $n$, the number of columns, $p$, the
+number of groups, $g$, and the largest and smallest group sizes in
+parentheses if the groups are not all size 1. The number of nonzero
+groups, $k$, is also displayed via inward-facing tick marks on the
+horizontal axes. The proportion of truly
+nonzero groups recovered in the first $k$ steps is denoted $k$-SP,
+where recall $k$ is the number of truly nonzero groups.
+
+Finally, most of the plots below show the following information. The
+horizontal axis is the step index for forward stepwise. The dashed
+line shows the proportion of iterations where a truly nonzero variable
+was added to the model at that step. Solid verticle lines show the
+middle 50\% of p-values calculated at that step, with red calculated
+on a null model with no nonzero groups, blue calculated on the non-null
+model, and green calculated on the non-null model from the usual
+$\chi^2$ significance test.
+The points in these lines are the average p-value at that step, and
+the triangles above and below show the 95\% and 5\% quantiles.
 
@@
 \label{fig:smallgauss}
-\includegraphics[width=0.5\textwidth]{../figs/gaussian_size1_n100_p200_g200_k10_lower1p8_upper1p9.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/gaussian_n100_p200_g200_k5_lower1pt5_upper2_corr0pt2_noisecorr0pt1.pdf}}
 \hspace{-15pt}
@@
 \label{fig:smallcat}
-\includegraphics[width=0.5\textwidth]{../figs/gaussian_size1_n100_p200_g200_k10_lower1p8_upper1p9_corr0p1.pdf}}
-\caption{\small \it The left panel shows results of a simulation with
-  an independent Gaussian design matrix and a signal vector having
-  three groups with non-zero coefficients of sizes one, two, and
-  three, and seven groups of noise variables with sizes varying from
-  one to ten, for a total of 30 variables. For the right panel we
-  repeated this setup, but group sizes corresponded to levels of
-  categorical variables. The groups of size one were increased to two,
-so there are 10 categorical variables with a total of 32 levels.}
+\includegraphics[width=0.5\textwidth]{../figs/categorical_n100_p100_g20_k4_lower4_upper5.pdf}}
+\caption{\small \it The left panel shows results for Gaussian design
+matrices with equicorrelation $\rho=0.2$ and equicorrelated noise with
+correlation 0.1. The right panel shows results for categorical matrices
+with all categories having 5 levels.}
 \end{center}
@@
 \subsection{Real data example}
+\label{sec:hiv}
 \todo Sanity check my analysis \\

### 2014-03-20 sim_cat.R

@@ -12,14 +12,14 @@ noisecorr = 0
 
 nsim = 400
 n = 100
-num.nonzero = 5
+num.nonzero = 4
 k = num.nonzero
 max.steps = 10
-upper.coeff = 3
-lower.coeff = 2
+upper.coeff = 5
+lower.coeff = 4
 Sigma = (1-noisecorr)*diag(rep(1,n)) + noisecorr
-#groups = sort(rep(1:20, 5))
-groups = sort(rep(1:20, 2))
+groups = sort(rep(1:20, 5))
+#groups = sort(rep(1:20, 2))
 p = length(groups)
 g = length(unique(groups))
 mult = sqrt(2*log(p)/n)
@@ -41,8 +41,8 @@ print(warnings())
 
 with(output.l, step_plot(TrueStep, null.p, signal.p, chi.p, num.nonzero, n, p, g, ugsizes, max.steps, upper.coeff, lower.coeff, max.beta, min.beta, fwd.power, design, fn.append))
 
-#groups = sort(c(rep(1:100, 4), rep(101:200, 2)))
-groups = sort(rep(1:30, 5))
+groups = sort(c(rep(1:100, 4), rep(101:200, 2)))
+#groups = sort(rep(1:30, 5))
 p = length(groups)
 g = length(unique(groups))
 mult = sqrt(2*log(p)/n)

### 2014-03-20 sim_small.R

@@ -10,15 +10,15 @@ fn.append = ''
 corr = 0 # nonzero only supported for gaussian design
 noisecorr = 0
 
-nsim = 100
+nsim = 400
 n = 100
-num.nonzero = 2
+num.nonzero = 4
 k = num.nonzero
-max.steps = 8
-upper.coeff = 0.8
-lower.coeff = 0.5
+max.steps = 10
+upper.coeff = 1
+lower.coeff = 0.7
 Sigma = (1-noisecorr)*diag(rep(1,n)) + noisecorr
-groups = 1:20
+groups = 1:50
 p = length(groups)
 g = length(unique(groups))
 mult = sqrt(2*log(p))
@@ -46,11 +46,11 @@ with(output.l, step_plot(TrueStep, null.p, signal.p, chi.p, num.nonzero, n, p, g
 
 
 groups = sort(c(rep(1:8, 5), 9:10))
-num.nonzero = 2
-max.steps = 8
+num.nonzero = 4
+max.steps = 10
 p = length(groups)
 g = length(unique(groups))
-mult = sqrt(2*log(g)) # or g?
+mult = sqrt(2*log(p)) # or g?
 upper = upper.coeff*mult
 lower = lower.coeff*mult
 

### 2014-03-21 paper/paper.tex
Paper: finished p-value calculation! Code: fwd_step uses Pg, group_lasso light edits, rerun some simulations.

@@
 \renewcommand{\algorithmicensure}{\textbf{Output:}}
+\renewcommand{\algorithmiccomment}[1]{\# #1}
 
@@
 power'' was calculated as the number of true variables in the active
-set after $k$ steps divided by $k$. Nonzero coefficients had magnitude
+set after $k$ steps divided by $k$. (\todo Make this consistent with
+$\gamma$ in simulations section)
+Nonzero coefficients had magnitude
 in a range of multiples of $\gamma := \sqrt{2\log(G)}$, e.g. in $[1.4
@@
 We compute an orthonormal basis by Gram-Schmidt and form a
-$p \times p_g$ matrix, which we denote $V_g$, by appending 0's in the
-coordinates corresponding to all groups $h \neq g$ and to one column
-in group $g$ (since Gram-Schmidt only produces $p_g-1$ columns).
+$p_g \times p_g$ matrix, which we denote $V_g$, by appending one column
+%0's in the coordinates corresponding to all groups $h \neq g$ and to
+of zeroes (since Gram-Schmidt only produces $p_g-1$ columns).
 With this we are ready to define the projection
@@
     \label{eq:proj}
-    P_g &= \Sigma XV_g (V_g^T X^T \Sigma X V_g)^\dagger V_g^TX^T \\
+    P_g &= \Sigma X_gV_g (V_g^T X_g^T \Sigma X_g V_g)^\dagger V_g^TX_g^T \\
    \end{aligned}
 \end{equation}
-and the conditional variance
+Also define $H_g = (I-P_g)\Sigma$ and the conditional variance
 
@@
     \label{eq:cvar}
-    \sigma^2_g &= y^TX_gX_g^T(I-P_g)\Sigma X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2) \\
+    \sigma^2_g &= y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2) \\
   \end{aligned}
 \end{equation}
-These simplify when $X_g$ is a single column, in which case $P_g = 0$
-and $\sigma^2_g = X_g^T\Sigma X_g^T$.
-
-Define $U_g = X_g^Ty/(w_g\norm{X_g^Ty}_2)$.
+These simplify when $X_g$ is a single column, in which case $P_g = 0$,
+$H_g = \Sigma$, and $\sigma^2_g = X_g^T\Sigma X_g^T/w_g^2$. We now
+describe how to compute our p-value in Algorithm~\ref{algo:pval}. Note
+that $F_{\chi^2_r}$ denotes the distribution function of a $\chi^2$
+random variable with $r$ degrees of freedom.
+%Define $U_g = X_g^Ty/(w_g\norm{X_g^Ty}_2)$.
+
 \begin{algorithm}
@@
   \begin{algorithmic}[1]
-    \REQUIRE Response $y$, design matrix $X$, index $g$ of next group to enter
-    \ENSURE P-value for testing if $g$ and all other remaining groups are noise
+    \REQUIRE Response $y$, grouped design matrix $X$ with weights, inactive set $A^c$, index $g$ of last group to enter active set $A$.
+    \ENSURE Our p-value for the group $g$ entering the model.
+    \STATE Compute $H_g$ and $\sigma^2_g$
+    \IF{$X_g$ is a single column}
+    \STATE $\sigma_g^2 \gets X_g^T\Sigma X_g/w_g^2$
+    \STATE $\tilde X_g \gets X_g / w_g \cdot \text{sign}(X_g^Ty)$
+    \ELSE
+    \STATE $P_g \gets \Sigma XV_g (V_g^T X^T \Sigma X V_g)^\dagger V_g^TX^T$
+    \STATE $H_g \gets (I - P_g)\Sigma$
+    \STATE $\sigma^2_g \gets y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2)$
+    \STATE $\tilde X_g \gets H_g X_g X_g^T y /(\norm{X_g^Ty}_2 w_g)$
+    \ENDIF
     \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$, $r \gets \text{rank}(X_g)$
-%    \IF{$X_g$ is a single column}
-%    \STATE $P_g \gets 0$ is an $n \times n$ matrix
-%    \STATE $\sigma^2_g \gets X_g^T \Sigma X_g$
-%    \ELSE
-    \STATE Compute $P_g$, $\sigma^2_g$ from equations \ref{eq:proj} and
-      \ref{eq:cvar}
-%    \ENDIF
-    \RETURN $A$
+%    \STATE \COMMENT{Compute the following two $p$-vectors}
+    \STATE $a \gets X^T(y - \lambda \tilde X_g)$
+    \STATE $b \gets X^T \tilde X_g$
+    \STATE Compute solution $(v_-, v_+)$ of \textsc{LinFrac}(a,b) optimization subproblem
+    \STATE $u \gets [F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(\lambda^2/\sigma_g^2)]/[F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(v^2_+/\sigma_g^2)]$
+    \RETURN $u$
+  \end{algorithmic}
+\end{algorithm}
+
+We next describe the \textsc{LinFrac} subproblem and its solution. We
+gave the problem this name because it originated in the form
+\[
+\maximize_{h \neq g, \norm{u_h}_2 = 1} \frac{u_h^TX_h^Ty - u_h^TX_h^TX_gX_g^Ty}
+{1 - u_g^TX_g^TX_hu_h}
+\]
+The solution we describe next is to a slightly different problem, and
+also incorporates the information that $g$ maximizes $\norm{X_g^Ty}_2/w_g$.
+For details see Appendix~\ref{app:closedform} or \cite{tests:adaptive}.
+
+\begin{algorithm}
+  \caption{The \textsc{LinFrac} optimization subproblem}
+  \label{algo:linfrac}
+  \begin{algorithmic}[1]
+    \REQUIRE The $p$-vectors $a, b$ in Algorithm~\ref{algo:pval}, weights, inactive set $A^c$, a small tolerance number
+%(we take $1e^{-10}$)
+    \ENSURE Solution pair $(v_-, v_+)$.
+    \FOR{$h$ in $A^c$}
+      \IF{$\norm{b_h}_2 == 0$ or $\norm{a_h}_2/\norm{b_h}_2 < \text{tol}$}
+        \STATE $(v^-_h, v^+_h) \gets (0, \infty)$
+      \ELSE
+        \STATE $\theta_c \gets a_h^Tb_h/(\norm{a_h}_2 \norm{b_h}_2)$
+        \STATE $\theta_s \gets \sqrt{1-\theta_c^2}$
+        \STATE $\theta \gets \arccos (\theta_c)$
+        \STATE $\phi_s \gets \theta_s \norm{b_h}_2 / w_h$
+        \IF{$\phi_s > 1$}
+          \STATE $(v^-_h, v^+_h) \gets (0, \infty)$
+        \ELSE
+          \STATE $\phi \gets \arcsin(\phi_s)$
+%          \STATE $\phi_2 \gets \pi - \phi_1$
+          \STATE $z_\pm \gets s_\pm \norm{a_h}_2 \cos(\phi) /(w_h - s_\pm \norm{b_h}_2 \cos(\theta - \phi))$ for $s_\pm = \pm 1$
+          \IF{$\norm{b_h}_2 < w_h$}
+            \STATE $(v_h^-, v_h^+) \gets (\max \{ z_+, z_- \}, \infty)$
+          \ELSE
+            \STATE $(v_h^-, v_h^+) \gets (\min \{ z_+, z_- \}, \max \{ z_+, z_- \})$
+          \ENDIF
+        \ENDIF
+      \ENDIF
+    \ENDFOR
+    \STATE $v^- \gets \min_h v_h^-$
+    \STATE $v^+ \gets \max_h v_h^+$
+    \RETURN $(v^-, v^+)$
   \end{algorithmic}
@@
 nonzero coefficients on the scale of $\gamma := \sqrt{2 \log(p)/n}$.
-(\todo Cite some paper on support recovery?). Coefficients within a
+(\todo Check this scaling. \todo Cite some paper on support recovery?).
+Coefficients within a
 nonzero group have roughly the same magnitude, and the magnitudes for
@@
 \begin{center}
-\subfigure[Gaussian design]{
-\label{fig:smallgauss}
+\subfigure[Gaussian with correlations]{
+\label{fig:gausscorr}
 \includegraphics[width=0.5\textwidth]{../figs/gaussian_n100_p200_g200_k5_lower1pt5_upper2_corr0pt2_noisecorr0pt1.pdf}}
 \hspace{-15pt}
-\subfigure[Categorical design]{
-\label{fig:smallcat}
-\includegraphics[width=0.5\textwidth]{../figs/categorical_n100_p100_g20_k4_lower4_upper5.pdf}}
+\subfigure[Gaussian with groups]{
+\label{fig:gaussgroups}
+\includegraphics[width=0.5\textwidth]{../figs/gaussian_n100_p200_g35_k5_lower2_upper2pt5.pdf}}
 \caption{\small \it The left panel shows results for Gaussian design
 matrices with equicorrelation $\rho=0.2$ and equicorrelated noise with
-correlation 0.1. The right panel shows results for categorical matrices
-with all categories having 5 levels.}
+correlation 0.1. The right panel shows results for an independent
+Gaussian design with groups of sizes 5 and 10.}
 \end{center}
@@
 \begin{center}
-\subfigure[Large Gaussian lasso]{
-\label{fig:largelasso}
-\includegraphics[width=0.5\textwidth]{../figs/gaussian_size5-10_n100_p200_g35_k10_lower1p8_upper1p9.pdf}}
+\subfigure[Real data categorical design]{
+\label{fig:hivnrti}
+\includegraphics[width=0.5\textwidth]{../figs/HIV_NRTI_n848_p48_g14_k4_lower21_upper25.pdf}}
 \hspace{-15pt}
-\subfigure[Large Gaussian group lasso]{
-\label{fig:largegrplasso}
-\includegraphics[width=0.5\textwidth]{../figs/gaussian_size5-10_n100_p200_g35_k10_lower1p8_upper1p9_corr0p1.pdf}}
-\caption{\small \it The left panel shows results of a simulation with
-  an independent Gaussian design matrix and a signal vector having
-  three groups with non-zero coefficients of sizes one, two, and
-  three, and seven groups of noise variables with sizes varying from
-  one to ten, for a total of 30 variables. For the right panel we
-  repeated this setup, but group sizes corresponded to levels of
-  categorical variables. The groups of size one were increased to two,
-so there are 10 categorical variables with a total of 32 levels.}
+\subfigure[Simulated categorical design]{
+\label{fig:gausscat}
+\includegraphics[width=0.5\textwidth]{../figs/categorical_n100_p100_g20_k4_lower4_upper5.pdf}}
+\caption{\small \it The left panel shows results for a categorical design
+taken from a real data set.
+The right panel shows results for categorical matrices with all categories having 5 levels.}
 \end{center}
@@
 \input{../large_sim_selection.tex}
-
-The large simulation has 50 groups, 25 of size 1, 10 of size 5, 10 of
-size 10, and 5 of size 15. Signal vectors were supported on random
-choices of 10 of these groups (changing in each realization), with
-non-zero signal magnitude around $\sqrt{2\log p}$ where $p = 250$
-(roughly 3.3). The design matrix had 200 rows and the simulation
-performed 100 realizations.
 
@@
 \section{Derivation of closed forms of statistic}
+\label{app:closedform}
 

### 2014-03-21 HIV/sim_HIV.R

@@ -4,26 +4,27 @@ source('generate_data.R')
 source('fwd_step_sim.R')
 source('tex_table.R')
 source('plots.R')
+library(plyr)
 
-PI = read.table("http://hivdb.stanford.edu/pages/published_analysis/genophenoPNAS2006/DATA/NRTI_DATA.txt", sep='\t', header=TRUE)
-db="PI"
-data=PI
-resps = 4:9
-## NRTI = read.table("http://hivdb.stanford.edu/pages/published_analysis/genophenoPNAS2006/DATA/PI_DATA.txt", sep = "\t", header = TRUE)
-## data=NRTI
-## db="NRTI"
-## resps = 4:10
+## PI = read.table("http://hivdb.stanford.edu/pages/published_analysis/genophenoPNAS2006/DATA/NRTI_DATA.txt", sep='\t', header=TRUE)
+## db="PI"
+## data=PI
+## resps = 4:9
+NRTI = read.table("http://hivdb.stanford.edu/pages/published_analysis/genophenoPNAS2006/DATA/PI_DATA.txt", sep = "\t", header = TRUE)
+data=NRTI
+db="NRTI"
+resps = 4:10
 
 design = paste0("HIV_", db)
-fn.append = ''
+fn.append = 'Pg'
 corr = 0 # nonzero only supported for gaussian design
 noisecorr = 0
-nsim = 100
+nsim = 200
 num.nonzero = 4
 k = num.nonzero
 max.steps = 10
-upper.coeff = 10
-lower.coeff = 7
+upper.coeff = 25
+lower.coeff = 21
 
 # Clean data, restrict to cleaned subset
 nresps = length(resps)
@@ -39,6 +40,7 @@ cnames = c()
 for (j in max(resps):ncol(data)) {
       newcol = clean_col(data[,j])
           if ((length(levels(newcol)) > 1) & (min(table(newcol)) >= 2)) {
+            newcol = mapvalues(newcol, from=levels(newcol), to=1:length(levels(newcol)))
                     cnames = c(cnames, names(data)[j])
                             Z = cbind(Z, newcol)
                   }
@@ -83,7 +85,7 @@ if (noisecorr != 0) {
   fn.append = paste0(fn.append, '_noisecorr', noisecorr)
 }
 
[…]

### 2014-03-21 fwd_step.R

@@ -42,30 +42,28 @@ add_group = function(X, Y, groups, weights, Sigma, active.set = 0, eff.p = 0, ca
   # Also using group_lasso.R
   p.value = pvalue(results$L, results$Mplus, results$Mminus, sqrt(results$var), kmax)
   new.active.set = update_active_set(active.set, imax)
+  inactive.set = setdiff(unique(groups), new.active.set)
 
   # Form new residual
   X.project = X
   Xgmax = X[, gmax]
-  Y.resid = lm(Y ~ Xgmax - 1)$residual
 
+##   Y.resid = lm(Y ~ Xgmax - 1)$residual
 ##   # If X[, gmax] is categorical, leave one column out
 ##   # Not necessary, Y already de-meaned
-  if ((is.element(imax, cat.groups)) & (is.matrix(Xgmax))) {
-    Xgmax = Xgmax[ , -1]
-  }
-##   Pgmax = Xgmax %*% ginv(Xgmax)
+##   if ((is.element(imax, cat.groups)) & (is.matrix(Xgmax))) {
+##     Xgmax = Xgmax[ , -1]
+##   }
+  Pgmax = Xgmax %*% ginv(Xgmax)
       
   # Project all other groups orthogonal to the one being added
-  for (gind in 1:max(groups)) {
-    if (gind != imax) {
-      group = groups == gind
-      #X.project[, group] = X[, group] - Pgmax %*% X[, group]
-      X.project[, group] = lsfit(Xgmax, X[, group], intercept=FALSE)$residuals
-    }
+  for (gind in inactive.set) {
+    group = groups == gind
+    X.project[, group] = X[, group] - Pgmax %*% X[, group]
+    #X.project[, group] = lsfit(Xgmax, X[, group], intercept=FALSE)$residuals
   }
 
-
-  #Y.resid = Y - Pgmax %*% Y
+  Y.resid = Y - Pgmax %*% Y
   
   return(list(test.output = results, var = results$var, p.value = p.value, added = imax, active.set = new.active.set, eff.p = new.eff.p, Y.update = Y.resid, X.update = X.project, grank=kmax))
 }

### 2014-03-21 glint/sim_glint.R

@@ -17,8 +17,8 @@ num.nonzero = 3
 k = num.nonzero
 num.main = 1
 max.steps = 10
-upper.coeff = 5
-lower.coeff = 3
+upper.coeff = 20
+lower.coeff = 18
 Sigma = (1-noisecorr)*diag(rep(1,n)) + noisecorr
 groups = 1:20
 p = length(groups)

### 2014-03-21 sim_small.R

@@ -10,13 +10,13 @@ fn.append = ''
 corr = 0 # nonzero only supported for gaussian design
 noisecorr = 0
 
-nsim = 400
+nsim = 100
 n = 100
 num.nonzero = 4
 k = num.nonzero
 max.steps = 10
-upper.coeff = 1
-lower.coeff = 0.7
+upper.coeff = 0.8
+lower.coeff = 0.2
 Sigma = (1-noisecorr)*diag(rep(1,n)) + noisecorr
 groups = 1:50
 p = length(groups)

### 2014-03-21 paper/paper.tex
Paper: some small edits

@@
 that begins with an empty model and sequentially adds the best predictor
-variable in each step. Classical significance tests fail
-when a model
-has been selected this way and tend to be anti-conservative.
-In the lasso setting, \cite{significance:lasso} derived a novel test
-statistic along with a correct asymptotic distribution, making possible
-valid inferences after model selection.
+variable in each step. The usual $\chi^2$ and $F$-tests for significance
+fail when a model has been selected this way and tend to be
+anti-conservative unless they are computed on a held-out validation
+dataset. In the lasso setting, \cite{significance:lasso} derived
+a novel test statistic and its asymptotic null distribution, making
+possible valid inferences after model selection using the full data.
 \cite{tests:adaptive} modified and extended those results to the
@@
 
-Some classical attempts to address this issue include Monte Carlo
+Some attempts to address this issue include Monte Carlo
 estimation of tables of adjusted $F$-statistic values \citep{mc:ftoenter},
@@
 measurements for all genes in a single functional pathway.
-To emphasize this we will use $g, h$ as indices rather than the usual
+To emphasize this we use $g, h$ as indices rather than the usual
 $i, j$ throughout. Since single variables can be considered groups of
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-where $\epsilon$ is noise. Unless otherwise specified we assume
-i.i.d. Gaussian noise $\epsilon \sim N(0, I_{n \times n})$ and that
-$\sigma$ is known.
-
-We allow for the possibility $p > n$ in which case \eqref{eq:gmodel} is
-generically underdetermined. In such cases it still may be possible to
+where $\epsilon$ is noise. We assume Gaussian noise
+$\epsilon \sim N(0, \Sigma)$ with known covariance matrix $\Sigma$.
+
+The model \eqref{eq:gmodel} is underdetermined when $p > n$.
+In such cases it still may be possible to
 estimate $\beta$ well if it is sparse--that is, if it has few nonzero
@@
 more difficult for the group $X_g$ to enter the model. The
-modeler can choose weights arbitrarily, but we will mostly use one
+modeler can choose weights arbitrarily, but we mostly use one
 particular choice, based on $p_g$, that we discuss later. With this we
@@
     \STATE $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$
-    \STATE $P \gets I - X_{g^*}X_{g^*}^\dagger$
+    \STATE $P_g \gets I_{n\times n} - X_{g^*}X_{g^*}^\dagger$
     \STATE $A \gets A \cup \{ g^* \}$, $A^c \gets A^c \backslash \{ g^* \}$
     \FORALL{$g \in A^c$}
-      \STATE $X_g \gets P X_g$
+      \STATE $X_g \gets P_g X_g$
     \ENDFOR
-    \STATE $r_s \gets P r_{s-1}$
+    \STATE $r_s \gets P_g r_{s-1}$
     \ENDFOR
@@
 
-In the common case where all groups are single variables this version
-coincides with the standard implementations of forward stepwise.
-We will now use forward stepwise to refer specifically to
+The active set $A$ contains variable groups chosen to be included in
+the model. Fitting $\hat \beta$ can be done by tracking the individual
+fits and projections, or by simply fitting a linear model on the submatrix
+of $X$ corresponding to $A$. 
+We now use forward stepwise to refer specifically to
 Algorithm~\ref{algo:fs} unless otherwise specified. Note that other
@@
 complicated. For the group lasso, $\hat \beta(\lambda)$ does not have
-piecewise linear components. To overcome this difficulty we will
+piecewise linear components. To overcome this difficulty we
 restrict our attention to the first group of variables to enter the
@@
 space is trivial and the desired orthogonal projection is the identity.
-Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$ will
-generally have dimension $p_g-1$.
+Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
+generally has dimension $p_g-1$.
 % have dimension $r-1$ for $r = \text{rank}(X_g)$. This holds except in
@@
   \label{algo:pval}
-  \begin{algorithmic}[1]
+  \begin{algorithmic}
     \REQUIRE Response $y$, grouped design matrix $X$ with weights, inactive set $A^c$, index $g$ of last group to enter active set $A$.
@@
     \ENDIF
-    \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$, $r \gets \text{rank}(X_g)$
+    \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$
 %    \STATE \COMMENT{Compute the following two $p$-vectors}
@@
     \STATE $b \gets X^T \tilde X_g$
-    \STATE Compute solution $(v_-, v_+)$ of \textsc{LinFrac}(a,b) optimization subproblem
+    \STATE Compute solution $(v_-, v_+)$ of \textsc{LinearFractional}(a,b) optimization subproblem
+    \STATE $r \gets \text{rank}(X_g)$
     \STATE $u \gets [F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(\lambda^2/\sigma_g^2)]/[F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(v^2_+/\sigma_g^2)]$
@@
 
-We next describe the \textsc{LinFrac} subproblem and its solution. We
+We next describe the \textsc{LinearFractional} subproblem and its solution. We
 gave the problem this name because it originated in the form
@@
 \begin{algorithm}
-  \caption{The \textsc{LinFrac} optimization subproblem}
+  \caption{The \textsc{LinearFractional} optimization subproblem}
   \label{algo:linfrac}
-  \begin{algorithmic}[1]
+  \begin{algorithmic}
     \REQUIRE The $p$-vectors $a, b$ in Algorithm~\ref{algo:pval}, weights, inactive set $A^c$, a small tolerance number
@@
 model by group lasso, we use forward stepwise on the expanded design
-matrix. The overlapping groups still guarantee that our fitted model will
-satisfy the strong hierarchy condition.
+matrix. The overlapping groups still guarantee that our fitted model
+satisfies the strong hierarchy condition.
 
@@
 analysis using forward stepwise, and report our P-value at each step.
-Categorical covariates are encoding as groups of binary variables using
+Categorical covariates are encoded as groups of binary variables using
 the full encoding, and weights are set to the square-root of the
@@
 Results from two data sets are displayed in Figure~\ref{fig:HIVPI} and
-Figure~\ref{fig:HIVNRTI}. Each panel shows results for a different drug,
+Figure~\ref{fig:HIVNRTI}. The PI data contains protease inhibitor
+resistance mutations, and the NRTI data contains nucleoside RT inhibitor
+resistance mutations.
+Each panel shows results for a different drug,
 with p-values plotted by step in forward stepwise. The solid black point
@@
 Note that $K_h = \emptyset$ if and only if $\norm{b_h} < w$.
-We will consider two cases.
+We consider two cases.
 First, suppose $\norm{b_h} > w$.

### 2014-03-21 group_lasso.R

@@ -14,7 +14,7 @@ trignometric_form = function(num, den, weight, tol=1.e-10) {
   normb = sqrt(sum(b^2))
 
   if (normb == 0) {
-#    stop("Something is wrong, norm(b) can't be zero!")
+#    stop("Something is wrong, norm(b) shouldn't be zero!")
     return(c(0, Inf))
   }
   

### 2014-04-03 paper/paper.tex
rough cut and paste of quadratic materila

@@
 
+\newcommand{\jonathan}[1]{{\color{blue}  #1}}
+
 \begin{document}
@@
 used for valid significance testing when computed on the same
-data used for model selection, eliminating the need for data splitting.
+data used for model selection, eliminating the need for data splitting or for situations in which
+data splitting is not appropriate.
 The present work iteratively applies the global null test
@@
 forward stepwise algorithm used in most of the paper.
-Section~\ref{sec:testing} briefly reviews
-the parts of \cite{significance:lasso} and \cite{tests:adaptive}
-relevant to our significance test, and describes the group lasso which
-we require for applying the test with grouped variables.
+Section~\ref{sec:testing}  reviews some recent work on
+post-selection inference: \cite{significance:lasso,tests:adaptive,lasso:fixed}
+relevant to our work here, and we describes a general framework
+for post-selection inference based on quadratic comparisons.
 Simulation results in Section~\ref{sec:simulations} show
@@
 When a group encodes a categorical variable as binary indicators for the
-levels of that variable, we use the full encoding with a column for every
+levels of that variable, by default we use the full encoding with a column for every
 level. Although this introduces collinearity, our method does not require
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-We assume throughout that the individual columns of $X$ are scaled to
+For convenience, we assume throughout that the individual columns of $X$ are scaled to
 have unit 2-norm (we do not scale groups, this is accomplished later
@@
 \label{eq:gmodel}
-y & = X \beta + \sigma \epsilon \\
+y | X& = X \beta + \sigma \epsilon \\
    & = \sum_{g=1}^G X_g \beta_g + \sigma \epsilon
@@
 where $\epsilon$ is noise. We assume Gaussian noise
-$\epsilon \sim N(0, \Sigma)$ with known covariance matrix $\Sigma$.
+$\epsilon | X \sim N(0, \Sigma)$ with known covariance matrix $\Sigma$.
 
@@
 
+\jonathan{Is $r_0=0$ or $y$? Should be defined in the algorithm. I also put a $y$ in below which I'm not sure
+is correct or not. }
+
 \begin{algorithm}
@@
     \FOR{$s=1$ to $steps$}
-    \STATE $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$
-    \STATE $P_g \gets I_{n\times n} - X_{g^*}X_{g^*}^\dagger$
+    \STATE $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T(y- r_{s-1})}_2 / w_g \}$
+    \STATE $P_{g^*} \gets I_{n\times n} - X_{g^*}X_{g^*}^\dagger$
     \STATE $A \gets A \cup \{ g^* \}$, $A^c \gets A^c \backslash \{ g^* \}$
-    \FORALL{$g \in A^c$}
-      \STATE $X_g \gets P_g X_g$
+    \FORALL{$h \in A^c$}
+      \STATE $X_h \gets P_{g^*} X_h$
     \ENDFOR
-    \STATE $r_s \gets P_g r_{s-1}$
+    \STATE $r_s \gets P_{g^*} r_{s-1}$
     \ENDFOR
@@
 Since we do not renormalize the columns after projecting the covariates
-(lines 6 to 8 above), and since we have weights, we are not computing
-correlations unless the design matrix is orthogonal and all weights are 1.
+(lines 6 to 8 above), and since we have weights, we are generally not computing
+correlations unless the design matrix is orthogonal and all weights are 1. We could
+renormalize the columns, though we choose not to.
 There are advantages and disadvantages to both criteria which we do not
@@
 a subset to include in the model. The most general form of variable
-selection is called subset selection, a procedure which searches among
+selection is  "best-subset" selection, a procedure which searches among
 all $2^G$ subsets of $G$ variable groups and picks the best model.
@@
 
+\jonathan{Is average power the same as TPR (or, more correctly, TPP)?}
+
 For various sparsity levels $k$, Algorithm~\ref{algo:fs} was run on
@@
 power'' was calculated as the number of true variables in the active
-set after $k$ steps divided by $k$. (\todo Make this consistent with
+set after $k$ steps divided by $k$:
+$$
+\text{Power}_k = \frac{\# \left\{g: g \in A_k, \beta_g \neq 0\right\}}{k}.
+$$ (\todo Make this consistent with
 $\gamma$ in simulations section)
@@
 
-Writing $\hat \beta(\lambda)$ for the lasso solution for a fixed value of $\lambda$, we need
-the following facts summarized in \cite{significance:lasso} (ref Tibs2012?).
+Writing $\hat \beta(\lambda)$ for the lasso solution for a fixed value of $\lambda$, we recall
+the following facts summarized in \cite{significance:lasso,tibshirani_lasso_uniqueness}.
 
@@
     changes in slope occurring at a finite number of $\lambda$ values
-    referred to as \emph{knots}. The knots depend on the data and are
+    referred to as \emph{knots}. 
+    \item The knots depend on the data and are
     usually written in order $\lambda_1 \geq \lambda_2 \geq \cdots
-    \geq \lambda_r \geq 0$. We follow this convention.
-
-  \item The \emph{active set} $A_k$ is a set of indices of variables
-    for which the corresponding coordinates of $\hat \beta(\lambda_k)$
-    are potentially nonzero. Any variable with index not in $A_k$ has
-    a zero coefficient in $\hat \beta(\lambda_k)$, but the converse is
-    not true. \todo{is this equicorrelation set?}
-
-  \item Path algorithms for computing lasso solutions proceed by
-    fitting models at a grid of $\lambda$ values. The active set
-    changes whenever $\lambda$ crosses a knot, and predictor variables
-    can both enter and leave the active set. However, at the first two
-    knots $\lambda_1$ and $\lambda_2$ no variable can leave the active
-    set. So the first two knots correspond to the first two variables
-    entering the model.
+    \geq \lambda_r \geq 0$.
 
@@
 
-\cite{significance:lasso} prove that under the null hypothesis that $A_k$
-contains all the strong predictor variables, the distribution of a
-test statistic $T_k \propto \lambda_k(\lambda_k - \lambda_{k+1})$ is
-asymptotically Exp(1). In the lasso case we know a lot about the knots
-and active set, but the group lasso picture is slightly more
-complicated. For the group lasso, $\hat \beta(\lambda)$ does not have
-piecewise linear components. To overcome this difficulty we
-restrict our attention to the first group of variables to enter the
-active set since the analysis then follows almost exactly as for the
-lasso.
+The covariance test (assuming $X$'s columns are of length one and $\Sigma = \sigma^2 I$ is at the
+first step by
+\begin{equation}
+\label{eq:covtest}
+T_1 = \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \overset{n,p \to \infty}{\to} \text{Exp}(1).
+\end{equation}
+In \cite{tests:adaptive} it is pointed out 
+\begin{equation}
+\label{test:exact}
+\exp\left(- \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \right) \approx \frac{1 - \Phi(\lambda_1/\sigma)}{1 - \Phi(\lambda_2 / \sigma)} \overset{D}{=} \text{Unif}(0,1).
+\end{equation}
+Further, the above test statistic can be understood as the survival function of $\lambda_1 = \|X^Ty\|_{\infty}$
+{\em conditional} on which variable achieves $\lambda_1$ as well as its sign.
+The limiting results about later steps in \cite{significance:lasso} can (formally) be interpreted as
+recursively applying $T_1$ to the variables that had not previously been selected by LARS.
+
+In the setting of grouped variables, we use an analogous test statistic to
+the right hand side of \cite{test:exact}. The test statistic is presented as an example
+in \cite{tests:adaptive} though we re-derive it here in simpler form. The test draws inspiration
+from the group LASSO.
 
@@
 
+There is a largest value of $\lambda$ such that $\hat{\beta}_{\lambda}=0$ which corresponds to $\lambda_1$
+in the LASSO setting:
+\begin{equation}
+\label{eq:lammax}
+\lambda_1^{\text{group}} = \max_{g} \frac{1}{w_g}\|X_g^Ty\|_2 . 
+\end{equation}
+Formally, the value above is just the dual norm of the penalty
+$$
+{\cal P}(\beta) = \sum_{g=1}^G w_g \| \beta_g \|_2
+$$
+hence it can be expressed as
+\begin{equation}
+\label{eq:lammax}
+\lambda_1^{\text{group}} = \max_{g} \frac{1}{w_g}\|X_g^Ty\|_2 = \frac{1}{w_{g^*}}\eta_{g^*}^TX_{g^*}^Ty . 
+\end{equation}
+where $g^*$ is the group that achieves the max and 
+$\eta_{g^*} = X_{g^*}^Ty / \|X_{g^*}^Ty\|_2$ is the $\ell_2$ unit vector that achieves $\|X_{g^*}^Ty\|_2$.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-Before considering the group lasso, we review some ingredients of the
-proof for the lasso. Let $J = \{ 1, 2, \ldots, p \}$ index variables
-and consider a stochastic process $f_{j,s} = sX_j^Ty$ defined on $T =
-J \times [ -1, 1 ]$. This stochastic process is simply a collection of
-linear combinations of $y$, hence it is Gaussian under the assumption
-of Gaussian errors. The \emph{Karush-Kuhn-Tuckher (KKT) conditions}
-(ref???) imply that $\lambda_1 = \max_j |X_j^Ty|$. By introducing the
-sign variable $s$, we can remove the absolute value and write
-$\lambda_1$ as the maximum of our Gaussian process
-
-\begin{equation}
-\lambda_1 = \max_{(j,s)} f_{j,s}
-\end{equation}
-
-We have exhibited the first knot as the maximum of a Gaussian
-process. We can do this for the second knot by introducing a new
-process. Let $(j_1, s_1)$ be the maximizer so that $\lambda_1 =
-s_1X_{j_1}^Ty$, and define 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{equation}
+\begin{figure}
+\begin{center}
+\resizebox{!}{3in}{\input{curved.pdf_tex}}
+\end{center}
+Selection conditioned on an arbitrary selection event.
+\end{figure}
+We are interested in inference after a selection procedure, say, $S$, for
+which the selection event is the intersection of a list of quadratic inequalities:
+
+$$
 \begin{aligned}
-  f^{(j_1,s_1)}_{(j,s)} & = \frac{ sX_j^T y - s X_j^T X_{j_1} X_{j_1}^Ty  } { 1 -  ss_1 X_j^TX_{j_1}} \\
-  & = \frac{ sX_j^T(I-P_{j_1}) y }{ 1 -  s_1 X_{j_1}^T X_js}
+\{S(y)=s\} = \cap_{i \in I(s)} \{y: y^TQ_iy + a_i^Ty\leq b_i \}.
 \end{aligned}
-\end{equation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-where $P_j$ is the projection onto the subspace spanned by $X_j$. We
-can think of this as a ``residual process'' after regressing out the
-maximum. Write $M = \max_{j \neq j_1, s} f^{(j_1,s_1)}_{(j,s)}$, the
-maximum of this residual process. It can be shown from the KKT
-conditions that $M  = \lambda_2$. To summarize, we have represented
-the first two knots of the lasso solution path as the maxima of some
-natural Gaussian processes. Distributional facts about Gaussian
-processes now allow us to make conclusions about the distribution of
-functions of the knots.
-
-\subsection{Group lasso}
-To extend this argument to the group lasso we need to define Gaussian
-processes that characterize the knots of the group lasso solution
-path.
-
-\todo Either use the simplified argument in the case of equal weights
-(equal group sizes), or finish adjusting the proof of $M = \lambda_2$
-to include the weights and use that version. The proof can go in an
-appendix.
-
-\todo Modify write-up to match notation with \cite{tests:adaptive} and
-then paste it in here
+$$
+
+The quadratic forms are not assumed non-negative definite, but we can, without loss of generality
+assume they are symmetric.
+
+If we consider forward stepwise with groups the selection procedure is just
+choosing the first group, $g^*$ and we have
+$$
+\{g^*(y)=g\} =  \cap_{h \neq g} \{y: y^T(X_hX_h^T/w_h^2 - X_gX_g^T/w_g^2)y \leq 0\}.
+$$
+
+We want to slice through the selection event along a ray with direction $\eta$ that passes through $y$. 
+That is, we need to find
+$$
+\begin{aligned}
+\left\{t: S(y+t\eta)=s \right\} &= \cap_{i \in I(s)} \left\{t: (y+t\eta)^T Q_i (y+t\eta) + a_i^T(y+t\eta) \leq b_i\right\}. \\
+\end{aligned}
+$$
+
+For any given $i \in I(s)$ we see that
+$$
+\begin{aligned}
+\left\{t: (y+t\eta)^T Q_i (y+t\eta) + a_i^T(y+t\eta) \leq b_i\right\} &= \left\{t: t^2 \cdot \eta^TQ_i\eta + t \cdot(2 y^TQ_i\eta + a_i^T\eta) + y^TQ_iy + a_i^Ty - b_i \leq 0 \right\} \\
+&= \text{Intervals}(Q_i,a_i,b_i,y,\eta)
+\end{aligned}
+$$
+
+Each value of $\text{Intervals}(Q_i,a_i,b_i,y,\eta)$ above is one of:
+- $\emptyset$, in which case the selection event is also the $\emptyset$ (so this never occurs);
+- $(-\infty, L(Q_i,a_i,b_i,y,\eta)] \cap [U(Q_i,a_i,b_i,y,\eta), \infty)$ if $\eta^TQ_i\eta < 0$;
+- $[L(Q_i,a_i,b_i,y,\eta), U(Q_i,a_i,b_i,y,\eta)]$ if $\eta^TQ_i\eta > 0$;
+- $(-\infty, L(Q_i,a_i,b_i,y,\eta)]$ if $\eta^TQ_i\eta=0$ and $2 y^TQ_i\eta + a_i^T\eta > 0$;
+- $[U(Q_i,a_i,b_i,y,\eta), \infty)$ if $\eta^TQ_i\eta=0$ and $2 y^TQ_i\eta + a_i^T\eta < 0$;
+- $(-\infty,\infty)$ if $\eta^TQ_i\eta=0, 2 y^TQ_i\eta + a_i^T\eta = 0$ and $y^TQ_iy + a_i^Ty - b_i \leq 0$.
 

### 2014-04-04 paper/paper.tex
working on quadratic selection

@@
 \usepackage{algorithm}
-\usepackage{algorithmic}
-\usepackage{graphics,amssymb,color}
-\usepackage{graphicx,subfigure}
+\usepackage{algorithmic,amssymb}
+\usepackage{graphicx,xcolor}
+\usepackage{subfigure}
 %\usepackage[boxed]{algorithm2e}
@@
 $\chi^2_k$ random variable. This generally does not work when the
-group to be added has been chosen by a method that uses the data,
+group to be added has been chosen by a method that uses the data \cite{olshen:flevel},
 and in particular it fails for forward stepwise procedures adding
@@
 In the setting of grouped variables, we use an analogous test statistic to
-the right hand side of \cite{test:exact}. The test statistic is presented as an example
+the right hand side of \eqref{test:exact}. The test statistic is presented as an example
 in \cite{tests:adaptive} though we re-derive it here in simpler form. The test draws inspiration
@@
 
-\todo Fix references below
-
-The group lasso estimator is discussed in (ref that guy's thesis) and
-(ref Yuan and Lin).  An important extension is the sparse group lasso
-(ref ???) which enforces sparsity in groups as well as sparsity of the
-coefficients within the groups.  For a survey on group lasso and
-related factor models see (ref???). \todo review some more literature
-and add a few more references here if they seem worthwhile.
-
 \begin{figure}
+\label{fig:polytope}
 \begin{center}
-\resizebox{!}{3in}{\input{curved.pdf_tex}}
+\resizebox{!}{3in}{\input{../figs/polytope.pdf_tex}}
 \end{center}
-Selection conditioned on an arbitrary selection event.
+A polytope selection event. Solving the LASSO \eqref{eq:lasso} for a fixed value
+of $\lambda$ yields both an active set $\hat{E}(y)$ and a collection of signs of 
+active variables $z_{\hat{E}(y)}$. The event $\{y: (\hat{E}(y), z_{\hat{E}(y)}) = (E, z_E)\}$
+can be written as a set of affine constraints.
 \end{figure}
-We are interested in inference after a selection procedure, say, $S$, for
-which the selection event is the intersection of a list of quadratic inequalities:
-
-$$
+
+\subsection{Selection events}
+
+The approach to post-selection inference described \cite{lasso:fixed}
+describes the selection event for the LASSO as a convex polytope stylized
+in Figure \ref{fig:polytope}. If all of our groups were of size 1, then
+the event that we observe a given set of variables
+in the forward stepwise path (with their signs as they enter) would also
+be described by a set of affine inequalities. This is described in more detail in \cite{exact:lars}.
+
+After selection, exact inference for a (randomly) chosen contrast $\eta(y)^T\mu$ of the mean vector
+$\mu$ is carried by considering the conditional distribution of 
+\begin{equation}
+y \sim N(\mu, \sigma^2 I) \bigl| Ay \leq b.
+\end{equation}
+See \cite{lasso:fixed,exact:lars} for further details.
+
+In the grouped variables setting, however, the event that $g^*(y)$, the first
+group chosen, is equal to some fixed group $g$ is generally not described by
+a set of affine constraints. The event can be described as
+\begin{equation}
+\label{eq:first:quadratic}
 \begin{aligned}
-\{S(y)=s\} = \cap_{i \in I(s)} \{y: y^TQ_iy + a_i^Ty\leq b_i \}.
+\{g^*(y)=g\} &= \{ \|X_g^Ty\|_2 /w_g \geq \|X_h^Ty\|_2 /w_h , \ \forall h \neq g\} \\
+&=  \{ y^T(X_g^TX_g)y / w_g - y^T(X_h^TX_h)y / w_h, \ \forall h \neq g\} \\
 \end{aligned}
-$$
+\end{equation}
+
+Hence, our selection event can be expressed as 
+the intersection of a list of quadratic inequalities. This non-affine selection event is stylized in Figure
+\ref{fig:curved}.
+In this section, we consider an arbitrary selection procedure
+$S$ that returns one of a set of possible outcomes $s \in {\cal S}$ by a set of quadratic inequalities.
+That is, for all possible outcomes $s \in {\cal S}$:
+\begin{equation}
+\label{eq:selection}
+\begin{aligned}
+\{S(y)=s\} = \cap_{i \neq I(s)} \{y: y^TQ_iy + a_i^Ty\leq b_i \}.
+\end{aligned}
+\end{equation}
 
@@
 assume they are symmetric.
-
-If we consider forward stepwise with groups the selection procedure is just
-choosing the first group, $g^*$ and we have
+As in \cite{lasso:exact} we will be interested in some randomly chosen contrast
+$\eta(S(y))^T\mu$. In our grouped selection procedure, the procedure $S(y)=g^*(y)$
+and $\eta$ is typically 
 $$
-\{g^*(y)=g\} =  \cap_{h \neq g} \{y: y^T(X_hX_h^T/w_h^2 - X_gX_g^T/w_g^2)y \leq 0\}.
+\eta(S(y)) = \eta(g^*(y)) = \frac{X_{g^*(y)}^Ty}{\|X_{g^*(y)}^Ty\|_2}.
 $$
 
-We want to slice through the selection event along a ray with direction $\eta$ that passes through $y$. 
+Having fixed $\eta$, we must slice through the selection event along a ray with direction $\eta$ that passes through $y$. 
 That is, we need to find
@@
 \begin{aligned}
-\left\{t: S(y+t\eta)=s \right\} &= \cap_{i \in I(s)} \left\{t: (y+t\eta)^T Q_i (y+t\eta) + a_i^T(y+t\eta) \leq b_i\right\}. \\
+\left\{t: S(y+t \cdot \eta)=s \right\} &= \cap_{i \in I(s)} \left\{t: (y+t \cdot \eta)^T Q_i (y+t \cdot \eta) + a_i^T(y+t \cdot \eta) \leq b_i\right\}. \\
 \end{aligned}
@@
 
-For any given $i \in I(s)$ we see that
+\begin{figure}
+\label{fig:curved}
+\begin{center}
+\resizebox{!}{3in}{\input{../figs/curved.pdf_edit_tex}}
+\end{center}
+A generic selection event, given by the intersection of a list
+of quadratic inequalities. \end{figure}
+
+The slice for any particular inequality is
 $$
 \begin{aligned}
-\left\{t: (y+t\eta)^T Q_i (y+t\eta) + a_i^T(y+t\eta) \leq b_i\right\} &= \left\{t: t^2 \cdot \eta^TQ_i\eta + t \cdot(2 y^TQ_i\eta + a_i^T\eta) + y^TQ_iy + a_i^Ty - b_i \leq 0 \right\} \\
-&= \text{Intervals}(Q_i,a_i,b_i,y,\eta)
+\lefteqn{
+\left\{t: (y+t\eta)^T Q_i (y+t \cdot \eta) + a_i^T(y+t \cdot \eta) \leq b_i\right\}} \\
+ & \qquad = \left\{t: t^2 \cdot \eta^TQ_i\eta + t \cdot(2 y^TQ_i\eta + a_i^T\eta) + y^TQ_iy + a_i^Ty - b_i \leq 0 \right\} \\
 \end{aligned}
 $$
-
-Each value of $\text{Intervals}(Q_i,a_i,b_i,y,\eta)$ above is one of:
-- $\emptyset$, in which case the selection event is also the $\emptyset$ (so this never occurs);
-- $(-\infty, L(Q_i,a_i,b_i,y,\eta)] \cap [U(Q_i,a_i,b_i,y,\eta), \infty)$ if $\eta^TQ_i\eta < 0$;
-- $[L(Q_i,a_i,b_i,y,\eta), U(Q_i,a_i,b_i,y,\eta)]$ if $\eta^TQ_i\eta > 0$;
-- $(-\infty, L(Q_i,a_i,b_i,y,\eta)]$ if $\eta^TQ_i\eta=0$ and $2 y^TQ_i\eta + a_i^T\eta > 0$;
-- $[U(Q_i,a_i,b_i,y,\eta), \infty)$ if $\eta^TQ_i\eta=0$ and $2 y^TQ_i\eta + a_i^T\eta < 0$;
-- $(-\infty,\infty)$ if $\eta^TQ_i\eta=0, 2 y^TQ_i\eta + a_i^T\eta = 0$ and $y^TQ_iy + a_i^Ty - b_i \leq 0$.
+This can be computed easily, and results in either the empty, a single interval (possibly infinite) or the union of 
+two disjoint infinite intervals. The intersection over all $I(s)$ is therefore easily computable.
+This yields an easily computable form for the slice
+\begin{equation}
+\label{eq:slice}
+\{g^*(y)=g, \eta(g^*(y))=\eta\}.
+\end{equation}
 
@@
 
+\jonathan{make the tick marks denoting the number of true groups larger -- can you not just shade the right hand part gray ? easy to do in python...}
 Finally, most of the plots below show the following information. The

### 2014-04-08 paper/paper.tex
quadratic almost done

@@
 A generic selection event, given by the intersection of a list
-of quadratic inequalities. \end{figure}
+of quadratic inequalities. Note that the event need not be simply connected. \end{figure}
 
@@
 \end{equation}
-
-\subsection{Relation to covariance test}
-
-Consider the simple case for the first covariate to enter the model.
-In this case the covariance test statistic is $T = \lambda_1(\lambda_1 - M)$.
-\todo Does $M = \lambda_2$? Check proof, now that statistic has changed
-
-This is Lemma 5 from \cite{significance:lasso}.
-
-\begin{itemize}
-
-  \item As in LTTT, $T = \lambda_1 (\lambda_1 - M)$ and $M = \lambda_2$
-  \item Convergence to the limiting Exp(1) distribution is too slow
-    \[
-      \frac{P(\chi_k / w_g \geq m + t/m)}{P(\chi_k / w_g \geq m)} \to e^{-t} \text{ as } m \to \infty
-    \]
-    (when the group achieving $\lambda_1$ is group $g$ and has rank $k$)
-  \item The limiting distribution only depends on $T$, but we also observe $M$
-  \item Let's just try the ratio (conditional $\chi_k$ tail probability) evaluated at $T$ and $M$ (it works better)
-  \item Going one step further, instead of using the approximation (see LTTT Proof of Lemma 5)
-    \[
-      \frac{M + \sqrt{M^2+4t}}{2} \approx M + \frac{t}{M}
-    \]
-    we can just use the left hand side
-  \item For $T = \lambda_1(\lambda_1 - M)$ the left hand side simplifies to $\lambda_1$
-  \item Now our p-value is
-    \[
-      \frac{P(\chi_k / w_g \geq \lambda_1)}{P(\chi_k / w_g \geq \lambda_2)}
-    \]   
-\end{itemize}
-
-\subsection{Exact p-value calculation}
-
-\todo Complete this section
-
-We now describe how to calculate our p-value, following the discussion of
-group lasso in \cite{tests:adaptive}. For the rest of this section let
-$g = \gstar$ be the index of
-the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
-%\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
-%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
-For a vector $u \in \real^p$, let $u_h$ denote the coordinates
-of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
-We can rearrange the columns of $X$ to group these adjacently, so that
-\[
-u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
-\]
-One step of the calculation will be to find an orthonormal basis for the
-linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
-\text{ for all } h \neq g \}$ so that we can project orthogonally to
-this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
-should be the case since $g$ maximizes the absolute value of this
-quantity), then the
-space is trivial and the desired orthogonal projection is the identity.
-Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
-generally has dimension $p_g-1$.
-% have dimension $r-1$ for $r = \text{rank}(X_g)$. This holds except in
-% degenerate cases such as $X_g = 0$ or $r \ll p_g$, but these cases are not
-% likely to occur since $\norm{X_g^T y}_2/w_g$ is supposed to be maximized.
-% We proceed by forming the $n \times p_g$ matrix
-% $X_g - X_g X_g^T y y^T X_g/\norm{X_g^Ty}_2^2$ and performing Gram-Schmidt
-% on the first $r-1$ columns. For precise theorems showing that this will
-% generally work, see \cite{outerp:indep}.
-We compute an orthonormal basis by Gram-Schmidt and form a
-$p_g \times p_g$ matrix, which we denote $V_g$, by appending one column
-%0's in the coordinates corresponding to all groups $h \neq g$ and to
-of zeroes (since Gram-Schmidt only produces $p_g-1$ columns).
-With this we are ready to define the projection
-
-\begin{equation}
+In the picture \eqref{eq:curved}, for every $\eta$, the slice \eqref{eq:slice} is a function
+of $P_{\eta}^{\perp}y$. This amounts to a proof of the following lemma.
+
+\begin{lemma}
+Suppose $S$ is a selection procedure, i.e. a map $S:\real^n \mapsto {\cal S}$
+such that for each $s \in S$ \eqref{eq:selection} holds 
+and we are given a matrix valued function $X_s \in \real^{n \times p(s)}$ of rank
+$k(s)$.
+Then, for every $\eta_s \in \real^{p(s)}$ with $\|\eta_s\|_2=1$, the
+slice
+\begin{equation}
+\label{eq:slice:general}
+\left\{t: S(y+t \cdot \eta_s)=s, \frac{X_s^Ty}{\|X_s^Ty\|_2}=\eta_s \right\}
+\end{equation}
+can be described by
+a finite union of closed intervals whose endpoints are functions of 
+$(P_s^{\perp}y, (P_s- (X_s\eta_s)(X_s\eta_s)^T)y)$ where
+$P_s^{\perp}$ is projection on the orthogonal column space of $X_s$.
+Denote this set by $E(s, \eta_s, P_s^{\perp}y, (P_s - X_s \eta_s(X_s\eta_s)^T)y)$ and also define
+$$
 \begin{aligned}
-    \label{eq:proj}
-    P_g &= \Sigma X_gV_g (V_g^T X_g^T \Sigma X_g V_g)^\dagger V_g^TX_g^T \\
+E_+(s, \eta_s, P_s^{\perp}y, (P_s - X_s \eta_s(X_s\eta_s)^T)y) &= \left\{y + t \cdot \eta_s: S(y+t \cdot \eta)=s, \frac{X_s^Ty}{\|X_s^Ty\|_2}=\eta_s \right\} \\
 \end{aligned}
-\end{equation}
-Also define $H_g = (I-P_g)\Sigma$ and the conditional variance
-
-\begin{equation}
-  \begin{aligned}
-    \label{eq:cvar}
-    \sigma^2_g &= y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2) \\
-  \end{aligned}
-\end{equation}
-These simplify when $X_g$ is a single column, in which case $P_g = 0$,
-$H_g = \Sigma$, and $\sigma^2_g = X_g^T\Sigma X_g^T/w_g^2$. We now
-describe how to compute our p-value in Algorithm~\ref{algo:pval}. Note
-that $F_{\chi^2_r}$ denotes the distribution function of a $\chi^2$
-random variable with $r$ degrees of freedom.
-%Define $U_g = X_g^Ty/(w_g\norm{X_g^Ty}_2)$.
-
-\begin{algorithm}
-  \caption{Computing p-value}
-  \label{algo:pval}
-  \begin{algorithmic}
-    \REQUIRE Response $y$, grouped design matrix $X$ with weights, inactive set $A^c$, index $g$ of last group to enter active set $A$.
-    \ENSURE Our p-value for the group $g$ entering the model.
-    \STATE Compute $H_g$ and $\sigma^2_g$
-    \IF{$X_g$ is a single column}
-    \STATE $\sigma_g^2 \gets X_g^T\Sigma X_g/w_g^2$
-    \STATE $\tilde X_g \gets X_g / w_g \cdot \text{sign}(X_g^Ty)$
-    \ELSE
-    \STATE $P_g \gets \Sigma XV_g (V_g^T X^T \Sigma X V_g)^\dagger V_g^TX^T$
-    \STATE $H_g \gets (I - P_g)\Sigma$
-    \STATE $\sigma^2_g \gets y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2)$
-    \STATE $\tilde X_g \gets H_g X_g X_g^T y /(\norm{X_g^Ty}_2 w_g)$
-    \ENDIF
-    \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$
-%    \STATE \COMMENT{Compute the following two $p$-vectors}
-    \STATE $a \gets X^T(y - \lambda \tilde X_g)$
-    \STATE $b \gets X^T \tilde X_g$
-    \STATE Compute solution $(v_-, v_+)$ of \textsc{LinearFractional}(a,b) optimization subproblem
-    \STATE $r \gets \text{rank}(X_g)$
-    \STATE $u \gets [F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(\lambda^2/\sigma_g^2)]/[F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(v^2_+/\sigma_g^2)]$
-    \RETURN $u$
-  \end{algorithmic}
-\end{algorithm}
-
-We next describe the \textsc{LinearFractional} subproblem and its solution. We
-gave the problem this name because it originated in the form
-\[
-\maximize_{h \neq g, \norm{u_h}_2 = 1} \frac{u_h^TX_h^Ty - u_h^TX_h^TX_gX_g^Ty}
-{1 - u_g^TX_g^TX_hu_h}
-\]
-The solution we describe next is to a slightly different problem, and
-also incorporates the information that $g$ maximizes $\norm{X_g^Ty}_2/w_g$.
-For details see Appendix~\ref{app:closedform} or \cite{tests:adaptive}.
-
-\begin{algorithm}
-  \caption{The \textsc{LinearFractional} optimization subproblem}
-  \label{algo:linfrac}
-  \begin{algorithmic}
-    \REQUIRE The $p$-vectors $a, b$ in Algorithm~\ref{algo:pval}, weights, inactive set $A^c$, a small tolerance number
-%(we take $1e^{-10}$)
-    \ENSURE Solution pair $(v_-, v_+)$.
-    \FOR{$h$ in $A^c$}
-      \IF{$\norm{b_h}_2 == 0$ or $\norm{a_h}_2/\norm{b_h}_2 < \text{tol}$}
-        \STATE $(v^-_h, v^+_h) \gets (0, \infty)$
-      \ELSE
-        \STATE $\theta_c \gets a_h^Tb_h/(\norm{a_h}_2 \norm{b_h}_2)$
-        \STATE $\theta_s \gets \sqrt{1-\theta_c^2}$
-        \STATE $\theta \gets \arccos (\theta_c)$
-        \STATE $\phi_s \gets \theta_s \norm{b_h}_2 / w_h$
-        \IF{$\phi_s > 1$}
-          \STATE $(v^-_h, v^+_h) \gets (0, \infty)$
-        \ELSE
-          \STATE $\phi \gets \arcsin(\phi_s)$
-%          \STATE $\phi_2 \gets \pi - \phi_1$
-          \STATE $z_\pm \gets s_\pm \norm{a_h}_2 \cos(\phi) /(w_h - s_\pm \norm{b_h}_2 \cos(\theta - \phi))$ for $s_\pm = \pm 1$
-          \IF{$\norm{b_h}_2 < w_h$}
-            \STATE $(v_h^-, v_h^+) \gets (\max \{ z_+, z_- \}, \infty)$
-          \ELSE
-            \STATE $(v_h^-, v_h^+) \gets (\min \{ z_+, z_- \}, \max \{ z_+, z_- \})$
-          \ENDIF
-        \ENDIF
-      \ENDIF
-    \ENDFOR
-    \STATE $v^- \gets \min_h v_h^-$
-    \STATE $v^+ \gets \max_h v_h^+$
-    \RETURN $(v^-, v^+)$
-  \end{algorithmic}
-\end{algorithm}
+$$
+\end{lemma}
+
+In turn, this allows to derive a test statistic to test the hypothesis
+$H_{0,s}:X_s^T\mu=0$ conditional on $s$ being selected.
+\begin{lemma}
+\label{eq:test:dbn}
+Conditional on $H_{0,s} \cap E_+(s, \eta_s, P_s^{\perp}y, (P_s - \eta_s\eta_s^T)y)$
+the following equality holds in distribution
+\begin{equation}
+\label{eq:chi:truncated}
+(X_s\eta_s)^Ty \overset{D}{=} \theta_s \chi_{k(s)} | (X_s\eta_s)^TE_+(s, \eta_s, P_s^{\perp}y, (P_s - \eta_s\eta_s^T)y) 
+\end{equation}
+where the scale 
+$$
+\theta_s = \sigma \frac{y^TP_sy}{y^TX_sX_s^Ty}.
+$$
+\end{lemma}
+
+{\bf Proof:}
+For any $s$, decompose the distribution of $y$ as
+$$
+(Z^{\perp}_s,Z_s) = (P_s^{\perp}y, X_s^Ty).
+$$
+Under $H_{0,s}:X_s^T\mu=0$, the density of $Z_s$ can be written as
+$$
+(2 \pi \Sigma_s \sigma^2)^{-k(s)/2} \exp \left(-\frac{1}{2 \sigma^2} z^T\Sigma_s^{-1}z \right)
+$$
+with $\Sigma_s = X_s^TX_s$.
+Transforming to polar coordinates $(R,U)$ yields a density
+$$
+(2 \pi \Sigma_s \sigma^2)^{-k(s)/2} r^{k(s)-1} \exp \left(-\frac{r^2}{2 \sigma^2} u^T\Sigma_s^{-1}u \right), \qquad r \geq 0, \|u\|_2=1.
+$$
+Conditioning on $U$ shows that $R=\|X_s^Ty\|_2|U$ has distribution proportional to a $\chi_{r(s)}$. The only thing to determine
+is the scale of the $\chi$, which must be $\sigma \|X_s\eta_s\|_2$.
+
+%\subsection{Relation to covariance test}
+%
+%Consider the simple case for the first covariate to enter the model.
+%In this case the covariance test statistic is $T = \lambda_1(\lambda_1 - M)$.
+%\todo Does $M = \lambda_2$? Check proof, now that statistic has changed
+%
+%This is Lemma 5 from \cite{significance:lasso}.
+%
+%\begin{itemize}
+%
+%  \item As in LTTT, $T = \lambda_1 (\lambda_1 - M)$ and $M = \lambda_2$
+%  \item Convergence to the limiting Exp(1) distribution is too slow
+%    \[
+%      \frac{P(\chi_k / w_g \geq m + t/m)}{P(\chi_k / w_g \geq m)} \to e^{-t} \text{ as } m \to \infty
+%    \]
+%    (when the group achieving $\lambda_1$ is group $g$ and has rank $k$)
+%  \item The limiting distribution only depends on $T$, but we also observe $M$
+%  \item Let's just try the ratio (conditional $\chi_k$ tail probability) evaluated at $T$ and $M$ (it works better)
+%  \item Going one step further, instead of using the approximation (see LTTT Proof of Lemma 5)
+%    \[
+%      \frac{M + \sqrt{M^2+4t}}{2} \approx M + \frac{t}{M}
+%    \]
+%    we can just use the left hand side
+%  \item For $T = \lambda_1(\lambda_1 - M)$ the left hand side simplifies to $\lambda_1$
+%  \item Now our p-value is
+%    \[
+%      \frac{P(\chi_k / w_g \geq \lambda_1)}{P(\chi_k / w_g \geq \lambda_2)}
+%    \]   
+%\end{itemize}
+%
+%
+%\subsection{Exact p-value calculation}
+%
+%\todo Complete this section
+%
+%We now describe how to calculate our p-value, following the discussion of
+%group lasso in \cite{tests:adaptive}. For the rest of this section let
+%$g = \gstar$ be the index of
+%the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
+%%\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
+%%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
+%For a vector $u \in \real^p$, let $u_h$ denote the coordinates
+%of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
+%We can rearrange the columns of $X$ to group these adjacently, so that
+%\[
+%u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
+%\]
+%One step of the calculation will be to find an orthonormal basis for the
+%linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
+%\text{ for all } h \neq g \}$ so that we can project orthogonally to
+%this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
+%should be the case since $g$ maximizes the absolute value of this
+%quantity), then the
+%space is trivial and the desired orthogonal projection is the identity.
+%Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
+%generally has dimension $p_g-1$.
+%% have dimension $r-1$ for $r = \text{rank}(X_g)$. This holds except in
+%% degenerate cases such as $X_g = 0$ or $r \ll p_g$, but these cases are not
+%% likely to occur since $\norm{X_g^T y}_2/w_g$ is supposed to be maximized.
+%% We proceed by forming the $n \times p_g$ matrix
+%% $X_g - X_g X_g^T y y^T X_g/\norm{X_g^Ty}_2^2$ and performing Gram-Schmidt
+%% on the first $r-1$ columns. For precise theorems showing that this will
+%% generally work, see \cite{outerp:indep}.
+%We compute an orthonormal basis by Gram-Schmidt and form a
+%$p_g \times p_g$ matrix, which we denote $V_g$, by appending one column
+%%0's in the coordinates corresponding to all groups $h \neq g$ and to
+%of zeroes (since Gram-Schmidt only produces $p_g-1$ columns).
+%With this we are ready to define the projection
+%
+%\begin{equation}
+%  \begin{aligned}
+%    \label{eq:proj}
+%    P_g &= \Sigma X_gV_g (V_g^T X_g^T \Sigma X_g V_g)^\dagger V_g^TX_g^T \\
+%   \end{aligned}
+%\end{equation}
+%Also define $H_g = (I-P_g)\Sigma$ and the conditional variance
+%
+%\begin{equation}
+%  \begin{aligned}
+%    \label{eq:cvar}
+%    \sigma^2_g &= y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2) \\
+%  \end{aligned}
+%\end{equation}
+%These simplify when $X_g$ is a single column, in which case $P_g = 0$,
+%$H_g = \Sigma$, and $\sigma^2_g = X_g^T\Sigma X_g^T/w_g^2$. We now
+%describe how to compute our p-value in Algorithm~\ref{algo:pval}. Note
+%that $F_{\chi^2_r}$ denotes the distribution function of a $\chi^2$
+%random variable with $r$ degrees of freedom.
+%%Define $U_g = X_g^Ty/(w_g\norm{X_g^Ty}_2)$.
+%
+%\begin{algorithm}
+%  \caption{Computing p-value}
+%  \label{algo:pval}
+%  \begin{algorithmic}
+%    \REQUIRE Response $y$, grouped design matrix $X$ with weights, inactive set $A^c$, index $g$ of last group to enter active set $A$.
+%    \ENSURE Our p-value for the group $g$ entering the model.
+%    \STATE Compute $H_g$ and $\sigma^2_g$
+%    \IF{$X_g$ is a single column}
+%    \STATE $\sigma_g^2 \gets X_g^T\Sigma X_g/w_g^2$
+%    \STATE $\tilde X_g \gets X_g / w_g \cdot \text{sign}(X_g^Ty)$
+%    \ELSE
+%    \STATE $P_g \gets \Sigma XV_g (V_g^T X^T \Sigma X V_g)^\dagger V_g^TX^T$
+%    \STATE $H_g \gets (I - P_g)\Sigma$
+%    \STATE $\sigma^2_g \gets y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2)$
+%    \STATE $\tilde X_g \gets H_g X_g X_g^T y /(\norm{X_g^Ty}_2 w_g)$
+%    \ENDIF
+%    \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$
+%%    \STATE \COMMENT{Compute the following two $p$-vectors}
+%    \STATE $a \gets X^T(y - \lambda \tilde X_g)$
+%    \STATE $b \gets X^T \tilde X_g$
+%    \STATE Compute solution $(v_-, v_+)$ of \textsc{LinearFractional}(a,b) optimization subproblem
+%    \STATE $r \gets \text{rank}(X_g)$
+%    \STATE $u \gets [F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(\lambda^2/\sigma_g^2)]/[F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(v^2_+/\sigma_g^2)]$
+%    \RETURN $u$
+%  \end{algorithmic}
+%\end{algorithm}
+%
+%We next describe the \textsc{LinearFractional} subproblem and its solution. We
+%gave the problem this name because it originated in the form
+%\[
+%\maximize_{h \neq g, \norm{u_h}_2 = 1} \frac{u_h^TX_h^Ty - u_h^TX_h^TX_gX_g^Ty}
+%{1 - u_g^TX_g^TX_hu_h}
+%\]
+%The solution we describe next is to a slightly different problem, and
+%also incorporates the information that $g$ maximizes $\norm{X_g^Ty}_2/w_g$.
+%For details see Appendix~\ref{app:closedform} or \cite{tests:adaptive}.
+%
+%\begin{algorithm}
+%  \caption{The \textsc{LinearFractional} optimization subproblem}
+%  \label{algo:linfrac}
+%  \begin{algorithmic}
+%    \REQUIRE The $p$-vectors $a, b$ in Algorithm~\ref{algo:pval}, weights, inactive set $A^c$, a small tolerance number
+%%(we take $1e^{-10}$)
+%    \ENSURE Solution pair $(v_-, v_+)$.
+%    \FOR{$h$ in $A^c$}
+%      \IF{$\norm{b_h}_2 == 0$ or $\norm{a_h}_2/\norm{b_h}_2 < \text{tol}$}
+%        \STATE $(v^-_h, v^+_h) \gets (0, \infty)$
+%      \ELSE
+%        \STATE $\theta_c \gets a_h^Tb_h/(\norm{a_h}_2 \norm{b_h}_2)$
+%        \STATE $\theta_s \gets \sqrt{1-\theta_c^2}$
+%        \STATE $\theta \gets \arccos (\theta_c)$
+%        \STATE $\phi_s \gets \theta_s \norm{b_h}_2 / w_h$
+%        \IF{$\phi_s > 1$}
+%          \STATE $(v^-_h, v^+_h) \gets (0, \infty)$
+%        \ELSE
+%          \STATE $\phi \gets \arcsin(\phi_s)$
+%%          \STATE $\phi_2 \gets \pi - \phi_1$
+%          \STATE $z_\pm \gets s_\pm \norm{a_h}_2 \cos(\phi) /(w_h - s_\pm \norm{b_h}_2 \cos(\theta - \phi))$ for $s_\pm = \pm 1$
+%          \IF{$\norm{b_h}_2 < w_h$}
+%            \STATE $(v_h^-, v_h^+) \gets (\max \{ z_+, z_- \}, \infty)$
+%          \ELSE
+%            \STATE $(v_h^-, v_h^+) \gets (\min \{ z_+, z_- \}, \max \{ z_+, z_- \})$
+%          \ENDIF
+%        \ENDIF
+%      \ENDIF
+%    \ENDFOR
+%    \STATE $v^- \gets \min_h v_h^-$
+%    \STATE $v^+ \gets \max_h v_h^+$
+%    \RETURN $(v^-, v^+)$
+%  \end{algorithmic}
+%\end{algorithm}
+%
 

### 2014-04-21 paper/paper.tex
Changed GAMSel to not be hierarchical

@@
 
-The covariance test (assuming $X$'s columns are of length one and $\Sigma = \sigma^2 I$ is at the
-first step by
+The covariance test (assuming groups of size one and $\Sigma = \sigma^2 I$ is given at the first step by
 \begin{equation}
@@
 
-The \emph{group lasso estimator} is a solution to the following convex
-problem
+The \emph{group lasso estimator} is a solution to the following penalized
+least-squares convex optimization problem
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \displaystyle \hat \beta_\lambda = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
-   \lambda \sum_{g=1}^G w_g \| \beta_g \|_2
+   \lambda {\cal P}(\beta) \\
+\end{aligned}
+\end{equation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+with the group penalty
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{equation}
+  \begin{aligned}
+  \label{eq:gpen}
+    {\cal P}(\beta)&= \sum_{g=1}^G w_g \| \beta_g \|_2 \\
   \end{aligned}
@@
 $w_g$ are usually taken to be $\sqrt {p_g}$ to normalize the penalty
-across groups.  Note that this includes the usual lasso estimator as a
+across groups with different sizes.  Note that this includes the usual lasso estimator as a
 special case when all of the groups are of size 1, since then the
-penalty term is the $L^1$-norm of $\beta$.
-
-There is a largest value of $\lambda$ such that $\hat{\beta}_{\lambda}=0$ which corresponds to $\lambda_1$
-in the LASSO setting:
+penalty term is the $L^1$-norm of $\beta$. The solution path of the group
+lasso has similar properties to the lasso case, however it is not
+generally piecewise linear.
+
+For sufficiently large $\lambda$ the solution to~\eqref{eq:gsoln} is forced
+to be zero. The smallest such $\lambda$ is denoted
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{equation}
+  \begin{aligned}
+   \lambda_1 &= \inf \{ \lambda \geq 0 : \hat \beta_{\lambda'} = 0 \text{ for all } \lambda' > \lambda \} \\
+  \end{aligned}
+\end{equation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+In the LASSO setting this is explicitly computable by
 \begin{equation}
@@
 \end{equation}
-Formally, the value above is just the dual norm of the penalty
-$$
-{\cal P}(\beta) = \sum_{g=1}^G w_g \| \beta_g \|_2
-$$
+The value above is also the dual norm of the penalty~\eqref{eq:gpen}
 hence it can be expressed as
@@
 \label{eq:lammax}
-\lambda_1^{\text{group}} = \max_{g} \frac{1}{w_g}\|X_g^Ty\|_2 = \frac{1}{w_{g^*}}\eta_{g^*}^TX_{g^*}^Ty . 
+\lambda_1^{\text{group}}
+% = \max_{g} \frac{1}{w_g}\|X_g^Ty\|_2
+ = \frac{1}{w_{g^*}}\eta_{g^*}^TX_{g^*}^Ty . 
 \end{equation}
 where $g^*$ is the group that achieves the max and 
-$\eta_{g^*} = X_{g^*}^Ty / \|X_{g^*}^Ty\|_2$ is the $\ell_2$ unit vector that achieves $\|X_{g^*}^Ty\|_2$.
+$\eta_{g^*} = X_{g^*}^Ty / \|X_{g^*}^Ty\|_2$ is the unit vector that achieves $\|X_{g^*}^Ty\|_2$.
 

### 2014-04-21 gamsel/fwd_gamsel_sim.R

@@ -7,7 +7,7 @@ source('generate_data.R')
 #source('pred_and_estim.R') not for glinternet
 #source('fwd_step/coherence.R')
 
-fwd_gamsel_simulation = function(n, Sigma, groups, num.nonzero, num.linear, lower, upper, nsim,  max.steps, design = 'gaussian', corr = 0, categorical = FALSE, predictions = FALSE, coherence = FALSE, fixed.data = NULL, cat.groups = NULL, alpha = .1) {
+fwd_gamsel_simulation = function(n, Sigma, groups, num.nonzero, num.linear, lower, upper, nsim,  max.steps, design = 'uniform', corr = 0, categorical = FALSE, predictions = FALSE, coherence = FALSE, fixed.data = NULL, cat.groups = NULL, alpha = .1) {
 
   # Initialize
   p = length(groups)
@@ -62,31 +62,23 @@ fwd_gamsel_simulation = function(n, Sigma, groups, num.nonzero, num.linear, lowe
     if (length(names(fixed.data)) > 0) {
       print("Using fixed design matrix")
       data = fixed.data
-      # Do not use predictions yet-- need to split to training/test
-#      data.test = fixed.data
     } else {
       if (design == 'gaussian') {
         X = gaussian_design(n, groups, col.normalize = FALSE, corr = corr)
-#        X.test = gaussian_design(n, groups, col.normalize = FALSE, corr = corr)
       } else {
         design_name = paste0(design, "_design")
         if (exists(design_name, mode = "function")) {
           design_fun = get(design_name)
           X = design_fun(n, groups)
-#          X.test = design_fun(n, groups)
         } else {
           stop(paste("Misspecified design matrix:", design))
         }
       }
       data = generate_gamsel(X, groups)
-#      data.test = generate_glinternet(X.test, groups)    
     }
     X = data$X
-#    X.test = data.test$X
     all.groups = data$all.groups
     G = length(unique(all.groups))
-#    weights = sqrt(rle(all.groups)$lengths)
-#    weights = sqrt(sapply(rle(all.groups)$lengths - 1, function(x) max(1, x)))
     weights = rep(1, G)
     spline.groups = data$spline.groups
[…]

### 2014-04-21 gamsel/generate_gamsel.R

@@ -49,10 +49,11 @@ generate_gamsel = function(X, groups, degrees = 3, col.normalize = FALSE) {
     if (sum(col) > 1) {
       stop("Currently only supports groups of size 1")
     }
-    spline.basis = bs(col_normalize(X[ ,col]), degree = degree)
-    X.out = cbind(X.out, X[, col])
+    Xvect = X[, col]
+    spline.basis = bs(Xvect, degree = degrees[g], Boundary.knots = c(0,1))
+    #X.out = cbind(X.out, X[, col])
     X.out = cbind(X.out, spline.basis)
-    all.groups = c(all.groups, rep(gnew, 1+ncol(spline.basis)))
+    all.groups = c(all.groups, rep(gnew, ncol(spline.basis))) #+1))
   }
   
   colnames(X.out) = NULL

### 2014-04-21 gamsel/sim_gamsel.R

@@ -5,22 +5,22 @@ source('gamsel/fwd_gamsel_sim.R')
 #source('tex_table.R')
 source('plots.R')
 
-design = 'gaussian'
+design = 'uniform'
 fn.append = ''
 main.append = ''
 corr = 0 # nonzero only supported for gaussian design
 noisecorr = 0
 
-nsim = 50
-n = 100
-num.nonzero = 5
+nsim = 100
+n = 80
+num.nonzero = 4
 k = num.nonzero
 num.linear = 2
 max.steps = 10
-upper.coeff = 10
-lower.coeff = 7
+upper.coeff = 3
+lower.coeff = 2
 Sigma = (1-noisecorr)*diag(rep(1,n)) + noisecorr
-groups = 1:25
+groups = 1:200
 p = length(groups)
 g = length(unique(groups))
 mult = sqrt(2*log(g))

### 2014-04-30 paper/paper.tex
Some paper edits and additions, gamsel modification

@@
 \newcommand{\jonathan}[1]{{\color{blue}  #1}}
+\newcommand{\josh}[1]{{\color{green} #1}}
 
@@
 is correct or not. }
+\josh{Yes, I forgot to specify $r_0 = y$, hence the $y$ is not required (I removed it)}
 
@@
     \ENSURE Ordered active set $A$ of variable groups included in the model at each step
-    \STATE $A \gets \emptyset$, $A^c \gets \{ 1, \ldots, G\}$
+    \STATE $A \gets \emptyset$, $A^c \gets \{ 1, \ldots, G\}$, $r_0 \gets y$
     \FOR{$s=1$ to $steps$}
-    \STATE $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T(y- r_{s-1})}_2 / w_g \}$
+    \STATE $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^Tr_{s-1}}_2 / w_g \}$
     \STATE $P_{g^*} \gets I_{n\times n} - X_{g^*}X_{g^*}^\dagger$
@@
 \jonathan{Is average power the same as TPR (or, more correctly, TPP)?}
+\josh{Yes, maybe I should have used that terminology instead?}
 
@@
 {\em conditional} on which variable achieves $\lambda_1$ as well as its sign.
-The limiting results about later steps in \cite{significance:lasso} can (formally) be interpreted as
+The limiting results about later steps in \cite{significance:lasso} can
+%(formally)
+be interpreted as
 recursively applying $T_1$ to the variables that had not previously been selected by LARS.
@@
 
-The \emph{group lasso estimator} is a solution to the following penalized
-least-squares convex optimization problem
+Recall the lasso estimator is given by
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{equation}
+\begin{aligned}
+\label{eq:lasso}
+\displaystyle \hat \beta_\lambda = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
+   \lambda \| \beta \|_1 \\
+\end{aligned}
+\end{equation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The \emph{group lasso estimator}, a generalization of the lasso,
+ is a solution to the following penalized
+least-squares convex problem
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-In the LASSO setting this is explicitly computable by
+For the LASSO,
+
+this is explicitly computable by
 \begin{equation}
@@
 \end{equation}
-where $g^*$ is the group that achieves the max and 
-$\eta_{g^*} = X_{g^*}^Ty / \|X_{g^*}^Ty\|_2$ is the unit vector that achieves $\|X_{g^*}^Ty\|_2$.
+where $g^*$ is the group that achieves the maximum and 
+$\eta_{g^*} = X_{g^*}^Ty / \|X_{g^*}^Ty\|_2$ is the unit vector that achieves the norm $\|X_{g^*}^Ty\|_2$.
 
@@
 
-\begin{figure}
-\label{fig:polytope}
-\begin{center}
-\resizebox{!}{3in}{\input{../figs/polytope.pdf_tex}}
-\end{center}
-A polytope selection event. Solving the LASSO \eqref{eq:lasso} for a fixed value
-of $\lambda$ yields both an active set $\hat{E}(y)$ and a collection of signs of 
-active variables $z_{\hat{E}(y)}$. The event $\{y: (\hat{E}(y), z_{\hat{E}(y)}) = (E, z_E)\}$
-can be written as a set of affine constraints.
-\end{figure}
+% \begin{figure}
+% \label{fig:polytope}
+% \begin{center}
+% \resizebox{!}{3in}{\input{../figs/polytope.pdf_tex}}
+% \end{center}
+% A polytope selection event. Solving the LASSO \eqref{eq:lasso} for a fixed value
+% of $\lambda$ yields both an active set $\hat{E}(y)$ and a collection of signs of 
+% active variables $z_{\hat{E}(y)}$. The event $\{y: (\hat{E}(y), z_{\hat{E}(y)}) = (E, z_E)\}$
+% can be written as a set of affine constraints.
+% \end{figure}
 
@@
 
-The approach to post-selection inference described \cite{lasso:fixed}
-describes the selection event for the LASSO as a convex polytope stylized
-in Figure \ref{fig:polytope}. If all of our groups were of size 1, then
+The approach to post-selection inference in \cite{lasso:fixed}
+describes the selection event for the LASSO as a convex polytope
+satisfying a list of affine constraints.
+%stylized in Figure \ref{fig:polytope}.
+If all of our groups were of size 1, then
 the event that we observe a given set of variables
-in the forward stepwise path (with their signs as they enter) would also
-be described by a set of affine inequalities. This is described in more detail in \cite{exact:lars}.
-
+in the forward stepwise path (with their signs as they enter) would similarly
+be given by a set of affine inequalities.
+%This is described in more detail in \cite{exact:lars}.
 After selection, exact inference for a (randomly) chosen contrast $\eta(y)^T\mu$ of the mean vector
-$\mu$ is carried by considering the conditional distribution of 
+$\mu$ could then be accomplished by analyzing the conditional distribution
 \begin{equation}
@@
 \end{equation}
-See \cite{lasso:fixed,exact:lars} for further details.
+See \cite{lasso:fixed,exact:lars} for further details on this approach.
 
 In the grouped variables setting, however, the event that $g^*(y)$, the first
-group chosen, is equal to some fixed group $g$ is generally not described by
-a set of affine constraints. The event can be described as
+group chosen, is equal to some fixed group $g$ is generally not given by
+a set of affine constraints. Rather, considering line 3 of Algorithm~\ref{algo:fs}, we see
 \begin{equation}
@@
 \{g^*(y)=g\} &= \{ \|X_g^Ty\|_2 /w_g \geq \|X_h^Ty\|_2 /w_h , \ \forall h \neq g\} \\
-&=  \{ y^T(X_g^TX_g)y / w_g - y^T(X_h^TX_h)y / w_h, \ \forall h \neq g\} \\
+&=  \{ y^T(X_g^TX_g)y / w_g^2 - y^T(X_h^TX_h)y / w_h^2 \geq 0, \ \forall h \neq g\} \\
 \end{aligned}
 \end{equation}
-
 Hence, our selection event can be expressed as 
@@
 \ref{fig:curved}.
-In this section, we consider an arbitrary selection procedure
+In the rest of this section, we consider an arbitrary selection procedure
 $S$ that returns one of a set of possible outcomes $s \in {\cal S}$ by a set of quadratic inequalities.
-That is, for all possible outcomes $s \in {\cal S}$:
+That is, for all possible outcomes $s \in {\cal S}$ there is a set of indices
+$I(s)$ such that
 \begin{equation}
@@
 \begin{aligned}
-\{S(y)=s\} = \cap_{i \neq I(s)} \{y: y^TQ_iy + a_i^Ty\leq b_i \}.
+\{S(y)=s\} = \cap_{i \in I(s)} \{y: y^TQ_iy + a_i^Ty\leq b_i \}.
 \end{aligned}
@@
 assume they are symmetric.
-As in \cite{lasso:exact} we will be interested in some randomly chosen contrast
+As in \cite{lasso:fixed} we will be interested in some randomly chosen contrast
 $\eta(S(y))^T\mu$. In our grouped selection procedure, the procedure $S(y)=g^*(y)$
@@
 
-Having fixed $\eta$, we must slice through the selection event along a ray with direction $\eta$ that passes through $y$. 
+Having fixed $\eta$, we proceed to study this contrast by slicing through the selection event along a ray with direction $\eta$ that passes through $y$. 
 That is, we need to find
@@
 \begin{aligned}
-\left\{t: S(y+t \cdot \eta)=s \right\} &= \cap_{i \in I(s)} \left\{t: (y+t \cdot \eta)^T Q_i (y+t \cdot \eta) + a_i^T(y+t \cdot \eta) \leq b_i\right\}. \\
+\lefteqn{
+\left\{t \in \mathbb R : S(y+t \cdot \eta)=s \right\} } \\
+ & \qquad = \cap_{i \in I(s)} \left\{t \in \mathbb R : (y+t \cdot \eta)^T Q_i (y+t \cdot \eta) + a_i^T(y+t \cdot \eta) \leq b_i\right\}. \\
 \end{aligned}
@@
 \lefteqn{
-\left\{t: (y+t\eta)^T Q_i (y+t \cdot \eta) + a_i^T(y+t \cdot \eta) \leq b_i\right\}} \\
- & \qquad = \left\{t: t^2 \cdot \eta^TQ_i\eta + t \cdot(2 y^TQ_i\eta + a_i^T\eta) + y^TQ_iy + a_i^Ty - b_i \leq 0 \right\} \\
+\left\{t \in \mathbb R : (y+t\cdot \eta)^T Q_i (y+t \cdot \eta) + a_i^T(y+t \cdot \eta) \leq b_i\right\}} \\
+ & \qquad = \left\{t \in \mathbb R : t^2 \cdot (\eta^TQ_i\eta) + t \cdot(2 y^TQ_i\eta + a_i^T\eta) + y^TQ_iy + a_i^Ty - b_i \leq 0 \right\} \\
 \end{aligned}
 $$
-This can be computed easily, and results in either the empty, a single interval (possibly infinite) or the union of 
-two disjoint infinite intervals. The intersection over all $I(s)$ is therefore easily computable.
-This yields an easily computable form for the slice
+This can be computed explicitly, and results in either the empty set, a single interval (possibly infinite) or the union of 
+two disjoint infinite intervals. The intersection over all $I(s)$ is therefore also computable, yielding a form for the slice
 \begin{equation}
@@
 \end{equation}
-In the picture \eqref{eq:curved}, for every $\eta$, the slice \eqref{eq:slice} is a function
+In the picture \ref{fig:curved}, for every $\eta$, the slice \eqref{eq:slice} is a function
 of $P_{\eta}^{\perp}y$. This amounts to a proof of the following lemma.
@@
 $P_s^{\perp}$ is projection on the orthogonal column space of $X_s$.
-Denote this set by $E(s, \eta_s, P_s^{\perp}y, (P_s - X_s \eta_s(X_s\eta_s)^T)y)$ and also define
+\end{lemma}
+Denote the set \eqref{eq:slice:general} by $E(s, \eta_s, P_s^{\perp}y, (P_s - X_s \eta_s(X_s\eta_s)^T)y)$ and also define
 $$
 \begin{aligned}
-E_+(s, \eta_s, P_s^{\perp}y, (P_s - X_s \eta_s(X_s\eta_s)^T)y) &= \left\{y + t \cdot \eta_s: S(y+t \cdot \eta)=s, \frac{X_s^Ty}{\|X_s^Ty\|_2}=\eta_s \right\} \\
+\lefteqn{
+E_+(s, \eta_s, P_s^{\perp}y, (P_s - X_s \eta_s(X_s\eta_s)^T)y) } \\
+ & \qquad = \left\{y + t \cdot \eta_s: S(y+t \cdot \eta)=s, \frac{X_s^Ty}{\|X_s^Ty\|_2}=\eta_s \right\} \\
 \end{aligned}
 $$
-\end{lemma}
 
@@
 Conditional on $H_{0,s} \cap E_+(s, \eta_s, P_s^{\perp}y, (P_s - \eta_s\eta_s^T)y)$
-the following equality holds in distribution
+we have the following distributional result
 \begin{equation}
@@
 {\bf Proof:}
-For any $s$, decompose the distribution of $y$ as
+For any fixed $s$, decompose $y$ as
 $$
@@
 $$
-Conditioning on $U$ shows that $R=\|X_s^Ty\|_2|U$ has distribution proportional to a $\chi_{r(s)}$. The only thing to determine
-is the scale of the $\chi$, which must be $\sigma \|X_s\eta_s\|_2$.
+Conditioning on $U$ shows that $R=\|X_s^Ty\|_2|U$ has distribution proportional to a $\chi_{k(s)}$. Finally, observe that 
+the scale of the $\chi$ above is given by $\theta_s$. \qed
+%$\sigma \|X_s\eta_s\|_2$.
 
@@
 \jonathan{make the tick marks denoting the number of true groups larger -- can you not just shade the right hand part gray ? easy to do in python...}
+\josh{Done, updated plots now have this form. Will soon finish updating all plots}
 Finally, most of the plots below show the following information. The
@@
 
-\input{../small_sim_selection.tex}
-
-\input{../small_sim_estimation.tex}
+\input{../tables/default_gaussian_nsim500_n100_p50_g50_k5_lower1pt5_upper2.tex}
+
+\input{../tables/default_gaussian_nsim500_n100_p50_g50_k5_lower1pt5_upper2_estimation.tex}
 
@@
 
-\input{../large_sim_selection.tex}
+\input{../tables/default_gaussian_nsim500_n100_p200_g200_k5_lower1pt5_upper2.tex}
 
@@
 
+\subsection{Nonlinear regression with splines}
+
+\begin{figure}
+\begin{center}
+\subfigure[Forward stepwise with splines]{
+\label{fig:spline:fwd}
+\includegraphics[width=0.5\textwidth]{../figs/fwdstep/gauss_glint_p20_n200_lower2p9_upper3p6.pdf}}
+\hspace{-15pt}
+\subfigure[P-value with splines]{
+\label{fig:spline:pval}
+\includegraphics[width=0.5\textwidth]{../figs/gamsel_uniform_nsim100_n200_p50_g50_k6_k02_lower10_upper15.pdf}}
+\caption{\small \it The left panel shows the power of forward stepwise
+for various sparsity levels. The right panel shows our p-value marginally
+over each step, with red indicating the global null case and blue
+indicating a signal with $k=6$ nonzero groups.}
+\label{fig:spline}
+\end{center}
+\end{figure}
+
+First we consider a simple extension of the usual linear regression setting.
+Given a design matrix $X$, we can augment this by adding additional columns given by nonlinear functions of the original covariates.
+Specifically, we compute spline basis vectors for each covariate and create a group given by this spline basis.
+Denote this new design matrix
+\begin{equation}
+\label{eq:splinemat}
+  \begin{aligned}
+    \tilde X &=  \begin{pmatrix} X_1 & \cdots & X_G & X^+_1 & \cdots & X^+_G  \end{pmatrix}\\
+    &= \begin{pmatrix} X & X^+ \end{pmatrix}
+  \end{aligned}
+\end{equation}
+with $\tilde G = 2G$ groups.
+For simplicity we assume the original groups are all size 1, and use the B-spline basis.
+In our simulations we generate original covariates uniformly in the interval [-1,1] and use cubic splines with boundary knots
+\begin{equation}
+  \begin{aligned}
+    X_i &\sim \text{Unif}[-1,1], \quad
+    X_i^+ = \text{CubicSplineBasis}(X_i)
+  \end{aligned}
+\end{equation}
+Nonzero coefficients are split between original covariates and spline groups.
+Simulation results are shown in Figure~\ref{fig:spline}.
+Note that the varying group sizes present some difficulty to forward stepwise, which tends to first try linear approximations to nonzero spline groups before adding the true spline group.
+
 \subsection{Glinternet for hierarchical interactions}
 \label{sec:glint}
-In regression settings with many variables, considering models with pairwise interactions can drastically increase model complexity.
+In regression settings with many variables, choosing among models with pairwise interactions can drastically increase model complexity.
 \cite{glint} propose a method called \textsc{glinternet} to reduce the statistical complexity of this problem.
@@
 interactions.
-This way each nonzero main effect with any nonzero interactions has
-exactly one nonzero interaction. Furthermore, each nonzero interaction
-coefficient has been inflated to be larger than the corresponding main
+%This way each nonzero main effect with any nonzero interactions has
+%exactly one nonzero interaction. 
+We also inflate each nonzero interaction
+coefficient to be larger than the corresponding main
 effect coefficients. This special case is favorable for our algorithm,
@@
 
-\subsection{Generalized additive model selection}
-\todo Finish code \\
-\todo Write section \\
-
-\subsection{Real data example}
+\subsection{HIVdb data example}
 \label{sec:hiv}
@@
 drugs in treating HIV. Quantitative measurements of drug
-response/resistance were regressed on categorical covariates encoded
+response/resistance were regressed on categorical covariates encoding
 the presence and type of certain genetic mutations. We attempt a similar
@@
 %\end{figure}
+
+We can also use the procedure described in Section~\ref{sec:glint} to fit a model with pairwise interactions between mutations.
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.9\textwidth]{../figs/HIV_PI_glint.pdf}
+\caption{\small \it Results from PI dataset}
+\label{fig:HIVPI:glint}
+\end{center}
+\end{figure}
+
+\begin{figure}
+\begin{center}
+\includegraphics[width=0.9\textwidth]{../figs/HIV_NRTI_glint.pdf}
+\caption{\small \it Results from NRTI dataset}
+\label{fig:HIVNRTI:glint}
+\end{center}
+\end{figure}
+
+\input{../tables/HIV_PI_glint.tex}
+
+\input{../tables/HIV_NRTI_glint.tex}
 

### 2014-05-01 paper/paper.tex
Reworked power simulations, began comparison with BIC

@@
 \end{equation}
-(\todo describe construction of $\beta$ after finding $A$)
 where we have written $\{ \beta_s : s = 1, \ldots, steps \}$ as
@@
 potentially sub-optimal sequence of models, under favorable conditions
-it can still perform well. There is a small segment of the
-compressed sensing literature \citep{donoho:pursuit, cai:wang:omp}
-with results stating that forward stepwise (often referred to in that
-literature as Orthogonal Matching Pursuit or OMP) can exactly select
-the true model
-under some stringent conditions involving quantities like the sparsity
-of the true model and the coherence of the design matrix. The
-coherence $\mu(X)$ of a matrix $X$ with columns scaled to have unit
-2-norm is defined as
-
+it can still perform well.
+There is a subset of the compressed sensing literature \citep{donoho:pursuit, cai:wang:omp} dedicated to forward stepwise (often referred to in that literature as Orthogonal Matching Pursuit or OMP).
+Typical results from these works establish that forward stepwise can exactly select the true model under some stringent conditions involving quantities like the sparsity of the true model and the coherence of the design matrix.
+The coherence $\mu(X)$ of a matrix $X$ with columns scaled to have unit 2-norm is defined as
 \begin{equation}
@@
 \end{equation}
-
-Recalling that $k$ is the sparsity of $\beta$, typical results in this
-literature say that if $k < (1/\mu + 1)/2$ and the nonzero
-coefficients of $\beta$ are sufficiently large then forward stepwise
-recovers $\beta$ with high probability. The coherence condition is
-necessary to guarantee exact recovery \citep{cai:wang:xu:sharp} in the
-sense that it is possible to construct counterexamples with
-$k = (1/\mu + 1)/2$. We refer the reader to the literature for details.
-For our purposes the
-conditions required to guarantee exact recovery are usually too
-stringent. Simulations show empirically that forward
-stepwise can work well even when it is not working perfectly, and that
-it does so under a wide range of conditions.
+Recalling that $k$ is the sparsity of $\beta$, it has been shown that if $k < (1/\mu + 1)/2$ and the nonzero coefficients of $\beta$ are sufficiently large then forward stepwise recovers $\beta$ with high probability.
+The coherence condition is necessary to guarantee exact recovery \citep{cai:wang:xu:sharp} in the sense that it is possible to construct counterexamples with $k = (1/\mu + 1)/2$.
+We refer the reader to \cite{donoho:pursuit, cai:wang:omp} for details.
+But for our purposes the conditions required to guarantee exact recovery are usually too stringent.
+Simulations show empirically that forward stepwise can work well even when it is not working perfectly, and that it does so under a wide range of conditions.
 
@@
 \begin{center}
-\subfigure[Ternary design]{
-\label{fig:fsternary}
-\includegraphics[width=0.5\textwidth]{../figs/fwdstep/ternary_p2n_n100_lower1p4_upper1p4.pdf}}
+\subfigure[Categorical design]{
+\label{fig:fscat}
+\includegraphics[width=0.5\textwidth]{../figs/tpr_default_categorical_nsim100_n100_p400_g200_sizes2-2_k24_lower2_upper4.pdf}}
 \hspace{-15pt}
-\subfigure[Gaussian design with correlation]{
+\subfigure[Correlated gaussian design]{
 \label{fig:fsgausscorr}
-\includegraphics[width=0.5\textwidth]{../figs/fwdstep/gaussian_p2n_n100_lower1p4_upper1p6_corr0p2.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/tpr_default_gaussian_nsim100_n100_p200_g200_k24_lower1pt2_upper2_corr0pt1.pdf}}
 \caption{\small \it The left panel shows the results of the simulation
-  using design matrices with i.i.d. ternary entries taking values $0,
-  \pm 1$ with equal probability. The left panel shows the results
-  with design matrices of standard Gaussian entries with
-  equi-correlation of 0.2. }
+  using a categorical design consisting of 200 binary factors.
+  The right panel shows the results with design matrices of Gaussian
+  entries with equi-correlation 0.1 between all columns. Plotted in
+  green are the TPP and average model sizes for models chosen by BIC.}
 \label{fig:fwdstepsim}
@@
 
-\jonathan{Is average power the same as TPR (or, more correctly, TPP)?}
-\josh{Yes, maybe I should have used that terminology instead?}
-
 For various sparsity levels $k$, Algorithm~\ref{algo:fs} was run on
-various data sets with \textit{steps} equal to $k$ and the ``average
-power'' was calculated as the number of true variables in the active
-set after $k$ steps divided by $k$:
-$$
-\text{Power}_k = \frac{\# \left\{g: g \in A_k, \beta_g \neq 0\right\}}{k}.
-$$ (\todo Make this consistent with
-$\gamma$ in simulations section)
-Nonzero coefficients had magnitude
-in a range of multiples of $\gamma := \sqrt{2\log(G)}$, e.g. in $[1.4
+various data sets with \textit{steps} equal to $k$.
+In the resulting active set $A_k$, the number of true variables divided by $k$ is the True Positive Proportion (TPP):
+\[
+\text{TPP}_k = \frac{\# \left\{g: g \in A_k, \beta_g \neq 0\right\}}{k}.
+\]
+Since we know $k$ in the simulations we can use $\text{TPP}_k$ as a measure of how well forward stepwise is performing.
+When it is close to 1 we are recovering most of the true variables before including false ones.
+Nonzero coefficients had magnitudes
+in a range of multiples of $\gamma := \sqrt{2\log(G)/n}$, e.g. in $[1.4
 \gamma, 1.6 \gamma]$. Results are shown in
@@
 recovery, but the performance, while not perfect, may still be good
-enough for some purposes.
+enough for some purposes. Finally, to serve as a comparison with existing
+commonly used model selection procedures, in Figure~\ref{fig:fwdstepsim}
+we show in green the TPP and average model sizes for models chosen by
+the \textit{step} function in \textit{R} using the BIC criterion.
 
@@
 \label{fig:spline:fwd}
-\includegraphics[width=0.5\textwidth]{../figs/fwdstep/gauss_glint_p20_n200_lower2p9_upper3p6.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/tpr_gamsel_uniform_nsim50_n100_p200_g200_k21_k01_lower5_upper10.pdf}}
 \hspace{-15pt}
@@
 \label{fig:glint:fwd}
-\includegraphics[width=0.5\textwidth]{../figs/fwdstep/gauss_glint_p20_n200_lower2p9_upper3p6.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/tpr_glint_categorical_nsim50_n100_p72_g36_sizes2-2_k15_k01_lower3_upper5.pdf}}
 \hspace{-15pt}
@@
 \label{fig:glint:pval}
-\includegraphics[width=0.5\textwidth]{../figs/gaussian_size1_n200_p40_g40_k6_lower2p5_upper3p9_glint.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/glint_gaussian_nsim100_n100_p30_g30_k6_k02_lower4_upper8.pdf}}
 \caption{\small \it The left panel shows the power of forward stepwise

### 2014-05-01 fwd_step.R

@@ -65,8 +65,50 @@ add_group = function(X, Y, groups, weights, Sigma, active.set = 0, eff.p = 0, ca
   return(list(test.output = results, var = results$var, p.value = p.value, added = imax, active.set = new.active.set, eff.p = new.eff.p, Y.update = Y.resid, X.update = X.project, grank=kmax))
 }
 
+
+# Bare bones version for power simulation without p-values
+add_group_nopval= function(X, Y, groups, weights, Sigma, active.set = 0, eff.p = 0, cat.groups = NULL) {
+    n = length(Y)
+    g.labels = unique(groups)
+    inactive.groups = setdiff(g.labels, active.set)
+
+    terms = sapply(inactive.groups, function(x) sqrt(sum((t(X[,groups==x]) %*% Y)^2))/weights[x])
+
+      imax = inactive.groups[which.max(terms)]
+    
+    gmax = groups == imax
+    grank = sum(gmax)
+
+    new.active.set = update_active_set(active.set, imax)
+    inactive.set = setdiff(inactive.groups, imax)
+    
+    # Effective number of parameters
+    new.eff.p = eff.p + sum(gmax)
+
+    if (new.eff.p >= n) {
+        stop("Too many variables added. Try decreasing max.steps")
+    }
+
+    # Form new residual
+    X.project = X
+    Xgmax = X[, gmax]
+    Pgmax = Xgmax %*% ginv(Xgmax)
+      
+    # Project all other groups orthogonal to the one being added
+    for (gind in inactive.set) {
+        group = groups == gind
+        X.project[, group] = X[, group] - Pgmax %*% X[, group]
+    }
+
+    Y.resid = Y - Pgmax %*% Y
+  
+    return(list(test.output = NA, var = NA, p.value = NA, added = imax, active.set = new.active.set, eff.p = new.eff.p, Y.update = Y.resid, X.update = X.project, grank=grank))
+}
+
+
+# Main forward stepwise function
 # Iterate add_group for max.steps
-forward_group = function(X, Y, groups, weights = 0, Sigma = NULL, max.steps = 0, cat.groups = NULL, tol=1e-10) {
+forward_group = function(X, Y, groups, weights = 0, Sigma = NULL, max.steps = 0, cat.groups = NULL, tol=1e-10, pval=TRUE) {
   n = length(Y)
   group.sizes = rle(groups)$lengths
 
[…]

### 2014-05-01 fwd_step/fstep_sim_gaussian.R

@@ -63,10 +63,7 @@ for (i in 1:length(ns)) {
   }
 
   power.MCavg = rowMeans(pow.mat)
-
   print(pow.mat)
-
-#  print(mu.mat)
 
   # Save plot
   plot.main = paste0("n = ", n, ", p = ", p, ", signal strength ", lower.coeff, "/", upper.coeff)

### 2014-05-01 sim_default.R

@@ -1,4 +1,5 @@
 source('simulation.R')
+source('tpr_simulation.R')
 source('tex_table.R')
 source('plots.R')
 
@@ -7,7 +8,7 @@ type = 'default'
 design = 'gaussian'
 estimation = TRUE
 n = 100
-groups = 1:200
+groups = 1:00
 #groups = 1:50
 k = 5
 upper = 2
@@ -30,7 +31,6 @@ output = run_simulation(
     estimation = estimation, verbose = TRUE)
 
 with(output, step_plot(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, max.steps, upper, lower, max.beta, min.beta, fwd.power, design, filename))
-
 
 ## ps.fname = paste0('figs/bysignal/', design, '_size1_n', n, '_p', p, '_g', p, '_k', k, '_lower', lower, '_upper', upper)
 ## if (corr != 0) {

### 2014-05-04 paper/paper.tex
Finished comparison of stopping rules

@@
 
-\jonathan{Is $r_0=0$ or $y$? Should be defined in the algorithm. I also put a $y$ in below which I'm not sure
-is correct or not. }
-\josh{Yes, I forgot to specify $r_0 = y$, hence the $y$ is not required (I removed it)}
+%\jonathan{Is $r_0=0$ or $y$? Should be defined in the algorithm. I also put a $y$ in below which I'm not sure
+%is correct or not. }
+%\josh{Yes, I forgot to specify $r_0 = y$, hence the $y$ is not required (I removed it)}
 
@@
   entries with equi-correlation 0.1 between all columns. Plotted in
-  green are the TPP and average model sizes for models chosen by BIC.}
+  green are the TPP and median model sizes (top) for models chosen by BIC,
+  and in red the same are shown (bottom) for RIC.}
 \label{fig:fwdstepsim}
@@
 commonly used model selection procedures, in Figure~\ref{fig:fwdstepsim}
-we show in green the TPP and average model sizes for models chosen by
-the \textit{step} function in \textit{R} using the BIC criterion.
+we show in green the TPP and median model sizes for models chosen by
+the \textit{step} function in \textit{R} using the BIC criterion, and
+the same for RIC are plotted in red.
 
@@
 
-\todo Compare to AIC and other classical stopping rules? \\
-
 We performed simulations with several classes of design matrices including
@@
 
-\jonathan{make the tick marks denoting the number of true groups larger -- can you not just shade the right hand part gray ? easy to do in python...}
-\josh{Done, updated plots now have this form. Will soon finish updating all plots}
 Finally, most of the plots below show the following information. The
@@
 
-\input{../tables/default_gaussian_nsim500_n100_p50_g50_k5_lower1pt5_upper2.tex}
-
-\input{../tables/default_gaussian_nsim500_n100_p50_g50_k5_lower1pt5_upper2_estimation.tex}
-
 \begin{figure}[h]
@@
 
-\input{../tables/default_gaussian_nsim500_n100_p200_g200_k5_lower1pt5_upper2.tex}
+%\input{../tables/default_gaussian_nsim500_n100_p200_g200_k5_lower1pt5_upper2.tex}
+
+\input{../tables/stop_comparison_easy.tex}
+
+\input{../tables/stop_comparison_hard.tex}
 
@@
 \label{fig:spline:fwd}
-\includegraphics[width=0.5\textwidth]{../figs/tpr_gamsel_uniform_nsim50_n100_p200_g200_k21_k01_lower5_upper10.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/tpr_gamsel_uniform_nsim100_n100_p200_g200_k18_k01_lower5_upper10.pdf}}
 \hspace{-15pt}
@@
 \label{fig:glint:fwd}
-\includegraphics[width=0.5\textwidth]{../figs/tpr_glint_categorical_nsim50_n100_p72_g36_sizes2-2_k15_k01_lower3_upper5.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/tpr_glint_categorical_nsim100_n100_p72_g36_sizes2-2_k15_k01_lower5_upper10.pdf}}
 \hspace{-15pt}

### 2014-05-04 selection.R

@@ -19,42 +19,68 @@ selection_stats = function(p.list, active.set, true.step, k, stop.rule, alpha = 
   } else {
     stop(paste("Misspecified stopping rule:", stop.rule))
   }
-#  if (stop.rule == "first") {
-#    stop.ind = stop_first(p.list, alpha)
-#  } else if (stop.rule == "forward") {
-#    stop.ind = stop_forward(p.list, alpha)
-#  } else if (stop.rule == "last") {
-#    stop.ind = stop_last(p.list, alpha)
   if (stop.ind == 0) {
-    stopped.set = c()
+    #stopped.set = c()
     S = 0
   } else {
-    stopped.set = active.set[1:stop.ind]
+    #stopped.set = active.set[1:stop.ind]
     S = sum(true.step[1:stop.ind])
   }
   R = stop.ind
   V = R - S
   Fdp = V/max(1, R)
   Power = S/max(1, k)
-  return(c(Fdp, R, S, V, Power))
+  return(c(R, Fdp, Power))
 }
 
 # names = null.p, signal.p, active.set, true.set
-sim_select_stats = function(p.lists, active.sets, true.steps, k, alpha = .1) {
-  first.mat = matrix(0, nrow = nrow(p.lists), ncol = 5)
+sim_select_stats = function(p.lists, active.sets, true.steps, k, alpha = .1, bic.list = NULL, ric.list = NULL) {
+  first.mat = matrix(0, nrow = nrow(p.lists), ncol = 3)
   forward.mat = first.mat
   last.mat = first.mat
+  bic.mat = first.mat
+  ric.mat = first.mat
+  oracle.mat = first.mat
   for (i in 1:nrow(p.lists)) {
+    oracle.mat[i, ] = selection_stats(p.lists[i, 1:k], active.sets[i, ], true.steps[i, ], k, stop.rule = "all", alpha)
     first.mat[i, ] = selection_stats(p.lists[i, ], active.sets[i, ], true.steps[i, ], k, stop.rule = "first", alpha)
     forward.mat[i, ] = selection_stats(p.lists[i, ], active.sets[i, ], true.steps[i, ], k, stop.rule = "forward", alpha)
     last.mat[i, ] = selection_stats(p.lists[i, ], active.sets[i, ], true.steps[i, ], k, stop.rule = "last", alpha)
+    if (length(bic.list) > 0) {
+      bics = bic.list[[i]]
[…]

### 2014-05-04 sim_default.R

@@ -3,17 +3,22 @@ source('tpr_simulation.R')
 source('tex_table.R')
 source('plots.R')
 
-nsim = 500
+nsim = 400
 type = 'default'
 design = 'gaussian'
-estimation = TRUE
+estimation = FALSE #TRUE
 n = 100
-groups = 1:00
-#groups = 1:50
-k = 5
+#groups = 1:500
+#groups = 1:400
+#groups = 1:200
+#groups = 1:100
+groups = 1:50
+
+k = 10
 upper = 2
 lower = 1.5
-max.steps = 10
+
+max.steps = min(n/2, max(groups)-1) #floor(max(groups)/3)
 corr = 0 # nonzero only supported for gaussian design
 noisecorr = 0
 
@@ -28,9 +33,12 @@ output = run_simulation(
     n = n, groups = groups, k = k,
     upper = upper, lower = lower,
     max.steps = max.steps,
+    corr = corr,
     estimation = estimation, verbose = TRUE)
 
-with(output, step_plot(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, max.steps, upper, lower, max.beta, min.beta, fwd.power, design, filename))
+output$main.append = paste0(", BIC: (", round(output$bic.tpp, 2), "/", output$bic.size, ")")
+
+with(output, step_plot(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, max.steps, upper, lower, max.beta, min.beta, fwd.power, design, filename, main.append))
 
 ## ps.fname = paste0('figs/bysignal/', design, '_size1_n', n, '_p', p, '_g', p, '_k', k, '_lower', lower, '_upper', upper)
 ## if (corr != 0) {
@@ -67,11 +75,10 @@ with(output, step_plot(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, m
 ## abline(v = 1, col = "gray")
 ## dev.off()
 
-upper = 3
-lower = 2
-groups = sort(c(rep(1:30, 5), rep(31:35, 10)))
+#groups = sort(c(rep(1:30, 5), rep(31:35, 10)))
 #groups = sort(c(rep(1:10, 3), rep(11:20, 2)))
-max.steps = 10
+#groups = 1:200
+k = 15
 
 output.g = run_simulation(
     nsim = nsim,
@@ -80,16 +87,19 @@ output.g = run_simulation(
     n = n, groups = groups, k = k,
[…]

### 2014-05-04 stop_rules.R

@@ -5,6 +5,12 @@
 # Input a list of p-values and alpha
 # Output largest index for model inclusion (0 if empty model)
 
+# Include everything
+# This one is a shortcut for AIC, BIC, and k-oracle
+stop_all = function(p.list, alpha = .1) {
+  return(length(p.list))
+}
+
 stop_first = function(p.list, alpha = .1) {
 
   p.big = which(p.list > alpha)

### 2014-05-04 tpr_simulation.R

@@ -23,7 +23,8 @@ run_tpr_simulation = function(
     verbose=TRUE,
     ...) {
 
-    klist = 3*(1:floor(kmax/3))
+    #klist = 3*(1:floor(kmax/3))
+    klist = c(1,5,10,15,20,25)
     klen = length(klist)
 
     if (type != "default") {
@@ -79,6 +80,8 @@ run_tpr_simulation = function(
     
     bic.tpp = tpp.mat
     bic.steps = tpp.mat
+    ric.tpp = tpp.mat
+    ric.steps = tpp.mat
     
     estimation.errs = matrix(0, nrow=1, ncol=klen)
     start.time = as.numeric(Sys.time())
@@ -226,18 +229,24 @@ run_tpr_simulation = function(
             results = forward_group(Xscaled, Y.beta, groups=all.groups, weights, Sigma, max.steps = k, cat.groups = cat.groups, pval=FALSE)
             as = results$active.set
 
-            # Using step() with BIC
+            # Using step() with BIC/RIC
             if (type == "default") {
               if (design == "gaussian") {
-
-                steps = min(2*k, k+10)
-                bic.as = step.as(X=X, Y=Y.beta, steps = steps, k=log(n))
+                dfmult = log(n)
+                steps = min(floor(n/2), max(groups)-1)
+                
+                bic.as = step.as(X=Xscaled, Y=Y.beta, steps = steps, k=dfmult)
                 b.tpp = length(intersect(bic.as, true.active))/k
                 bic.tpp[i,j] = b.tpp
                 bic.steps[i,j] = length(bic.as)
 
+                dfmult = 2*log(p)
+                ric.as = step.as(X=Xscaled, Y=Y.beta, steps = steps, k=dfmult)
+                r.tpp = length(intersect(ric.as, true.active))/k
+                ric.tpp[i,j] = r.tpp
+                ric.steps[i,j] = length(ric.as)
+
               }
-              
             }
             
             
@@ -264,7 +273,11 @@ run_tpr_simulation = function(
     # Summarize some results
     TPR = colMeans(tpp.mat)
     bic.TPR = colMeans(bic.tpp)
-    bic.size = as.character(round(colMeans(bic.steps), 1))
+    bic.size = as.character(apply(bic.steps, 2, median))
+    if (p > n) {
+      ric.TPR = colMeans(ric.tpp)
[…]

### 2014-05-06 paper/paper.tex
Almost done

@@
 % "Title of the paper"
-\title{A significance test in forward stepwise regression}
+\title{A significance test in forward stepwise model selection}
 
@@
 \begin{abstract}
-% perhaps de-emphasize group lasso and emphasize forward stepwise?
-  We extend the methods developed by \cite{significance:lasso} and
+% perhaps de-emphasize group LASSO and emphasize forward stepwise?
+  We extend the methods developed by \cite{significance:LASSO} and
   \cite{tests:adaptive} on significance tests for penalized
@@
 
-Forward stepwise regression is a classical model selection procedure
+Consider the regression setting with a single response variable $Y$
+and collection of predictor or covariate variables denoted $\mathcal X$.
+One often wishes to choose a subset of variables $X \subset \mathcal X$
+for modeling the response, assuming the remaining variables are
+irrelevent and can be discarded without much loss of predictive
+or explanatory power. Doing this in a structured and principled way
+usually requires algorithmic methods for choosing $X$. One such method,
+forward stepwise regression, is a procedure
 that begins with an empty model and sequentially adds the best predictor
-variable in each step. The usual $\chi^2$ and $F$-tests for significance
+variable in each step. Because of the stochastic nature of this
+algorithm--making use of the data to choose variables--
+the usual $\chi^2$ and $F$-tests for significance
 fail when a model has been selected this way and tend to be
 anti-conservative unless they are computed on a held-out validation
-dataset. In the lasso setting, \cite{significance:lasso} derived
+dataset. This is one instance of the general problem of conducting
+inference and model selection using the same data, a problem
+of central importance on which some recent progress has been made.
+
+In the LASSO setting, \cite{significance:LASSO} derived
 a novel test statistic and its asymptotic null distribution, making
@@
 \cite{tests:adaptive} modified and extended those results to the
-group lasso \citep{grouplasso} and other penalized regression
+group LASSO \citep{groupLASSO} and other penalized regression
 problems, but only under the global null hypothesis.
@@
 used for valid significance testing when computed on the same
-data used for model selection, eliminating the need for data splitting or for situations in which
-data splitting is not appropriate.
+data used for model selection, eliminating the need for data splitting
+or for situations in which data splitting is not appropriate. For example,
+when categorical covariates contain rare levels with very few observations
+it can be difficult or impossible to split the sample with an adequate
+number of such observations occurring in each split.
+
 The present work iteratively applies the global null test
@@
 
-In Section~\ref{sec:stepwise} we establish notation and describe the
+In the next section we establish notation and describe the
 forward stepwise algorithm used in most of the paper.
 Section~\ref{sec:testing} reviews some recent work on
-post-selection inference: \cite{significance:lasso,tests:adaptive,lasso:fixed}
-relevant to our work here, and we describes a general framework
+post-selection inference
+\citep{significance:LASSO,tests:adaptive,LASSO:fixed}
+relevant to our work here, and describes a general framework
 for post-selection inference based on quadratic comparisons.
@@
 to be widely used by practitioners.
-For example, search results on PubMed for forward stepwise
-(summarized in Figure~\ref{fig:pubmed}) show that many recent
+For example, search results on PubMed for forward stepwise,
+summarized in Figure~\ref{fig:pubmed}, show that many recent
 papers mention the method and there is an increasing trend over time.
@@
 We allow forward stepwise selection to add groups of variables in each
-step, not only in the case of binary encoding for categorical
+step, not only in the case of dummy variable encoding for categorical
 variables but also for any grouping purpose. For example, groups of
@@
 measurements for all genes in a single functional pathway.
-To emphasize this we use $g, h$ as indices rather than the usual
+To emphasize this we use $g, h$ as covariate indices rather than the usual
 $i, j$ throughout. Since single variables can be considered groups of
-size 1, our general exposition includes non-grouped situations as a special
+size 1, this includes non-grouped situations as a special
 case. Our variant of forward stepwise uses a different method than
@@
 % Do we need to assume $p_g \ll n$ or $\text{rank}(X_g) = p_g$?
-When a group encodes a categorical variable as binary indicators for the
-levels of that variable, by default we use the full encoding with a column for every
-level. Although this introduces collinearity, our method does not require
-the design matrix to have full rank.
-
-Define $p = \sum_{g=1}^Gp_g$,
+When a group encodes a categorical variable as indicators for the
+levels of that variable, by default we use the full encoding with a
+column for every level. Although this introduces collinearity, our method
+does not require the design matrix to have full rank.
+
+Denote $p = \sum_{g=1}^Gp_g$,
 the total number of individual variables, so $p = G$ in the
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-For convenience, we assume throughout that the individual columns of $X$ are scaled to
-have unit 2-norm (we do not scale groups, this is accomplished later
-by weighting).
+We normalize each group by its Frobenius norm, but this is not required
+especially since we also allow weights, $w_g$, for each group. These
+weights act like penalties or costs, so increasing $w_g$ makes it
+more difficult for the group $X_g$ to enter the model. The
+modeler can choose weights arbitrarily, but we set them all to be
+1 since we already normalize by the Frobenius norm.
+
 With each group we associate the $p_g \times 1$ coefficient vector
@@
 of such nonzero groups as the \textit{sparsity} of the model, and
-denote this $k := \# \{ g : \beta_g \neq 0 \}$.
-
-Before we describe the forward stepwise procedure we require one last
-ingredient. To each group of variables we assign a weight $w_g$. These
-weights act like penalties or costs, so increasing $w_g$ makes it
-more difficult for the group $X_g$ to enter the model. The
-modeler can choose weights arbitrarily, but we mostly use one
-particular choice, based on $p_g$, that we discuss later. With this we
-are ready to describe the forward stepwise procedure.
+denote this $k := \# \{ g : \beta_g \neq 0 \}$. With this notation
+we are now ready to describe our procedure concretely.
 
@@
 \textit{steps} should be at most $\min (n, G) - 1$, but it is
-computationally desirable to set it as low as possible while still being
-larger than the (unknown) sparsity of $\beta$.
-This way forward stepwise has a chance of recovering all the nonzero
-coefficients of $\beta$ and then terminating without performing much
+computationally desirable to set it as low as possible while
+safely larger than the sparsity of $\beta$. 
+Then forward stepwise may recover all the nonzero
+coefficients of $\beta$ and terminate without performing much
 additional computation.
+Of course the sparsity is usually unknown, so this requires guesswork.
 In our implementation we treat the active set $A$ as an ordered list to easily track the order of groups entering the model.
-
-% \begin{algorithm}
-% \DontPrintSemicolon
-% \KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
-% \KwResult{Active set $A$ of variable groups included in the model}
-% \SetKwFunction{lsfitResidual}{lsfitResidual}
-% \caption{\small \it Forward stepwise procedure with groups and weights}
-% \BlankLine
-% $A \gets \emptyset$\;
-% $A^c \gets \{ 1, \ldots, G \}$\;
-% $r_0 \gets y$\;
-% \For{$s \gets 1$ \KwTo $steps$}{
-%   $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
-%   $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
-%   $A \gets A \cup \{ g^* \}$\;
-%   $A^c \gets A^c \backslash \{ g^* \}$\;
-%   \For{$g \in A^c$}{
-%     $X_g \gets P X_g$\;
-%   }
-%   $r_s \gets P r_{s-1}$\;
-% }
-% \KwRet{$A$}
-% \label{algo:fs}
-% \end{algorithm}
-
-%\jonathan{Is $r_0=0$ or $y$? Should be defined in the algorithm. I also put a $y$ in below which I'm not sure
-%is correct or not. }
-%\josh{Yes, I forgot to specify $r_0 = y$, hence the $y$ is not required (I removed it)}
 
@@
 of $X$ corresponding to $A$. 
-We now use forward stepwise to refer specifically to
-Algorithm~\ref{algo:fs} unless otherwise specified. Note that other
+ Note that other
 implementations of forward stepwise use different criteria for choosing
@@
 There are advantages and disadvantages to both criteria which we do not
-discuss. Our choice was motivated by the group lasso result in
+discuss. Our choice was motivated by the group LASSO result in
 \cite{tests:adaptive}, but we believe other criteria can be handled
 with appropriate modifications.
+We now use forward stepwise to refer specifically to
+Algorithm~\ref{algo:fs} unless otherwise specified.
 
@@
 variable selection: from a potentially large set of variables it chooses
-a subset to include in the model. The most general form of variable
-selection is  "best-subset" selection, a procedure which searches among
-all $2^G$ subsets of $G$ variable groups and picks the best model.
+a subset to include in the model. The most ambitious form of variable
+selection is  ``best-subset'' selection, a procedure which picks the
+best model among all $2^G$ subsets of the $G$ possible groups.
 This exhaustive search is computationally infeasible when $G$ is
@@
 smaller set of potential models, with cardinality at most
-\textit{steps}. However
+\textit{steps} (which, recall, is less than $\min(n,G)$). However
 it is a greedy algorithm, so the set of models it produces may not
@@
 where we have written $\{ \beta_s : s = 1, \ldots, steps \}$ as
-the sequence of models outputted by forward stepwise. The function
+the sequence of models output by forward stepwise. The function
 $\pen(\beta)$ is a penalty on model complexity usually taken to be the
@@
 statistics have also been used, so it is natural to consider using
-the new test statistics of \cite{significance:lasso} or
+the new test statistics of \cite{significance:LASSO} or
 \cite{tests:adaptive}
 to choose a model. \cite{sequential:fdr} examined some stopping rules
-using the asymptotic p-values of \cite{significance:lasso} and showed
+using the asymptotic p-values of \cite{significance:LASSO} and showed
 their stopping rules control false discovery rate--the expected
@@
 \end{equation}
-Recalling that $k$ is the sparsity of $\beta$, it has been shown that if $k < (1/\mu + 1)/2$ and the nonzero coefficients of $\beta$ are sufficiently large then forward stepwise recovers $\beta$ with high probability.
+Denoting $k$ as the sparsity of $\beta$, the literature establishes
+that if $k < (1/\mu + 1)/2$ and the nonzero coefficients of $\beta$ are sufficiently large then forward stepwise recovers $\beta$ with high probability.
 The coherence condition is necessary to guarantee exact recovery \citep{cai:wang:xu:sharp} in the sense that it is possible to construct counterexamples with $k = (1/\mu + 1)/2$.
 We refer the reader to \cite{donoho:pursuit, cai:wang:omp} for details.
-But for our purposes the conditions required to guarantee exact recovery are usually too stringent.
+For our purposes the conditions required to guarantee exact recovery are usually too stringent.
 Simulations show empirically that forward stepwise can work well even when it is not working perfectly, and that it does so under a wide range of conditions.
@@
 \begin{center}
-\subfigure[Categorical design]{
+\subfigure[Categorical designs]{
 \label{fig:fscat}
@@
 \hspace{-15pt}
-\subfigure[Correlated gaussian design]{
+\subfigure[Gaussian designs]{
 \label{fig:fsgausscorr}
-\includegraphics[width=0.5\textwidth]{../figs/tpr_default_gaussian_nsim100_n100_p200_g200_k24_lower1pt2_upper2_corr0pt1.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/tpr_default_gaussian_nsim200_n100_p200_g200_k25_lower1pt5_upper2.pdf}}
 \caption{\small \it The left panel shows the results of the simulation
@@
   The right panel shows the results with design matrices of Gaussian
-  entries with equi-correlation 0.1 between all columns. Plotted in
-  green are the TPP and median model sizes (top) for models chosen by BIC,
-  and in red the same are shown (bottom) for RIC.}
+  entries. Plotted in green are the TPP (dashed line) and median model
+  sizes (top) for models chosen by BIC, and in purple the same are shown
+  (dot-dash line, bottom) for RIC.}
 \label{fig:fwdstepsim}
@@
 
-For various sparsity levels $k$, Algorithm~\ref{algo:fs} was run on
-various data sets with \textit{steps} equal to $k$.
-In the resulting active set $A_k$, the number of true variables divided by $k$ is the True Positive Proportion (TPP):
+For various sparsity levels $k$, we applied Algorithm~\ref{algo:fs} to
+various data sets with \textit{steps} set to $k$.
+In the resulting active set $A_k$, the number of true variables divided by $k$ is the $k$-True Positive Proportion ($k$-TPP):
 \[
@@
 \]
-Since we know $k$ in the simulations we can use $\text{TPP}_k$ as a measure of how well forward stepwise is performing.
+In some plots this has been denoted $k$-SP instead of $k$-TPP (\todo fix).
+Since we know $k$ in the simulations we can use $k$-TPP as a measure of how well forward stepwise is performing.
 When it is close to 1 we are recovering most of the true variables before including false ones.
-Nonzero coefficients had magnitudes
+In our simulations nonzero coefficients have magnitudes
 in a range of multiples of $\gamma := \sqrt{2\log(G)/n}$, e.g. in $[1.4
@@
 enough for some purposes. Finally, to serve as a comparison with existing
-commonly used model selection procedures, in Figure~\ref{fig:fwdstepsim}
+commonly used model selection procedures, in Figure~\ref{fig:fwdstepsim}(b)
 we show in green the TPP and median model sizes for models chosen by
 the \textit{step} function in \textit{R} using the BIC criterion, and
-the same for RIC are plotted in red.
+the same for RIC are plotted in purple.
 
@@
 
-\section{Significance testing: from lasso to group lasso}
+\section{Significance testing: from LASSO to group LASSO}
 \label{sec:testing}
-
-\todo Reorganize and update this section \\
 
@@
 $\chi^2_k$ random variable. This generally does not work when the
-group to be added has been chosen by a method that uses the data \cite{olshen:flevel},
+group to be added has been chosen by a method that uses the data
+\citep{olshen:flevel},
 and in particular it fails for forward stepwise procedures adding
 the ``best'' (e.g. most highly correlated) predictor in each step. In
-that case, the test statistic (drop in RSS) does not match the
+that case, the drop in RSS does not match its
 theoretical null distribution even when the null hypothesis is
-true. \cite{significance:lasso} introduced a new test statistic based
-on the knots in the lasso solution path. They derived a simple
+true. \cite{significance:LASSO} introduced a new test statistic based
+on the knots in the LASSO solution path. They derived a simple
 asymptotic null distribution, proved a type of convergence under broad
@@
 of combining inference with model selection.  \cite{tests:adaptive}
-extended that work to the group lasso \citep{grouplasso} and other
+extended that work to the group LASSO \citep{groupLASSO} and other
 problems, and modified the test statistic to one with an exact finite
@@
 
-Writing $\hat \beta(\lambda)$ for the lasso solution for a fixed value of $\lambda$, we recall
-the following facts summarized in \cite{significance:lasso,tibshirani_lasso_uniqueness}.
+To describe the previous work we require some facts about the solution path
+of the LASSO.
+Writing $\hat \beta(\lambda)$ for the LASSO solution for a fixed value of $\lambda$, we recall
+the following facts summarized in \cite{significance:LASSO,tibshirani_LASSO_uniqueness}.
 
@@
   \item The vector valued function $\hat \beta(\lambda)$ is a
-    continuous function of $\lambda$. For the lasso path, the
+    continuous function of $\lambda$. For the LASSO path, the
     coordinates of $\hat \beta(\lambda)$ are piecewise linear with
@@
 {\em conditional} on which variable achieves $\lambda_1$ as well as its sign.
-The limiting results about later steps in \cite{significance:lasso} can
+The limiting results about later steps in \cite{significance:LASSO} can
 %(formally)
@@
 
-Recall the lasso estimator is given by
+Recall the LASSO estimator is given by
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \begin{aligned}
-\label{eq:lasso}
+\label{eq:LASSO}
 \displaystyle \hat \beta_\lambda = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The \emph{group lasso estimator}, a generalization of the lasso,
+The \emph{group LASSO estimator}, a generalization of the LASSO,
  is a solution to the following penalized
@@
 $w_g$ are usually taken to be $\sqrt {p_g}$ to normalize the penalty
-across groups with different sizes.  Note that this includes the usual lasso estimator as a
-special case when all of the groups are of size 1, since then the
+across groups with different sizes.  This can also be accomplished
+by scaling the corresponding submatrices by their Frobenius norms
+and setting all weights equal.
+Note that this includes the usual LASSO estimator as a
+special case when all groups are of size 1, since then the
 penalty term is the $L^1$-norm of $\beta$. The solution path of the group
-lasso has similar properties to the lasso case, however it is not
+LASSO has similar properties to the LASSO case, however it is not
 generally piecewise linear.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-For the LASSO,
-
-this is explicitly computable by
+For the group LASSO this is explicitly computable by
 \begin{equation}
@@
 where $g^*$ is the group that achieves the maximum and 
-$\eta_{g^*} = X_{g^*}^Ty / \|X_{g^*}^Ty\|_2$ is the unit vector that achieves the norm $\|X_{g^*}^Ty\|_2$.
+$\eta_{g^*} = X_{g^*}^Ty / \|X_{g^*}^Ty\|_2$ is the unit vector that achieves the norm $\|X_{g^*}^Ty\|_2$. One of the key conceptual points about our
+test is that it conditions on both the maximizer $g^*$ and on the unit
+vector $\eta_{g^*}$. Most of the p-value computations will be done in
+terms of these quantities.
 
@@
 % \end{center}
-% A polytope selection event. Solving the LASSO \eqref{eq:lasso} for a fixed value
+% A polytope selection event. Solving the LASSO \eqref{eq:LASSO} for a fixed value
 % of $\lambda$ yields both an active set $\hat{E}(y)$ and a collection of signs of 
@@
 
-The approach to post-selection inference in \cite{lasso:fixed}
+Before giving the full description of the p-value, we turn to another
+previous work and describe how to extend the framework there to the
+setting with groups.
+The approach to post-selection inference in \cite{LASSO:fixed}
 describes the selection event for the LASSO as a convex polytope
@@
 %stylized in Figure \ref{fig:polytope}.
-If all of our groups were of size 1, then
+If all groups are of size 1, then
 the event that we observe a given set of variables
@@
 \end{equation}
-See \cite{lasso:fixed,exact:lars} for further details on this approach.
-
-In the grouped variables setting, however, the event that $g^*(y)$, the first
+See \cite{LASSO:fixed,exact:lars} for further details on this approach.
+
+With non-trivial groups, however, the event that $g^*(y)$, the first
 group chosen, is equal to some fixed group $g$ is generally not given by
@@
 \end{equation}
-
 The quadratic forms are not assumed non-negative definite, but we can, without loss of generality
 assume they are symmetric.
-As in \cite{lasso:fixed} we will be interested in some randomly chosen contrast
-$\eta(S(y))^T\mu$. In our grouped selection procedure, the procedure $S(y)=g^*(y)$
+As in \cite{LASSO:fixed} we will be interested in some randomly chosen contrast
+$\eta(S(y))^T\mu$.
+In our grouped selection procedure, the selection rule is $S(y)=g^*(y)$
 and $\eta$ is typically 
@@
 Conditional on $H_{0,s} \cap E_+(s, \eta_s, P_s^{\perp}y, (P_s - \eta_s\eta_s^T)y)$
-we have the following distributional result
+we have the following truncated $\chi$ distributional result
 \begin{equation}
@@
 $$
+and the vertical bar $|$ here denotes restriction to an interval.
 \end{lemma}
@@
 %
-%This is Lemma 5 from \cite{significance:lasso}.
+%This is Lemma 5 from \cite{significance:LASSO}.
 %
@@
 %
-%\subsection{Exact p-value calculation}
-%
-%\todo Complete this section
-%
-%We now describe how to calculate our p-value, following the discussion of
-%group lasso in \cite{tests:adaptive}. For the rest of this section let
-%$g = \gstar$ be the index of
-%the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
-%%\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
-%%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
-%For a vector $u \in \real^p$, let $u_h$ denote the coordinates
-%of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
-%We can rearrange the columns of $X$ to group these adjacently, so that
-%\[
-%u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
-%\]
-%One step of the calculation will be to find an orthonormal basis for the
-%linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
-%\text{ for all } h \neq g \}$ so that we can project orthogonally to
-%this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
-%should be the case since $g$ maximizes the absolute value of this
-%quantity), then the
-%space is trivial and the desired orthogonal projection is the identity.
-%Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
-%generally has dimension $p_g-1$.
-%% have dimension $r-1$ for $r = \text{rank}(X_g)$. This holds except in
-%% degenerate cases such as $X_g = 0$ or $r \ll p_g$, but these cases are not
-%% likely to occur since $\norm{X_g^T y}_2/w_g$ is supposed to be maximized.
-%% We proceed by forming the $n \times p_g$ matrix
-%% $X_g - X_g X_g^T y y^T X_g/\norm{X_g^Ty}_2^2$ and performing Gram-Schmidt
-%% on the first $r-1$ columns. For precise theorems showing that this will
-%% generally work, see \cite{outerp:indep}.
-%We compute an orthonormal basis by Gram-Schmidt and form a
-%$p_g \times p_g$ matrix, which we denote $V_g$, by appending one column
-%%0's in the coordinates corresponding to all groups $h \neq g$ and to
-%of zeroes (since Gram-Schmidt only produces $p_g-1$ columns).
-%With this we are ready to define the projection
-%
-%\begin{equation}
-%  \begin{aligned}
-%    \label{eq:proj}
-%    P_g &= \Sigma X_gV_g (V_g^T X_g^T \Sigma X_g V_g)^\dagger V_g^TX_g^T \\
-%   \end{aligned}
-%\end{equation}
-%Also define $H_g = (I-P_g)\Sigma$ and the conditional variance
-%
-%\begin{equation}
-%  \begin{aligned}
-%    \label{eq:cvar}
-%    \sigma^2_g &= y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2) \\
-%  \end{aligned}
-%\end{equation}
-%These simplify when $X_g$ is a single column, in which case $P_g = 0$,
-%$H_g = \Sigma$, and $\sigma^2_g = X_g^T\Sigma X_g^T/w_g^2$. We now
-%describe how to compute our p-value in Algorithm~\ref{algo:pval}. Note
-%that $F_{\chi^2_r}$ denotes the distribution function of a $\chi^2$
-%random variable with $r$ degrees of freedom.
-%%Define $U_g = X_g^Ty/(w_g\norm{X_g^Ty}_2)$.
-%
-%\begin{algorithm}
-%  \caption{Computing p-value}
-%  \label{algo:pval}
-%  \begin{algorithmic}
-%    \REQUIRE Response $y$, grouped design matrix $X$ with weights, inactive set $A^c$, index $g$ of last group to enter active set $A$.
-%    \ENSURE Our p-value for the group $g$ entering the model.
-%    \STATE Compute $H_g$ and $\sigma^2_g$
-%    \IF{$X_g$ is a single column}
-%    \STATE $\sigma_g^2 \gets X_g^T\Sigma X_g/w_g^2$
-%    \STATE $\tilde X_g \gets X_g / w_g \cdot \text{sign}(X_g^Ty)$
-%    \ELSE
-%    \STATE $P_g \gets \Sigma XV_g (V_g^T X^T \Sigma X V_g)^\dagger V_g^TX^T$
-%    \STATE $H_g \gets (I - P_g)\Sigma$
-%    \STATE $\sigma^2_g \gets y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2)$
-%    \STATE $\tilde X_g \gets H_g X_g X_g^T y /(\norm{X_g^Ty}_2 w_g)$
-%    \ENDIF
-%    \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$
-%%    \STATE \COMMENT{Compute the following two $p$-vectors}
-%    \STATE $a \gets X^T(y - \lambda \tilde X_g)$
-%    \STATE $b \gets X^T \tilde X_g$
-%    \STATE Compute solution $(v_-, v_+)$ of \textsc{LinearFractional}(a,b) optimization subproblem
-%    \STATE $r \gets \text{rank}(X_g)$
-%    \STATE $u \gets [F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(\lambda^2/\sigma_g^2)]/[F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(v^2_+/\sigma_g^2)]$
-%    \RETURN $u$
-%  \end{algorithmic}
-%\end{algorithm}
-%
-%We next describe the \textsc{LinearFractional} subproblem and its solution. We
-%gave the problem this name because it originated in the form
-%\[
-%\maximize_{h \neq g, \norm{u_h}_2 = 1} \frac{u_h^TX_h^Ty - u_h^TX_h^TX_gX_g^Ty}
-%{1 - u_g^TX_g^TX_hu_h}
-%\]
-%The solution we describe next is to a slightly different problem, and
-%also incorporates the information that $g$ maximizes $\norm{X_g^Ty}_2/w_g$.
-%For details see Appendix~\ref{app:closedform} or \cite{tests:adaptive}.
-%
-%\begin{algorithm}
-%  \caption{The \textsc{LinearFractional} optimization subproblem}
-%  \label{algo:linfrac}
-%  \begin{algorithmic}
-%    \REQUIRE The $p$-vectors $a, b$ in Algorithm~\ref{algo:pval}, weights, inactive set $A^c$, a small tolerance number
-%%(we take $1e^{-10}$)
-%    \ENSURE Solution pair $(v_-, v_+)$.
-%    \FOR{$h$ in $A^c$}
-%      \IF{$\norm{b_h}_2 == 0$ or $\norm{a_h}_2/\norm{b_h}_2 < \text{tol}$}
-%        \STATE $(v^-_h, v^+_h) \gets (0, \infty)$
-%      \ELSE
-%        \STATE $\theta_c \gets a_h^Tb_h/(\norm{a_h}_2 \norm{b_h}_2)$
-%        \STATE $\theta_s \gets \sqrt{1-\theta_c^2}$
-%        \STATE $\theta \gets \arccos (\theta_c)$
-%        \STATE $\phi_s \gets \theta_s \norm{b_h}_2 / w_h$
-%        \IF{$\phi_s > 1$}
-%          \STATE $(v^-_h, v^+_h) \gets (0, \infty)$
-%        \ELSE
-%          \STATE $\phi \gets \arcsin(\phi_s)$
-%%          \STATE $\phi_2 \gets \pi - \phi_1$
-%          \STATE $z_\pm \gets s_\pm \norm{a_h}_2 \cos(\phi) /(w_h - s_\pm \norm{b_h}_2 \cos(\theta - \phi))$ for $s_\pm = \pm 1$
-%          \IF{$\norm{b_h}_2 < w_h$}
-%            \STATE $(v_h^-, v_h^+) \gets (\max \{ z_+, z_- \}, \infty)$
-%          \ELSE
-%            \STATE $(v_h^-, v_h^+) \gets (\min \{ z_+, z_- \}, \max \{ z_+, z_- \})$
-%          \ENDIF
-%        \ENDIF
-%      \ENDIF
-%    \ENDFOR
-%    \STATE $v^- \gets \min_h v_h^-$
-%    \STATE $v^+ \gets \max_h v_h^+$
-%    \RETURN $(v^-, v^+)$
-%  \end{algorithmic}
-%\end{algorithm}
-%
+
+\subsection{Computation of p-value}
+
+\begin{algorithm}
+ \caption{Computing p-value}
+ \label{algo:pval}
+ \begin{algorithmic}
+   \REQUIRE Response $y$, grouped design matrix $X$ with weights, inactive set $A^c$, index $g$ of last group to enter active set $A$.
+   \ENSURE Our p-value for the group $g$ entering the model.
+   \STATE Compute $H_g$ and $\sigma^2_g$
+   \IF{$X_g$ is a single column}
+   \STATE $\sigma_g^2 \gets X_g^T\Sigma X_g/w_g^2$
+   \STATE $\tilde X_g \gets X_g / w_g \cdot \text{sign}(X_g^Ty)$
+   \ELSE
+   \STATE $P_g \gets \Sigma XV_g (V_g^T X^T \Sigma X V_g)^\dagger V_g^TX^T$
+   \STATE $H_g \gets (I - P_g)\Sigma$
+   \STATE $\sigma^2_g \gets y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2)$
+   \STATE $\tilde X_g \gets H_g X_g X_g^T y /(\norm{X_g^Ty}_2 w_g)$
+   \ENDIF
+   \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$
+   \STATE \COMMENT{Compute the following two $p$-vectors}
+   \STATE $a \gets X^T(y - \lambda \tilde X_g)$
+   \STATE $b \gets X^T \tilde X_g$
+   \STATE Compute solution $(v_-, v_+)$ of \textsc{LinearFractional}(a,b) optimization subproblem
+   \STATE $r \gets \text{rank}(X_g)$
+   \STATE $u \gets [F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(\lambda^2/\sigma_g^2)]/[F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(v^2_+/\sigma_g^2)]$
+   \RETURN $u$
+ \end{algorithmic}
+\end{algorithm}
+
+We now describe how to calculate our p-value, following the discussion of
+group LASSO in \cite{tests:adaptive}. For the rest of this section let
+$g = \gstar$ be the index of
+the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
+%$\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
+%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
+For a vector $u \in \real^p$, let $u_h$ denote the coordinates
+of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
+We can rearrange the columns of $X$ to group these adjacently, so that
+\[
+u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
+\]
+One step of the calculation will be to find an orthonormal basis for the
+linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
+\text{ for all } h \neq g \}$ so that we can project orthogonally to
+this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
+should be the case since $g$ maximizes the absolute value of this
+quantity), then the
+space is trivial and the desired orthogonal projection is the identity.
+Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
+generally has dimension $p_g-1$.
+
+%We proceed by forming the $n \times p_g$ matrix
+%$X_g - X_g X_g^T y y^T X_g/\norm{X_g^Ty}_2^2$ and performing Gram-Schmidt
+%on the first $r-1$ columns. For precise theorems showing that this will
+%generally work, see \cite{outerp:indep}.
+We compute an orthonormal basis by Gram-Schmidt and form a
+$p_g \times p_g$ matrix, which we denote $V_g$, by appending
+0's in the coordinates corresponding to all groups $h \neq g$ and
+an additional column of zeroes (since Gram-Schmidt only produces $p_g-1$).
+We can now define the projection
+
+\begin{equation}
+ \begin{aligned}
+   \label{eq:proj}
+   P_g &= \Sigma X_gV_g (V_g^T X_g^T \Sigma X_g V_g)^\dagger V_g^TX_g^T. \\
+  \end{aligned}
+\end{equation}
+Also define $H_g = (I-P_g)\Sigma$ and the conditional variance
+
+\begin{equation}
+ \begin{aligned}
+   \label{eq:cvar}
+   \sigma^2_g &= y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2) \\
+ \end{aligned}
+\end{equation}
+These simplify when $X_g$ is a single column, in which case $P_g = 0$,
+$H_g = \Sigma$, and $\sigma^2_g = X_g^T\Sigma X_g^T/w_g^2$.
+Note that $F_{\chi^2_r}$ denotes the distribution function of a $\chi^2$
+random variable with $r$ degrees of freedom. This completes the
+description of Algorithm~\ref{algo:pval} which we use to compute our
+p-value. Additional details, including the statement and solution of
+the \textsc{LinearFractional} subproblem, appear in Appendix~\ref{app:closedform}.
 
@@
 nonzero coefficients on the scale of $\gamma := \sqrt{2 \log(p)/n}$.
-(\todo Check this scaling. \todo Cite some paper on support recovery?).
+(\todo Cite some paper on support recovery?).
 Coefficients within a
@@
 parentheses if the groups are not all size 1. The number of nonzero
-groups, $k$, is also displayed via inward-facing tick marks on the
-horizontal axes. The proportion of truly
+groups, $k$, is also displayed graphically by shading with the portion
+of the plot showing steps after $k$ shaded gray.
+The proportion of truly
 nonzero groups recovered in the first $k$ steps is denoted $k$-SP,
@@
 \label{fig:gausscorr}
-\includegraphics[width=0.5\textwidth]{../figs/gaussian_n100_p200_g200_k5_lower1pt5_upper2_corr0pt2_noisecorr0pt1.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/default_gaussian_nsim400_n100_p200_g200_k10_lower1pt5_upper2_corr0pt2.pdf}}
 \hspace{-15pt}
@@
 \label{fig:gaussgroups}
-\includegraphics[width=0.5\textwidth]{../figs/gaussian_n100_p200_g35_k5_lower2_upper2pt5.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/default_gaussian_nsim500_n100_p200_g35_sizes5-10_k5_lower2_upper3.pdf}}
 \caption{\small \it The left panel shows results for Gaussian design
@@
 \label{fig:gausscat}
-\includegraphics[width=0.5\textwidth]{../figs/categorical_n100_p100_g20_k4_lower4_upper5.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/default_categorical_nsim500_n100_p80_g40_sizes2-2_k5_lower2_upper4.pdf}}
 \caption{\small \it The left panel shows results for a categorical design
@@
 \begin{center}
-\subfigure[Forward stepwise with splines]{
+\subfigure[Power of forward stepwise with splines]{
 \label{fig:spline:fwd}
@@
 \hspace{-15pt}
-\subfigure[P-value with splines]{
+\subfigure[P-value with spline groups]{
 \label{fig:spline:pval}
@@
 \caption{\small \it The left panel shows the power of forward stepwise
-for various sparsity levels. The right panel shows our p-value marginally
-over each step, with red indicating the global null case and blue
-indicating a signal with $k=6$ nonzero groups.}
+in this setting for various sparsity levels, with the dashed line
+showing a liberal definition of power that allows mistakes when
+choosing between a non-zero spline group and its corresponding linear
+term.
+The right panel shows our p-value marginally
+over each step, with red indicating the global null case, blue
+indicating a signal with $k=6$ nonzero groups, and green showing
+the nominal $\chi^2$ p-values also computed on the non-null case.}
 \label{fig:spline}
@@
 Note that the varying group sizes present some difficulty to forward stepwise, which tends to first try linear approximations to nonzero spline groups before adding the true spline group.
+In Figure~\ref{fig:spline}(b) the average proportion of truly nonzero spline groups included in the first $k$ steps appears in parentheses next to the $k$-SP (top right).
 
@@
 They accomplish this by creating a design matrix with both main effects alone and also with groups including main effects with their first order interactions.
-Then they fit a group lasso model with the expanded design matrix.
+Then they fit a group LASSO model with the expanded design matrix.
 Because interaction terms only appear in groups along with their respective main effects, the hierarchy condition holds for the fitted model.
@@
 \end{equation}
-where $X_{g:h}$ is the submatrix including $X_g$, $X_h$, and all
+where $X_{g:h}$ is the submatrix encoding the interaction between $X_g$
+and $X_h$. For example, if both of these are categorical variables,
+then $X_{g:h}$ consists of all
 $p_gp_h$  column multiples between columns in group $g$ and columns
@@
 \]
-are two groups each containing two columns, then
+are two categorical variables each with two levels, then
 \[
-X_{g:h} = \begin{pmatrix} X_g & X_h & X_{g1} * X_{h1} & X_{g1} * X_{h2} & X_{g1} * X_{h1} & X_{g2} * X_{h2} \end{pmatrix}
+X_{g:h} = \begin{pmatrix} X_{g1} * X_{h1} & X_{g1} * X_{h2} & X_{g1} * X_{h1} & X_{g2} * X_{h2} \end{pmatrix}
 \]
@@
 (the $i$th entry of $X_{g1} * X_{h1}$ is the $i$th entry of $X_{g1}$ times
-the $i$th entry of $X_{h1}$). Note that this expanded matrix has
+the $i$th entry of $X_{h1}$). For more details see \cite{glint}.
+Note that this expanded matrix has
 $\tilde G = G + \binom{G}{2} = O(G^2)$ groups. We refer to the first
-$G$ of these as main effects or main effect groups, and the remaining
+$G$ of these as main effect groups and the remaining
 as interaction groups. Finally, instead of fitting a
-model by group lasso, we use forward stepwise on the expanded design
+model by group LASSO, we use forward stepwise on the expanded design
 matrix. The overlapping groups still guarantee that our fitted model
@@
 but our purpose here is merely to demonstrate the flexibility of the
-hypothesis test and not to propose a general procedure for models with
+hypothesis test and not to propose an optimal procedure for models with
 interactions.
@@
 using the group structure we impose, and not in terms of the original
-main effects and interactions. However, the dotted line shows a more
+main effects and interactions. However, the dashed line shows a more
 forgiving definition of power where we are rewarded for discovering part
 of a nonzero group, i.e. for discovering only one main effect from a
-true interaction group. In the right panel we see that the global null
-result still holds (red vertical lines), and that with a signal the
-p-value is small until the signal has been recovered (blue). The
-interaction power is the proportion of nonzero interaction groups
-that were discovered and is labeled ``Int. Power.''
+true interaction group. The proportion of nonzero interaction groups
+that were discovered in the first $k$-steps is shown in parentheses
+after the $k$-SP (top right).
 
@@
 \begin{center}
-\subfigure[Forward stepwise on glinternet]{
+\subfigure[Power of forward stepwise with interactions]{
 \label{fig:glint:fwd}
@@
 \hspace{-15pt}
-\subfigure[P-value with glinternet]{
+\subfigure[P-value with \textsc{glinternet}]{
 \label{fig:glint:pval}
@@
 \caption{\small \it The left panel shows the power of forward stepwise
-for various sparsity levels. The right panel shows our p-value marginally
-over each step, with red indicating the global null case and blue
-indicating a signal with $k=6$ nonzero groups.}
+on the \textsc{glinternet} problem
+for various sparsity levels. The dashed line shows a more forgiving
+definition of power described in the section.
+The right panel shows our p-value marginally
+over each step, with red indicating the global null case, blue
+indicating a signal with $k=6$ nonzero groups, and green showing
+the nominal $\chi^2$ p-values.}
 \label{fig:glint}
@@
 \label{sec:hiv}
-\todo Sanity check my analysis \\
-\todo For rare mutations try binary wildtype/non-wt? \\
-\todo Finish writing section \\
 
@@
 for each drug response, restricting to the subset of the data with
-no missing values and requiring all categorical variables to have 2 or
-greater observations in each level.
+no missing values.
 
@@
 \includegraphics[width=0.9\textwidth]{../figs/HIV_PI.pdf}
-\caption{\small \it Results from PI dataset}
+\caption{\small \it Forward stepwise results from PI dataset}
 \label{fig:HIVPI}
@@
 \includegraphics[width=0.9\textwidth]{../figs/HIV_NRTI.pdf}
-\caption{\small \it Results from NRTI dataset}
+\caption{\small \it Forward stepwise results from NRTI dataset}
 \label{fig:HIVNRTI}
@@
 Each panel shows results for a different drug,
-with p-values plotted by step in forward stepwise. The solid black point
-is the last p-value to fall under the 0.1 threshold and points after it
-are colored red. If the \textit{last} stopping rule were used then the
-p-values for including variables are the black points, and the resulting
-models are shown in Tables 4 and 5. (\todo Fix this ref after re-running)
+with p-values plotted by step in forward stepwise. 
+The \textit{last} stopping rule is applied and the region of the plot
+corresponding to the chosen model is unshaded, with the remainder shaded.
+The chosen variables are shown in Tables 3 and 4.
+(\todo Fix this ref after final re-run of code)
 
@@
 
-We can also use the procedure described in Section~\ref{sec:glint} to fit a model with pairwise interactions between mutations.
+We can also use the \textsc{Glinternet} procedure described in Section~\ref{sec:glint} to fit a model with pairwise interactions.
+Results from this are shown in Figures~\ref{fig:HIVPI:glint} and \ref{fig:HIVNRTI:glint} and Tables 5 and 6.
 
@@
 \includegraphics[width=0.9\textwidth]{../figs/HIV_PI_glint.pdf}
-\caption{\small \it Results from PI dataset}
+\caption{\small \it \textsc{Glinternet} results from PI dataset}
 \label{fig:HIVPI:glint}
@@
 \includegraphics[width=0.9\textwidth]{../figs/HIV_NRTI_glint.pdf}
-\caption{\small \it Results from NRTI dataset}
+\caption{\small \it \textsc{Glinternet} results from NRTI dataset}
 \label{fig:HIVNRTI:glint}
@@
 \appendix
-\section{Derivation of closed forms of statistic}
+\section{Additional details on closed-form computations}
 \label{app:closedform}
 
-\todo Clean this section \\
-\todo Make $M=\lambda_2$ a special case corollary \\
+We next describe the \textsc{LinearFractional} subproblem and its solution. We
+gave the problem this name because it originated in the form
+\[
+\maximize_{h \neq g, \norm{u_h}_2 = 1} \frac{u_h^TX_h^Ty - u_h^TX_h^TX_gX_g^Ty}
+{1 - u_g^TX_g^TX_hu_h}
+\]
+The solution we describe next is to a slightly different problem, and
+also incorporates the information that $g$ maximizes $\norm{X_g^Ty}_2/w_g$.
+%For details see \cite{tests:adaptive}.
+
+\begin{algorithm}
+ \caption{The \textsc{LinearFractional} optimization subproblem}
+ \label{algo:linfrac}
+ \begin{algorithmic}
+   \REQUIRE The $p$-vectors $a, b$ in Algorithm~\ref{algo:pval}, weights, inactive set $A^c$, a small tolerance number
+(we take $1e^{-10}$)
+   \ENSURE Solution pair $(v_-, v_+)$.
+   \FOR{$h$ in $A^c$}
+     \IF{$\norm{b_h}_2 == 0$ or $\norm{a_h}_2/\norm{b_h}_2 < \text{tol}$}
+       \STATE $(v^-_h, v^+_h) \gets (0, \infty)$
+     \ELSE
+       \STATE $\theta_c \gets a_h^Tb_h/(\norm{a_h}_2 \norm{b_h}_2)$
+       \STATE $\theta_s \gets \sqrt{1-\theta_c^2}$
+       \STATE $\theta \gets \arccos (\theta_c)$
+       \STATE $\phi_s \gets \theta_s \norm{b_h}_2 / w_h$
+       \IF{$\phi_s > 1$}
+         \STATE $(v^-_h, v^+_h) \gets (0, \infty)$
+       \ELSE
+         \STATE $\phi \gets \arcsin(\phi_s)$
+         \STATE $\phi_2 \gets \pi - \phi_1$
+         \STATE $z_\pm \gets s_\pm \norm{a_h}_2 \cos(\phi) /(w_h - s_\pm \norm{b_h}_2 \cos(\theta - \phi))$ for $s_\pm = \pm 1$
+         \IF{$\norm{b_h}_2 < w_h$}
+           \STATE $(v_h^-, v_h^+) \gets (\max \{ z_+, z_- \}, \infty)$
+         \ELSE
+           \STATE $(v_h^-, v_h^+) \gets (\min \{ z_+, z_- \}, \max \{ z_+, z_- \})$
+         \ENDIF
+       \ENDIF
+     \ENDIF
+   \ENDFOR
+   \STATE $v^- \gets \min_h v_h^-$
+   \STATE $v^+ \gets \max_h v_h^+$
+   \RETURN $(v^-, v^+)$
+ \end{algorithmic}
+\end{algorithm}
+
+% \todo Clean this section \\
+% \todo Make $M=\lambda_2$ a special case corollary \\
 
@@
 
-I haven't been able to algebraically simplify these yet to a form that allows my $\lambda_2$ proof below to go through. The version below just assumes $\norm{b_h} < w = 1$.
-
-\hrule
-\hrule
-We can rewrite this by rationalizing the denominator:
-\[
-M =
-\frac
-{ a_h^T b_h +  \sqrt{ a_h^T a_h (1-b_h^Tb_h) + (a_h^T b_h)^2 }  }
-{ 1-b_h^Tb_h }
-\]
-Leaving $M$ in this form we now consider $\lambda_2$. The KKT conditions for the group lasso problem give (directly from Yuan and Lin)
-\begin{align*}
-\norm{ X_g^T(y - X \beta) } & \leq \lambda w_g \quad \forall \beta_g \neq 0 \\
-\beta_g & = \left( 1 - \frac{ \lambda w_g }{ \norm{S_g} } \right)_+ S_g
-\end{align*}
-where $S_g = X_g^T(y - X \beta_{-g})$. These expressions simplify for $\lambda_2 \leq \lambda < \lambda_1$. Letting $g$ be the index of the first group to enter we now have $S_g = X_g^Ty$ 
-\begin{align*}
-\beta_g & = \left( 1 - \frac{ \lambda w_g }{ \norm{X_g^Ty} } \right) X_g^Ty \\
-\lambda w_h & \geq \norm{ X_h^Ty - X_h^T P_g y (1-\lambda w_g / \norm{X_g^Ty})  } \quad \forall h \neq g
-\end{align*}
-Now let $\lambda = \lambda_2$, so the inequality above is strict for all other groups and equality is attained for the second group to enter. Hence $\lambda_2$ satisfies
-\begin{align*}
-\lambda_2 & = \max_{h \neq g} 
-\norm{ X_h^T(I - P_g) y - \lambda_2 ( w_g X_h^T P_g y / \norm{X_g^Ty} )  } / w_h\\
-& = \max_{h \neq g} \norm{ a_h - \lambda_2 b_h } 
-\end{align*}
-
-Write $\lambda_{2,h}$ as the positive root of the quadratic equation obtained by squaring the norm, so $\lambda_{2,h}^2 = \norm{ a_h - \lambda_2 b_h }^2$.
+% I haven't been able to algebraically simplify these yet to a form that allows my $\lambda_2$ proof below to go through. The version below just assumes $\norm{b_h} < w = 1$.
+
+% \hrule
+% \hrule
+% We can rewrite this by rationalizing the denominator:
+% \[
+% M =
+% \frac
+% { a_h^T b_h +  \sqrt{ a_h^T a_h (1-b_h^Tb_h) + (a_h^T b_h)^2 }  }
+% { 1-b_h^Tb_h }
+% \]
+% Leaving $M$ in this form we now consider $\lambda_2$. The KKT conditions for the group LASSO problem give (directly from Yuan and Lin)
+% \begin{align*}
+% \norm{ X_g^T(y - X \beta) } & \leq \lambda w_g \quad \forall \beta_g \neq 0 \\
+% \beta_g & = \left( 1 - \frac{ \lambda w_g }{ \norm{S_g} } \right)_+ S_g
+% \end{align*}
+% where $S_g = X_g^T(y - X \beta_{-g})$. These expressions simplify for $\lambda_2 \leq \lambda < \lambda_1$. Letting $g$ be the index of the first group to enter we now have $S_g = X_g^Ty$ 
+% \begin{align*}
+% \beta_g & = \left( 1 - \frac{ \lambda w_g }{ \norm{X_g^Ty} } \right) X_g^Ty \\
+% \lambda w_h & \geq \norm{ X_h^Ty - X_h^T P_g y (1-\lambda w_g / \norm{X_g^Ty})  } \quad \forall h \neq g
+% \end{align*}
+% Now let $\lambda = \lambda_2$, so the inequality above is strict for all other groups and equality is attained for the second group to enter. Hence $\lambda_2$ satisfies
+% \begin{align*}
+% \lambda_2 & = \max_{h \neq g} 
+% \norm{ X_h^T(I - P_g) y - \lambda_2 ( w_g X_h^T P_g y / \norm{X_g^Ty} )  } / w_h\\
+% & = \max_{h \neq g} \norm{ a_h - \lambda_2 b_h } 
+% \end{align*}
+
+% Write $\lambda_{2,h}$ as the positive root of the quadratic equation obtained by squaring the norm, so $\lambda_{2,h}^2 = \norm{ a_h - \lambda_2 b_h }^2$.
 Then $\lambda_2 = \max_{h \neq g} \lambda_{2,h}$.
 Solving this (note that $b_h^Tb_h < 1$), we find
-\[
-\lambda_{2,h} = \frac
-{ a_h^T b_h +  \sqrt{ a_h^T a_h (1-b_h^Tb_h) + (a_h^T b_h)^2 }  }
-{ 1-b_h^Tb_h }
-\]
-Hence when $h$ is the index of the second group to enter, we have $\lambda_2 = \lambda_{2,h} = M$.
+% \[
+% \lambda_{2,h} = \frac
+% { a_h^T b_h +  \sqrt{ a_h^T a_h (1-b_h^Tb_h) + (a_h^T b_h)^2 }  }
+% { 1-b_h^Tb_h }
+% \]
+% Hence when $h$ is the index of the second group to enter, we have $\lambda_2 = \lambda_{2,h} = M$.
 

### 2014-05-06 HIV/glint_HIV.R

@@ -3,7 +3,7 @@ db = "NRTI"
 
 binary.encoding = TRUE
 
-max.steps = 50
+max.steps = 40
 
 setwd('..')
 source('fwd_step.R')
@@ -32,9 +32,9 @@ clean_col = function(column) {
 }
 
 cnames = c()
-for (j in max(resps):ncol(data)) {
+for (j in (max(resps)+1):ncol(data)) {
     newcol = clean_col(data[,j])
-    if ((length(levels(newcol)) > 1) & (min(table(newcol)) >= 2)) {
+    if (length(unique(newcol)) > 1) {
         cnames = c(cnames, names(data)[j])
         Z = cbind(Z, newcol)
     }
@@ -57,14 +57,14 @@ for (respind in 1:nresps) {
     Y = log(Z[subrows,respind])
     Y = Y - mean(Y)
     n = length(Y)
-    # Now restrict to gene variants with >= 2 observations
     X = matrix(NA, nrow=n)
+    X.cat = X
     groups = c()
     g = 1
     cnames = c()
     for (j in (nresps+1):ncol(Z)) {
         newcol = Z[subrows, j]
-        if ((length(levels(newcol)) > 1) & (min(table(newcol)) >= 2)) {
+        if (length(unique(newcol)) > 1) {
             cnames = c(cnames, names(Z)[j])
             submat = model.matrix(~ Z[subrows,j] - 1)
             if (binary.encoding) {
@@ -75,14 +75,20 @@ for (respind in 1:nresps) {
             } else {
                 colnames(submat) = paste0(names(Z)[j], levels(Z[,j]))
             }
-            if (min(colSums(submat)) > 2) {
-                groups = c(groups, rep(g, ncol(submat)))
-                g = g + 1
-                X = cbind(X, submat)
-            }
+            groups = c(groups, rep(g, ncol(submat)))
+            g = g + 1
+            X = cbind(X, submat)
+            X.cat = cbind.data.frame(X.cat, as.factor(newcol))
         }
     }
     X = X[ , -1]
+    X.cat = X.cat[, -1]
+    colnames(X.cat) = cnames
+    full.model = lm(Y ~ . - 1, data=data.frame(X.cat), na.action = na.omit)
+    s2 = sum(full.model$residuals^2)
+    s2 = s2/full.model$df.residual
+    Sigma = diag(rep(s2,n))
+    
[…]

### 2014-05-06 HIV/sim_HIV.R

@@ -1,30 +1,36 @@
+#db = "NRTI"
+db = "PI"
+
+binary.encoding = TRUE
+
 setwd('..')
 source('fwd_step.R')
 source('generate_data.R')
-source('fwd_step_sim.R')
+source('simulation.R')
 source('tex_table.R')
 source('plots.R')
 library(plyr)
 
-## PI = read.table("http://hivdb.stanford.edu/pages/published_analysis/genophenoPNAS2006/DATA/NRTI_DATA.txt", sep='\t', header=TRUE)
-## db="PI"
-## data=PI
-## resps = 4:9
-NRTI = read.table("http://hivdb.stanford.edu/pages/published_analysis/genophenoPNAS2006/DATA/PI_DATA.txt", sep = "\t", header = TRUE)
-data=NRTI
-db="NRTI"
-resps = 4:10
+if (db == "NRTI") {
+  data = read.table("http://hivdb.stanford.edu/pages/published_analysis/genophenoPNAS2006/DATA/NRTI_DATA.txt", sep='\t', header=TRUE)
+  resps = 4:9
+} else {
+  data = read.table("http://hivdb.stanford.edu/pages/published_analysis/genophenoPNAS2006/DATA/PI_DATA.txt", sep = "\t", header = TRUE)
+  resps = 4:10
+}
 
-design = paste0("HIV_", db)
-fn.append = 'Pg'
+nsim = 10
+type = "default"
+design = "fixed"
+fn.append = '' #'Pg'
 corr = 0 # nonzero only supported for gaussian design
 noisecorr = 0
-nsim = 200
-num.nonzero = 4
+nsim = 10
+num.nonzero = 5
 k = num.nonzero
-max.steps = 10
-upper.coeff = 25
-lower.coeff = 21
+max.steps = 15
+upper = 10
+lower = 5
 
 # Clean data, restrict to cleaned subset
 nresps = length(resps)
@@ -37,25 +43,32 @@ clean_col = function(column) {
 }
 
 cnames = c()
-for (j in max(resps):ncol(data)) {
-      newcol = clean_col(data[,j])
-          if ((length(levels(newcol)) > 1) & (min(table(newcol)) >= 2)) {
-            newcol = mapvalues(newcol, from=levels(newcol), to=1:length(levels(newcol)))
-                    cnames = c(cnames, names(data)[j])
-                            Z = cbind(Z, newcol)
-                  }
-    }
[…]

### 2014-05-06 fwd_step.R

@@ -137,6 +137,7 @@ forward_group = function(X, Y, groups, weights = 0, Sigma = NULL, max.steps = 0,
   chi.pvals = c()
   c.vars = c()
   Ls = c()
+  ranks = c()
   Ynorms = c()
 
   Y.update = Y
@@ -153,6 +154,7 @@ forward_group = function(X, Y, groups, weights = 0, Sigma = NULL, max.steps = 0,
       active.set = output$active.set
       imax = output$added
       grank = output$grank
+      ranks = c(ranks, grank)
       eff.p = output$eff.p
       RSS = sum(Y.update^2)
       Y.update = output$Y.update

### 2014-05-06 group_lasso.R

@@ -24,6 +24,7 @@ trignometric_form = function(num, den, weight, tol=1.e-10) {
   
   Ctheta = sum(a*b) / (norma*normb)
   # R might believe 1 > 1 and return NaN in sqrt() below
+  # so we truncate to the interval first
   Ctheta = min(max(Ctheta, -1), 1)
   Stheta = sqrt(1-Ctheta^2)
   theta = acos(Ctheta)
@@ -71,9 +72,10 @@ group_lasso_knot <- function(X, Y, groups, weights, Sigma, active.set=0) {
   Uwhich = U[which]
   Xmax = X[,which]
   kmax = sum(which)
-  kmaxrank = kmax
   if (length(dim(Xmax)) == 2) {
     kmaxrank = rankMatrix(Xmax)[1]
+  } else {
+    kmaxrank = 1
   }
 
   eta = rep(0, p)
@@ -114,6 +116,7 @@ group_lasso_knot <- function(X, Y, groups, weights, Sigma, active.set=0) {
     conditional_variance = sum(Xeta^2)
   }
   # use formula above display (42) in tests:adaptive
+  
   Xeta = Xeta / conditional_variance
   C_X = t(X) %*% Xeta
 

### 2014-05-06 simulation.R

@@ -138,7 +138,7 @@ run_simulation = function(
             all.groups = groups
             b.data = beta_staircase(groups, k, upper.scaled, lower.scaled, cat.groups=cat.groups, rand.sign=TRUE, perturb=TRUE)
             group.sizes = rle(groups)$lengths
-            weights = sqrt(group.sizes)
+            #weights = sqrt(group.sizes)
             
         } else {
             # Nomenclature: "special" means splines or interactions
@@ -168,20 +168,21 @@ run_simulation = function(
                 X.test = data.test$X
             }
 
-            # Frobenius normalized, do not need weights
-            weights = rep(1, G)
-            
-
-            
             # Generate special coefficient vector
             b.data = beta_fun(groups=groups, all.groups=all.groups, special.groups=special.groups, default.groups = default.groups, k=k, num.default=k0, upper=upper.scaled, lower=lower.scaled, cat.groups=cat.groups)
             # Special active set for glint/gamsel
             true.special = b.data$true.special
                 
         } 
+        # Frobenius normalized, do not need weights
+        weights = rep(1, G)
+            
+        if (length(cat.groups) > 0) {
+          Xscaled = frob_normalize(X, all.groups)
+        } else {
+          Xscaled = groupwise_center_scale(X, all.groups)
+        }
 
-        Xscaled = scale(X, center = TRUE, scale = FALSE)
-        Xscaled = frob_normalize(Xscaled, all.groups)
         if (estimation) {
             Xscaled.test = frob_normalize(X.test, all.groups)
         }
@@ -213,6 +214,7 @@ run_simulation = function(
 #####################################################
         # De-mean before passing to solvers #
         Y.beta = Y.beta - mean(Y.beta)
+        centered.noise = noise - mean(noise)
         # Do we really want to do this?     #
 #####################################################
         
@@ -223,7 +225,7 @@ run_simulation = function(
         }
         
         # Null results
-        results = forward_group(Xscaled, noise, groups=all.groups, weights=weights, Sigma, max.steps = max.steps, cat.groups = cat.groups)
+        results = forward_group(Xscaled, centered.noise, groups=all.groups, weights=weights, Sigma, max.steps = max.steps, cat.groups = cat.groups)
         P.mat[i, ] = results$p.vals
         AS.mat[i, ] = results$active.set
 
@@ -253,10 +255,6 @@ run_simulation = function(
             ric.list[[i]] = c(length(ric.as), ric.true.step)
           }
[…]

### 2014-05-08 paper/paper.tex
Small example, note on power, various edits

@@
 % perhaps de-emphasize group LASSO and emphasize forward stepwise?
-  We extend the methods developed by \cite{significance:LASSO} and
+  We extend the methods developed by \cite{significance:lasso} and
   \cite{tests:adaptive} on significance tests for penalized
@@
 
-In the LASSO setting, \cite{significance:LASSO} derived
+In the LASSO setting, \cite{significance:lasso} derived
 a novel test statistic and its asymptotic null distribution, making
@@
 \cite{tests:adaptive} modified and extended those results to the
-group LASSO \citep{groupLASSO} and other penalized regression
+group LASSO \citep{grouplasso} and other penalized regression
 problems, but only under the global null hypothesis.
@@
 
-In the next section we establish notation and describe the
-forward stepwise algorithm used in most of the paper.
+As a small illustrative example, consider a response $Y$ with ten
+categorical predictors $X_g$ having between 2 and 4 levels. The true 
+relationship is that $Y$ only depends on $X_1$ and $X_9$. Specifically,
+if observation $i$ has covariates $X_{1,i} = j, X_{9,i} = k$ then
+\[
+    Y_i = \beta_{1,j} + \beta_{9,k} + \epsilon_i , \qquad
+    \epsilon_i \sim N(0,1)
+\]
+$X_1$ has three levels with $\beta_1 = (1, .5, -1)$, $X_9$
+has two levels with $\beta_9 = (.5, -.5)$. We generated a random
+categorical design matrix $X$ and created an instance of this example
+with $n = 40$ observations. Running forward stepwise, we calculated
+both our p-value and the usual $\chi^2$ p-value at each step. In
+Table~\ref{tab:example} we see that forward stepwise chooses the
+truly nonzero variables first and both p-values for these are small.
+However, once forward stepwise begins adding noise variables, the
+$\chi^2$ p-value remains relatively small, potentially leading to 
+incorrect inferences.
+
+\input{../tables/small.tex}
+
+In the next section we establish general notation and describe the
+forward stepwise algorithm used throughout the paper.
 Section~\ref{sec:testing} reviews some recent work on
 post-selection inference
-\citep{significance:LASSO,tests:adaptive,LASSO:fixed}
+\citep{significance:lasso,tests:adaptive,lasso:fixed}
 relevant to our work here, and describes a general framework
@@
 stepwise tailored to models with interactions and generalized additive
-models, as well as to a real data example.
-(\todo Change this paragraph if necessary as the sections are completed)
+models, as well as to a real data example involving genomic prediction
+of individual drug responses to treat HIV.
+
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \subsection{Performance of forward stepwise}
+\label{sec:fs:performance}
 
@@
 statistics have also been used, so it is natural to consider using
-the new test statistics of \cite{significance:LASSO} or
+the new test statistics of \cite{significance:lasso} or
 \cite{tests:adaptive}
 to choose a model. \cite{sequential:fdr} examined some stopping rules
-using the asymptotic p-values of \cite{significance:LASSO} and showed
+using the asymptotic p-values of \cite{significance:lasso} and showed
 their stopping rules control false discovery rate--the expected
@@
 \[
-\text{TPP}_k = \frac{\# \left\{g: g \in A_k, \beta_g \neq 0\right\}}{k}.
+k\text{-TPP} = \frac{\# \left\{g: g \in A_k, \beta_g \neq 0\right\}}{k}.
 \]
-In some plots this has been denoted $k$-SP instead of $k$-TPP (\todo fix).
 Since we know $k$ in the simulations we can use $k$-TPP as a measure of how well forward stepwise is performing.
@@
 
-\section{Significance testing: from LASSO to group LASSO}
+\section{Significance testing with model selection}
 \label{sec:testing}
@@
 theoretical null distribution even when the null hypothesis is
-true. \cite{significance:LASSO} introduced a new test statistic based
+true. \cite{significance:lasso} introduced a new test statistic based
 on the knots in the LASSO solution path. They derived a simple
@@
 of combining inference with model selection.  \cite{tests:adaptive}
-extended that work to the group LASSO \citep{groupLASSO} and other
+extended that work to the group LASSO \citep{grouplasso} and other
 problems, and modified the test statistic to one with an exact finite
@@
 Writing $\hat \beta(\lambda)$ for the LASSO solution for a fixed value of $\lambda$, we recall
-the following facts summarized in \cite{significance:LASSO,tibshirani_LASSO_uniqueness}.
+the following facts summarized in \cite{significance:lasso,tibshirani_LASSO_uniqueness}.
 
@@
 {\em conditional} on which variable achieves $\lambda_1$ as well as its sign.
-The limiting results about later steps in \cite{significance:LASSO} can
+The limiting results about later steps in \cite{significance:lasso} can
 %(formally)
@@
 \begin{aligned}
-\label{eq:LASSO}
+\label{eq:lasso}
 \displaystyle \hat \beta_\lambda = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
@@
 \begin{equation}
-\label{eq:lammax}
+%\label{eq:lammax}
 \lambda_1^{\text{group}} = \max_{g} \frac{1}{w_g}\|X_g^Ty\|_2 . 
@@
 % \end{center}
-% A polytope selection event. Solving the LASSO \eqref{eq:LASSO} for a fixed value
+% A polytope selection event. Solving the LASSO \eqref{eq:lasso} for a fixed value
 % of $\lambda$ yields both an active set $\hat{E}(y)$ and a collection of signs of 
@@
 
-\subsection{Selection events}
+\subsection{Quadratic selection events}
 
@@
 setting with groups.
-The approach to post-selection inference in \cite{LASSO:fixed}
+The approach to post-selection inference in \cite{lasso:fixed}
 describes the selection event for the LASSO as a convex polytope
@@
 \end{equation}
-See \cite{LASSO:fixed,exact:lars} for further details on this approach.
+See \cite{lasso:fixed,exact:lars} for further details on this approach.
 
@@
 assume they are symmetric.
-As in \cite{LASSO:fixed} we will be interested in some randomly chosen contrast
+As in \cite{lasso:fixed} we will be interested in some randomly chosen contrast
 $\eta(S(y))^T\mu$.
@@
 %
-%This is Lemma 5 from \cite{significance:LASSO}.
+%This is Lemma 5 from \cite{significance:lasso}.
 %
@@
    \STATE Compute $H_g$ and $\sigma^2_g$
-   \IF{$X_g$ is a single column}
+   \IF{$p_g = 1$}
    \STATE $\sigma_g^2 \gets X_g^T\Sigma X_g/w_g^2$
@@
 
-We now describe how to calculate our p-value, following the discussion of
-group LASSO in \cite{tests:adaptive}. For the rest of this section let
-$g = \gstar$ be the index of
+Lemma \ref{eq:test:dbn} gives an explicit characterization of our p-value
+that can also be used to calculate it in practice. However, in this
+section we describe an alternative derivation following the discussion of
+group LASSO in \cite{tests:adaptive}. The implementation described here
+was used in all simulations.
+For the rest of this section let $g = \gstar$ be the index of
 the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
@@
 
-We performed simulations with several classes of design matrices including
-Gaussian, categorical, and categorical from a real data set. Gaussian
+We performed simulations with several classes of random design matrices
+including Gaussian and categorical, and with fixed categorical designs
+from the HIV data set of Section~\ref{sec:hiv}. Gaussian
 design matrices are generated either independently or with some global
@@
 number of observations in any level was at least five. Finally, categorical
-varables were encoded as groups of 0-1 vectors using the full encoding.
-The categorical data matrices from real data are described in
-Section~\ref{sec:hiv}
+variables were encoded as groups of 0-1 vectors using the full encoding.
 
 Since we are interested in variable selection we generate signals with
-nonzero coefficients on the scale of $\gamma := \sqrt{2 \log(p)/n}$.
-(\todo Cite some paper on support recovery?).
+nonzero coefficients on the scale of $\gamma := \sqrt{2 \log(G)/n}$,
+where recall $G$ is the number of groups.
 Coefficients within a
-nonzero group have roughly the same magnitude, and the magnitudes for
-each nonzero group range from a lower limit times $\gamma$ to an upper
-limit times $\gamma$. In plots, the limits are listed next to ``beta:''
-and the numbers in parentheses are the largest and smallest 2-norms of
-nonzero coefficient groups respectively. Each plot also displays the
+nonzero group have roughly the same magnitude and are normalized
+so that $\| \beta_g \|_2$ has the same scale independent of $p_g$.
+The magnitudes for different
+nonzero groups range from a lower limit times $\gamma$ to an upper
+limit times $\gamma$. In text above each plot, the limits are listed
+next to ``beta:'' and the numbers in parentheses are the largest
+and smallest 2-norms of nonzero coefficient groups respectively.
+Each plot also displays the
 number of observations or rows, $n$, the number of columns, $p$, the
@@
 The proportion of truly
-nonzero groups recovered in the first $k$ steps is denoted $k$-SP,
+nonzero groups recovered in the first $k$ steps is denoted $k$-TPP,
 where recall $k$ is the number of truly nonzero groups.
-
-Finally, most of the plots below show the following information. The
-horizontal axis is the step index for forward stepwise. The dashed
-line shows the proportion of iterations where a truly nonzero variable
-was added to the model at that step. Solid verticle lines show the
-middle 50\% of p-values calculated at that step, with red calculated
-on a null model with no nonzero groups, blue calculated on the non-null
-model, and green calculated on the non-null model from the usual
-$\chi^2$ significance test.
-The points in these lines are the average p-value at that step, and
-the triangles above and below show the 95\% and 5\% quantiles.
 
@@
 \label{fig:gausscorr}
-\includegraphics[width=0.5\textwidth]{../figs/default_gaussian_nsim400_n100_p200_g200_k10_lower1pt5_upper2_corr0pt2.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/default_gaussian_nsim400_n100_p200_g200_k10_lower1pt5_upper2_corr0pt2_noisecorr0pt1.pdf}}
 \hspace{-15pt}
@@
 
+We show two types of plots, one containing the same information as in
+Figure~\ref{fig:fwdstepsim}. The other plots each show a scenario with
+a fixed sparsity level and contain the following information. The
+horizontal axis is the step index for forward stepwise. The dashed
+line shows the proportion of iterations where a truly nonzero variable
+was added to the model at that step. Solid verticle lines are essentially
+narrow boxplots, showing the middle 50\% of p-values calculated at that
+step with circular points in these lines as the average and triangles
+showing the 95\% and 5\% quantiles. Different color verticle lines
+represent the following: red are calculated using our p-value on a null
+model with no nonzero groups, blue are calculated on the non-null
+model, and green calculated on the non-null model using the usual
+$\chi^2$ significance test.
+
 \begin{figure}[h]
@@
 \label{fig:hivnrti}
-\includegraphics[width=0.5\textwidth]{../figs/HIV_NRTI_n848_p48_g14_k4_lower21_upper25.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/default_fixed_nsim100_n639_p354_g177_sizes2-2_k5_lower5_upper10.pdf}}
 \hspace{-15pt}
@@
 
-%\input{../tables/default_gaussian_nsim500_n100_p200_g200_k5_lower1pt5_upper2.tex}
+The performance of forward stepwise in these plots is characterized by how
+close the dashed line is to a step function. If forward stepwise finds all
+truly nonzero variables first this line will be close to 1 until the step
+index reaches $k$ and then it will be close to 0. When this is the case,
+our p-value tends to be small while forward stepwise is selecting truly
+nonzero variables, uniform on the step where the dashed line goes to 0,
+and then progressively larger afterward. A reasonable stopping rule based
+on this behavior is to pick the model including all variables up to and
+including the last one with a significant p-value. We call this the
+\textit{last} stopping rule, and it selects the first $\hat k_\textit{last}$
+variables in the forward stepwise path where, if
+$p_j$ denotes our p-value calculated at the $j$-th step, then
+\begin{equation}
+  \label{eq:klast}
+  \begin{aligned}
+    \hat k_\textit{last} & = \max \{ j \leq steps : p_j < \alpha \}. \\
+  \end{aligned}
+\end{equation}
+To test this stopping rule we compare it with several others which we now
+describe. The \textit{oracle} stopping rule assumes knowledge of the
+sparsity $k$ and always picks the first $k$ variables. The \textit{first}
+rule stops immediately before the first variable to yield a p-value
+larger than $\alpha$, and since this requires rejection of the global 
+null in order to include the first variable
+it controls the family-wise error rate. However it does not seem to have
+good power.
+The \textit{forward}
+stopping rule is defined in \cite{sequential:fdr} and has the desirable
+property that it controls the false discovery rate. Finally, \textit{RIC}
+and \textit{BIC}
+are the risk inflation criterion and Bayesian information criterion
+described in Section~\ref{sec:fs:performance}. Tables~\ref{tab:easy} and
+\ref{tab:hard} show simulation results with $\alpha = 0.1$.
+To summarize these results, \textit{forward} is the only rule which
+controls FDR, \textit{BIC} selects models that are far too large,
+\textit{RIC} has low FDR and good power when the sparsity is low,
+and \textit{last} is comparable to \textit{RIC},
+with perhaps more power in the higher dimensional setting
+(and a correspondingly larger FDR when the signal is also weak).
 
@@
 \label{fig:spline:pval}
-\includegraphics[width=0.5\textwidth]{../figs/gamsel_uniform_nsim100_n200_p50_g50_k6_k02_lower10_upper15.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/gamsel_uniform_nsim200_n200_p50_g50_k6_k02_lower5_upper10.pdf}}
 \caption{\small \it The left panel shows the power of forward stepwise
@@
 Note that the varying group sizes present some difficulty to forward stepwise, which tends to first try linear approximations to nonzero spline groups before adding the true spline group.
-In Figure~\ref{fig:spline}(b) the average proportion of truly nonzero spline groups included in the first $k$ steps appears in parentheses next to the $k$-SP (top right).
+In Figure~\ref{fig:spline}(b) the average proportion of truly nonzero spline groups included in the first $k$ steps appears in parentheses next to the $k$-TPP (top right).
 
@@
 that were discovered in the first $k$-steps is shown in parentheses
-after the $k$-SP (top right).
+after the $k$-TPP (top right).
 
@@
 corresponding to the chosen model is unshaded, with the remainder shaded.
-The chosen variables are shown in Tables 3 and 4.
-(\todo Fix this ref after final re-run of code)
+The chosen variables are shown in Tables~\ref{tab:hivPI} and
+\ref{tab:hivNRTI}.
 
@@
 We can also use the \textsc{Glinternet} procedure described in Section~\ref{sec:glint} to fit a model with pairwise interactions.
-Results from this are shown in Figures~\ref{fig:HIVPI:glint} and \ref{fig:HIVNRTI:glint} and Tables 5 and 6.
+Results from this are shown in Figures~\ref{fig:HIVPI:glint} and \ref{fig:HIVNRTI:glint} and Tables~\ref{tab:glintPI} and \ref{tab:glintNRTI}.
 
@@
 
-\todo Brief summary \\
-\todo After finising everything else, move some points of discussion
-here \\
-\todo Mention ongoing work, like tracking all the constraints to
-get an exact p-value at every step. \\
+Under the global null hypothesis $\beta = 0$, we have a test statistic
+which we can use against the alternative of including the ``best''
+predictor (the one chosen by forward stepwise).
+This test statistic has an exact, finite sample distribution.
+If we choose to include the variable, we are no longer in the global null
+setting. However, by orthogonalizing both the response and the remaining
+predictors with respect to the included variable, it is reasonable to
+iterate the global null test. While not fully theoretically justified,
+this method seems to work well in all our simulations. By this we mean that
+the p-value tends to be small on the step when including the last truly
+nonzero predictor, uniform on the following step, and subsequently
+stochastically larger than uniform. This is in contrast to p-values
+calculated from traditional variable-inclusion tests like the $\chi^2$
+test, which tend to be smaller than uniform long after the last truly
+nonzero predictor has been included.
+
+When adding the next variable in forward stepwise, our hypothesis test
+roughly depends on the improvement gained by this variable compared to
+the next best variable. Thus, the p-value is small when there is a large
+gap between the variable to be included and all the remaining variables.
+If multiple predictors have truly nonzero coefficients that are close
+in magnitude, the p-value may be large until forward stepwise reaches
+the last one. This motivated us to consider the \textit{last} stopping
+rule for model selection \eqref{eq:klast}. In simulations we
+found this stopping rule to have good performance in terms of power
+(true positive rate), comparable to that of RIC \citep{RIC}.
+But it does not control the false discovery rate
+unless the truly nonzero coefficients are large enough to guarantee
+forward stepwise picks the corresponding variables before picking too
+many noise variables.
+
+The present work calculates p-values at each step ignoring the constraints
+imposed by previous steps. Future work adjusting for all previous steps
+is in progress, and may be able to give exact p-values at a step chosen
+stochastically by procedures like BIC and RIC. The authors hope to
+eventually release an R package implementing these methods.
 
@@
 \appendix
-\section{Additional details on closed-form computations}
+\section{Linear fractional subproblem}
 \label{app:closedform}
 
-We next describe the \textsc{LinearFractional} subproblem and its solution. We
-gave the problem this name because it originated in the form
+This appendix describes the \textsc{LinearFractional} optimization
+subproblem and its solution.
+The problem was named as it originated in the form
 \[
-\maximize_{h \neq g, \norm{u_h}_2 = 1} \frac{u_h^TX_h^Ty - u_h^TX_h^TX_gX_g^Ty}
-{1 - u_g^TX_g^TX_hu_h}
+\maximize_{h \neq g, \norm{u_h}_2 = 1}
+\frac{u_h^TX_h^Ty - u_h^TX_h^T\tilde X_g \tilde X_g^Ty}
+{1 - u_g^T \tilde X_g^TX_hu_h}
 \]
-The solution we describe next is to a slightly different problem, and
+The solution we describe next is to a slightly different problem which
 also incorporates the information that $g$ maximizes $\norm{X_g^Ty}_2/w_g$.
-%For details see \cite{tests:adaptive}.
+Although the logic seems a bit complicated, it mainly involves ruling out
+several cases as infeasible. The infeasibilities are precisely those given
+by the characterization of the global maximizer in \cite{tests:adaptive}.
+After ruling out degenerate cases, we obtain the following solution
+by transforming to trigonometric coordinates and using calculus.
 
 \begin{algorithm}
- \caption{The \textsc{LinearFractional} optimization subproblem}
+ \caption{The \textsc{LinearFractional} subproblem}
  \label{algo:linfrac}
@@
 
-For the sake of completeness this appendix provides full derivations
-of closed forms for the quantities required to compute our
-statistic. We require two facts from \cite{tests:adaptive}
-
-A point $\eta \in \K$ maximizes $f_\eta$ over a convex
-set $\K$ if and only if the following conditions hold: 
-\begin{equation}
-\label{eq:maxcond}
-\grad f_{|T_{\eta}\K} = 0, \quad
-\tf^{\eta}_{\eta} \geq \V^+_{\eta}, \quad
-\tf^{\eta}_{\eta} \leq \V^-_{\eta}, \quad \text{and} \quad
-\V^0_{\eta} \leq 0.
-\end{equation}
-The same equivalence holds true even when $\K$ is only locally
-convex.
-
-Write $M^{\pm}_{\eta, h}$ and $M^0_{\eta, h}$ as the corresponding suprema and infima over the restricted parameter set $S_h$.
+% For the sake of completeness this appendix provides full derivations
+% of closed forms for the quantities required to compute our
+% statistic. We require two facts from \cite{tests:adaptive}
+
+% A point $\eta \in \K$ maximizes $f_\eta$ over a convex
+% set $\K$ if and only if the following conditions hold: 
+% \begin{equation}
+% \label{eq:maxcond}
+% \grad f_{|T_{\eta}\K} = 0, \quad
+% \tf^{\eta}_{\eta} \geq \V^+_{\eta}, \quad
+% \tf^{\eta}_{\eta} \leq \V^-_{\eta}, \quad \text{and} \quad
+% \V^0_{\eta} \leq 0.
+% \end{equation}
+% The same equivalence holds true even when $\K$ is only locally
+% convex.
+
+% Write $M^{\pm}_{\eta, h}$ and $M^0_{\eta, h}$ as the corresponding suprema and infima over the restricted parameter set $S_h$.
 Note that the characterization of the global maximizer (Lemma 1, general paper) implies $M^0_{\eta, h} \leq 0$ and $M^+_{\eta, h} \leq M^-_{\eta, h}$ for all $h \neq g$.
@@
 
-Now fix $h$, let $P_g = X_gX_g^T$ and define
-\begin{align*}
-a_h &= w_h^{-1} X_h^T (I-P_g) y \\
-b_h &= w_g w_h^{-1} X_h^T P_g y / \norm{X_g^T y} \\
-c_h &= a_h^T b_h / (\norm{a_h} \norm{b_h}) \\
-K_h   &= \{ x : \norm{x} = 1, b_h^Tx > w \}
-\end{align*}
-Note that $K_h = \emptyset$ if and only if $\norm{b_h} < w$.
+% Now fix $h$, let $P_g = X_gX_g^T$ and define
+% \begin{align*}
+% a_h &= w_h^{-1} X_h^T (I-P_g) y \\
+% b_h &= w_g w_h^{-1} X_h^T P_g y / \norm{X_g^T y} \\
+% c_h &= a_h^T b_h / (\norm{a_h} \norm{b_h}) \\
+% K_h   &= \{ x : \norm{x} = 1, b_h^Tx > w \}
+% \end{align*}
+% Note that $K_h = \emptyset$ if and only if $\norm{b_h} < w$.
 We consider two cases.
@@
 
-Now, if $a_h$ is in the polar cone of $K_h$ (and ruling out the probability zero case where $a_h^Tx = 0$ on the boundary of $K_h$), then the numerator $a_h^Tx < 0$ on $K_h$ so there is a finite positive infimum in the interior $K_h$.
+% Now, if $a_h$ is in the polar cone of $K_h$ (and ruling out the probability zero case where $a_h^Tx = 0$ on the boundary of $K_h$), then the numerator $a_h^Tx < 0$ on $K_h$ so there is a finite positive infimum in the interior $K_h$.
 Similarly there is a finite positive supremum on the interior of $K_h^c$ (the numerator is negative near the boundary of $K_h$ by continuity).
 
-To attain the infimum on $K_h$ requires making $x$ close to $b_h$ and simultaneously making the numerator closer to zero, so $x$ should be on the side of $b_h$ that is closer to $a_h$.
+% To attain the infimum on $K_h$ requires making $x$ close to $b_h$ and simultaneously making the numerator closer to zero, so $x$ should be on the side of $b_h$ that is closer to $a_h$.
 To attain the supremum on $K_h^c$ requires making $x$ simultaneously close to $a_h$ (so that $a_h^Tx > 0$) and $b_h$ (to make the denominator small).
@@
 
-We have verified that solving both cases requires finding the maxima of the trigonometric form $\norm{a_h} \cos (\psi) / (w - \norm{b_h} \cos(\theta - \psi))$ of the linear fraction on the interiors of $K_h$ (if it is nonempty) and $K_h^c$.
+% We have verified that solving both cases requires finding the maxima of the trigonometric form $\norm{a_h} \cos (\psi) / (w - \norm{b_h} \cos(\theta - \psi))$ of the linear fraction on the interiors of $K_h$ (if it is nonempty) and $K_h^c$.
 The derivative is proportional to
-\[
-\frac{\norm{b_h} \sin(\theta) - w \sin(\psi)}{w - \norm{b_h} \cos(\theta - \psi)}
-\]
-Since we have ruled out the boundary of $K_h$, we can ignore the denominator and focus on critical points corresponding to the angles $\psi^{\pm}$ (symmetric about $\pi/2$) where $\sin(\psi^{\pm}) = (\norm{b_h} / w) \sin (\theta)$.
+% \[
+% \frac{\norm{b_h} \sin(\theta) - w \sin(\psi)}{w - \norm{b_h} \cos(\theta - \psi)}
+% \]
+% Since we have ruled out the boundary of $K_h$, we can ignore the denominator and focus on critical points corresponding to the angles $\psi^{\pm}$ (symmetric about $\pi/2$) where $\sin(\psi^{\pm}) = (\norm{b_h} / w) \sin (\theta)$.
 That is,
-\[
-\psi^{\pm} \in \arcsin \frac{\norm{b_h}}{w} \sqrt{1-c_h^2}
-\]
-Note that these critical points only exist if $\sin(\theta) < w/\norm{b_h}$, and since we have argued that the infimum and supremum are attained it follows that this condition is met in our case.
+% \[
+% \psi^{\pm} \in \arcsin \frac{\norm{b_h}}{w} \sqrt{1-c_h^2}
+% \]
+% Note that these critical points only exist if $\sin(\theta) < w/\norm{b_h}$, and since we have argued that the infimum and supremum are attained it follows that this condition is met in our case.
 Let $\psi^+$ denote the smaller of the two angles.
@@
 
-Finally we calculate the values of the linear fraction at $\psi^{\pm}$. Let us first record some facts:
-\begin{align*}
-0 < \psi^+ < \pi/2, \quad \psi^- = \pi - \psi^+ \\
-\cos(\theta) = c_h, \quad \sin(\theta) = \sqrt{1-c_h^2} \\
-\sin(\psi^+) = \sin(\psi^-) = \frac{\norm{b_h}}{w} \sin(\theta) \\
-\cos(\psi^+) = -\cos(\psi^-) = \sqrt{1 - (\norm{b_h}/w)^2(1-c_h^2)}
-\end{align*}
-Using the angle difference formula,
-\begin{align*}
-\cos(\theta - \psi^+) & = \frac{\norm{b_h}}{w} (1-c_h^2) +  c_h \cos(\psi^+) \\
-& = \frac{\norm{b_h}}{w} (1-c_h^2) +  c_h \cos(\psi^+)
-\end{align*}
-Hence
-\begin{align*}
-M^+_{\eta, h} & = \frac{ \norm{a_h} c_h }{ w - (\norm{b_h}^2/w) (1-c_h^2) - \norm{b_h} c_h \cos(\psi^+) } \\
-% & = 
-% & = \frac{ w \norm{a} a_h^Tb_h / \norm{b_h} }{ w^2 - \norm{a}\norm{b}^2 + (a^Tb)/\norm{a} - a^Tb \cos \psi^+ }
-\end{align*}
-And $M^-_{\eta, h}$ is $\infty$ when $\norm{b_h} < w$ and otherwise given by
-\begin{align*}
-M^+_{\eta, h} & = \frac{ \norm{a_h} c_h }{ w - (\norm{b_h}^2/w) (1-c_h^2) + \norm{b_h} c_h \cos(\psi^+) } \\
-\end{align*}
+% Finally we calculate the values of the linear fraction at $\psi^{\pm}$. Let us first record some facts:
+% \begin{align*}
+% 0 < \psi^+ < \pi/2, \quad \psi^- = \pi - \psi^+ \\
+% \cos(\theta) = c_h, \quad \sin(\theta) = \sqrt{1-c_h^2} \\
+% \sin(\psi^+) = \sin(\psi^-) = \frac{\norm{b_h}}{w} \sin(\theta) \\
+% \cos(\psi^+) = -\cos(\psi^-) = \sqrt{1 - (\norm{b_h}/w)^2(1-c_h^2)}
+% \end{align*}
+% Using the angle difference formula,
+% \begin{align*}
+% \cos(\theta - \psi^+) & = \frac{\norm{b_h}}{w} (1-c_h^2) +  c_h \cos(\psi^+) \\
+% & = \frac{\norm{b_h}}{w} (1-c_h^2) +  c_h \cos(\psi^+)
+% \end{align*}
+% Hence
+% \begin{align*}
+% M^+_{\eta, h} & = \frac{ \norm{a_h} c_h }{ w - (\norm{b_h}^2/w) (1-c_h^2) - \norm{b_h} c_h \cos(\psi^+) } \\
+% % & = 
+% % & = \frac{ w \norm{a} a_h^Tb_h / \norm{b_h} }{ w^2 - \norm{a}\norm{b}^2 + (a^Tb)/\norm{a} - a^Tb \cos \psi^+ }
+% \end{align*}
+% And $M^-_{\eta, h}$ is $\infty$ when $\norm{b_h} < w$ and otherwise given by
+% \begin{align*}
+% M^+_{\eta, h} & = \frac{ \norm{a_h} c_h }{ w - (\norm{b_h}^2/w) (1-c_h^2) + \norm{b_h} c_h \cos(\psi^+) } \\
+% \end{align*}
 
@@
 
+\section{A note on power}
+\label{app:power}
+
+Although we have shown through simulations in a variety of settings that
+our test has power, here we give a theoretical result in one specific case
+where it is easy to prove something. Namely, consider the case where all
+groups have size 1, and the design matrix is orthonormal. By a
+transformation we can reduce the identity design matrix case
+\[
+y_i = \mu_i + \epsilon_i, \qquad \epsilon_i \sim N(0,1)
+\]
+without loss of generality. In this simple case our p-value can be
+calculated as
+\[
+\frac{\tilde \Phi(\lambda_1)}{\tilde \Phi(\lambda_2)}
+\]
+where $\lambda_1$ and $\lambda_2$ are the first two knots in the LASSO
+solution path and $\tilde \Phi = 1 - \Phi$ is the survival function of
+a standard Gaussian.
+
+The simplest possible alternative hypothesis
+is the 1-sparse case where a single $\mu_i$ is nonzero. Let
+$\mu_1 = r \sqrt{2\log(p)}$ and all other $\mu_i = 0$. If $r$ is some
+constant greater than 1, then with high probability the first
+knot $\lambda_1$ will be achieved by $y_1$. 
+Applying the normal tail bounds
+\[
+\frac{\tilde \Phi(xr)}
+{\tilde\Phi(x)} \leq \frac{\phi(rx)/rx}{(1-1/x^2)\phi(x)/x}
+\]
+with $x = \sqrt{2\log(p)}$ we obtain, by continuity,
+\[
+ \frac{\tilde \Phi(\lambda_1)}{\tilde \Phi(\lambda_2)}
+\lessapprox \frac{p^{1-r^2}}{r\left(1-\frac{1}{2\log(p)}\right)}.
+\]
+For any $r > 1$ the upper bound goes to 0 as $p \to \infty$, so
+in this case the test has asymptotic full power at the same threshold
+as Bonferroni. This is the best possible threshold for asymptotic power
+against the 1-sparse alternative.
+
 \end{document}

### 2014-05-08 HIV/sim_HIV.R

@@ -1,7 +1,7 @@
-#db = "NRTI"
-db = "PI"
+db = "NRTI"
+#db = "PI"
 
-binary.encoding = TRUE
+binary.encoding = FALSE
 
 setwd('..')
 source('fwd_step.R')
@@ -19,18 +19,17 @@ if (db == "NRTI") {
   resps = 4:10
 }
 
-nsim = 10
+nsim = 400
 type = "default"
 design = "fixed"
 fn.append = '' #'Pg'
 corr = 0 # nonzero only supported for gaussian design
 noisecorr = 0
-nsim = 10
 num.nonzero = 5
 k = num.nonzero
-max.steps = 15
-upper = 10
-lower = 5
+max.steps = 10
+upper = 8
+lower = 4
 
 # Clean data, restrict to cleaned subset
 nresps = length(resps)

### 2014-05-08 sim_default.R

@@ -15,8 +15,8 @@ groups = 1:200
 #groups = 1:50
 
 k = 10
-upper = 2
-lower = 1.5
+upper = 3
+lower = 2
 
 max.steps = 15
 #max.steps = min(n/2, max(groups)-1) #floor(max(groups)/3)
@@ -77,9 +77,12 @@ with(output, step_plot(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, m
 ## abline(v = 1, col = "gray")
 ## dev.off()
 
-#groups = sort(c(rep(1:30, 5), rep(31:35, 10)))
+groups = sort(c(rep(1:30, 5), rep(31:35, 10)))
 #groups = sort(c(rep(1:10, 3), rep(11:20, 2)))
 #groups = 1:200
+corr = 0
+noisecorr = 0
+max.steps = 10
 k = 5
 
 output.g = run_simulation(
@@ -93,13 +96,13 @@ output.g = run_simulation(
     noisecorr = noisecorr,
     estimation = estimation, verbose = TRUE)
 
-output.g$main.append = paste0(", RIC: (", round(output.g$ric.tpp, 2), "/", output.g$ric.size, ")")
+#output.g$main.append = paste0(", RIC: (", round(output.g$ric.tpp, 2), "/", output.g$ric.size, ")")
 
-with(output.g, step_plot(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, max.steps, upper, lower, max.beta, min.beta, fwd.power, design, filename, main.append))
+with(output.g, step_plot(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, max.steps, upper, lower, max.beta, min.beta, fwd.power, design, filename))
 
 caption = "Evaluation of model selection using several stopping rules based on our p-values."
 results.l = with(output, sim_select_stats(signal.p, active.set, true.step, k, bic.list = bic.list, ric.list = ric.list))
-results.g = with(output.g, sim_select_stats(signal.p, active.set, true.step, k, bic.list = bic.list, ric.list = ric.list))
+results.g = with(output.g, sim_select_stats(signal.p, active.set, true.step, k))
 
 #rownames(results.l) = paste0("*", rownames(results.l))
 #rownames(results.g) = paste0("**", rownames(results.g))

### 2014-05-08 sim_special.R

@@ -12,18 +12,18 @@ if (type == 'gamsel') {
     design = 'uniform'
     n = 200
     groups = 1:50
-    nsim = 100
+    nsim = 200
 } else {
     design = 'gaussian'
     n = 100
     groups = 1:30
-    nsim = 100
+    nsim = 200
 }
 
 k = 6
 k0 = 2
-upper = 15
-lower = 10
+upper = 10
+lower = 5
 max.steps = 18
 
 

### 2014-05-08 simulation.R

@@ -234,7 +234,7 @@ run_simulation = function(
 
         # Comparison with BIC/RIC
         if (type == "default") {
-          if (design == "gaussian") {
+          if (p == G) {
 
             steps = min(floor(n/2), max(groups)-1)
             
@@ -315,9 +315,11 @@ run_simulation = function(
         num.recovered.special = rowSums(special.recover.mat[, 1:k])
         special.fwd.power = mean(num.recovered.special) / length(true.special)
     } else {
-      if (design == "gaussian") {
+      if (p == G) {
         bic.size = median(bic.size)
         bic.tpp = mean(bic.tpp)
+        ric.size = median(ric.size)
+        ric.tpp = mean(ric.tpp)
       }
     }
     fwd.power = mean(num.recovered.groups) / k
@@ -341,7 +343,7 @@ run_simulation = function(
     if (type != "default") {
         outlist$special.fwd.power = special.fwd.power
     } else {
-      if (design == "gaussian") {
+      if (p == G) {
         outlist$bic.list = bic.list
         outlist$bic.tpp = bic.tpp
         outlist$bic.size = bic.size

### 2014-05-08 small.R

@@ -1,0 +1,31 @@
+set.seed(1)
+source("generate_data.R")
+source("fwd_step.R")
+source("tex_table.R")
+
+n = 40
+groups = c(1,1,1,2,2,3,3,3,3,4,4,5,5,6,6,7,7,7,8,8,9,9,10,10,10,10)
+group.labels = unique(groups)
+max.steps = 8
+weights = rep(1, length(group.labels))
+X = categorical_design(n, groups)
+beta = rep(0, length(groups))
+beta[groups == 1] = c(1,.5,-1)
+beta[groups == 9] = c(.5,-.5)
+
+Y = X %*% beta + rnorm(n)
+Y = Y - mean(Y)
+Sigma = diag(rep(1,n))
+Xnormed = frob_normalize(X, groups)
+
+results = forward_group(Xnormed, Y, groups, weights, Sigma, max.steps = max.steps, cat.groups = group.labels)
+
+step = 1:max.steps
+Variable = results$active.set
+pval = results$p.vals
+chi.pval = results$chi.pvals
+matrix = rbind(Variable, pval, chi.pval)
+colnames(matrix) = step
+texfile = "tables/small.tex"
+caption = "A small illustrative example"
+tex_table(texfile, matrix, caption)

### 2014-05-12 paper/paper.tex
quadratic decisions algorithm

@@
 $$
+Note that for each fixed $g$ and $\eta \in \real^g$ with $\|\eta\|_2=1$, the event 
+$$
+\left\{y:\frac{X_g^Ty}{\|X_g^Ty\|_2} = \eta \right\} = \left\{y: \|X_g^Ty\|^2_2 \leq (X_g\eta)^Ty \right\},
+$$
+is defined by another quadratic inequality that can be appended to $I(s)$.
 
@@
 
+We now present an explicit algorithm for computing this slice in quadratic selection procedures.
+
+\begin{algorithm}
+ \caption{Truncation interval for quadratic decisions}
+ \label{algo:quadratic}
+ \begin{algorithmic}
+   \REQUIRE Response $y$, state $s$ list of quadratic inequalities $\{Q_i, a_i, b_i: i \in I(s)\}$, direction
+   of interest $\eta$ with $\|\eta\|_2=1$.
+   \ENSURE The set $\{t: (y+t \cdot \eta)^TQ_i(y+t \cdot \eta) + a_i^T(y+t \cdot \eta) \leq b_i  \ \forall i \in I(s) \}$.
+      \STATE Initialize  interval: $M= (-\infty,\infty)$, $U=-\infty, L=\infty$
+    \FOR{$i$ in $I(s)$}
+    \STATE{$a=\eta^TQ_i\eta, b=2y^TQ_i\eta + a_i^T\eta, c=y^TQ_iy + a_i^Ty - b_i$}
+    \IF{$a \neq 0$}
+    \IF{$b^2-4ac > 0$}
+    \STATE Stop: $y$ does not satisfy inequalities!
+    \ELSIF{$a > 0$}
+     \STATE $M \gets M \cap [(-b-\sqrt{b^2-4ac})/2a, (-b+\sqrt{b^2-4ac})/2a]$;
+     \ELSIF{$a < 0$}
+     \STATE $L \gets \min(L,(-b-\sqrt{b^2-4ac})/2a)$,
+     \STATE $U \gets \max(U,(-b+\sqrt{b^2-4ac})/2a).$;
+    \ENDIF
+    \ELSE
+    \IF{$b > 0$}
+   \STATE $M \gets M \cap (-\infty, -c/b]$;
+   \ELSIF{$b < 0$}
+   \STATE $M \gets M \cap [-c/b,\infty)$;
+   \ELSIF{$c > 0$}
+   \STATE Stop: $y$ does not satisfy inequalities!
+   \ENDIF
+   \ENDIF
+   \ENDFOR
+   \RETURN $M \cap (L,U)^c.$
+ \end{algorithmic}
+\end{algorithm}
+
 In turn, this allows to derive a test statistic to test the hypothesis
@@
 the scale of the $\chi$ above is given by $\theta_s$. \qed
+
+\subsubsection{The $T\chi$ test statistic}
+
+This leads us to what we call the {\em  $T\chi$ test statistic} which is determined by
+a degrees of freedom parameter $r$ as well as a truncation set
+$M \subset \real$, a scale parameter $\theta$ and an observed value $t$.
+The test statistic is just the survival function of a $\chi_r$ random variable
+truncated to the set $M$ evaluated at our observed 
+norm $\|X_g^Ty\|_2$ (after scaling by the appropriate scale $\theta$).
+\begin{equation}
+\label{eq:tchi}
+T\chi(t, r, \theta,  M) = \frac{\int_{M /\theta \cap [t/\theta,\infty)} F_{\chi_r}(dz)}{\int_{M / \theta} F_{\chi_r}(dz)}.
+\end{equation}
+
 %$\sigma \|X_s\eta_s\|_2$.
@@
 
+Lemma \ref{eq:test:dbn} and Algorithm \ref{algo:quadratic} give an explicit characterization of our p-value
+that can also be used to calculate it in practice. It can also be applied to different
+quadratic selection procedures besides the variant of forward
+stepwise we consider here. For our purposes,
+this truncation interval 
+section was derived as a special case of the
+Kac-Rice test  defined in \cite{tests:adaptive} applied to the first step of the group LASSO path.
+
+The implementation described in Algorithm \ref{algo:pval}
+was used in all simulations.
+For the rest of this section let $g = \gstar$ be the index of
+the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
+%$\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
+%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
+For a vector $u \in \real^p$, let $u_h$ denote the coordinates
+of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
+We can rearrange the columns of $X$ to group these adjacently, so that
+\[
+u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
+\]
+One step of the calculation will be to find an orthonormal basis for the
+linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
+\text{ for all } h \neq g \}$ so that we can project orthogonally to
+this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
+should be the case since $g$ maximizes the absolute value of this
+quantity), then the
+space is trivial and the desired orthogonal projection is the identity.
+Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
+generally has dimension $p_g-1$.
+
 \begin{algorithm}
@@
    \STATE $b \gets X^T \tilde X_g$
-   \STATE Compute solution $(v_-, v_+)$ of \textsc{LinearFractional}(a,b) optimization subproblem
+   \STATE Compute solution $(v_+, v_-)$ of \textsc{LinearFractional}(a,b) optimization subproblem
    \STATE $r \gets \text{rank}(X_g)$
-   \STATE $u \gets [F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(\lambda^2/\sigma_g^2)]/[F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(v^2_+/\sigma_g^2)]$
-   \RETURN $u$
+   \RETURN $T\chi(\|X_g^Ty\|_2, r, \sigma_g,  (v_-,v_+)).$
  \end{algorithmic}
 \end{algorithm}
-
-Lemma \ref{eq:test:dbn} gives an explicit characterization of our p-value
-that can also be used to calculate it in practice. However, in this
-section we describe an alternative derivation following the discussion of
-group LASSO in \cite{tests:adaptive}. The implementation described here
-was used in all simulations.
-For the rest of this section let $g = \gstar$ be the index of
-the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
-%$\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
-%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
-For a vector $u \in \real^p$, let $u_h$ denote the coordinates
-of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
-We can rearrange the columns of $X$ to group these adjacently, so that
-\[
-u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
-\]
-One step of the calculation will be to find an orthonormal basis for the
-linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
-\text{ for all } h \neq g \}$ so that we can project orthogonally to
-this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
-should be the case since $g$ maximizes the absolute value of this
-quantity), then the
-space is trivial and the desired orthogonal projection is the identity.
-Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
-generally has dimension $p_g-1$.
 
@@
    \ENDFOR
-   \STATE $v^- \gets \min_h v_h^-$
-   \STATE $v^+ \gets \max_h v_h^+$
-   \RETURN $(v^-, v^+)$
+   \STATE $v_- \gets \max_h v_h^-$
+   \STATE $v_+ \gets \min_h v_h^+$
+   \RETURN $(v_-, v_+)$
  \end{algorithmic}

### 2014-05-12 paper/paper.tex
abstract update

@@
   the global null hypothesis test on the residual in a forward stepwise
-  procedure. While not fully theoretically justified, the resulting method
+  procedure. The resulting method
   has the computational strengths of stepwise selection and partially
   solves the problem of inflated test statistics due to model selection.
-  We illustrate the flexibility of this method by applying it to
+  In describing our variant of forward stepwise selection, we introduce
+  a new class of selection procedures, described by quadratic inequalities and
+  provide an algorithm that can be used for exact post-selection inference for such problems.
+
+  We illustrate the flexibility of our procedure by applying it to
   novel specialized applications of forward stepwise to a hierarchically
-  constrained interactions model and generalized additive models.
+  constrained interactions model and recently described flexible additive models
+  that choose between linear and nonlinear effects for each variable.
 \end{abstract}

### 2014-05-12 paper/paper.tex
Paper edits, adding figs/tables

--- 2014-05-08 paper/paper.tex
@@
 % "Title of the paper"
-\title{A significance test in forward stepwise model selection}
+\title{A significance test for forward stepwise model selection}
 
@@
 \begin{abstract}
-% perhaps de-emphasize group LASSO and emphasize forward stepwise?
-  We extend the methods developed by \cite{significance:lasso} and
+  We apply the methods developed by \cite{significance:lasso} and
   \cite{tests:adaptive} on significance tests for penalized
-  regression to stepwise model selection by iteratively applying
-  the global null hypothesis test on the residual in a forward stepwise
-  procedure. While not fully theoretically justified, the resulting method
-  has the computational strengths of stepwise selection and partially
-  solves the problem of inflated test statistics due to model selection.
+  regression to forward stepwise model selection. We introduce a new
+  test statistic for settings with grouped variables, allowing us to
+  handle categorical variables and factor models. This
+  statistic has a truncated $\chi$ distribution under the global null,
+  and we apply it in forward stepwise iteratively on the residual after
+  each step. The resulting method
+  has the computational strengths of stepwise selection and addresses
+  the problem of invalid test statistics due to model selection.
   We illustrate the flexibility of this method by applying it to
-  novel specialized applications of forward stepwise to a hierarchically
-  constrained interactions model and generalized additive models.
+  several specialized applications of forward stepwise including a
+  hierarchical interactions model and regression with polynomial splines.
 \end{abstract}
@@
 variable in each step. Because of the stochastic nature of this
-algorithm--making use of the data to choose variables--
-the usual $\chi^2$ and $F$-tests for significance
-fail when a model has been selected this way and tend to be
+algorithm---making use of the data to choose variables---the
+usual $\chi^2$ and $F$-tests for significance
+fail when a model has been selected this way. These tests will be
 anti-conservative unless they are computed on a held-out validation
-dataset. This is one instance of the general problem of conducting
-inference and model selection using the same data, a problem
+dataset. This problem is one instance of the general problem of
+conducting inference and model selection using the same data, a problem
 of central importance on which some recent progress has been made.
@@
 used for valid significance testing when computed on the same
-data used for model selection, eliminating the need for data splitting
-or for situations in which data splitting is not appropriate. For example,
-when categorical covariates contain rare levels with very few observations
+data used for model selection, eliminating the need for data splitting.
+This is especially important in
+situations in which data splitting is not appropriate. For example,
+when categorical covariates have levels with very few observations
 it can be difficult or impossible to split the sample with an adequate
-number of such observations occurring in each split.
+number of such observations occurring in each split. Furthermore, even
+if sample splitting is possible, it sacrifices accuracy in the model
+selection procedure and power in the subsequent inferences.
 
@@
 
-As a small illustrative example, consider a response $Y$ with ten
+As a small illustrative example of what we gain from this method,
+consider a response $Y$ with ten
 categorical predictors $X_g$ having between 2 and 4 levels. The true 
@@
 stepwise itself performs well, and that
-various stopping rules using our test statistic--including
-some from \cite{sequential:fdr}--appear promising. In
+various stopping rules using our test statistic---including
+some from \cite{sequential:fdr}---appear promising. In
 Section~\ref{sec:applications} we apply
@@
 size 1, this includes non-grouped situations as a special
-case. Our variant of forward stepwise uses a different method than
-usual for choosing which group to add because we are using the
-same objective in the forward stepwise procedure that was used to
-derive the test statistics in penalized regression settings.
-
-Denote $y \in \real^n$ for $n$ i.i.d. measurements of the outcome variable.
+case.
+%Our variant of forward stepwise uses a different method than
+%usual for choosing which group to add because we are using the
+%same objective in the forward stepwise procedure that was used to
+%derive the test statistics in penalized regression settings.
+
+Let $y \in \real^n$ be $n$ i.i.d. measurements of the outcome variable.
 Let an integer $G \geq 2$ be the number of groups of explanatory variables.
@@
 
-Denote $p = \sum_{g=1}^Gp_g$,
+Denote by $p = \sum_{g=1}^Gp_g$
 the total number of individual variables, so $p = G$ in the
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-We normalize each group by its Frobenius norm, but this is not required
-especially since we also allow weights, $w_g$, for each group. These
+We also allow weights $w_g$ for each group. These
 weights act like penalties or costs, so increasing $w_g$ makes it
 more difficult for the group $X_g$ to enter the model. The
-modeler can choose weights arbitrarily, but we set them all to be
-1 since we already normalize by the Frobenius norm.
+modeler can choose weights arbitrarily for calibration purposes,
+but throughout we set them all constant (equal to 1) and normalize
+groups by the Frobenius norm of their corresponding submatrices.
 
@@
 \label{eq:gmodel}
-y | X& = X \beta + \sigma \epsilon \\
+y & = X \beta + \sigma \epsilon \\
    & = \sum_{g=1}^G X_g \beta_g + \sigma \epsilon
@@
 
-\subsection{Description of the algorithm}
+\subsection{Description of the forward stepwise algorithm}
 
 First, the user must specify the maximum number of steps allowed,
-which we denote \textit{steps} (consistent with the \textit{R} function
-for stepwise). To enable our test statistic computations,
+which we denote \textit{steps}.
+%(consistent with the \textit{R} function for stepwise).
+To enable our test statistic computations,
 \textit{steps} should be at most $\min (n, G) - 1$, but it is
@@
 of the LASSO.
-Writing $\hat \beta(\lambda)$ for the LASSO solution for a fixed value of $\lambda$, we recall
-the following facts summarized in \cite{significance:lasso,tibshirani_LASSO_uniqueness}.
+Recall the LASSO estimator is given by
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{equation}
+\begin{aligned}
+\label{eq:lasso}
+\displaystyle \hat \beta(\lambda) = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
+   \lambda \| \beta \|_1 \\
+\end{aligned}
+\end{equation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The following facts are summarized in \cite{significance:lasso,tibshirani_LASSO_uniqueness}.
 
@@
 
-The covariance test (assuming groups of size one and $\Sigma = \sigma^2 I$ is given at the first step by
+Assuming groups of size one and $\Sigma = \sigma^2 I$, the covariance test is given at the first step by
 \begin{equation}
@@
 \end{equation}
+This is a hypothesis test for including the first variable in the LASSO
+path, with large values of the test statistic being evidence against the
+global null.
 In \cite{tests:adaptive} it is pointed out 
@@
 \end{equation}
-Further, the above test statistic can be understood as the survival function of $\lambda_1 = \|X^Ty\|_{\infty}$
+The right hand side has an exact, finite sample null distribution.
+Further, this new statistic can be understood as the survival function of $\lambda_1 = \|X^Ty\|_{\infty}$
 {\em conditional} on which variable achieves $\lambda_1$ as well as its sign.
@@
 recursively applying $T_1$ to the variables that had not previously been selected by LARS.
-
-In the setting of grouped variables, we use an analogous test statistic to
-the right hand side of \eqref{test:exact}. The test statistic is presented as an example
-in \cite{tests:adaptive} though we re-derive it here in simpler form. The test draws inspiration
+In forward stepwise with groups of size 1, we recursively apply the right hand side test statistic in the same manner.
+With larger groups, we use an analogous test statistic to the right hand side of \eqref{test:exact}.
+The test statistic is presented as an example in \cite{tests:adaptive} though we re-derive it here in simpler form.
+The test draws inspiration
 from the group LASSO.
 
-Recall the LASSO estimator is given by
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{equation}
-\begin{aligned}
-\label{eq:lasso}
-\displaystyle \hat \beta_\lambda = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
-   \lambda \| \beta \|_1 \\
-\end{aligned}
-\end{equation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 The \emph{group LASSO estimator}, a generalization of the LASSO,
@@
 special case when all groups are of size 1, since then the
-penalty term is the $L^1$-norm of $\beta$. The solution path of the group
+penalty term is the $\ell_1$-norm of $\beta$. The solution path of the group
 LASSO has similar properties to the LASSO case, however it is not
@@
 $$
+This choice of $\eta$ reflects our goal of calculating a p-value for inclusion of the variable $g^*$ chosen by forward stepwise.
 
@@
 $$
-
 \begin{figure}
-\label{fig:curved}
 \begin{center}
@@
 \end{center}
+\caption{ \em
 A generic selection event, given by the intersection of a list
-of quadratic inequalities. Note that the event need not be simply connected. \end{figure}
-
+of quadratic inequalities. Note that the event need not be simply connected.
+}
+\label{fig:curved}
+\end{figure}
 The slice for any particular inequality is
-$$
+\begin{equation}
 \begin{aligned}
+\label{eq:quadslice}
 \lefteqn{
@@
 \end{aligned}
-$$
+\end{equation}
 This can be computed explicitly, and results in either the empty set, a single interval (possibly infinite) or the union of 
@@
 \end{equation}
-In the picture \ref{fig:curved}, for every $\eta$, the slice \eqref{eq:slice} is a function
+In Figure \ref{fig:curved}, for every $\eta$, the slice \eqref{eq:slice} is a function
 of $P_{\eta}^{\perp}y$. This amounts to a proof of the following lemma.
@@
 the scale of the $\chi$ above is given by $\theta_s$. \qed
+
+Lemma \ref{eq:test:dbn} gives an explicit characterization of our p-value
+that can also be used to calculate it in practice.
+An implementation only needs to compute the intersection of all the intervals implied by each quadratic constraint \eqref{eq:quadslice} and then evaluate the $\chi$ distribution truncated to the intersection of these intervals at the observed value.
+For an alternative
+derivation including a detailed algorithm description for the implementation used throughout this paper see Appendix~\ref{app:algo}.
+
+After adding the first group in forward stepwise and computing its p-value, we orthogonalize the response $y$ and all remaining groups.
+This imposes some constraints on $y$ that in principle should be used when computing p-values at subsequent steps.
+For now we ignore these constraints and iterate the global null test, but work on incorporating all the known constraints is ongoing.
+We believe that the effect of not tracking all constraints is what causes the nominal p-value to be increasingly stochastically larger than uniform further down the forward stepwise path, and only truly uniform at the first step where all remaining variables are noise.
+
 %$\sigma \|X_s\eta_s\|_2$.
@@
 
-\subsection{Computation of p-value}
-
-\begin{algorithm}
- \caption{Computing p-value}
- \label{algo:pval}
- \begin{algorithmic}
-   \REQUIRE Response $y$, grouped design matrix $X$ with weights, inactive set $A^c$, index $g$ of last group to enter active set $A$.
-   \ENSURE Our p-value for the group $g$ entering the model.
-   \STATE Compute $H_g$ and $\sigma^2_g$
-   \IF{$p_g = 1$}
-   \STATE $\sigma_g^2 \gets X_g^T\Sigma X_g/w_g^2$
-   \STATE $\tilde X_g \gets X_g / w_g \cdot \text{sign}(X_g^Ty)$
-   \ELSE
-   \STATE $P_g \gets \Sigma XV_g (V_g^T X^T \Sigma X V_g)^\dagger V_g^TX^T$
-   \STATE $H_g \gets (I - P_g)\Sigma$
-   \STATE $\sigma^2_g \gets y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2)$
-   \STATE $\tilde X_g \gets H_g X_g X_g^T y /(\norm{X_g^Ty}_2 w_g)$
-   \ENDIF
-   \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$
-   \STATE \COMMENT{Compute the following two $p$-vectors}
-   \STATE $a \gets X^T(y - \lambda \tilde X_g)$
-   \STATE $b \gets X^T \tilde X_g$
-   \STATE Compute solution $(v_-, v_+)$ of \textsc{LinearFractional}(a,b) optimization subproblem
-   \STATE $r \gets \text{rank}(X_g)$
-   \STATE $u \gets [F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(\lambda^2/\sigma_g^2)]/[F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(v^2_+/\sigma_g^2)]$
-   \RETURN $u$
- \end{algorithmic}
-\end{algorithm}
-
-Lemma \ref{eq:test:dbn} gives an explicit characterization of our p-value
-that can also be used to calculate it in practice. However, in this
-section we describe an alternative derivation following the discussion of
-group LASSO in \cite{tests:adaptive}. The implementation described here
-was used in all simulations.
-For the rest of this section let $g = \gstar$ be the index of
-the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
-%$\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
-%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
-For a vector $u \in \real^p$, let $u_h$ denote the coordinates
-of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
-We can rearrange the columns of $X$ to group these adjacently, so that
-\[
-u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
-\]
-One step of the calculation will be to find an orthonormal basis for the
-linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
-\text{ for all } h \neq g \}$ so that we can project orthogonally to
-this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
-should be the case since $g$ maximizes the absolute value of this
-quantity), then the
-space is trivial and the desired orthogonal projection is the identity.
-Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
-generally has dimension $p_g-1$.
-
-%We proceed by forming the $n \times p_g$ matrix
-%$X_g - X_g X_g^T y y^T X_g/\norm{X_g^Ty}_2^2$ and performing Gram-Schmidt
-%on the first $r-1$ columns. For precise theorems showing that this will
-%generally work, see \cite{outerp:indep}.
-We compute an orthonormal basis by Gram-Schmidt and form a
-$p_g \times p_g$ matrix, which we denote $V_g$, by appending
-0's in the coordinates corresponding to all groups $h \neq g$ and
-an additional column of zeroes (since Gram-Schmidt only produces $p_g-1$).
-We can now define the projection
-
-\begin{equation}
- \begin{aligned}
-   \label{eq:proj}
-   P_g &= \Sigma X_gV_g (V_g^T X_g^T \Sigma X_g V_g)^\dagger V_g^TX_g^T. \\
-  \end{aligned}
-\end{equation}
-Also define $H_g = (I-P_g)\Sigma$ and the conditional variance
-
-\begin{equation}
- \begin{aligned}
-   \label{eq:cvar}
-   \sigma^2_g &= y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2) \\
- \end{aligned}
-\end{equation}
-These simplify when $X_g$ is a single column, in which case $P_g = 0$,
-$H_g = \Sigma$, and $\sigma^2_g = X_g^T\Sigma X_g^T/w_g^2$.
-Note that $F_{\chi^2_r}$ denotes the distribution function of a $\chi^2$
-random variable with $r$ degrees of freedom. This completes the
-description of Algorithm~\ref{algo:pval} which we use to compute our
-p-value. Additional details, including the statement and solution of
-the \textsc{LinearFractional} subproblem, appear in Appendix~\ref{app:closedform}.
-
 \section{Simulations}
@@
 \label{fig:gaussgroups}
-\includegraphics[width=0.5\textwidth]{../figs/default_gaussian_nsim500_n100_p200_g35_sizes5-10_k5_lower2_upper3.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/default_gaussian_nsim400_n100_p200_g35_sizes5-10_k5_lower2_upper3.pdf}}
 \caption{\small \it The left panel shows results for Gaussian design
@@
 \label{fig:hivnrti}
-\includegraphics[width=0.5\textwidth]{../figs/default_fixed_nsim100_n639_p354_g177_sizes2-2_k5_lower5_upper10.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/default_fixed_nsim400_n848_p537_g82_sizes2-24_k5_lower4_upper8.pdf}}
 \hspace{-15pt}
@@
 \label{fig:gausscat}
-\includegraphics[width=0.5\textwidth]{../figs/default_categorical_nsim500_n100_p80_g40_sizes2-2_k5_lower2_upper4.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/default_categorical_nsim500_n100_p80_g40_sizes2-2_k10_lower2_upper4.pdf}}
 \caption{\small \it The left panel shows results for a categorical design
@@
 Given a design matrix $X$, we can augment this by adding additional columns given by nonlinear functions of the original covariates.
-Specifically, we compute spline basis vectors for each covariate and create a group given by this spline basis.
-Denote this new design matrix
+Specifically, for each original covariate $X_g$ we compute a submatrix of spline basis vectors $X_g^s$.
+Adding these groups of spline basis vectors to the original design matrix yields a new design matrix
 \begin{equation}
@@
   \begin{aligned}
-    \tilde X &=  \begin{pmatrix} X_1 & \cdots & X_G & X^+_1 & \cdots & X^+_G  \end{pmatrix}\\
-    &= \begin{pmatrix} X & X^+ \end{pmatrix}
+    \tilde X &=  \begin{pmatrix} X_1 & \cdots & X_G & X^s_1 & \cdots & X^s_G  \end{pmatrix}\\
   \end{aligned}
@@
   \begin{aligned}
-    X_i &\sim \text{Unif}[-1,1], \quad
-    X_i^+ = \text{CubicSplineBasis}(X_i)
+    X_g &\sim \text{Unif}[-1,1], \quad
+    X_g^s = \text{CubicSplineBasis}(X_g).
   \end{aligned}
@@
 \label{fig:glint:pval}
-\includegraphics[width=0.5\textwidth]{../figs/glint_gaussian_nsim100_n100_p30_g30_k6_k02_lower4_upper8.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/glint_gaussian_nsim200_n100_p30_g30_k6_k02_lower5_upper10.pdf}}
 \caption{\small \it The left panel shows the power of forward stepwise
@@
 analysis using forward stepwise, and report our P-value at each step.
-Categorical covariates are encoded as groups of binary variables using
+Categorical covariates are encoded as groups of dummy variables using
 the full encoding, and weights are set to the square-root of the
@@
 with p-values plotted by step in forward stepwise. 
-The \textit{last} stopping rule is applied and the region of the plot
+The \textit{last} stopping rule \eqref{eq:klast} is applied and the region of the plot
 corresponding to the chosen model is unshaded, with the remainder shaded.
@@
 is in progress, and may be able to give exact p-values at a step chosen
-stochastically by procedures like BIC and RIC. The authors hope to
-eventually release an R package implementing these methods.
+stochastically by procedures like BIC and RIC. The authors will
+release an R package implementing these methods.
 
@@
 \appendix
-\section{Linear fractional subproblem}
-\label{app:closedform}
-
-This appendix describes the \textsc{LinearFractional} optimization
+\section{P-value algorithm}
+\label{app:algo}
+
+\begin{algorithm}
+ \caption{Computing p-value}
+ \label{algo:pval}
+ \begin{algorithmic}
+   \REQUIRE Response $y$, grouped design matrix $X$ with weights, inactive set $A^c$, index $g$ of last group to enter active set $A$.
+   \ENSURE Our p-value for the group $g$ entering the model.
+   \STATE Compute $H_g$ and $\sigma^2_g$
+   \IF{$p_g = 1$}
+   \STATE $\sigma_g^2 \gets X_g^T\Sigma X_g/w_g^2$
+   \STATE $\tilde X_g \gets X_g / w_g \cdot \text{sign}(X_g^Ty)$
+   \ELSE
+   \STATE $P_g \gets \Sigma XV_g (V_g^T X^T \Sigma X V_g)^\dagger V_g^TX^T$
+   \STATE $H_g \gets (I - P_g)\Sigma$
+   \STATE $\sigma^2_g \gets y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2)$
+   \STATE $\tilde X_g \gets H_g X_g X_g^T y /(\norm{X_g^Ty}_2 w_g)$
+   \ENDIF
+   \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$
+   \STATE \COMMENT{Compute the following two $p$-vectors}
+   \STATE $a \gets X^T(y - \lambda \tilde X_g)$
+   \STATE $b \gets X^T \tilde X_g$
+   \STATE Compute solution $(v_-, v_+)$ of \textsc{LinearFractional}(a,b) optimization subproblem
+   \STATE $r \gets \text{rank}(X_g)$
+   \STATE $u \gets [F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(\lambda^2/\sigma_g^2)]/[F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(v^2_+/\sigma_g^2)]$
+   \RETURN $u$
+ \end{algorithmic}
+\end{algorithm}
+
+In this section we describe an alternative derivation following the
+discussion of group LASSO in \cite{tests:adaptive}. The implementation
+described here was used in all simulations.
+For the rest of this section let $g = \gstar$ be the index of
+the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
+%$\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
+%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
+For a vector $u \in \real^p$, let $u_h$ denote the coordinates
+of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
+We can rearrange the columns of $X$ to group these adjacently, so that
+\[
+u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
+\]
+One step of the calculation will be to find an orthonormal basis for the
+linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
+\text{ for all } h \neq g \}$ so that we can project orthogonally to
+this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
+should be the case since $g$ maximizes the absolute value of this
+quantity), then the
+space is trivial and the desired orthogonal projection is the identity.
+Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
+generally has dimension $p_g-1$.
+
+%We proceed by forming the $n \times p_g$ matrix
+%$X_g - X_g X_g^T y y^T X_g/\norm{X_g^Ty}_2^2$ and performing Gram-Schmidt
+%on the first $r-1$ columns. For precise theorems showing that this will
+%generally work, see \cite{outerp:indep}.
+We compute an orthonormal basis by Gram-Schmidt and form a
+$p_g \times p_g$ matrix, which we denote $V_g$, by appending
+0's in the coordinates corresponding to all groups $h \neq g$ and
+an additional column of zeroes (since Gram-Schmidt only produces $p_g-1$).
+We can now define the projection
+
+\begin{equation}
+ \begin{aligned}
+   \label{eq:proj}
+   P_g &= \Sigma X_gV_g (V_g^T X_g^T \Sigma X_g V_g)^\dagger V_g^TX_g^T. \\
+  \end{aligned}
+\end{equation}
+Also define $H_g = (I-P_g)\Sigma$ and the conditional variance
+
+\begin{equation}
+ \begin{aligned}
+   \label{eq:cvar}
+   \sigma^2_g &= y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2) \\
+ \end{aligned}
+\end{equation}
+These simplify when $X_g$ is a single column, in which case $P_g = 0$,
+$H_g = \Sigma$, and $\sigma^2_g = X_g^T\Sigma X_g^T/w_g^2$.
+Note that $F_{\chi^2_r}$ denotes the distribution function of a $\chi^2$
+random variable with $r$ degrees of freedom.
+
+Next we describe the \textsc{LinearFractional} optimization
 subproblem and its solution.

### 2014-05-12 tables/small.tex
Paper edits, adding figs/tables

% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Wed May  7 22:07:35 2014
\begin{table}[ht]
\centering
\begin{tabular}{r|rrrrrrrr}
  \hline
Step & 1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 \\ 
  \hline
Variable & 1 & 9 & 2 & 8 & 4 & 7 & 3 & 10 \\ 
T$\chi$ p-value & 0.00 & 0.01 & 0.15 & 0.50 & 0.77 & 0.81 & 0.84 & 0.92 \\ 
$\chi^2$ p-value & 0.00 & 0.00 & 0.04 & 0.21 & 0.29 & 0.56 & 0.73 & 0.82 \\ 
   \hline
\end{tabular}
\caption{\em Results of forward stepwise on this example}
\label{tab:example}
\end{table}

### 2014-05-12 sim_categorical.R

@@ -10,7 +10,7 @@ estimation = FALSE
 n = 100
 groups = sort(rep(1:40, 2))
 
-k = 5
+k = 10
 upper = 4
 lower = 2
 max.steps = 15
@@ -65,6 +65,7 @@ with(output, step_plot(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, m
 ## dev.off()
 
 groups = sort(c(rep(1:20, 4), rep(21:40, 8)))
+k = 5
 max.steps = 10
 
 output.g = run_simulation(

### 2014-05-12 sim_special.R

@@ -1,5 +1,5 @@
-#type = 'glint'
-type = 'gamsel'
+type = 'glint'
+#type = 'gamsel'
 
 
 source('simulation.R')

### 2014-05-12 paper/paper.tex

@@
 $$
-This choice of $\eta$ reflects our goal of calculating a p-value for inclusion of the variable $g^*$ chosen by forward stepwise.
+Note that for each fixed $g$ and $\eta \in \real^g$ with $\|\eta\|_2=1$, the event 
+$$
+\left\{y:\frac{X_g^Ty}{\|X_g^Ty\|_2} = \eta \right\} = \left\{y: \|X_g^Ty\|^2_2 \leq (X_g\eta)^Ty \right\},
+$$
+is defined by another quadratic inequality that can be appended to $I(s)$.
 
@@
 
+We now present an explicit algorithm for computing this slice in quadratic selection procedures.
+
+\begin{algorithm}
+ \caption{Truncation interval for quadratic decisions}
+ \label{algo:quadratic}
+ \begin{algorithmic}
+   \REQUIRE Response $y$, state $s$ list of quadratic inequalities $\{Q_i, a_i, b_i: i \in I(s)\}$, direction
+   of interest $\eta$ with $\|\eta\|_2=1$.
+   \ENSURE The set $\{t: (y+t \cdot \eta)^TQ_i(y+t \cdot \eta) + a_i^T(y+t \cdot \eta) \leq b_i  \ \forall i \in I(s) \}$.
+      \STATE Initialize  interval: $M= (-\infty,\infty)$, $U=-\infty, L=\infty$
+    \FOR{$i$ in $I(s)$}
+    \STATE{$a=\eta^TQ_i\eta, b=2y^TQ_i\eta + a_i^T\eta, c=y^TQ_iy + a_i^Ty - b_i$}
+    \IF{$a \neq 0$}
+    \IF{$b^2-4ac > 0$}
+    \STATE Stop: $y$ does not satisfy inequalities!
+    \ELSIF{$a > 0$}
+     \STATE $M \gets M \cap [(-b-\sqrt{b^2-4ac})/2a, (-b+\sqrt{b^2-4ac})/2a]$;
+     \ELSIF{$a < 0$}
+     \STATE $L \gets \min(L,(-b-\sqrt{b^2-4ac})/2a)$,
+     \STATE $U \gets \max(U,(-b+\sqrt{b^2-4ac})/2a).$;
+    \ENDIF
+    \ELSE
+    \IF{$b > 0$}
+   \STATE $M \gets M \cap (-\infty, -c/b]$;
+   \ELSIF{$b < 0$}
+   \STATE $M \gets M \cap [-c/b,\infty)$;
+   \ELSIF{$c > 0$}
+   \STATE Stop: $y$ does not satisfy inequalities!
+   \ENDIF
+   \ENDIF
+   \ENDFOR
+   \RETURN $(M \cap (L,U)^c) + \eta^Ty.$
+ \end{algorithmic}
+\end{algorithm}
+
 In turn, this allows to derive a test statistic to test the hypothesis
@@
 
-Lemma \ref{eq:test:dbn} gives an explicit characterization of our p-value
-that can also be used to calculate it in practice.
-An implementation only needs to compute the intersection of all the intervals implied by each quadratic constraint \eqref{eq:quadslice} and then evaluate the $\chi$ distribution truncated to the intersection of these intervals at the observed value.
-For an alternative
-derivation including a detailed algorithm description for the implementation used throughout this paper see Appendix~\ref{app:algo}.
-
-After adding the first group in forward stepwise and computing its p-value, we orthogonalize the response $y$ and all remaining groups.
-This imposes some constraints on $y$ that in principle should be used when computing p-values at subsequent steps.
-For now we ignore these constraints and iterate the global null test, but work on incorporating all the known constraints is ongoing.
-We believe that the effect of not tracking all constraints is what causes the nominal p-value to be increasingly stochastically larger than uniform further down the forward stepwise path, and only truly uniform at the first step where all remaining variables are noise.
+\subsubsection{The $T\chi$ test statistic}
+
+This leads us to what we call the {\em  $T\chi$ test statistic} which is determined by
+a degrees of freedom parameter $r$ as well as a truncation set
+$M \subset \real$, a scale parameter $\theta$ and an observed value $t$.
+The test statistic is just the survival function of a $\chi_r$ random variable
+truncated to the set $M$ evaluated at our observed 
+norm $\|X_g^Ty\|_2$ (after scaling by the appropriate scale $\theta$).
+\begin{equation}
+\label{eq:tchi}
+T\chi(t, r, \theta,  M) = \frac{\int_{M /\theta \cap [t/\theta,\infty)} F_{\chi_r}(dz)}{\int_{M / \theta} F_{\chi_r}(dz)}.
+\end{equation}
 

### 2014-05-13 paper/paper.tex
Edits, including captions

@@
   \cite{tests:adaptive} on significance tests for penalized
-  regression to forward stepwise model selection. We introduce a new
-  test statistic for settings with grouped variables, allowing us to
-  handle categorical variables and factor models. This
-  statistic has a truncated $\chi$ distribution under the global null,
-  and we apply it in forward stepwise iteratively on the residual after
+  regression to forward stepwise model selection. A general framework
+  for selection procedures described by quadratic inequalities includes
+  a variant of forward stepwise with grouped variables, allowing us to
+  handle categorical variables and factor models. We provide an algorithm
+  to compute a new statistic with an exact null distribution conditional
+  on the outcome of the model selection procedure. This new
+  statistic, which we denote $T\chi$,
+  has a truncated $\chi$ distribution under the global null.
+  We apply this test in forward stepwise iteratively on the residual after
   each step. The resulting method
@@
   several specialized applications of forward stepwise including a
-  hierarchical interactions model and regression with polynomial splines.
+  hierarchical interactions model and a recently described additive model
+  that adaptively chooses between linear and nonlinear effects for
+  each variable. 
 \end{abstract}
@@
 categorical design matrix $X$ and created an instance of this example
-with $n = 40$ observations. Running forward stepwise, we calculated
-both our p-value and the usual $\chi^2$ p-value at each step. In
+with $n = 40$ observations. Running forward stepwise for eight steps,
+we calculated both the new $T\chi$ $p$-value and the usual
+$\chi^2$ $p$-value at each step. In
 Table~\ref{tab:example} we see that forward stepwise chooses the
-truly nonzero variables first and both p-values for these are small.
+truly nonzero variables first and both $p$-values for these are small.
 However, once forward stepwise begins adding noise variables, the
-$\chi^2$ p-value remains relatively small, potentially leading to 
+$\chi^2$ $p$-value remains relatively small, potentially leading to 
 incorrect inferences.
@@
 \includegraphics[width=0.9\textwidth]{../figs/pubmed.pdf}
-\caption{\small \it Forward stepwise enjoys widespread use among
+\caption{\small \em Forward stepwise enjoys widespread use among
 practitioners}
@@
 to choose a model. \cite{sequential:fdr} examined some stopping rules
-using the asymptotic p-values of \cite{significance:lasso} and showed
+using the asymptotic $p$-values of \cite{significance:lasso} and showed
 their stopping rules control false discovery rate--the expected
@@
 \subfigure[Gaussian designs]{
-\label{fig:fsgausscorr}
+\label{fig:fsgauss}
 \includegraphics[width=0.5\textwidth]{../figs/tpr_default_gaussian_nsim200_n100_p200_g200_k25_lower1pt5_upper2.pdf}}
-\caption{\small \it The left panel shows the results of the simulation
+\caption{\small \em
+  For various sparsity levels $k$, the True Positive Proportion of the
+  model selected by forward stepwise at $k$ steps.
+  The left panel shows the results of the simulation
   using a categorical design consisting of 200 binary factors.
@@
 enough for some purposes. Finally, to serve as a comparison with existing
-commonly used model selection procedures, in Figure~\ref{fig:fwdstepsim}(b)
+commonly used model selection procedures, in Figure~\ref{fig:fsgauss}
 we show in green the TPP and median model sizes for models chosen by
@@
 
+\subsection{Background}
 In the ordinary least squares setting, a significance test for a
@@
 \label{test:exact}
-\exp\left(- \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \right) \approx \frac{1 - \Phi(\lambda_1/\sigma)}{1 - \Phi(\lambda_2 / \sigma)} \overset{D}{=} \text{Unif}(0,1).
+\exp\left(- \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \right) \approx \frac{1 - \Phi(\lambda_1/\sigma)}{1 - \Phi(\lambda_2 / \sigma)} \overset{D}{=} \text{Unif}([0,1]).
 \end{equation}
@@
 
-The \emph{group LASSO estimator}, a generalization of the LASSO,
- is a solution to the following penalized
+The \emph{group LASSO estimator} \citep{grouplasso,bakin:grouplasso},
+a generalization of the LASSO, is a solution to the following penalized
 least-squares convex problem
@@
 test is that it conditions on both the maximizer $g^*$ and on the unit
-vector $\eta_{g^*}$. Most of the p-value computations will be done in
+vector $\eta_{g^*}$. Most of the $p$-value computations will be done in
 terms of these quantities.
@@
 
-\subsection{Quadratic selection events}
-
-Before giving the full description of the p-value, we turn to another
-previous work and describe how to extend the framework there to the
-setting with groups.
+\subsection{Quadratic framework for derivation of test statistic}
+
+In this section we derive the new $p$-value and describe how to compute it.
+First we give a brief summary of what follows; readers not interested in the full derivation can skip the rest of this section after they understand the summary.
+The event $E_g$ that forward stepwise Algorithm~\ref{algo:fs} selects a given group $g$ is equivalent to a set of quadratic inequalities involving $y$ (see \eqref{eq:first:quadratic} below).
+To derive a statistic related to the group $g$ being included, we choose a direction vector $\eta(g)$ in the relevant direction and study the distribution of $\eta(g)^Ty$ restricted to the event $E_g$.
+In Figure~\ref{fig:curved} below the event $E_g$ is the shaded region, and the direction $\eta$ determines a slice through this event.
+Finally, we solve for the amounts $t$ by which $y$ can be translated in the directions $\pm \eta$ and still satisfy the constraints imposed by $E_g$.
+These are the points where the slice intersects the boundary of $E_g$.
+Let $M \subset \mathbb R$ be the set of $t$ such that $y + t\cdot\eta \in E_g$.
+With all of these quantities and a little more work we find that the observed norm $\| X_g^Ty \|_2$ has a $\chi$ distribution truncated to the set $M$, and with a computable scale parameter.
+Applying the appropriate CDF transform \eqref{eq:tchi} with these quantities yields our $p$-value.
+
+We now give the full derivation of the $p$-value, beginning with a previous work and then extending the framework there to the setting with groups.
 The approach to post-selection inference in \cite{lasso:fixed}
@@
 
-With non-trivial groups, however, the event that $g^*(y)$, the first
-group chosen, is equal to some fixed group $g$ is generally not given by
-a set of affine constraints. Rather, considering line 3 of Algorithm~\ref{algo:fs}, we see
+However, with group sizes larger than 1, the event that the first group
+chosen $g^*(y)$ is equal to some fixed group $g$ is no longer given by
+a set of affine constraints. Rather, considering line 3 of
+Algorithm~\ref{algo:fs}, we see
 \begin{equation}
@@
 In the rest of this section, we consider an arbitrary selection procedure
-$S$ that returns one of a set of possible outcomes $s \in {\cal S}$ by a set of quadratic inequalities.
+$S$ that returns one of a set of possible outcomes $s \in {\cal S}$ determined by a set of quadratic inequalities.
 That is, for all possible outcomes $s \in {\cal S}$ there is a set of indices
@@
 As in \cite{lasso:fixed} we will be interested in some randomly chosen contrast
-$\eta(S(y))^T\mu$.
+$\eta(s)^T\mu$.
 In our grouped selection procedure, the selection rule is $S(y)=g^*(y)$
-and $\eta$ is typically 
+and the value of $\eta$ on $\{ g^* = g \}$ is
 $$
-\eta(S(y)) = \eta(g^*(y)) = \frac{X_{g^*(y)}^Ty}{\|X_{g^*(y)}^Ty\|_2}.
+\eta(s) = \eta(g) = \frac{X_{g}^Ty}{\|X_{g}^Ty\|_2}.
 $$
-Note that for each fixed $g$ and $\eta \in \real^g$ with $\|\eta\|_2=1$, the event 
+This choice of $\eta$ reflects our goal of calculating a $p$-value for inclusion of the variable $g^*$ chosen by forward stepwise.
+Note that for each fixed $g$ and any $\eta \in \real^g$ with $\|\eta\|_2=1$, the event 
 $$
@@
 \caption{ \em
-A generic selection event, given by the intersection of a list
-of quadratic inequalities. Note that the event need not be simply connected.
+A quadratic selection event, given by the intersection of a list
+of quadratic inequalities. The selection event and chosen contrast
+$\eta$ determine a slice through the event with computable limits
+where the boundary of the event is reached.
+Note that the event need not be simply connected.
 }
@@
 
-In turn, this allows to derive a test statistic to test the hypothesis
+In turn, this allows us to derive a test statistic to test the hypothesis
 $H_{0,s}:X_s^T\mu=0$ conditional on $s$ being selected.
@@
 \label{eq:test:dbn}
-Conditional on $H_{0,s} \cap E_+(s, \eta_s, P_s^{\perp}y, (P_s - \eta_s\eta_s^T)y)$
+For $y \sim N(0, \sigma^2 I)$,
+conditional on the {\em event}
+\[
+H_{0,s} \cap E_+(s, \eta_s, P_s^{\perp}y, (P_s - \eta_s\eta_s^T)y)
+\]
 we have the following truncated $\chi$ distributional result
@@
 \end{equation}
-where the scale 
+where the vertical bar $|$ here denotes restriction to an interval, and the scale is given by
 $$
@@
 $$
-and the vertical bar $|$ here denotes restriction to an interval.
 \end{lemma}
@@
 
-\subsubsection{The $T\chi$ test statistic}
-
-This leads us to what we call the {\em  $T\chi$ test statistic} which is determined by
-a degrees of freedom parameter $r$ as well as a truncation set
-$M \subset \real$, a scale parameter $\theta$ and an observed value $t$.
-The test statistic is just the survival function of a $\chi_r$ random variable
-truncated to the set $M$ evaluated at our observed 
-norm $\|X_g^Ty\|_2$ (after scaling by the appropriate scale $\theta$).
+\subsection{The $T\chi$ test statistic}
+
+With Lemma \ref{eq:test:dbn} and Algorithm \ref{algo:quadratic} we are finally ready to define our test statistic, which we term the {\em truncated $\chi$ test statistic}, denoted $T\chi$.
+This statistic is determined by a degrees of freedom parameter $r$ as well as a truncation set $M \subset \real$, a scale parameter $\theta$ and an observed value $t$.
+The test statistic is given by the survival function of a $\chi_r$ random variable truncated to the set $M$ evaluated at our observed norm $\|X_g^Ty\|_2$ (after scaling by the appropriate scale $\theta$).
 \begin{equation}
@@
 \end{equation}
-
-%$\sigma \|X_s\eta_s\|_2$.
-
-%\subsection{Relation to covariance test}
-%
-%Consider the simple case for the first covariate to enter the model.
-%In this case the covariance test statistic is $T = \lambda_1(\lambda_1 - M)$.
-%\todo Does $M = \lambda_2$? Check proof, now that statistic has changed
-%
-%This is Lemma 5 from \cite{significance:lasso}.
-%
-%\begin{itemize}
-%
-%  \item As in LTTT, $T = \lambda_1 (\lambda_1 - M)$ and $M = \lambda_2$
-%  \item Convergence to the limiting Exp(1) distribution is too slow
-%    \[
-%      \frac{P(\chi_k / w_g \geq m + t/m)}{P(\chi_k / w_g \geq m)} \to e^{-t} \text{ as } m \to \infty
-%    \]
-%    (when the group achieving $\lambda_1$ is group $g$ and has rank $k$)
-%  \item The limiting distribution only depends on $T$, but we also observe $M$
-%  \item Let's just try the ratio (conditional $\chi_k$ tail probability) evaluated at $T$ and $M$ (it works better)
-%  \item Going one step further, instead of using the approximation (see LTTT Proof of Lemma 5)
-%    \[
-%      \frac{M + \sqrt{M^2+4t}}{2} \approx M + \frac{t}{M}
-%    \]
-%    we can just use the left hand side
-%  \item For $T = \lambda_1(\lambda_1 - M)$ the left hand side simplifies to $\lambda_1$
-%  \item Now our p-value is
-%    \[
-%      \frac{P(\chi_k / w_g \geq \lambda_1)}{P(\chi_k / w_g \geq \lambda_2)}
-%    \]   
-%\end{itemize}
-%
-%
+For simplicity we always use this distributional transform rather than reporting the observed values of $(t,r,\theta,M)$. Hence, $T\chi$ itself is our $p$-value, with
+\[
+T\chi \sim \text{Unif}([0,1])
+\]
+under the global null.
+
+After adding the first group in forward stepwise and computing the $T\chi$ $p$-value, we orthogonalize the response $y$ and all remaining groups with respect to the group just added.
+This imposes further constraints that in principle should be used when computing $p$-values at subsequent steps.
+For now we ignore these constraints and iterate the global null test, but work on incorporating all known constraints is ongoing.
+We believe that the effect of not tracking all constraints is what causes the nominal $p$-value to be increasingly stochastically larger than uniform further down the forward stepwise path (as can be seen in Figure~\ref{fig:pval1}), and only truly uniform at the first step where all remaining variables are noise.
+
+For an alternate derivation that follows more closely \cite{tests:adaptive}, and an algorithm that includes weights and an arbitrary (known) covariance matrix $\Sigma$ see Appendix~\ref{app:algo}.
 
@@
 
+To study the behavior of the $T\chi$ test after taking steps in the forward stepwise path and to understand its power to detect various departures from the global null we conduct simulations in a wide variety of settings including both $p > n$ and $p < n$ problems.
 We performed simulations with several classes of random design matrices
@@
 from the HIV data set of Section~\ref{sec:hiv}. Gaussian
-design matrices are generated either independently or with some global
+design matrices were generated either independently or with some global
 correlation $\rho > 0$ between all pairs of columns. Categorical matrices
@@
 nonzero group have roughly the same magnitude and are normalized
-so that $\| \beta_g \|_2$ has the same scale independent of $p_g$.
+so that $\| \beta_g \|_2$ has scale $\gamma$ independently of $p_g$.
 The magnitudes for different
@@
 limit times $\gamma$. In text above each plot, the limits are listed
-next to ``beta:'' and the numbers in parentheses are the largest
+next to ``beta'' and the numbers in parentheses are the largest
 and smallest 2-norms of nonzero coefficient groups respectively.
@@
 \includegraphics[width=0.5\textwidth]{../figs/default_gaussian_nsim400_n100_p200_g35_sizes5-10_k5_lower2_upper3.pdf}}
-\caption{\small \it The left panel shows results for Gaussian design
+\caption{\small \em
+Forward stepwise with $T\chi$ simulation results.
+Dashed line is the local True Positive Rate at each step averaged over
+400 simulation realizations.
+Vertical lines
+are boxplots of $p$-values marginally at each step, with red computed
+using $T\chi$ on a null model, blue using $T\chi$ on a sparse alternative
+model, and green using $\chi^2$ on the same. The shaded region indicates
+the sparsity of the alternative.
+Left panel shows results for Gaussian design
 matrices with equicorrelation $\rho=0.2$ and equicorrelated noise with
-correlation 0.1. The right panel shows results for an independent
+correlation 0.1. Right panel shows results for an independent
 Gaussian design with groups of sizes 5 and 10.}
+\label{fig:pval1}
 \end{center}
@@
 was added to the model at that step. Solid verticle lines are essentially
-narrow boxplots, showing the middle 50\% of p-values calculated at that
+narrow boxplots, showing the middle 50\% of $p$-values calculated at that
 step with circular points in these lines as the average and triangles
 showing the 95\% and 5\% quantiles. Different color verticle lines
-represent the following: red are calculated using our p-value on a null
-model with no nonzero groups, blue are calculated on the non-null
-model, and green calculated on the non-null model using the usual
-$\chi^2$ significance test.
+represent the following: red are calculated using the $T\chi$ $p$-value
+on a null model with no nonzero groups, blue are calculated using the
+$T\chi$ $p$-value on the non-null model, and green calculated on the
+non-null model using the usual $\chi^2$ significance test.
 
@@
 \includegraphics[width=0.5\textwidth]{../figs/default_categorical_nsim500_n100_p80_g40_sizes2-2_k10_lower2_upper4.pdf}}
-\caption{\small \it The left panel shows results for a categorical design
+\caption{\small \em
+As in Figure \ref{fig:pval1}.
+The left panel shows results for a categorical design
 taken from a real data set.
-The right panel shows results for categorical matrices with all categories having 5 levels.}
+The right panel shows results for categorical matrices with all categories having 5 levels.
+The null cases being larger than uniform are likely the result of numerical error as these problems are poorly conditioned.
+}
 \end{center}
@@
 index reaches $k$ and then it will be close to 0. When this is the case,
-our p-value tends to be small while forward stepwise is selecting truly
+the $T\chi$ $p$-value tends to be small while forward stepwise is selecting truly
 nonzero variables, uniform on the step where the dashed line goes to 0,
@@
 on this behavior is to pick the model including all variables up to and
-including the last one with a significant p-value. We call this the
+including the last one with a significant $p$-value. We call this the
 \textit{last} stopping rule, and it selects the first $\hat k_\textit{last}$
 variables in the forward stepwise path where, if
-$p_j$ denotes our p-value calculated at the $j$-th step, then
+$p_j$ denotes the $T\chi$ $p$-value calculated at the $j$-th step, then
 \begin{equation}
@@
 sparsity $k$ and always picks the first $k$ variables. The \textit{first}
-rule stops immediately before the first variable to yield a p-value
+rule stops immediately before the first variable to yield a $p$-value
 larger than $\alpha$, and since this requires rejection of the global 
@@
 \hspace{-15pt}
-\subfigure[P-value with spline groups]{
+\subfigure[$p$-value with spline groups]{
 \label{fig:spline:pval}
@@
 term.
-The right panel shows our p-value marginally
+The right panel shows the $T\chi$ $p$-value marginally
 over each step, with red indicating the global null case, blue
 indicating a signal with $k=6$ nonzero groups, and green showing
-the nominal $\chi^2$ p-values also computed on the non-null case.}
+the nominal $\chi^2$ $p$-values also computed on the non-null case.}
 \label{fig:spline}
@@
   \begin{aligned}
-    X_g &\sim \text{Unif}[-1,1], \quad
+    X_g &\sim \text{Unif}([-1,1]), \quad
     X_g^s = \text{CubicSplineBasis}(X_g).
@@
 \hspace{-15pt}
-\subfigure[P-value with \textsc{glinternet}]{
+\subfigure[$p$-value with \textsc{glinternet}]{
 \label{fig:glint:pval}
@@
 definition of power described in the section.
-The right panel shows our p-value marginally
+The right panel shows the $T\chi$ $p$-value marginally
 over each step, with red indicating the global null case, blue
 indicating a signal with $k=6$ nonzero groups, and green showing
-the nominal $\chi^2$ p-values.}
+the nominal $\chi^2$ $p$-values.}
 \label{fig:glint}
@@
 the presence and type of certain genetic mutations. We attempt a similar
-analysis using forward stepwise, and report our P-value at each step.
+analysis using forward stepwise, and report the $T\chi$ $p$-value at each step.
 Categorical covariates are encoded as groups of dummy variables using
@@
 Each panel shows results for a different drug,
-with p-values plotted by step in forward stepwise. 
+with $p$-values plotted by step in forward stepwise. 
 The \textit{last} stopping rule \eqref{eq:klast} is applied and the region of the plot
@@
 this method seems to work well in all our simulations. By this we mean that
-the p-value tends to be small on the step when including the last truly
+the $T\chi$ $p$-value tends to be small on the step when including the last truly
 nonzero predictor, uniform on the following step, and subsequently
-stochastically larger than uniform. This is in contrast to p-values
+stochastically larger than uniform. This is in contrast to $p$-values
 calculated from traditional variable-inclusion tests like the $\chi^2$
@@
 roughly depends on the improvement gained by this variable compared to
-the next best variable. Thus, the p-value is small when there is a large
+the next best variable. Thus, the $T\chi$ $p$-value is small when there is a large
 gap between the variable to be included and all the remaining variables.
 If multiple predictors have truly nonzero coefficients that are close
-in magnitude, the p-value may be large until forward stepwise reaches
+in magnitude, the $p$-value may be large until forward stepwise reaches
 the last one. This motivated us to consider the \textit{last} stopping
@@
 
-The present work calculates p-values at each step ignoring the constraints
+The present work calculates $p$-values at each step ignoring the constraints
 imposed by previous steps. Future work adjusting for all previous steps
-is in progress, and may be able to give exact p-values at a step chosen
+is in progress, and may be able to give exact $p$-values at a step chosen
 stochastically by procedures like BIC and RIC. The authors will
@@
 \appendix
-\section{P-value algorithm}
+\section{Alternate derivation and algorithm for $T\chi$ test}
 \label{app:algo}
@@
 \begin{algorithm}
- \caption{Computing p-value}
+ \caption{Computing the $T\chi$ $p$-value}
  \label{algo:pval}
@@
    \REQUIRE Response $y$, grouped design matrix $X$ with weights, inactive set $A^c$, index $g$ of last group to enter active set $A$.
-   \ENSURE Our p-value for the group $g$ entering the model.
+   \ENSURE $T\chi$ $p$-value for the group $g$ entering the model.
    \STATE Compute $H_g$ and $\sigma^2_g$
@@
 groups have size 1, and the design matrix is orthonormal. By a
-transformation we can reduce the identity design matrix case
+transformation we can reduce to the identity design case where $n=p$,
+$X=I$, so
 \[
-y_i = \mu_i + \epsilon_i, \qquad \epsilon_i \sim N(0,1)
+y_i = \mu_i + \epsilon_i, \qquad \epsilon_i \sim N(0,1).
 \]
-without loss of generality. In this simple case our p-value can be
-calculated as
+In this simple case the $T\chi$ $p$-value can be calculated as
 \[

### 2014-05-14 paper/paper.tex
Paper possibly ready for arxiv

@@
 \usepackage{subfigure}
+%\usepackage{placeins}
 %\usepackage[boxed]{algorithm2e}
@@
 
-\newcommand{\jonathan}[1]{{\color{blue}  #1}}
-\newcommand{\josh}[1]{{\color{green} #1}}
-
 \begin{document}
@@
 % indicate corresponding author with \corref{}
-%\author{\fnms{Jonathan} \snm{Taylor}\corref{Jonathan E. Taylor}\ead[label=e1]{jonathan.taylor@stanford.edu}\thanksref{t1}
-%\and \fnms{Joshua} \snm{Loftus}}
-%\thankstext{t1}{Supported in part by NSF grant DMS 1208857 and AFOSR grant 113039.}
+\author{\fnms{Jonathan} \snm{Taylor}\corref{Jonathan E. Taylor}\ead[label=e1]{jonathan.taylor@stanford.edu}\thanksref{t1}
+\and \fnms{Joshua} \snm{Loftus}}
+\thankstext{t1}{Supported in part by NSF grant DMS 1208857 and AFOSR grant 113039.}
 
@@
 
+\renewcommand{\abstractname}{Acknowledgements}
+\begin{abstract}
+The authors would like to thank Robert Tibshirani and Trevor Hastie for many helpful comments.
+\end{abstract}
+
 %%%
@@
 
-As a small illustrative example of what we gain from this method,
+As an illustrative example of what we gain from this method,
 consider a response $Y$ with ten
@@
 with $n = 40$ observations. Running forward stepwise for eight steps,
-we calculated both the new $T\chi$ $p$-value and the usual
-$\chi^2$ $p$-value at each step. In
-Table~\ref{tab:example} we see that forward stepwise chooses the
+we calculated the new $T\chi$ $p$-value, the usual
+$\chi^2$ $p$-value based on drop in RSS at each step,
+and an estimate of an exact $p$-value based on comparing the norm
+achieved by the variable being added to the model, $\| X_{g^*}^T y \|_2$,
+to quantiles of the Monte Carlo sample
+\[
+\max_{g \in A^c} \| X_g^T z_m \|_2, \quad z_m \sim N(0, I) \quad \text{for } m = 1,\ldots,M
+\]
+with $A^c$ being the set of variables not yet included
+(updated \textit{after} computing this estimate, so that $g^* \in A^c$).
+
+\input{../tables/small.tex}
+
+In Table~\ref{tab:example} we see that forward stepwise chooses the
 truly nonzero variables first and both $p$-values for these are small.
 However, once forward stepwise begins adding noise variables, the
-$\chi^2$ $p$-value remains relatively small, potentially leading to 
-incorrect inferences.
-
-\input{../tables/small.tex}
+$\chi^2$ $p$-value remains small, potentially leading to 
+incorrect inferences. The $T\chi$ and MC-estimated exact $p$-values
+do not suffer from this selection effect.
+%However, the MC-estimate takes substantially more computational time to evaluate than the $T\chi$ statistic.
 
@@
 models, as well as to a real data example involving genomic prediction
-of individual drug responses to treat HIV.
+of individual drug responses and resistances for various mutations of HIV.
 
@@
 \includegraphics[width=0.9\textwidth]{../figs/pubmed.pdf}
-\caption{\small \em Forward stepwise enjoys widespread use among
-practitioners}
+\caption{\small \em PubMed search results show continued widespread usage of forward stepwise}
 \label{fig:pubmed}
@@
 \label{fig:fscat}
-\includegraphics[width=0.5\textwidth]{../figs/tpr_default_categorical_nsim100_n100_p400_g200_sizes2-2_k24_lower2_upper4.pdf}}
+\includegraphics[width=0.6\textwidth]{../figs/tpr_default_categorical_nsim100_n100_p400_g200_sizes2-2_k24_lower2_upper4.pdf}}
 \hspace{-15pt}
@@
 \label{fig:fsgauss}
-\includegraphics[width=0.5\textwidth]{../figs/tpr_default_gaussian_nsim200_n100_p200_g200_k25_lower1pt5_upper2.pdf}}
+\includegraphics[width=0.6\textwidth]{../figs/tpr_default_gaussian_nsim200_n100_p200_g200_k25_lower1pt5_upper2.pdf}}
 \caption{\small \em
@@
 
-\subsection{Quadratic framework for derivation of test statistic}
+\subsection{Quadratic framework and derivation of test statistic}
 
@@
 $$
-\begin{figure}
+\begin{figure}[!htp]
 \begin{center}
@@
 
-\begin{figure}[h]
+\begin{figure}[!htp]
 \begin{center}
@@
 \label{fig:gausscorr}
-\includegraphics[width=0.5\textwidth]{../figs/default_gaussian_nsim400_n100_p200_g200_k10_lower1pt5_upper2_corr0pt2_noisecorr0pt1.pdf}}
+\includegraphics[width=0.6\textwidth]{../figs/default_gaussian_nsim400_n100_p200_g200_k10_lower1pt5_upper2_corr0pt2_noisecorr0pt1.pdf}}
 \hspace{-15pt}
@@
 \label{fig:gaussgroups}
-\includegraphics[width=0.5\textwidth]{../figs/default_gaussian_nsim400_n100_p200_g35_sizes5-10_k5_lower2_upper3.pdf}}
+\includegraphics[width=0.6\textwidth]{../figs/default_gaussian_nsim400_n100_p200_g35_sizes5-10_k5_lower2_upper3.pdf}}
 \caption{\small \em
@@
 \label{fig:hivnrti}
-\includegraphics[width=0.5\textwidth]{../figs/default_fixed_nsim400_n848_p537_g82_sizes2-24_k5_lower4_upper8.pdf}}
+\includegraphics[width=0.6\textwidth]{../figs/default_fixed_nsim400_n848_p537_g82_sizes2-24_k5_lower4_upper8.pdf}}
 \hspace{-15pt}
@@
 \label{fig:gausscat}
-\includegraphics[width=0.5\textwidth]{../figs/default_categorical_nsim500_n100_p80_g40_sizes2-2_k10_lower2_upper4.pdf}}
+\includegraphics[width=0.6\textwidth]{../figs/default_categorical_nsim500_n100_p80_g40_sizes2-2_k10_lower2_upper4.pdf}}
 \caption{\small \em
@@
 
-\begin{figure}
+\begin{figure}[!htp]
 \begin{center}
-\subfigure[Power of forward stepwise with splines]{
+\subfigure[Power of forward stepwise with spline groups]{
 \label{fig:spline:fwd}
-\includegraphics[width=0.5\textwidth]{../figs/tpr_gamsel_uniform_nsim100_n100_p200_g200_k18_k01_lower5_upper10.pdf}}
+\includegraphics[width=0.6\textwidth]{../figs/tpr_gamsel_uniform_nsim100_n100_p200_g200_k18_k01_lower5_upper10.pdf}}
 \hspace{-15pt}
-\subfigure[$p$-value with spline groups]{
+\subfigure[$T\chi$ $p$-values with spline groups]{
 \label{fig:spline:pval}
-\includegraphics[width=0.5\textwidth]{../figs/gamsel_uniform_nsim200_n200_p50_g50_k6_k02_lower5_upper10.pdf}}
-\caption{\small \it The left panel shows the power of forward stepwise
+\includegraphics[width=0.6\textwidth]{../figs/gamsel_uniform_nsim200_n200_p50_g50_k6_k02_lower5_upper10.pdf}}
+\caption{\small \it The first panel shows the power of forward stepwise
 in this setting for various sparsity levels, with the dashed line
@@
 term.
-The right panel shows the $T\chi$ $p$-value marginally
+The second panel shows the $T\chi$ $p$-value marginally
 over each step, with red indicating the global null case, blue
@@
 
-First we consider a simple extension of the usual linear regression setting.
-Given a design matrix $X$, we can augment this by adding additional columns given by nonlinear functions of the original covariates.
+First we consider a simple extension of the usual linear regression setting to include nonlinear effects.
+Given a design matrix $X$, we augment this by adding additional columns given by nonlinear functions of the original covariates.
 Specifically, for each original covariate $X_g$ we compute a submatrix of spline basis vectors $X_g^s$.
@@
 \end{equation}
-with $\tilde G = 2G$ groups.
+with $\tilde G = 2G$ groups. Thus, at each step the procedure chooses between including a linear effect for each covariate or including a full spline basis if the true relationship is nonlinear.
 For simplicity we assume the original groups are all size 1, and use the B-spline basis.
@@
 Note that the varying group sizes present some difficulty to forward stepwise, which tends to first try linear approximations to nonzero spline groups before adding the true spline group.
-In Figure~\ref{fig:spline}(b) the average proportion of truly nonzero spline groups included in the first $k$ steps appears in parentheses next to the $k$-TPP (top right).
+In Figure~\ref{fig:spline:pval} the average proportion of truly nonzero spline groups included in the first $k$ steps appears in parentheses next to the $k$-TPP (top right).
 
@@
 model by group LASSO, we use forward stepwise on the expanded design
-matrix. The overlapping groups still guarantee that our fitted model
+matrix and call the resulting procedure FS-\textsc{glinternet}.
+The overlapping groups still guarantee that our fitted model
 satisfies the strong hierarchy condition.
@@
 
-\begin{figure}
+\begin{figure}[!htp]
 \begin{center}
-\subfigure[Power of forward stepwise with interactions]{
+\subfigure[Power of FS-\textsc{glinternet} procedure]{
 \label{fig:glint:fwd}
-\includegraphics[width=0.5\textwidth]{../figs/tpr_glint_categorical_nsim100_n100_p72_g36_sizes2-2_k15_k01_lower5_upper10.pdf}}
+\includegraphics[width=0.6\textwidth]{../figs/tpr_glint_categorical_nsim100_n100_p72_g36_sizes2-2_k15_k01_lower5_upper10.pdf}}
 \hspace{-15pt}
-\subfigure[$p$-value with \textsc{glinternet}]{
+\subfigure[$T\chi$ $p$-values with FS-\textsc{glinternet}]{
 \label{fig:glint:pval}
-\includegraphics[width=0.5\textwidth]{../figs/glint_gaussian_nsim200_n100_p30_g30_k6_k02_lower5_upper10.pdf}}
-\caption{\small \it The left panel shows the power of forward stepwise
+\includegraphics[width=0.6\textwidth]{../figs/glint_gaussian_nsim200_n100_p30_g30_k6_k02_lower5_upper10.pdf}}
+\caption{\small \it The first panel shows the power of forward stepwise
 on the \textsc{glinternet} problem
@@
 definition of power described in the section.
-The right panel shows the $T\chi$ $p$-value marginally
+The second panel shows the $T\chi$ $p$-value marginally
 over each step, with red indicating the global null case, blue
@@
 response/resistance were regressed on categorical covariates encoding
-the presence and type of certain genetic mutations. We attempt a similar
+the presence and type of genetic mutations in certain regions of the HIV
+genome. We attempt a similar
 analysis using forward stepwise, and report the $T\chi$ $p$-value at each step.
 Categorical covariates are encoded as groups of dummy variables using
-the full encoding, and weights are set to the square-root of the
-number of levels for that covariate. We perform forward stepwise once
+the full encoding, and these groups are normalized by their Frobenius
+norm. We perform forward stepwise once
 for each drug response, restricting to the subset of the data with
@@
 
-\begin{figure}
+\begin{figure}[!htp]
 \begin{center}
-\includegraphics[width=0.9\textwidth]{../figs/HIV_PI.pdf}
+\includegraphics[width=0.95\textwidth]{../figs/HIV_PI.pdf}
 \caption{\small \it Forward stepwise results from PI dataset}
@@
 
-\begin{figure}
+\begin{figure}[!htp]
 \begin{center}
-\includegraphics[width=0.9\textwidth]{../figs/HIV_NRTI.pdf}
+\includegraphics[width=0.95\textwidth]{../figs/HIV_NRTI.pdf}
 \caption{\small \it Forward stepwise results from NRTI dataset}
@@
 Figure~\ref{fig:HIVNRTI}. The PI data contains protease inhibitor
-resistance mutations, and the NRTI data contains nucleoside RT inhibitor
-resistance mutations.
+mutations, and the NRTI data contains nucleoside RT inhibitor mutations.
 Each panel shows results for a different drug,
@@
 corresponding to the chosen model is unshaded, with the remainder shaded.
-The chosen variables are shown in Tables~\ref{tab:hivPI} and
+The first several chosen variables are shown in Tables~\ref{tab:hivPI} and
 \ref{tab:hivNRTI}.
@@
 
-\begin{figure}
+\begin{figure}[!htp]
 \begin{center}
-\includegraphics[width=0.9\textwidth]{../figs/HIV_PI_glint.pdf}
+\includegraphics[width=0.95\textwidth]{../figs/HIV_PI_glint.pdf}
 \caption{\small \it \textsc{Glinternet} results from PI dataset}
@@
 
-\begin{figure}
+\begin{figure}[!htp]
 \begin{center}
-\includegraphics[width=0.9\textwidth]{../figs/HIV_NRTI_glint.pdf}
+\includegraphics[width=0.95\textwidth]{../figs/HIV_NRTI_glint.pdf}
 \caption{\small \it \textsc{Glinternet} results from NRTI dataset}
@@
 
-\newpage
-
 \bibliographystyle{ims}
@@
 \bibliography{paper}
-
-\newpage
 
@@
 
+\section{A note on parallel computation}
+\label{app:computation}
+
+One strength of forward stepwise is that it can be computed in parallel, hence can handle situations with many groups of predictor variables.
+We now describe, at very a high level, how to accomplish this.
+Suppose $m > 1$ machines are available for use.
+Partition the set of predictor variable groups, along with their respective weights, into $m$ disjoint blocks and store each block on its own machine.
+Now at each step of Algorithm~\ref{algo:fs}, the residual $r_{s-1}$ is broadcast to all machines, and each machine computes $\| X_g^T r_{s-1} \|_2/w_g$ for all the groups $g$ stored on that machine.
+Each machine reports its own maximum and maximizer $g^*$, and the global maximizer is found.
+The projector matrix $P_{g^*}$ is computed and broadcast to all machines which in turn use it to project all their predictors.
+
+We can also verify that Algorithms \ref{algo:pval} and \ref{algo:linfrac} only require block matrix operations with blocks given by individual groups
+or subproblems computable separately for each group. Hence, the $T\chi$ $p$-value can also be computed in parallel this way.
+
 \end{document}

### 2014-05-14 small.R

@@ -3,11 +3,14 @@ source("generate_data.R")
 source("fwd_step.R")
 source("tex_table.R")
 
+max.steps = 8
 n = 40
 groups = c(1,1,1,2,2,3,3,3,3,4,4,5,5,6,6,7,7,7,8,8,9,9,10,10,10,10)
 group.labels = unique(groups)
-max.steps = 8
+cat.groups = group.labels
+group.sizes = rle(groups)$lengths
 weights = rep(1, length(group.labels))
+
 X = categorical_design(n, groups)
 beta = rep(0, length(groups))
 beta[groups == 1] = c(1,.5,-1)
@@ -18,14 +21,77 @@ Y = Y - mean(Y)
 Sigma = diag(rep(1,n))
 Xnormed = frob_normalize(X, groups)
 
-results = forward_group(Xnormed, Y, groups, weights, Sigma, max.steps = max.steps, cat.groups = group.labels)
+active.set = 0
+eff.p = 0
+p.vals = c()
+chi.pvals = c()
+MC.pvals = c()
+c.vars = c()
+Ls = c()
+ranks = c()
+Ynorms = c()
+
+MC_pval = function(L, X, inactive.groups, inactive.labels, inactive.inds, M=200) {
+  n = nrow(X)
+  Ls = c()
+  for (iter in 1:M) {
+    epsilon = rnorm(n)
+    U = t(X[, inactive.inds]) %*% epsilon
+    terms = sapply(inactive.labels, function(x) sqrt(sum(U[inactive.groups==x]^2)))
+    Ls = c(Ls, max(terms))
+  }
+  q = sum(Ls > L)/M
+  return(q)
+}
+
+Y.update = Y
+X.update = X
+MC.time = 0
+Tx.time = 0
+
+for (i in 1:max.steps) {
+  # Tchi statistic computation
+  before.Tx.time = as.numeric(Sys.time())
+  output = add_group(X.update, Y.update, groups, weights, Sigma, active.set, eff.p, cat.groups = cat.groups)
+  Tx.time = Tx.time + as.numeric(Sys.time()) - before.Tx.time
+  L = output$test.output[[1]]
+  inactive.labels = setdiff(group.labels, active.set)
+  inactive.inds = sapply(groups, function(x) is.element(x, inactive.labels))
+  inactive.groups = groups[inactive.inds]
+  # MC statistic computation
+  before.MC.time = as.numeric(Sys.time())
[…]

### 2014-05-15 paper/paper.tex
Fixing rectangles

@@
 \usepackage{subfigure}
-%\usepackage{placeins}
-%\usepackage[boxed]{algorithm2e}
 
@@
 
-% "Title of the paper"
 \title{A significance test for forward stepwise model selection}
 
-%%%
 %%% \runtitle{} - I don't know what this is
-%%%
 
 % indicate corresponding author with \corref{}
-\author{\fnms{Jonathan} \snm{Taylor}\corref{Jonathan E. Taylor}\ead[label=e1]{jonathan.taylor@stanford.edu}\thanksref{t1}
-\and \fnms{Joshua} \snm{Loftus}}
+\author{\fnms{Joshua} \snm{Loftus}
+\and \fnms{Jonathan} \snm{Taylor}\corref{Jonathan E. Taylor}\ead[label=e1]{jonathan.taylor@stanford.edu}\thanksref{t1}}
 \thankstext{t1}{Supported in part by NSF grant DMS 1208857 and AFOSR grant 113039.}
@@
 
-%%%
 %%% \runauthor{Taylor} - I don't know what this is
-%%%
 
@@
 
-%%%
-%%% - Gotta change these??????????
-%%%
+%%% - Check these
 \begin{keyword}[class=AMS]
@@
 do not suffer from this selection effect.
-%However, the MC-estimate takes substantially more computational time to evaluate than the $T\chi$ statistic.
+However, the MC-estimate takes substantially more computational time to evaluate than the $T\chi$ statistic.
 
@@
 case.
-%Our variant of forward stepwise uses a different method than
-%usual for choosing which group to add because we are using the
-%same objective in the forward stepwise procedure that was used to
-%derive the test statistics in penalized regression settings.
 
@@
 the number of individual variables or columns in group $g$.
-% Do we need to assume $p_g \ll n$ or $\text{rank}(X_g) = p_g$?
 When a group encodes a categorical variable as indicators for the
@@
 which we denote \textit{steps}.
-%(consistent with the \textit{R} function for stepwise).
 To enable our test statistic computations,
@@
 The limiting results about later steps in \cite{significance:lasso} can
-%(formally)
 be interpreted as
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%% Related works %%%%%%%%%%%%%%%%%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% \begin{figure}
-% \label{fig:polytope}
-% \begin{center}
-% \resizebox{!}{3in}{\input{../figs/polytope.pdf_tex}}
-% \end{center}
-% A polytope selection event. Solving the LASSO \eqref{eq:lasso} for a fixed value
-% of $\lambda$ yields both an active set $\hat{E}(y)$ and a collection of signs of 
-% active variables $z_{\hat{E}(y)}$. The event $\{y: (\hat{E}(y), z_{\hat{E}(y)}) = (E, z_E)\}$
-% can be written as a set of affine constraints.
-% \end{figure}
-
 \subsection{Quadratic framework and derivation of test statistic}
@@
 satisfying a list of affine constraints.
-%stylized in Figure \ref{fig:polytope}.
 If all groups are of size 1, then
@@
 be given by a set of affine inequalities.
-%This is described in more detail in \cite{exact:lars}.
 After selection, exact inference for a (randomly) chosen contrast $\eta(y)^T\mu$ of the mean vector
@@
 
-%However, we must first modify our implementation of forward stepwise to prevent the possibility of including main effect groups after the main effects are already included in the model as part of an interaction group.
-To do this we compute the residual using the whole model in each step, rather than by subtracting the effect of the current group being included.
-For a set of group indices $A \subseteq \{ 1, \ldots G \}$ we write $X_A$ for the submatrix of the design matrix containing all the groups in $A$.
-% \begin{algorithm}
-% \DontPrintSemicolon
-% \caption{\small \it Modification to compute residual from full model}
-% \BlankLine
-% \CommentSty{\# Everything the same as before, additionally:}\;
-% %  $P_A \gets I - X_AX_A^\dagger$\;
-% %  $r_s \gets P_A r_{s-1}$\;
-% \If{$g^*$ is an interaction group}{
-%   $g_1, g_2 \gets$ main effects groups of $g^*$\;
-%   A^c \gets A^c \slash \{ g_1, g_2 \}\;
-% }%\;
-% %$r_s \gets y - \text{lsfit}(X_A, y)$\;
-% \BlankLine
-% \label{algo:fs:full}
-% \end{algorithm}
-% \begin{algorithm}
-% \DontPrintSemicolon
-% \KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ groups of variables}
-% \KwResult{Active set $A$ of variable groups included in the model}
-% \SetKwFunction{lsfitResidual}{lsfitResidual}
-% \caption{\small \it Forward stepwise for glinternet}
-% \BlankLine
-% Compute design matrix~(\ref{eq:glintmat}) and expanded grouping $\tilde G$\;
-% $A \gets \emptyset$\;
-% $A^c \gets \{ 1, \ldots, G \}$\;
-% $r_0 \gets y$\;
-% \For{$s \gets 1$ \KwTo $steps$}{
-%   $g^* \gets \argmax_{g \in A^c} \{ \norm{X_g^T r_{s-1}}_2 / w_g \}$\;
-%   $P \gets I - X_{g^*}X_{g^*}^\dagger$\;
-%   $A \gets A \cup \{ g^* \}$\;
-%   $A^c \gets A^c \backslash \{ g^* \}$\;
-%   \For{$g \in A^c$}{
-%     $X_g \gets P X_g$\;
-%   }
-%   $r_s \gets P r_{s-1}$\;
-% }
-% \KwRet{$A$}
-% \label{algo:fsgl}
-% \end{algorithm}
-
 To demonstrate this method by simulation, we constructed signals which have
@@
 interactions.
-%This way each nonzero main effect with any nonzero interactions has
-%exactly one nonzero interaction. 
 We also inflate each nonzero interaction
@@
 \begin{center}
-\includegraphics[width=0.95\textwidth]{../figs/HIV_PI.pdf}
+\includegraphics[width=0.95\textwidth]{../figs/HIV_PI_plot.pdf}
 \caption{\small \it Forward stepwise results from PI dataset}
@@
 \begin{center}
-\includegraphics[width=0.95\textwidth]{../figs/HIV_NRTI.pdf}
+\includegraphics[width=0.95\textwidth]{../figs/HIV_NRTI_plot.pdf}
 \caption{\small \it Forward stepwise results from NRTI dataset}
@@
 
-%\begin{figure}
-%\begin{center}
 \input{../tables/HIV_PI.tex}
-%\label{tab:HIVPI}
-%\end{center}
-%\end{figure}
-
-%\begin{figure}
-%\begin{center}
+
 \input{../tables/HIV_NRTI.tex}
-%\label{tab:HIVNRTI}
-%\end{center}
-%\end{figure}
 
@@
 \begin{center}
-\includegraphics[width=0.95\textwidth]{../figs/HIV_PI_glint.pdf}
+\includegraphics[width=0.95\textwidth]{../figs/HIV_PI_glint_plot.pdf}
 \caption{\small \it \textsc{Glinternet} results from PI dataset}
@@
 \begin{center}
-\includegraphics[width=0.95\textwidth]{../figs/HIV_NRTI_glint.pdf}
+\includegraphics[width=0.95\textwidth]{../figs/HIV_NRTI_glint_plot.pdf}
 \caption{\small \it \textsc{Glinternet} results from NRTI dataset}
@@
 the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
-%$\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
-%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
 For a vector $u \in \real^p$, let $u_h$ denote the coordinates
@@
 
-%We proceed by forming the $n \times p_g$ matrix
-%$X_g - X_g X_g^T y y^T X_g/\norm{X_g^Ty}_2^2$ and performing Gram-Schmidt
-%on the first $r-1$ columns. For precise theorems showing that this will
-%generally work, see \cite{outerp:indep}.
 We compute an orthonormal basis by Gram-Schmidt and form a
@@
 
-% \todo Clean this section \\
-% \todo Make $M=\lambda_2$ a special case corollary \\
-
-% For the sake of completeness this appendix provides full derivations
-% of closed forms for the quantities required to compute our
-% statistic. We require two facts from \cite{tests:adaptive}
-
-% A point $\eta \in \K$ maximizes $f_\eta$ over a convex
-% set $\K$ if and only if the following conditions hold: 
-% \begin{equation}
-% \label{eq:maxcond}
-% \grad f_{|T_{\eta}\K} = 0, \quad
-% \tf^{\eta}_{\eta} \geq \V^+_{\eta}, \quad
-% \tf^{\eta}_{\eta} \leq \V^-_{\eta}, \quad \text{and} \quad
-% \V^0_{\eta} \leq 0.
-% \end{equation}
-% The same equivalence holds true even when $\K$ is only locally
-% convex.
-
-% Write $M^{\pm}_{\eta, h}$ and $M^0_{\eta, h}$ as the corresponding suprema and infima over the restricted parameter set $S_h$.
-Note that the characterization of the global maximizer (Lemma 1, general paper) implies $M^0_{\eta, h} \leq 0$ and $M^+_{\eta, h} \leq M^-_{\eta, h}$ for all $h \neq g$.
-This will be used below to eliminate some degenerate cases of the optimization sub-problem on each $S_h$.
-
-% Now fix $h$, let $P_g = X_gX_g^T$ and define
-% \begin{align*}
-% a_h &= w_h^{-1} X_h^T (I-P_g) y \\
-% b_h &= w_g w_h^{-1} X_h^T P_g y / \norm{X_g^T y} \\
-% c_h &= a_h^T b_h / (\norm{a_h} \norm{b_h}) \\
-% K_h   &= \{ x : \norm{x} = 1, b_h^Tx > w \}
-% \end{align*}
-% Note that $K_h = \emptyset$ if and only if $\norm{b_h} < w$.
-We consider two cases.
-First, suppose $\norm{b_h} > w$.
-In this case we must rule out one sub-case, when $a_h$ is not in the polar cone of $K_h$.
-If this occurs, then on at least one side of the boundary of $K_h$ we have $a_h^Tx > 0$ there, so that the infimum inside $K$ is $-\infty$ but the supremum outside $K$ is $+\infty$.
-This violates the characterization of the global maximizer of the original process as described above, so this case cannot occur.
-
-% Now, if $a_h$ is in the polar cone of $K_h$ (and ruling out the probability zero case where $a_h^Tx = 0$ on the boundary of $K_h$), then the numerator $a_h^Tx < 0$ on $K_h$ so there is a finite positive infimum in the interior $K_h$.
-Similarly there is a finite positive supremum on the interior of $K_h^c$ (the numerator is negative near the boundary of $K_h$ by continuity).
-
-% To attain the infimum on $K_h$ requires making $x$ close to $b_h$ and simultaneously making the numerator closer to zero, so $x$ should be on the side of $b_h$ that is closer to $a_h$.
-To attain the supremum on $K_h^c$ requires making $x$ simultaneously close to $a_h$ (so that $a_h^Tx > 0$) and $b_h$ (to make the denominator small).
-In summary, both the infimum and supremum are attained with $x$ between $a_h$ and $b_h$, when the angle $\theta$ between $a_h$ and $b_h$ is larger than both the angle $\psi$ between $a_h$ and $x$ and the angle $\phi$ between $x$ and $b_h$.
-This allows the simplification $\phi = \theta - \psi$.
-This is easier to verify for the case when $\norm{b_h} < w$.
-
-% We have verified that solving both cases requires finding the maxima of the trigonometric form $\norm{a_h} \cos (\psi) / (w - \norm{b_h} \cos(\theta - \psi))$ of the linear fraction on the interiors of $K_h$ (if it is nonempty) and $K_h^c$.
-The derivative is proportional to
-% \[
-% \frac{\norm{b_h} \sin(\theta) - w \sin(\psi)}{w - \norm{b_h} \cos(\theta - \psi)}
-% \]
-% Since we have ruled out the boundary of $K_h$, we can ignore the denominator and focus on critical points corresponding to the angles $\psi^{\pm}$ (symmetric about $\pi/2$) where $\sin(\psi^{\pm}) = (\norm{b_h} / w) \sin (\theta)$.
-That is,
-% \[
-% \psi^{\pm} \in \arcsin \frac{\norm{b_h}}{w} \sqrt{1-c_h^2}
-% \]
-% Note that these critical points only exist if $\sin(\theta) < w/\norm{b_h}$, and since we have argued that the infimum and supremum are attained it follows that this condition is met in our case.
-Let $\psi^+$ denote the smaller of the two angles.
-If $K_h$ is nonempty, then $\psi^-$ gives the infimum in $K_h$ and $\psi^+$ the supremum on $K_h^c$.
-This is necessary since if $x$ is in $K_h$ then its angle from $a_h$ is larger than $\pi/2$ because $a_h$ is in the polar cone of $K_h$.
-If $K_h$ is empty, then the supremum is still attained at $\psi^+$ since the numerator is negative at $\psi^- > \pi/2$ and the denominator is always positive.
-
-% Finally we calculate the values of the linear fraction at $\psi^{\pm}$. Let us first record some facts:
-% \begin{align*}
-% 0 < \psi^+ < \pi/2, \quad \psi^- = \pi - \psi^+ \\
-% \cos(\theta) = c_h, \quad \sin(\theta) = \sqrt{1-c_h^2} \\
-% \sin(\psi^+) = \sin(\psi^-) = \frac{\norm{b_h}}{w} \sin(\theta) \\
-% \cos(\psi^+) = -\cos(\psi^-) = \sqrt{1 - (\norm{b_h}/w)^2(1-c_h^2)}
-% \end{align*}
-% Using the angle difference formula,
-% \begin{align*}
-% \cos(\theta - \psi^+) & = \frac{\norm{b_h}}{w} (1-c_h^2) +  c_h \cos(\psi^+) \\
-% & = \frac{\norm{b_h}}{w} (1-c_h^2) +  c_h \cos(\psi^+)
-% \end{align*}
-% Hence
-% \begin{align*}
-% M^+_{\eta, h} & = \frac{ \norm{a_h} c_h }{ w - (\norm{b_h}^2/w) (1-c_h^2) - \norm{b_h} c_h \cos(\psi^+) } \\
-% % & = 
-% % & = \frac{ w \norm{a} a_h^Tb_h / \norm{b_h} }{ w^2 - \norm{a}\norm{b}^2 + (a^Tb)/\norm{a} - a^Tb \cos \psi^+ }
-% \end{align*}
-% And $M^-_{\eta, h}$ is $\infty$ when $\norm{b_h} < w$ and otherwise given by
-% \begin{align*}
-% M^+_{\eta, h} & = \frac{ \norm{a_h} c_h }{ w - (\norm{b_h}^2/w) (1-c_h^2) + \norm{b_h} c_h \cos(\psi^+) } \\
-% \end{align*}
-
-% I haven't been able to algebraically simplify these yet to a form that allows my $\lambda_2$ proof below to go through. The version below just assumes $\norm{b_h} < w = 1$.
-
-% \hrule
-% \hrule
-% We can rewrite this by rationalizing the denominator:
-% \[
-% M =
-% \frac
-% { a_h^T b_h +  \sqrt{ a_h^T a_h (1-b_h^Tb_h) + (a_h^T b_h)^2 }  }
-% { 1-b_h^Tb_h }
-% \]
-% Leaving $M$ in this form we now consider $\lambda_2$. The KKT conditions for the group LASSO problem give (directly from Yuan and Lin)
-% \begin{align*}
-% \norm{ X_g^T(y - X \beta) } & \leq \lambda w_g \quad \forall \beta_g \neq 0 \\
-% \beta_g & = \left( 1 - \frac{ \lambda w_g }{ \norm{S_g} } \right)_+ S_g
-% \end{align*}
-% where $S_g = X_g^T(y - X \beta_{-g})$. These expressions simplify for $\lambda_2 \leq \lambda < \lambda_1$. Letting $g$ be the index of the first group to enter we now have $S_g = X_g^Ty$ 
-% \begin{align*}
-% \beta_g & = \left( 1 - \frac{ \lambda w_g }{ \norm{X_g^Ty} } \right) X_g^Ty \\
-% \lambda w_h & \geq \norm{ X_h^Ty - X_h^T P_g y (1-\lambda w_g / \norm{X_g^Ty})  } \quad \forall h \neq g
-% \end{align*}
-% Now let $\lambda = \lambda_2$, so the inequality above is strict for all other groups and equality is attained for the second group to enter. Hence $\lambda_2$ satisfies
-% \begin{align*}
-% \lambda_2 & = \max_{h \neq g} 
-% \norm{ X_h^T(I - P_g) y - \lambda_2 ( w_g X_h^T P_g y / \norm{X_g^Ty} )  } / w_h\\
-% & = \max_{h \neq g} \norm{ a_h - \lambda_2 b_h } 
-% \end{align*}
-
-% Write $\lambda_{2,h}$ as the positive root of the quadratic equation obtained by squaring the norm, so $\lambda_{2,h}^2 = \norm{ a_h - \lambda_2 b_h }^2$.
-Then $\lambda_2 = \max_{h \neq g} \lambda_{2,h}$.
-Solving this (note that $b_h^Tb_h < 1$), we find
-% \[
-% \lambda_{2,h} = \frac
-% { a_h^T b_h +  \sqrt{ a_h^T a_h (1-b_h^Tb_h) + (a_h^T b_h)^2 }  }
-% { 1-b_h^Tb_h }
-% \]
-% Hence when $h$ is the index of the second group to enter, we have $\lambda_2 = \lambda_{2,h} = M$.
-
 \section{A note on power}

### 2014-05-15 HIV/fwdstep_HIV.R

@@ -1,5 +1,5 @@
-db = "NRTI"
-#db = "PI"
+#db = "NRTI"
+db = "PI"
 
 binary.encoding = FALSE
 
@@ -122,7 +122,7 @@ for (respind in 1:nresps) {
 print(A.sets)
 print(sapply(A.sets, length))
 
-filename = paste0("figs/HIV_", db, ".pdf")
+filename = paste0("figs/HIV_", db, "_plot.pdf")
 pdf(filename)
 par(mfrow=c(2,3))
 for (i in 1:nresps) {
@@ -152,7 +152,7 @@ for (i in 1:nresps) {
     
     xvals = 1:length(yvals)
     plot(xvals, yvals, ylim=c(0,1), xlab="Step", ylab="P-values", main=names(A.sets)[i])
-    rect(k+.5, -.05, max.steps+.8, 1.05, col = "gray94", border = NA)
+    rect(k+.5, 0, max.steps+.8, 1, col = "gray94", border = NA)
     points(xvals, yvals)
     
     if (length(nonsignif) > 0) {

### 2014-05-15 HIV/glint_HIV.R

@@ -1,5 +1,5 @@
-#db = "PI"
-db = "NRTI"
+db = "PI"
+#db = "NRTI"
 
 binary.encoding = TRUE
 
@@ -172,7 +172,7 @@ for (i in 1:nresps) {
     
     xvals = 1:length(yvals)
     plot(xvals, yvals, ylim=c(0,1), xlab="Step", ylab="P-values", main=names(A.sets)[i])
-    rect(k+.5, -.05, max.steps+.8, 1.05, col = "gray94", border = NA)
+    rect(k+.5, 0, max.steps+.8, 1, col = "gray94", border = NA)
     points(xvals, yvals)
     
     if (length(nonsignif) > 0) {

### 2014-10-07 paper/paper.tex
Lots of changes, refactoring to package

--- 2014-05-13 paper/paper.tex
@@
 \begin{abstract}
-  We apply the methods developed by \cite{significance:lasso} and
+% perhaps de-emphasize group LASSO and emphasize forward stepwise?
+  We extend the methods developed by \cite{significance:lasso} and
   \cite{tests:adaptive} on significance tests for penalized
-  regression to forward stepwise model selection. A general framework
-  for selection procedures described by quadratic inequalities includes
-  a variant of forward stepwise with grouped variables, allowing us to
-  handle categorical variables and factor models. We provide an algorithm
-  to compute a new statistic with an exact null distribution conditional
-  on the outcome of the model selection procedure. This new
-  statistic, which we denote $T\chi$,
-  has a truncated $\chi$ distribution under the global null.
-  We apply this test in forward stepwise iteratively on the residual after
-  each step. The resulting method
-  has the computational strengths of stepwise selection and addresses
-  the problem of invalid test statistics due to model selection.
-  We illustrate the flexibility of this method by applying it to
-  several specialized applications of forward stepwise including a
-  hierarchical interactions model and a recently described additive model
-  that adaptively chooses between linear and nonlinear effects for
-  each variable. 
+  regression to stepwise model selection by iteratively applying
+  the global null hypothesis test on the residual in a forward stepwise
+  procedure. The resulting method
+  has the computational strengths of stepwise selection and partially
+  solves the problem of inflated test statistics due to model selection.
+  In describing our variant of forward stepwise selection, we introduce
+  a new class of selection procedures, described by quadratic inequalities and
+  provide an algorithm that can be used for exact post-selection inference for such problems.
+
+  We illustrate the flexibility of our procedure by applying it to
+  novel specialized applications of forward stepwise to a hierarchically
+  constrained interactions model and recently described flexible additive models
+  that choose between linear and nonlinear effects for each variable.
 \end{abstract}
@@
 variable in each step. Because of the stochastic nature of this
-algorithm---making use of the data to choose variables---the
-usual $\chi^2$ and $F$-tests for significance
-fail when a model has been selected this way. These tests will be
+algorithm--making use of the data to choose variables--
+the usual $\chi^2$ and $F$-tests for significance
+fail when a model has been selected this way and tend to be
 anti-conservative unless they are computed on a held-out validation
-dataset. This problem is one instance of the general problem of
-conducting inference and model selection using the same data, a problem
+dataset. This is one instance of the general problem of conducting
+inference and model selection using the same data, a problem
 of central importance on which some recent progress has been made.
@@
 used for valid significance testing when computed on the same
-data used for model selection, eliminating the need for data splitting.
-This is especially important in
-situations in which data splitting is not appropriate. For example,
-when categorical covariates have levels with very few observations
+data used for model selection, eliminating the need for data splitting
+or for situations in which data splitting is not appropriate. For example,
+when categorical covariates contain rare levels with very few observations
 it can be difficult or impossible to split the sample with an adequate
-number of such observations occurring in each split. Furthermore, even
-if sample splitting is possible, it sacrifices accuracy in the model
-selection procedure and power in the subsequent inferences.
+number of such observations occurring in each split.
 
@@
 
-As a small illustrative example of what we gain from this method,
-consider a response $Y$ with ten
+As a small illustrative example, consider a response $Y$ with ten
 categorical predictors $X_g$ having between 2 and 4 levels. The true 
@@
 categorical design matrix $X$ and created an instance of this example
-with $n = 40$ observations. Running forward stepwise for eight steps,
-we calculated both the new $T\chi$ $p$-value and the usual
-$\chi^2$ $p$-value at each step. In
+with $n = 40$ observations. Running forward stepwise, we calculated
+both our p-value and the usual $\chi^2$ p-value at each step. In
 Table~\ref{tab:example} we see that forward stepwise chooses the
-truly nonzero variables first and both $p$-values for these are small.
+truly nonzero variables first and both p-values for these are small.
 However, once forward stepwise begins adding noise variables, the
-$\chi^2$ $p$-value remains relatively small, potentially leading to 
+$\chi^2$ p-value remains relatively small, potentially leading to 
 incorrect inferences.
@@
 stepwise itself performs well, and that
-various stopping rules using our test statistic---including
-some from \cite{sequential:fdr}---appear promising. In
+various stopping rules using our test statistic--including
+some from \cite{sequential:fdr}--appear promising. In
 Section~\ref{sec:applications} we apply
@@
 \includegraphics[width=0.9\textwidth]{../figs/pubmed.pdf}
-\caption{\small \em Forward stepwise enjoys widespread use among
+\caption{\small \it Forward stepwise enjoys widespread use among
 practitioners}
@@
 size 1, this includes non-grouped situations as a special
-case.
-%Our variant of forward stepwise uses a different method than
-%usual for choosing which group to add because we are using the
-%same objective in the forward stepwise procedure that was used to
-%derive the test statistics in penalized regression settings.
-
-Let $y \in \real^n$ be $n$ i.i.d. measurements of the outcome variable.
+case. Our variant of forward stepwise uses a different method than
+usual for choosing which group to add because we are using the
+same objective in the forward stepwise procedure that was used to
+derive the test statistics in penalized regression settings.
+
+Denote $y \in \real^n$ for $n$ i.i.d. measurements of the outcome variable.
 Let an integer $G \geq 2$ be the number of groups of explanatory variables.
@@
 
-Denote by $p = \sum_{g=1}^Gp_g$
+Denote $p = \sum_{g=1}^Gp_g$,
 the total number of individual variables, so $p = G$ in the
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-We also allow weights $w_g$ for each group. These
+We normalize each group by its Frobenius norm, but this is not required
+especially since we also allow weights, $w_g$, for each group. These
 weights act like penalties or costs, so increasing $w_g$ makes it
 more difficult for the group $X_g$ to enter the model. The
-modeler can choose weights arbitrarily for calibration purposes,
-but throughout we set them all constant (equal to 1) and normalize
-groups by the Frobenius norm of their corresponding submatrices.
+modeler can choose weights arbitrarily, but we set them all to be
+1 since we already normalize by the Frobenius norm.
 
@@
 
-\subsection{Description of the forward stepwise algorithm}
+\subsection{Description of the algorithm}
 
 First, the user must specify the maximum number of steps allowed,
-which we denote \textit{steps}.
-%(consistent with the \textit{R} function for stepwise).
-To enable our test statistic computations,
+which we denote \textit{steps} (consistent with the \textit{R} function
+for stepwise). To enable our test statistic computations,
 \textit{steps} should be at most $\min (n, G) - 1$, but it is
@@
 to choose a model. \cite{sequential:fdr} examined some stopping rules
-using the asymptotic $p$-values of \cite{significance:lasso} and showed
+using the asymptotic p-values of \cite{significance:lasso} and showed
 their stopping rules control false discovery rate--the expected
@@
 \subfigure[Gaussian designs]{
-\label{fig:fsgauss}
+\label{fig:fsgausscorr}
 \includegraphics[width=0.5\textwidth]{../figs/tpr_default_gaussian_nsim200_n100_p200_g200_k25_lower1pt5_upper2.pdf}}
-\caption{\small \em
-  For various sparsity levels $k$, the True Positive Proportion of the
-  model selected by forward stepwise at $k$ steps.
-  The left panel shows the results of the simulation
+\caption{\small \it The left panel shows the results of the simulation
   using a categorical design consisting of 200 binary factors.
@@
 enough for some purposes. Finally, to serve as a comparison with existing
-commonly used model selection procedures, in Figure~\ref{fig:fsgauss}
+commonly used model selection procedures, in Figure~\ref{fig:fwdstepsim}(b)
 we show in green the TPP and median model sizes for models chosen by
@@
 
-\subsection{Background}
 In the ordinary least squares setting, a significance test for a
@@
 of the LASSO.
-Recall the LASSO estimator is given by
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{equation}
-\begin{aligned}
-\label{eq:lasso}
-\displaystyle \hat \beta(\lambda) = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
-   \lambda \| \beta \|_1 \\
-\end{aligned}
-\end{equation}
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-The following facts are summarized in \cite{significance:lasso,tibshirani_LASSO_uniqueness}.
+Writing $\hat \beta(\lambda)$ for the LASSO solution for a fixed value of $\lambda$, we recall
+the following facts summarized in \cite{significance:lasso,tibshirani_LASSO_uniqueness}.
 
@@
 
-Assuming groups of size one and $\Sigma = \sigma^2 I$, the covariance test is given at the first step by
+The covariance test (assuming groups of size one and $\Sigma = \sigma^2 I$ is given at the first step by
 \begin{equation}
@@
 \end{equation}
-This is a hypothesis test for including the first variable in the LASSO
-path, with large values of the test statistic being evidence against the
-global null.
 In \cite{tests:adaptive} it is pointed out 
@@
 \label{test:exact}
-\exp\left(- \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \right) \approx \frac{1 - \Phi(\lambda_1/\sigma)}{1 - \Phi(\lambda_2 / \sigma)} \overset{D}{=} \text{Unif}([0,1]).
-\end{equation}
-The right hand side has an exact, finite sample null distribution.
-Further, this new statistic can be understood as the survival function of $\lambda_1 = \|X^Ty\|_{\infty}$
+\exp\left(- \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \right) \approx \frac{1 - \Phi(\lambda_1/\sigma)}{1 - \Phi(\lambda_2 / \sigma)} \overset{D}{=} \text{Unif}(0,1).
+\end{equation}
+Further, the above test statistic can be understood as the survival function of $\lambda_1 = \|X^Ty\|_{\infty}$
 {\em conditional} on which variable achieves $\lambda_1$ as well as its sign.
@@
 recursively applying $T_1$ to the variables that had not previously been selected by LARS.
-In forward stepwise with groups of size 1, we recursively apply the right hand side test statistic in the same manner.
-With larger groups, we use an analogous test statistic to the right hand side of \eqref{test:exact}.
-The test statistic is presented as an example in \cite{tests:adaptive} though we re-derive it here in simpler form.
-The test draws inspiration
+
+In the setting of grouped variables, we use an analogous test statistic to
+the right hand side of \eqref{test:exact}. The test statistic is presented as an example
+in \cite{tests:adaptive} though we re-derive it here in simpler form. The test draws inspiration
 from the group LASSO.
 
-The \emph{group LASSO estimator} \citep{grouplasso,bakin:grouplasso},
-a generalization of the LASSO, is a solution to the following penalized
+Recall the LASSO estimator is given by
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{equation}
+\begin{aligned}
+\label{eq:lasso}
+\displaystyle \hat \beta_\lambda = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
+   \lambda \| \beta \|_1 \\
+\end{aligned}
+\end{equation}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+The \emph{group LASSO estimator}, a generalization of the LASSO,
+ is a solution to the following penalized
 least-squares convex problem
@@
 special case when all groups are of size 1, since then the
-penalty term is the $\ell_1$-norm of $\beta$. The solution path of the group
+penalty term is the $L^1$-norm of $\beta$. The solution path of the group
 LASSO has similar properties to the LASSO case, however it is not
@@
 test is that it conditions on both the maximizer $g^*$ and on the unit
-vector $\eta_{g^*}$. Most of the $p$-value computations will be done in
+vector $\eta_{g^*}$. Most of the p-value computations will be done in
 terms of these quantities.
@@
 
-\subsection{Quadratic framework for derivation of test statistic}
-
-In this section we derive the new $p$-value and describe how to compute it.
-First we give a brief summary of what follows; readers not interested in the full derivation can skip the rest of this section after they understand the summary.
-The event $E_g$ that forward stepwise Algorithm~\ref{algo:fs} selects a given group $g$ is equivalent to a set of quadratic inequalities involving $y$ (see \eqref{eq:first:quadratic} below).
-To derive a statistic related to the group $g$ being included, we choose a direction vector $\eta(g)$ in the relevant direction and study the distribution of $\eta(g)^Ty$ restricted to the event $E_g$.
-In Figure~\ref{fig:curved} below the event $E_g$ is the shaded region, and the direction $\eta$ determines a slice through this event.
-Finally, we solve for the amounts $t$ by which $y$ can be translated in the directions $\pm \eta$ and still satisfy the constraints imposed by $E_g$.
-These are the points where the slice intersects the boundary of $E_g$.
-Let $M \subset \mathbb R$ be the set of $t$ such that $y + t\cdot\eta \in E_g$.
-With all of these quantities and a little more work we find that the observed norm $\| X_g^Ty \|_2$ has a $\chi$ distribution truncated to the set $M$, and with a computable scale parameter.
-Applying the appropriate CDF transform \eqref{eq:tchi} with these quantities yields our $p$-value.
-
-We now give the full derivation of the $p$-value, beginning with a previous work and then extending the framework there to the setting with groups.
+\subsection{Quadratic selection events}
+
+Before giving the full description of the p-value, we turn to another
+previous work and describe how to extend the framework there to the
+setting with groups.
 The approach to post-selection inference in \cite{lasso:fixed}
@@
 
-However, with group sizes larger than 1, the event that the first group
-chosen $g^*(y)$ is equal to some fixed group $g$ is no longer given by
-a set of affine constraints. Rather, considering line 3 of
-Algorithm~\ref{algo:fs}, we see
+With non-trivial groups, however, the event that $g^*(y)$, the first
+group chosen, is equal to some fixed group $g$ is generally not given by
+a set of affine constraints. Rather, considering line 3 of Algorithm~\ref{algo:fs}, we see
 \begin{equation}
@@
 In the rest of this section, we consider an arbitrary selection procedure
-$S$ that returns one of a set of possible outcomes $s \in {\cal S}$ determined by a set of quadratic inequalities.
+$S$ that returns one of a set of possible outcomes $s \in {\cal S}$ by a set of quadratic inequalities.
 That is, for all possible outcomes $s \in {\cal S}$ there is a set of indices
@@
 As in \cite{lasso:fixed} we will be interested in some randomly chosen contrast
-$\eta(s)^T\mu$.
+$\eta(S(y))^T\mu$.
 In our grouped selection procedure, the selection rule is $S(y)=g^*(y)$
-and the value of $\eta$ on $\{ g^* = g \}$ is
-$$
-\eta(s) = \eta(g) = \frac{X_{g}^Ty}{\|X_{g}^Ty\|_2}.
-$$
-This choice of $\eta$ reflects our goal of calculating a $p$-value for inclusion of the variable $g^*$ chosen by forward stepwise.
-Note that for each fixed $g$ and any $\eta \in \real^g$ with $\|\eta\|_2=1$, the event 
+and $\eta$ is typically 
+$$
+\eta(S(y)) = \eta(g^*(y)) = \frac{X_{g^*(y)}^Ty}{\|X_{g^*(y)}^Ty\|_2}.
+$$
+Note that for each fixed $g$ and $\eta \in \real^g$ with $\|\eta\|_2=1$, the event 
 $$
@@
 $$
+
 \begin{figure}
+\label{fig:curved}
 \begin{center}
@@
 \end{center}
-\caption{ \em
-A quadratic selection event, given by the intersection of a list
-of quadratic inequalities. The selection event and chosen contrast
-$\eta$ determine a slice through the event with computable limits
-where the boundary of the event is reached.
-Note that the event need not be simply connected.
-}
-\label{fig:curved}
-\end{figure}
+A generic selection event, given by the intersection of a list
+of quadratic inequalities. Note that the event need not be simply connected. \end{figure}
+
 The slice for any particular inequality is
-\begin{equation}
+$$
 \begin{aligned}
-\label{eq:quadslice}
 \lefteqn{
@@
 \end{aligned}
-\end{equation}
+$$
 This can be computed explicitly, and results in either the empty set, a single interval (possibly infinite) or the union of 
@@
 \end{equation}
-In Figure \ref{fig:curved}, for every $\eta$, the slice \eqref{eq:slice} is a function
+In the picture \ref{fig:curved}, for every $\eta$, the slice \eqref{eq:slice} is a function
 of $P_{\eta}^{\perp}y$. This amounts to a proof of the following lemma.
@@
 
-In turn, this allows us to derive a test statistic to test the hypothesis
+In turn, this allows to derive a test statistic to test the hypothesis
 $H_{0,s}:X_s^T\mu=0$ conditional on $s$ being selected.
@@
 \label{eq:test:dbn}
-For $y \sim N(0, \sigma^2 I)$,
-conditional on the {\em event}
-\[
-H_{0,s} \cap E_+(s, \eta_s, P_s^{\perp}y, (P_s - \eta_s\eta_s^T)y)
-\]
+Conditional on $H_{0,s} \cap E_+(s, \eta_s, P_s^{\perp}y, (P_s - \eta_s\eta_s^T)y)$
 we have the following truncated $\chi$ distributional result
@@
 \end{equation}
-where the vertical bar $|$ here denotes restriction to an interval, and the scale is given by
+where the scale 
 $$
@@
 
-\subsection{The $T\chi$ test statistic}
-
-With Lemma \ref{eq:test:dbn} and Algorithm \ref{algo:quadratic} we are finally ready to define our test statistic, which we term the {\em truncated $\chi$ test statistic}, denoted $T\chi$.
-This statistic is determined by a degrees of freedom parameter $r$ as well as a truncation set $M \subset \real$, a scale parameter $\theta$ and an observed value $t$.
-The test statistic is given by the survival function of a $\chi_r$ random variable truncated to the set $M$ evaluated at our observed norm $\|X_g^Ty\|_2$ (after scaling by the appropriate scale $\theta$).
+\subsubsection{The $T\chi$ test statistic}
+
+This leads us to what we call the {\em  $T\chi$ test statistic} which is determined by
+a degrees of freedom parameter $r$ as well as a truncation set
+$M \subset \real$, a scale parameter $\theta$ and an observed value $t$.
+The test statistic is just the survival function of a $\chi_r$ random variable
+truncated to the set $M$ evaluated at our observed 
+norm $\|X_g^Ty\|_2$ (after scaling by the appropriate scale $\theta$).
 \begin{equation}
@@
 \end{equation}
-For simplicity we always use this distributional transform rather than reporting the observed values of $(t,r,\theta,M)$. Hence, $T\chi$ itself is our $p$-value, with
+
+%$\sigma \|X_s\eta_s\|_2$.
+
+%\subsection{Relation to covariance test}
+%
+%Consider the simple case for the first covariate to enter the model.
+%In this case the covariance test statistic is $T = \lambda_1(\lambda_1 - M)$.
+%\todo Does $M = \lambda_2$? Check proof, now that statistic has changed
+%
+%This is Lemma 5 from \cite{significance:lasso}.
+%
+%\begin{itemize}
+%
+%  \item As in LTTT, $T = \lambda_1 (\lambda_1 - M)$ and $M = \lambda_2$
+%  \item Convergence to the limiting Exp(1) distribution is too slow
+%    \[
+%      \frac{P(\chi_k / w_g \geq m + t/m)}{P(\chi_k / w_g \geq m)} \to e^{-t} \text{ as } m \to \infty
+%    \]
+%    (when the group achieving $\lambda_1$ is group $g$ and has rank $k$)
+%  \item The limiting distribution only depends on $T$, but we also observe $M$
+%  \item Let's just try the ratio (conditional $\chi_k$ tail probability) evaluated at $T$ and $M$ (it works better)
+%  \item Going one step further, instead of using the approximation (see LTTT Proof of Lemma 5)
+%    \[
+%      \frac{M + \sqrt{M^2+4t}}{2} \approx M + \frac{t}{M}
+%    \]
+%    we can just use the left hand side
+%  \item For $T = \lambda_1(\lambda_1 - M)$ the left hand side simplifies to $\lambda_1$
+%  \item Now our p-value is
+%    \[
+%      \frac{P(\chi_k / w_g \geq \lambda_1)}{P(\chi_k / w_g \geq \lambda_2)}
+%    \]   
+%\end{itemize}
+%
+%
+
+\subsection{Computation of p-value}
+
+Lemma \ref{eq:test:dbn} and Algorithm \ref{algo:quadratic} give an explicit characterization of our p-value
+that can also be used to calculate it in practice. It can also be applied to different
+quadratic selection procedures besides the variant of forward
+stepwise we consider here. For our purposes,
+this truncation interval 
+section was derived as a special case of the
+Kac-Rice test  defined in \cite{tests:adaptive} applied to the first step of the group LASSO path.
+
+The implementation described in Algorithm \ref{algo:pval}
+was used in all simulations.
+For the rest of this section let $g = \gstar$ be the index of
+the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
+%$\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
+%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
+For a vector $u \in \real^p$, let $u_h$ denote the coordinates
+of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
+We can rearrange the columns of $X$ to group these adjacently, so that
 \[
-T\chi \sim \text{Unif}([0,1])
+u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
 \]
-under the global null.
-
-After adding the first group in forward stepwise and computing the $T\chi$ $p$-value, we orthogonalize the response $y$ and all remaining groups with respect to the group just added.
-This imposes further constraints that in principle should be used when computing $p$-values at subsequent steps.
-For now we ignore these constraints and iterate the global null test, but work on incorporating all known constraints is ongoing.
-We believe that the effect of not tracking all constraints is what causes the nominal $p$-value to be increasingly stochastically larger than uniform further down the forward stepwise path (as can be seen in Figure~\ref{fig:pval1}), and only truly uniform at the first step where all remaining variables are noise.
-
-For an alternate derivation that follows more closely \cite{tests:adaptive}, and an algorithm that includes weights and an arbitrary (known) covariance matrix $\Sigma$ see Appendix~\ref{app:algo}.
+One step of the calculation will be to find an orthonormal basis for the
+linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
+\text{ for all } h \neq g \}$ so that we can project orthogonally to
+this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
+should be the case since $g$ maximizes the absolute value of this
+quantity), then the
+space is trivial and the desired orthogonal projection is the identity.
+Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
+generally has dimension $p_g-1$.
+
+\begin{algorithm}
+ \caption{Computing p-value}
+ \label{algo:pval}
+ \begin{algorithmic}
+   \REQUIRE Response $y$, grouped design matrix $X$ with weights, inactive set $A^c$, index $g$ of last group to enter active set $A$.
+   \ENSURE Our p-value for the group $g$ entering the model.
+   \STATE Compute $H_g$ and $\sigma^2_g$
+   \IF{$p_g = 1$}
+   \STATE $\sigma_g^2 \gets X_g^T\Sigma X_g/w_g^2$
+   \STATE $\tilde X_g \gets X_g / w_g \cdot \text{sign}(X_g^Ty)$
+   \ELSE
+   \STATE $P_g \gets \Sigma XV_g (V_g^T X^T \Sigma X V_g)^\dagger V_g^TX^T$
+   \STATE $H_g \gets (I - P_g)\Sigma$
+   \STATE $\sigma^2_g \gets y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2)$
+   \STATE $\tilde X_g \gets H_g X_g X_g^T y /(\norm{X_g^Ty}_2 w_g)$
+   \ENDIF
+   \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$
+   \STATE \COMMENT{Compute the following two $p$-vectors}
+   \STATE $a \gets X^T(y - \lambda \tilde X_g)$
+   \STATE $b \gets X^T \tilde X_g$
+   \STATE Compute solution $(v_+, v_-)$ of \textsc{LinearFractional}(a,b) optimization subproblem
+   \STATE $r \gets \text{rank}(X_g)$
+   \RETURN $T\chi(\|X_g^Ty\|_2, r, \sigma_g,  (v_-,v_+)).$
+ \end{algorithmic}
+\end{algorithm}
+
+%We proceed by forming the $n \times p_g$ matrix
+%$X_g - X_g X_g^T y y^T X_g/\norm{X_g^Ty}_2^2$ and performing Gram-Schmidt
+%on the first $r-1$ columns. For precise theorems showing that this will
+%generally work, see \cite{outerp:indep}.
+We compute an orthonormal basis by Gram-Schmidt and form a
+$p_g \times p_g$ matrix, which we denote $V_g$, by appending
+0's in the coordinates corresponding to all groups $h \neq g$ and
+an additional column of zeroes (since Gram-Schmidt only produces $p_g-1$).
+We can now define the projection
+
+\begin{equation}
+ \begin{aligned}
+   \label{eq:proj}
+   P_g &= \Sigma X_gV_g (V_g^T X_g^T \Sigma X_g V_g)^\dagger V_g^TX_g^T. \\
+  \end{aligned}
+\end{equation}
+Also define $H_g = (I-P_g)\Sigma$ and the conditional variance
+
+\begin{equation}
+ \begin{aligned}
+   \label{eq:cvar}
+   \sigma^2_g &= y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2) \\
+ \end{aligned}
+\end{equation}
+These simplify when $X_g$ is a single column, in which case $P_g = 0$,
+$H_g = \Sigma$, and $\sigma^2_g = X_g^T\Sigma X_g^T/w_g^2$.
+Note that $F_{\chi^2_r}$ denotes the distribution function of a $\chi^2$
+random variable with $r$ degrees of freedom. This completes the
+description of Algorithm~\ref{algo:pval} which we use to compute our
+p-value. Additional details, including the statement and solution of
+the \textsc{LinearFractional} subproblem, appear in Appendix~\ref{app:closedform}.
 
@@
 
-To study the behavior of the $T\chi$ test after taking steps in the forward stepwise path and to understand its power to detect various departures from the global null we conduct simulations in a wide variety of settings including both $p > n$ and $p < n$ problems.
 We performed simulations with several classes of random design matrices
@@
 from the HIV data set of Section~\ref{sec:hiv}. Gaussian
-design matrices were generated either independently or with some global
+design matrices are generated either independently or with some global
 correlation $\rho > 0$ between all pairs of columns. Categorical matrices
@@
 nonzero group have roughly the same magnitude and are normalized
-so that $\| \beta_g \|_2$ has scale $\gamma$ independently of $p_g$.
+so that $\| \beta_g \|_2$ has the same scale independent of $p_g$.
 The magnitudes for different
@@
 limit times $\gamma$. In text above each plot, the limits are listed
-next to ``beta'' and the numbers in parentheses are the largest
+next to ``beta:'' and the numbers in parentheses are the largest
 and smallest 2-norms of nonzero coefficient groups respectively.
@@
 \label{fig:gaussgroups}
-\includegraphics[width=0.5\textwidth]{../figs/default_gaussian_nsim400_n100_p200_g35_sizes5-10_k5_lower2_upper3.pdf}}
-\caption{\small \em
-Forward stepwise with $T\chi$ simulation results.
-Dashed line is the local True Positive Rate at each step averaged over
-400 simulation realizations.
-Vertical lines
-are boxplots of $p$-values marginally at each step, with red computed
-using $T\chi$ on a null model, blue using $T\chi$ on a sparse alternative
-model, and green using $\chi^2$ on the same. The shaded region indicates
-the sparsity of the alternative.
-Left panel shows results for Gaussian design
+\includegraphics[width=0.5\textwidth]{../figs/default_gaussian_nsim500_n100_p200_g35_sizes5-10_k5_lower2_upper3.pdf}}
+\caption{\small \it The left panel shows results for Gaussian design
 matrices with equicorrelation $\rho=0.2$ and equicorrelated noise with
-correlation 0.1. Right panel shows results for an independent
+correlation 0.1. The right panel shows results for an independent
 Gaussian design with groups of sizes 5 and 10.}
-\label{fig:pval1}
 \end{center}
@@
 was added to the model at that step. Solid verticle lines are essentially
-narrow boxplots, showing the middle 50\% of $p$-values calculated at that
+narrow boxplots, showing the middle 50\% of p-values calculated at that
 step with circular points in these lines as the average and triangles
 showing the 95\% and 5\% quantiles. Different color verticle lines
-represent the following: red are calculated using the $T\chi$ $p$-value
-on a null model with no nonzero groups, blue are calculated using the
-$T\chi$ $p$-value on the non-null model, and green calculated on the
-non-null model using the usual $\chi^2$ significance test.
+represent the following: red are calculated using our p-value on a null
+model with no nonzero groups, blue are calculated on the non-null
+model, and green calculated on the non-null model using the usual
+$\chi^2$ significance test.
 
@@
 \label{fig:hivnrti}
-\includegraphics[width=0.5\textwidth]{../figs/default_fixed_nsim400_n848_p537_g82_sizes2-24_k5_lower4_upper8.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/default_fixed_nsim100_n639_p354_g177_sizes2-2_k5_lower5_upper10.pdf}}
 \hspace{-15pt}
@@
 \label{fig:gausscat}
-\includegraphics[width=0.5\textwidth]{../figs/default_categorical_nsim500_n100_p80_g40_sizes2-2_k10_lower2_upper4.pdf}}
-\caption{\small \em
-As in Figure \ref{fig:pval1}.
-The left panel shows results for a categorical design
+\includegraphics[width=0.5\textwidth]{../figs/default_categorical_nsim500_n100_p80_g40_sizes2-2_k5_lower2_upper4.pdf}}
+\caption{\small \it The left panel shows results for a categorical design
 taken from a real data set.
-The right panel shows results for categorical matrices with all categories having 5 levels.
-The null cases being larger than uniform are likely the result of numerical error as these problems are poorly conditioned.
-}
+The right panel shows results for categorical matrices with all categories having 5 levels.}
 \end{center}
@@
 index reaches $k$ and then it will be close to 0. When this is the case,
-the $T\chi$ $p$-value tends to be small while forward stepwise is selecting truly
+our p-value tends to be small while forward stepwise is selecting truly
 nonzero variables, uniform on the step where the dashed line goes to 0,
@@
 on this behavior is to pick the model including all variables up to and
-including the last one with a significant $p$-value. We call this the
+including the last one with a significant p-value. We call this the
 \textit{last} stopping rule, and it selects the first $\hat k_\textit{last}$
 variables in the forward stepwise path where, if
-$p_j$ denotes the $T\chi$ $p$-value calculated at the $j$-th step, then
+$p_j$ denotes our p-value calculated at the $j$-th step, then
 \begin{equation}
@@
 sparsity $k$ and always picks the first $k$ variables. The \textit{first}
-rule stops immediately before the first variable to yield a $p$-value
+rule stops immediately before the first variable to yield a p-value
 larger than $\alpha$, and since this requires rejection of the global 
@@
 \hspace{-15pt}
-\subfigure[$p$-value with spline groups]{
+\subfigure[P-value with spline groups]{
 \label{fig:spline:pval}
@@
 term.
-The right panel shows the $T\chi$ $p$-value marginally
+The right panel shows our p-value marginally
 over each step, with red indicating the global null case, blue
 indicating a signal with $k=6$ nonzero groups, and green showing
-the nominal $\chi^2$ $p$-values also computed on the non-null case.}
+the nominal $\chi^2$ p-values also computed on the non-null case.}
 \label{fig:spline}
@@
 Given a design matrix $X$, we can augment this by adding additional columns given by nonlinear functions of the original covariates.
-Specifically, for each original covariate $X_g$ we compute a submatrix of spline basis vectors $X_g^s$.
-Adding these groups of spline basis vectors to the original design matrix yields a new design matrix
+Specifically, we compute spline basis vectors for each covariate and create a group given by this spline basis.
+Denote this new design matrix
 \begin{equation}
@@
   \begin{aligned}
-    \tilde X &=  \begin{pmatrix} X_1 & \cdots & X_G & X^s_1 & \cdots & X^s_G  \end{pmatrix}\\
+    \tilde X &=  \begin{pmatrix} X_1 & \cdots & X_G & X^+_1 & \cdots & X^+_G  \end{pmatrix}\\
+    &= \begin{pmatrix} X & X^+ \end{pmatrix}
   \end{aligned}
@@
   \begin{aligned}
-    X_g &\sim \text{Unif}([-1,1]), \quad
-    X_g^s = \text{CubicSplineBasis}(X_g).
+    X_i &\sim \text{Unif}[-1,1], \quad
+    X_i^+ = \text{CubicSplineBasis}(X_i)
   \end{aligned}
@@
 \hspace{-15pt}
-\subfigure[$p$-value with \textsc{glinternet}]{
+\subfigure[P-value with \textsc{glinternet}]{
 \label{fig:glint:pval}
-\includegraphics[width=0.5\textwidth]{../figs/glint_gaussian_nsim200_n100_p30_g30_k6_k02_lower5_upper10.pdf}}
+\includegraphics[width=0.5\textwidth]{../figs/glint_gaussian_nsim100_n100_p30_g30_k6_k02_lower4_upper8.pdf}}
 \caption{\small \it The left panel shows the power of forward stepwise
@@
 definition of power described in the section.
-The right panel shows the $T\chi$ $p$-value marginally
+The right panel shows our p-value marginally
 over each step, with red indicating the global null case, blue
 indicating a signal with $k=6$ nonzero groups, and green showing
-the nominal $\chi^2$ $p$-values.}
+the nominal $\chi^2$ p-values.}
 \label{fig:glint}
@@
 the presence and type of certain genetic mutations. We attempt a similar
-analysis using forward stepwise, and report the $T\chi$ $p$-value at each step.
-Categorical covariates are encoded as groups of dummy variables using
+analysis using forward stepwise, and report our P-value at each step.
+Categorical covariates are encoded as groups of binary variables using
 the full encoding, and weights are set to the square-root of the
@@
 Each panel shows results for a different drug,
-with $p$-values plotted by step in forward stepwise. 
-The \textit{last} stopping rule \eqref{eq:klast} is applied and the region of the plot
+with p-values plotted by step in forward stepwise. 
+The \textit{last} stopping rule is applied and the region of the plot
 corresponding to the chosen model is unshaded, with the remainder shaded.
@@
 this method seems to work well in all our simulations. By this we mean that
-the $T\chi$ $p$-value tends to be small on the step when including the last truly
+the p-value tends to be small on the step when including the last truly
 nonzero predictor, uniform on the following step, and subsequently
-stochastically larger than uniform. This is in contrast to $p$-values
+stochastically larger than uniform. This is in contrast to p-values
 calculated from traditional variable-inclusion tests like the $\chi^2$
@@
 roughly depends on the improvement gained by this variable compared to
-the next best variable. Thus, the $T\chi$ $p$-value is small when there is a large
+the next best variable. Thus, the p-value is small when there is a large
 gap between the variable to be included and all the remaining variables.
 If multiple predictors have truly nonzero coefficients that are close
-in magnitude, the $p$-value may be large until forward stepwise reaches
+in magnitude, the p-value may be large until forward stepwise reaches
 the last one. This motivated us to consider the \textit{last} stopping
@@
 
-The present work calculates $p$-values at each step ignoring the constraints
+The present work calculates p-values at each step ignoring the constraints
 imposed by previous steps. Future work adjusting for all previous steps
-is in progress, and may be able to give exact $p$-values at a step chosen
-stochastically by procedures like BIC and RIC. The authors will
-release an R package implementing these methods.
+is in progress, and may be able to give exact p-values at a step chosen
+stochastically by procedures like BIC and RIC. The authors hope to
+eventually release an R package implementing these methods.
 
@@
 \appendix
-\section{Alternate derivation and algorithm for $T\chi$ test}
-\label{app:algo}
-
-\begin{algorithm}
- \caption{Computing the $T\chi$ $p$-value}
- \label{algo:pval}
- \begin{algorithmic}
-   \REQUIRE Response $y$, grouped design matrix $X$ with weights, inactive set $A^c$, index $g$ of last group to enter active set $A$.
-   \ENSURE $T\chi$ $p$-value for the group $g$ entering the model.
-   \STATE Compute $H_g$ and $\sigma^2_g$
-   \IF{$p_g = 1$}
-   \STATE $\sigma_g^2 \gets X_g^T\Sigma X_g/w_g^2$
-   \STATE $\tilde X_g \gets X_g / w_g \cdot \text{sign}(X_g^Ty)$
-   \ELSE
-   \STATE $P_g \gets \Sigma XV_g (V_g^T X^T \Sigma X V_g)^\dagger V_g^TX^T$
-   \STATE $H_g \gets (I - P_g)\Sigma$
-   \STATE $\sigma^2_g \gets y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2)$
-   \STATE $\tilde X_g \gets H_g X_g X_g^T y /(\norm{X_g^Ty}_2 w_g)$
-   \ENDIF
-   \STATE $\lambda \gets \norm{X_g^Ty}_2/w_g$
-   \STATE \COMMENT{Compute the following two $p$-vectors}
-   \STATE $a \gets X^T(y - \lambda \tilde X_g)$
-   \STATE $b \gets X^T \tilde X_g$
-   \STATE Compute solution $(v_-, v_+)$ of \textsc{LinearFractional}(a,b) optimization subproblem
-   \STATE $r \gets \text{rank}(X_g)$
-   \STATE $u \gets [F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(\lambda^2/\sigma_g^2)]/[F_{\chi^2_r}(v^2_-/\sigma_g^2) - F_{\chi^2_r}(v^2_+/\sigma_g^2)]$
-   \RETURN $u$
- \end{algorithmic}
-\end{algorithm}
-
-In this section we describe an alternative derivation following the
-discussion of group LASSO in \cite{tests:adaptive}. The implementation
-described here was used in all simulations.
-For the rest of this section let $g = \gstar$ be the index of
-the group attaining the maximum on line 3 of Algorithm~\ref{algo:fs}.
-%$\lambda = \norm{X_g^Ty}_2/w_g$ be the value of the maximum, and
-%$r = \text{rank}(X_g)$ be the rank of the maximizing group.
-For a vector $u \in \real^p$, let $u_h$ denote the coordinates
-of $u$ corresponding to the columns of group $h$ in the design matrix $X$.
-We can rearrange the columns of $X$ to group these adjacently, so that
-\[
-u^TX^T = \begin{pmatrix} u_1^T X_1^T & u_2^T X_2^T & \cdots & u_G^TX_G^T \end{pmatrix}
-\]
-One step of the calculation will be to find an orthonormal basis for the
-linear space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0
-\text{ for all } h \neq g \}$ so that we can project orthogonally to
-this space. If $X_g$ is a single column and $X_g^Ty \neq 0$ (which
-should be the case since $g$ maximizes the absolute value of this
-quantity), then the
-space is trivial and the desired orthogonal projection is the identity.
-Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$
-generally has dimension $p_g-1$.
-
-%We proceed by forming the $n \times p_g$ matrix
-%$X_g - X_g X_g^T y y^T X_g/\norm{X_g^Ty}_2^2$ and performing Gram-Schmidt
-%on the first $r-1$ columns. For precise theorems showing that this will
-%generally work, see \cite{outerp:indep}.
-We compute an orthonormal basis by Gram-Schmidt and form a
-$p_g \times p_g$ matrix, which we denote $V_g$, by appending
-0's in the coordinates corresponding to all groups $h \neq g$ and
-an additional column of zeroes (since Gram-Schmidt only produces $p_g-1$).
-We can now define the projection
-
-\begin{equation}
- \begin{aligned}
-   \label{eq:proj}
-   P_g &= \Sigma X_gV_g (V_g^T X_g^T \Sigma X_g V_g)^\dagger V_g^TX_g^T. \\
-  \end{aligned}
-\end{equation}
-Also define $H_g = (I-P_g)\Sigma$ and the conditional variance
-
-\begin{equation}
- \begin{aligned}
-   \label{eq:cvar}
-   \sigma^2_g &= y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2) \\
- \end{aligned}
-\end{equation}
-These simplify when $X_g$ is a single column, in which case $P_g = 0$,
-$H_g = \Sigma$, and $\sigma^2_g = X_g^T\Sigma X_g^T/w_g^2$.
-Note that $F_{\chi^2_r}$ denotes the distribution function of a $\chi^2$
-random variable with $r$ degrees of freedom.
-
-Next we describe the \textsc{LinearFractional} optimization
+\section{Linear fractional subproblem}
+\label{app:closedform}
+
+This appendix describes the \textsc{LinearFractional} optimization
 subproblem and its solution.
@@
 groups have size 1, and the design matrix is orthonormal. By a
-transformation we can reduce to the identity design case where $n=p$,
-$X=I$, so
+transformation we can reduce the identity design matrix case
 \[
-y_i = \mu_i + \epsilon_i, \qquad \epsilon_i \sim N(0,1).
+y_i = \mu_i + \epsilon_i, \qquad \epsilon_i \sim N(0,1)
 \]
-In this simple case the $T\chi$ $p$-value can be calculated as
+without loss of generality. In this simple case our p-value can be
+calculated as
 \[

### 2014-10-07 HIV/fwdstep_HIV.R

@@ -1,5 +1,5 @@
-#db = "NRTI"
-db = "PI"
+db = "NRTI"
+#db = "PI"
 
 binary.encoding = FALSE
 
@@ -122,7 +122,7 @@ for (respind in 1:nresps) {
 print(A.sets)
 print(sapply(A.sets, length))
 
-filename = paste0("figs/HIV_", db, "_plot.pdf")
+filename = paste0("figs/HIV_", db, ".pdf")
 pdf(filename)
 par(mfrow=c(2,3))
 for (i in 1:nresps) {
@@ -152,7 +152,7 @@ for (i in 1:nresps) {
     
     xvals = 1:length(yvals)
     plot(xvals, yvals, ylim=c(0,1), xlab="Step", ylab="P-values", main=names(A.sets)[i])
-    rect(k+.5, 0, max.steps+.8, 1, col = "gray94", border = NA)
+    rect(k+.5, -.05, max.steps+.8, 1.05, col = "gray94", border = NA)
     points(xvals, yvals)
     
     if (length(nonsignif) > 0) {

### 2014-10-07 HIV/glint_HIV.R

@@ -1,5 +1,5 @@
-db = "PI"
-#db = "NRTI"
+#db = "PI"
+db = "NRTI"
 
 binary.encoding = TRUE
 
@@ -172,7 +172,7 @@ for (i in 1:nresps) {
     
     xvals = 1:length(yvals)
     plot(xvals, yvals, ylim=c(0,1), xlab="Step", ylab="P-values", main=names(A.sets)[i])
-    rect(k+.5, 0, max.steps+.8, 1, col = "gray94", border = NA)
+    rect(k+.5, -.05, max.steps+.8, 1.05, col = "gray94", border = NA)
     points(xvals, yvals)
     
     if (length(nonsignif) > 0) {

### 2014-10-07 plots.R

@@ -1,4 +1,4 @@
-
+library(ggplot2)
 
 step_plot = function(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, max.steps, upper.coeff, lower.coeff, max.beta, min.beta, fwd.power, design, filename,  main.append = "") {
 
@@ -6,27 +6,28 @@ step_plot = function(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, max
     #filename = paste0('figs/', design, "_n", n, '_p', p, '_g', g,
     #    '_k', k, '_lower', lower.coeff, '_upper', upper.coeff)
     #filename = paste0(filename, fn.append)
-    #filename = paste0(gsub(".", "pt", filename, fixed=TRUE), ".pdf")
-    filename = paste0("figs/", filename, ".pdf")
+    #filename = paste0(gsub(".", "pt", filename, fixed=TRUE), ".eps")
+    filename = paste0("art/", filename, ".eps")
 
-    bar.quantiles <- c(.25, .75)
-    point.quantiles <- c(.05, .95)
-    null.Pvals <- colMeans(null.p)
-    null.Pvals.bar <- apply(null.p, 2, function(col) quantile(col, probs = bar.quantiles))
-    null.Pvals.point <- apply(null.p, 2, function(col) quantile(col, probs = point.quantiles))
+    ## bar.quantiles <- c(.25, .75)
+    ## point.quantiles <- c(.05, .95)
+    ## null.Pvals <- colMeans(null.p)
+    ## null.Pvals.bar <- apply(null.p, 2, function(col) quantile(col, probs = bar.quantiles))
+    ## null.Pvals.point <- apply(null.p, 2, function(col) quantile(col, probs = point.quantiles))
 
-    Pvals = colMeans(signal.p)
-    Pvals.bar <- apply(signal.p, 2, function(col) quantile(col, probs = bar.quantiles))
-    Pvals.point <- apply(signal.p, 2, function(col) quantile(col, probs = point.quantiles))
+    ## Pvals = colMeans(signal.p)
+    ## Pvals.bar <- apply(signal.p, 2, function(col) quantile(col, probs = bar.quantiles))
+    ## Pvals.point <- apply(signal.p, 2, function(col) quantile(col, probs = point.quantiles))
 
-    Chivals = colMeans(chi.p)
[…]

### 2014-10-07 selection.R

@@ -1,4 +1,26 @@
 
+
+selection_criteria <- function(model, truth) {
+  if ((length(model) == 1) && (model == 0)) {
+    #out <- c(0, 0, 0)
+    return(list(R = 0, TPR = 0, PPV = 0))
+  } else {
+    TP <- sum(model %in% truth)
+    P <- length(truth)
+    R <- length(model)
+    return(list(R = R, TPR = TP/P, PPV = TP/R))
+    #out <- c(R, TP/P, TP/R)
+  }
+#  names(out) <- c("R", "TPR", "PPV")
+#  return(out)
+}
+
+
+
+# --------------------------------------------------------
+# old stuff below
+# --------------------------------------------------------
+
 source('stop_rules.R')
 
 # Input incl.list, p.list, true.active.set

### 2014-10-07 sim_categorical.R

@@ -1,101 +1,0 @@
-source('simulation.R')
-source('tex_table.R')
-source('plots.R')
-
-nsim = 500
-type = 'default'
-design = 'categorical'
-estimation = FALSE
-
-n = 100
-groups = sort(rep(1:40, 2))
-
-k = 10
-upper = 4
-lower = 2
-max.steps = 15
-noisecorr = 0
-
-output = run_simulation(
-    nsim = nsim,
-    type = type,
-    design = design,
-    n = n, groups = groups, k = k,
-    upper = upper, lower = lower,
-    max.steps = max.steps,
-    cat.groups = unique(groups),
-    estimation = estimation, verbose = TRUE)
-
-with(output, step_plot(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, max.steps, upper, lower, max.beta, min.beta, fwd.power, design, filename))
-
-
-## ps.fname = paste0('figs/bysignal/', design, '_size1_n', n, '_p', p, '_g', p, '_k', k, '_lower', lower, '_upper', upper)
-## if (corr != 0) {
-##   ps.fname = paste0(ps.fname, '_corr', corr)
-## }
-## if (noisecorr != 0) {
-##   ps.fname = paste0(ps.fname, '_noisecorr', noisecorr)
-## }
-## ps.fname = paste0(ps.fname, '.pdf')
-## k = output$k
-## #psr = t(apply(output$psr.mat, 1, cumsum)/k)
-## psr = output$psr.mat
-## print(psr)
-## l = nrow(psr)
-## m = ncol(psr)
-## for (j in 1:nrow(psr)) {
-##     inds = which(psr[j,] >= 1)
-##       if (length(inds) > 1) {
-##             inds = inds[-1]
-##                 psr[j,inds] = psr[j,inds] + cumsum(rep(1/(3*k), length(inds)))
-##           }
-##   }
-## psr = psr + matrix(0.02*rnorm(l*m), nrow=l)
-## pvals = output$signal.p
-## plot.main = paste0("n = ", n, ", p = ", p, ", signal strength ", lower, "/", upper)
-## pdf(ps.fname)
-## plot(psr[1, ], pvals[1, ], pch = ".", cex = 2, xlim=c(-0.1, max(psr) + .1), ylim=c(-0.1, 1.1), xlab = "Proportion of signal recovered", ylab = "P-values", main = plot.main)
-## for (j in 2:l) {
[…]

### 2014-10-07 sim_default.R

@@ -1,27 +1,26 @@
-source('simulation.R')
-source('tpr_simulation.R')
+source('simulations.R')
 source('tex_table.R')
 source('plots.R')
 
-nsim = 400
-type = 'default'
+nsim = 100
 design = 'gaussian'
 estimation = FALSE #TRUE
 n = 100
 #groups = 1:500
 #groups = 1:400
 groups = 1:200
+#groups <- sort(rep(1:100, 2))
 #groups = 1:100
 #groups = 1:50
 
-k = 10
-upper = 3
+k = 5
+upper = 4
 lower = 2
 
-max.steps = 15
+max.steps = 10
 #max.steps = min(n/2, max(groups)-1) #floor(max(groups)/3)
-corr = 0.2 # nonzero only supported for gaussian design
-noisecorr = 0.1
+corr = 0 # nonzero only supported for gaussian design
+noisecorr = 0
 
 if ((corr != 0) & (design != 'gaussian')) {
     stop("Feature correlation only supported for gaussian design")
@@ -29,18 +28,21 @@ if ((corr != 0) & (design != 'gaussian')) {
 
 output = run_simulation(
     nsim = nsim,
-    type = type,
     design = design,
     n = n, groups = groups, k = k,
     upper = upper, lower = lower,
     max.steps = max.steps,
     corr = corr,
     noisecorr = noisecorr,
-    estimation = estimation, verbose = TRUE)
+    estimation = estimation)
 
-output$main.append = paste0(", RIC: (", round(output$ric.tpp, 2), "/", output$ric.size, ")")
+#output$main.append = paste0(", RIC: (", round(output$ric.tpp, 2), "/", output$ric.size, ")")
 
-with(output, step_plot(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, max.steps, upper, lower, max.beta, min.beta, fwd.power, design, filename, main.append))
+with(output, step_plot(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, max.steps, upper, lower, max.beta, min.beta, fwd.power, design, filename)) #, main.append))
+
+warnings()
+
+stop('only one')
 
 ## ps.fname = paste0('figs/bysignal/', design, '_size1_n', n, '_p', p, '_g', p, '_k', k, '_lower', lower, '_upper', upper)
 ## if (corr != 0) {
[…]

### 2014-10-07 sim_special.R

@@ -1,120 +1,0 @@
-type = 'glint'
-#type = 'gamsel'
-
-
-source('simulation.R')
-source('tex_table.R')
-source('plots.R')
-
-estimation = FALSE
-
-if (type == 'gamsel') {
-    design = 'uniform'
-    n = 200
-    groups = 1:50
-    nsim = 200
-} else {
-    design = 'gaussian'
-    n = 100
-    groups = 1:30
-    nsim = 200
-}
-
-k = 6
-k0 = 2
-upper = 10
-lower = 5
-max.steps = 18
-
-
-corr = 0 # nonzero only supported for gaussian design
-noisecorr = 0
-
-p = length(groups)
-
-if ((corr != 0) & (design != 'gaussian')) {
-    stop("Feature correlation only supported for gaussian design")
-}
-
-output = run_simulation(
-    nsim = nsim,
-    type = type,
-    design = design,
-    n = n, groups = groups, k = k, k0 = k0,
-    upper = upper, lower = lower,
-    max.steps = max.steps,
-    estimation = estimation, verbose = TRUE)
-
-special.power = paste0("(", round(output$special.fwd.power, 2), ")")
-with(output, step_plot(TrueStep, null.p, signal.p, chi.p, k, n, p, g, ugsizes, max.steps, upper, lower, max.beta, min.beta, fwd.power, design, filename, main.append = special.power))
-
-ps.fname = paste0("figs/bysignal/", output$filename, ".pdf")
-k = output$k
-#psr = t(apply(output$psr.mat, 1, cumsum)/k)
-psr = output$psr.mat
-l = nrow(psr)
-m = ncol(psr)
-for (j in 1:nrow(psr)) {
-    inds = which(psr[j,] >= 1)
-      if (length(inds) > 1) {
-            inds = inds[-1]
-                psr[j,inds] = psr[j,inds] + cumsum(rep(1/(3*k), length(inds)))
-          }
-  }
-psr = psr + matrix(0.02*rnorm(l*m), nrow=l)
-pvals = output$signal.p
-plot.main = paste0("n = ", n, ", p = ", p, ", signal strength ", lower, "/", upper)
-pdf(ps.fname)
-plot(psr[1, ], pvals[1, ], pch = ".", cex = 2, xlim=c(-0.1, max(psr) + .1), ylim=c(-0.1, 1.1), xlab = "Proportion of signal recovered", ylab = "P-values", main = plot.main)
-for (j in 2:l) {
-    inds = which(psr[j,] <= 2)
[…]

### 2014-10-07 sim_tpr.R

@@ -1,64 +1,0 @@
-
-source('tpr_simulation.R')
-
-groups = 1:200
-#groups = 1:500
-#groups = 1:50
-#groups = sort(c(rep(1:26, 2), rep(27:38, 4)))
-#groups = sort(rep(1:36, 2))
-#groups = sort(rep(1:50, 3))
-#groups = sort(rep(1:200, 2))
-
-type = 'default'
-design = 'gaussian'
-#design = 'categorical'
-#design = 'ternary'
-#design = 'uniform'
-
-corr = 0
-noisecorr = 0
-#corr = 0
-#noisecorr = 0
-
-nsim = 200
-n = 100
-
-kmax = 25
-upper = 2
-lower = 1.5
-
-### No editing below ###
-
-if (type == 'gamsel') {
-  design = 'uniform'
-}
-
-if (design == 'categorical') {
-  cat.groups = unique(groups)
-} else {
-  cat.groups = NULL
-}
-
-if (design != 'gaussian') {
-  corr = 0
-}
-
-estimation = FALSE
-
-
-if ((corr != 0) & (design != 'gaussian')) {
-    stop("Feature correlation only supported for gaussian design")
-}
-
-output = run_tpr_simulation(
-    nsim = nsim,
-    type = type,
-    design = design,
-    n = n, groups = groups, kmax = kmax,
-    upper = upper, lower = lower,
-    cat.groups = cat.groups,
-    noisecorr = noisecorr,
-    corr = corr,
-    estimation = estimation, verbose = TRUE)
-
-

### 2014-10-07 biostatistics/paper.tex
Updated paper

\documentclass[oupdraft]{bio}
%\usepackage[colorlinks=true, urlcolor=citecolor, linkcolor=citecolor, citecolor=citecolor]{hyperref}
\usepackage{natbib}
\usepackage{amsmath,amssymb}
\usepackage[plain,noend]{algorithm2e}
\usepackage{xspace}
% Add history information for the article if required
% \history{Received August 1, 2010;
% revised October 1, 2010;
% accepted for publication November 1, 2010}

\def\aic{\textsc{aic}\xspace}
\def\bic{\textsc{bic}\xspace}
\def\ric{\textsc{ric}\xspace}
\def\forward{\textsc{forward}\xspace}
\def\last{\textsc{last}\xspace}
\def\T{{ \mathrm{\scriptscriptstyle T} }}
\def\v{{\varepsilon}}
\def\real{{\mathbb R}}
\def\pen{{\mathcal P}}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\argmin}{arg\,min}
\newcommand{\norm}[1]{\lVert #1 \rVert}
\newcommand{\vecsp}{\mathcal{L}}
\newcommand{\maximize}{\mathop{\mathrm{maximize}}}

\newenvironment{definition}[1][Definition]{\begin{trivlist}
\item[\hskip \labelsep {\bfseries #1}]}{\end{trivlist}}

\SetAlCapSkip{1em}

\begin{document}

% Title of paper
\title{A significance test for forward stepwise model selection}

% List of authors, with corresponding author marked by asterisk
\author{JOSHUA R. LOFTUS$^\ast$, JONATHAN E. TAYLOR\\[4pt]
% Author addresses
\textit{Department of Statistics,
Stanford University,
290 Serra Mall, Stanford, CA,
United States}
\\[2pt]
% E-mail address for correspondence
{joftius@stanford.edu}}

% Running headers of paper:
\markboth%
% First field is the short list of authors
{J. R. Loftus and J. E. Taylor}
% Second field is the short title of the paper
{Forward stepwise significance test}

\maketitle

% Add a footnote for the corresponding author if one has been
% identified in the author list
\footnotetext{To whom correspondence should be addressed.}

\begin{abstract}
{We propose a new significance test for variables
chosen in forward stepwise model selection.
The test statistic has an
exact truncated chi distribution at the first step conditional on the
variable chosen. Iterating the test on the residual after each step
yields a sequence of conditional $p$-values which we interpret as
tests for the inclusion of each variable conditional on the model
chosen up to that point. Simulations show the test is powerful and
becomes increasingly conservative after forward stepwise has added all
the truly nonzero variables. The resulting method has the
computational strengths of stepwise selection and addresses the
problem of invalid test statistics due to model selection.}
{Forward stepwise; Model selection; Significance test}
\end{abstract}

\section{Introduction}

Forward stepwise regression is a regression model selection procedure
that begins with an empty model and sequentially adds the best
predictor variable in each step. Since forward stepwise uses the data
to choose variables, the usual $\chi^2$ and $F$-tests for significance
are invalid when a model has been selected this way
\citep{olshen:flevel}.
These tests will
be anti-conservative unless they are computed on a held-out validation
dataset. This is an instance of the general problem of conducting
inference and model selection using the same data. In the lasso setting,
\cite{significance:lasso} derived a novel test
statistic with an asymptotic exponential distribution under the null.
\cite{tests:adaptive} modified and extended those
results to the group lasso \citep{bakin:grouplasso,grouplasso} and other penalized regression problems, but only under the global null.
The present work iteratively applies the global null test of
\cite{tests:adaptive} on the residual in each step in forward stepwise
and works out the details necessary for models with grouped
variables.
These tests can be used for
valid significance testing when computed on the same data used for
model selection, eliminating the need for data splitting. This is
important when data splitting is not appropriate, for
example, when categorical covariates have levels with very few
observations.

Forward stepwise adds the ``best'' variable into the model, and our test is
based roughly on comparing the improvement gained from adding this variable
to the improvement possible using the {\em second best} variable.
For intuition, consider the simple case when $y_i \sim N(\mu_i,1)$
for $i = 1,\ldots,n$
are independent and the design matrix $X = I$. Forward stepwise chooses the
largest $|y_i|$ to add to the model. Under the null, this value will be
on the order of $\sqrt{2 \log (n)}$. This is highly significant for a
$N(0,1)$ random variable, so the $\chi^2$ test will fail.
However, the second largest $|y_i|$ will be only slightly smaller, and our
test uses this information to form a valid test for the largest. Under a
sparse alternative in order for any test to have power the nonzero
$\mu_i$ must be larger than $\sqrt{2 \log (n)}$. In this case, 
the gap in improvement between the last truly nonzero $\mu_i$ added by
forward stepwise and the remaining noise variables should be large,
hence detectible by our test.

In the general setting without $X = I$, our statistic can be understood 
as the survival function of the random variable
\begin{equation}
\label{eq:lam1}
\lambda_1 = \|X^Ty\|_{\infty}
\end{equation}
{\em conditional} on which variable achieves the maximum and its sign.
Under the null, and ignoring the maximum for a moment, this quantity is
a $\chi$ random variable with some degrees of freedom $r$ and scale
$\theta$ depending on $X$. Accounting for the maximum by conditioning requires
a truncation set $M \subset \real^+$ given by the fact that we observe
$\lambda_1 \geq m$, where $m$ is determined by the ``second best'' variable.
With these ingredients and the relevant survival function we now define
our test.
\begin{definition}
The {\em truncated $\chi$ test statistic} is
\begin{equation}
\label{eq:tchi}
T\chi(\lambda_1, r, \theta,  M) = \frac{\int_{M /\theta \cap [\lambda_1/\theta,\infty)} F_{\chi_r}(dz)}{\int_{M / \theta} F_{\chi_r}(dz)}
\end{equation}
with explicit formulas for $r$, $\theta$, and $M$ given in
Algorithms~\ref{algo:pval}--\ref{algo:linfrac}.
\end{definition}

The main result in \cite{tests:adaptive} shows that
$T\chi \sim \text{Unif}([0,1])$ under the global null. In forward stepwise
we iterate this test on the residual after each step, and show in simulations
that it performs well.

\section{Forward stepwise model selection}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Background and notation}

Forward stepwise regression is a classical method, reviewed in
\cite{classical:selection}, which continues to be widely used by
practitioners despite the fact that it invalidates inferences using
the standard $\chi^2$ or $F$-tests. Some attempts to address this
issue include Monte Carlo estimation of tables of adjusted
$F$-statistic values \citep{mc:ftoenter} and permutation tests
\citep{permutation:stop}. Aside from the works this paper is based on,
there have been other recent attempts to do inference after model
selection. Most of these make use of subsampling
\citep{meinshausen:buhlmann} or data splitting
\citep{wasserman:roeder}. Our approach allows use of the full data and
does not require the extra computation involved in subsampling.

Our variant of forward stepwise is described in Algorithm~\ref{algo:fs}.
We allow arbitrary predefined groups of variables, so our method can be
used with categorical variables or factor models.
To emphasize this we use $g, h$ as covariate indices rather than
the usual $i, j$ throughout. Since single variables can be considered
groups of size 1, this includes non-grouped situations as a special
case. Let $y \in \real^n$ be $n$ measurements of the outcome
variable, not necessarily independent.
Let an integer $G \geq 2$ be the number of groups of
explanatory variables. For each $1 \leq g \leq G$ the design matrix
encoding the $g$th group is the $n \times p_g$ matrix denoted $X_g$,
where $p_g$ is the number of individual variables or columns in group
$g$. With categorical variables by default we use the full encoding with
one column for every level.
Denote by $p = \sum_{g=1}^Gp_g$ the total number of individual
variables, so $p = G$ in the case where all groups have size 1. Let
$X$ be the matrix constructed by column-binding the $X_g$, that is
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation*}
X = \begin{pmatrix} X_1 & X_2 & \cdots & X_G  \end{pmatrix}
\end{equation*}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
We also allow weights $w_g$ for each group. These
weights act like penalties or costs, so increasing $w_g$ makes it
more difficult for the group $X_g$ to enter the model. The
modeler can choose weights arbitrarily for calibration purposes,
for example to account for groups having different sizes.
Throughout we set all weights to 1 and normalize
groups by the Frobenius norm of their corresponding submatrices.

With each group we associate the $p_g \times 1$ coefficient vector
$\beta_g$, and write $\beta$ for the $p \times 1$ vector constructed
by stacking all of the $\beta_g$ in order.  Finally, our model for the
response is
\begin{equation}
\label{eq:gmodel}
y  = X \beta + \sigma \epsilon  = \sum_{g=1}^G X_g \beta_g + \sigma \epsilon
\end{equation}
where $\epsilon$ is noise. We assume Gaussian noise
$\epsilon | X \sim N(0, \Sigma)$ with known covariance matrix $\Sigma$.
The model \eqref{eq:gmodel} is underdetermined when $p > n$.
In such cases it still may be possible to
estimate $\beta$ well if it is sparse, that is it has few nonzero
entries. In the rest of this paper we refer to variable groups $X_g$
as noise groups if $\beta_g$ is a zero vector and as true or signal
groups if $\beta_g$ has any nonzero entries. We refer to the number
of such nonzero groups as the sparsity of the model, and
denote this $k := \# \{ g : \beta_g \neq 0 \}$.

Before using Algorithm~\ref{algo:fs}
the user must specify the maximum number of steps allowed,
which we denote \textit{steps}.
To enable $T\chi$ statistic computations,
\textit{steps} should be at most $\min (n, G) - 1$. It is
computationally desirable to set it as low as possible while
remaining safely larger than the sparsity of $\beta$
so that forward stepwise may recover all the nonzero
coefficients of $\beta$ and terminate without performing much
additional computation. Since the sparsity is usually unknown this
requires some guesswork.

\begin{figure}
\begin{algorithm}[H]
  \caption{Forward stepwise variant with groups and weights}
  \label{algo:fs}
    \SetAlgoLined
    \DontPrintSemicolon
    \KwData{An $n$ vector $y$ and $n \times p$ matrix $X$ of $G$ variable groups with weights $w_g$}
    \KwResult{Ordered active set $A$ of variable groups included in the model at each step}
    \Begin{
    $A \leftarrow \emptyset$, $A^c \leftarrow \{ 1, \ldots, G\}$, $r_0 \leftarrow y$\;
    \For{$s=1$ \KwTo $steps$}{
        $g^* \leftarrow \argmax_{g \in A^c} \{ \lVert X_g^Tr_{s-1} \rVert_2 / w_g \}$\;
        $P_{g^*} \leftarrow I_{n\times n} - X_{g^*}X_{g^*}^\dagger$\;
        $A \leftarrow A \cup \{ g^* \}$, $A^c \leftarrow A^c \backslash \{ g^* \}$\;
        \For{$h \in A^c$}{
            $X_h \leftarrow P_{g^*} X_h$\;
        }
        $r_s \leftarrow P_{g^*} r_{s-1}$\;
    }
    \KwRet{$A$}
    }
\end{algorithm}
\end{figure}

The active set $A$ contains variable groups chosen to be included in
the model. Fitting $\hat \beta$ can be done by tracking the individual
fits and projections, or by refitting a linear model on the submatrix
of $X$ corresponding to $A$. Other implementations of forward stepwise
use different criteria for choosing the next variable to add, such as
the correlation with the residual. Since we do not renormalize the
columns after projecting the covariates, and
since we have weights, we are not generally computing correlations
unless the design matrix is orthogonal and all weights are 1.
We use forward stepwise to refer specifically to
Algorithm~\ref{algo:fs}.

\section{Significance testing with model selection}
\label{sec:testing}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Previous work}

To describe previous work we require some facts about the solution
path of the lasso summarized in
\cite{significance:lasso,tibshirani_lasso_uniqueness}.
Recall that the lasso estimator is 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
\begin{aligned}
\label{eq:lasso}
\displaystyle \hat \beta(\lambda) = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
   \lambda \| \beta \|_1. \\
\end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{itemize}
\item The vector valued function $\hat \beta(\lambda)$ is a continuous
function of $\lambda$. For the lasso path, the coordinates of $\hat
\beta(\lambda)$ are piecewise linear with changes in slope occurring
at a finite number of $\lambda$ values referred to as \emph{knots}.

\item The knots depend on the data and are usually written in order
$\lambda_1 \geq \lambda_2 \geq \cdots \geq \lambda_r \geq 0$.
\end{itemize}

\cite{significance:lasso} introduced the \emph{covariance test} which
is based on the lasso solution path knots. They derived a simple
asymptotic null distribution, proved a type of convergence under broad
``minimum growth'' conditions, and demonstrated in simulations that
the test statistic closely matches its asymptotic distribution even in
finite samples. Assuming $\Sigma = \sigma^2 I$, the covariance test is
given at the first step by
\begin{equation}
\label{eq:covtest}
T_1 = \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \overset{n,p \to \infty}{\to} \text{Exp}(1).
\end{equation}
This is a hypothesis test for including the first variable in the lasso
path, with large values of the test statistic being evidence against the
global null. 
The limiting results about later steps in \cite{significance:lasso} can be interpreted as recursively applying $T_1$ to the variables that had not previously been selected by lars.
\cite{tests:adaptive} observed
\begin{equation}
\label{test:exact}
\exp\left(- \frac{\lambda_1(\lambda_1-\lambda_2)}{\sigma^2} \right)
\approx
\frac{1 - \Phi(\lambda_1/\sigma)}{1 - \Phi(\lambda_2 / \sigma)} \overset{D}{=} \text{Unif}([0,1])
\end{equation}
with the exact, finite sample distribution on the right hand side holding
under the global null.
Further, their test is also general enough to handle the group lasso and other penalized regression methods. The group lasso estimator \citep{bakin:grouplasso,grouplasso}, a generalization of the lasso, is a solution to the following penalized least-squares convex problem
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
\begin{aligned}
\label{eq:gsoln}
\displaystyle \hat \beta_\lambda = \argmin_{\beta \in \real^p} \frac{1}{2} \| y - X \beta \|_2^2 +
   \lambda {\cal P}(\beta) \\
\end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
with the group penalty
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \begin{aligned}
  \label{eq:gpen}
    {\cal P}(\beta)&= \sum_{g=1}^G w_g \| \beta_g \|_2 \\
  \end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
The parameter $\lambda \geq 0$ enforces sparsity in groups: for large
$\lambda$ most of the $\beta_g$ will be zero vectors. The weights
$w_g$ are usually taken to be $\sqrt {p_g}$ to normalize the penalty
across groups with different sizes.
This includes the usual lasso estimator as a
special case when all groups are of size 1, since then the
penalty term is the $\ell_1$-norm of $\beta$. The solution path of the group
lasso has similar properties to the lasso case, however it is not
generally piecewise linear.

The relation of lasso and group lasso to the present test is that
$T\chi$ is based on $\lambda_1$. Recall that
for sufficiently large $\lambda$ the solution to~\eqref{eq:gsoln} is forced
to be zero. The smallest such $\lambda$ is denoted
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{equation}
  \begin{aligned}
   \lambda_1 &= \inf \{ \lambda \geq 0 : \hat \beta_{\lambda'} = 0 \text{ for all } \lambda' > \lambda \} \\
  \end{aligned}
\end{equation}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
For the group lasso this is explicitly computable by
\begin{equation}
%\label{eq:lammax}
\lambda_1^{\text{group}} = \max_{g} \frac{1}{w_g}\|X_g^Ty\|_2
 = \frac{1}{w_{g^*}}\eta_{g^*}^TX_{g^*}^Ty .
\end{equation}
where $g^*$ is the group that achieves the maximum and
$\eta_{g^*} = X_{g^*}^Ty / \|X_{g^*}^Ty\|_2$ is the unit vector that
achieves the norm $\|X_{g^*}^Ty\|_2$. 
The $T\chi$ test conditions on both the maximizer $g^*$ and
on the unit vector $\eta_{g^*}$.

\subsection{The $T\chi$ test statistic}

For the rest of this section let $g = g^*$ be the index of the group attaining the maximum in Algorithm~\ref{algo:fs}. Computing the $T\chi$ statistic when this group has size larger than 1 requires a bit of extra work. For groups of size one, the sign $\eta_g = \pm 1$, but for larger groups $\eta_g$ is a unit vector, and this is roughly why we need to form the projection matrix in \eqref{eq:proj}.

The underlying Gaussian process theory in \cite{tests:adaptive} studies the process $\eta \mapsto \eta^TX^Ty$ with $\eta$ in some parameter space. With nontrivial groups the parameter space is a manifold with nontrivial tangent spaces. In this case studying the maximum and maximizer necessitates forming an orthogonal projection to the tangent space $\vecsp_g = \{ u \in \real^p : u_g^T X_g^T y = 0, u_h = 0 \text{ for all } h \neq g \}$. If $X_g$ is a single column and $X_g^Ty \neq 0$, which should be the case since $g$ maximizes the norm of this quantity, then the space is trivial and the desired orthogonal projection is the identity. Otherwise, if $X_g$ has $p_g > 1$ columns the space $\vecsp_g$ generally has dimension $p_g-1$. We compute an orthonormal basis by Gram-Schmidt and form a $p_g \times p_g$ matrix, which we denote $V_g$, by appending an additional column of zeroes since Gram-Schmidt only produces $p_g-1$. We can now define the $n \times n$ projection
\begin{equation}
   \label{eq:proj}
   P_g = \Sigma X_gV_g (V_g^T X_g^T \Sigma X_g V_g)^\dagger V_g^TX_g^T. \\
\end{equation}
Also define $H_g = (I-P_g)\Sigma$, and the conditional variance
\begin{equation}
   \label{eq:cvar}
   \sigma^2_g = y^TX_gX_g^T H_g X_gX_g^Ty / (w_g^2 \norm{X_g^Ty}_2^2) \\
\end{equation}
which will ultimately be the scale $\theta$ of the truncated $\chi$.
These simplify when $X_g$ is a single column, in which case $P_g = 0$,
$H_g = \Sigma$, and $\sigma^2_g = X_g^T\Sigma X_g^T/w_g^2$.
Finally, with $F_{\chi^2_r}$ denoting the distribution function of a $\chi^2$ random variable with $r$ degrees of freedom, we are ready to describe the computation of $T\chi$ in Algorithm~\ref{algo:pval}.

\begin{figure}
\begin{algorithm}[H]
 \caption{Computing the $T\chi$ $p$-value} \label{algo:pval}
 \SetAlgoLined
 \DontPrintSemicolon
   \KwData{Residual $r$, grouped design matrix $X$ with weights, known covariance $\Sigma$, inactive set $A^c$, index $g$ of last group to enter active set $A$.}
   \KwResult{$T\chi$ $p$-value for the group $g$ entering the model.}\;
   \Begin{
   \eIf{$p_g = 1$}{
     $\sigma_g^2 \leftarrow X_g^T\Sigma X_g/w_g^2$\;
     $U_g \leftarrow X_g / w_g \cdot \text{sign}(X_g^Tr)$\;
   }{
     $H_g \leftarrow (I - P_g)\Sigma$ \tcp*{using $P_g$ from \eqref{eq:proj}}
     $\theta \leftarrow \sigma^2_g$ \tcp*{see \eqref{eq:cvar}}
     $U_g \leftarrow H_g X_g X_g^T r /(\norm{X_g^Tr}_2 w_g)$\;
   }
   $\lambda \leftarrow \norm{X_g^Tr}_2/w_g$\;
   $a \leftarrow X^T(y - \lambda U_g)$\;
   $b \leftarrow X^T U_g$\;
   $(v_-, v_+) \leftarrow \textsc{LinearFractional}(a,b)$
     \tcp*{Algorithm~\ref{algo:linfrac}}
   $r \leftarrow \text{rank}(X_g)$\;
   \KwRet{$[F_{\chi^2_r}(v^2_-/\theta) - F_{\chi^2_r}(\lambda^2/\theta)]/[F_{\chi^2_r}(v^2_-/\theta) - F_{\chi^2_r}(v^2_+/\theta)]$}
   }
\end{algorithm}
\end{figure}
The \textsc{LinearFractional} optimization subproblem referred to in
Algorithm~\ref{algo:pval} is, intuitively, finding the ``second best''
value that we use to evaluate the significance of $\lambda_1$. It returns
the truncation set $M = (v_-, v_+)$. 
Up to some constraints, the problem is finding the maximum and minimum of
\begin{equation}
\label{eqn:linfrac}
\frac{u_h^T a_h }{1 - u_h^T b_h} \qquad h \neq g, \norm{u_h}_2 = 1
\end{equation}
on two different sets.
The constraints, given by Lemma 1 in \cite{tests:adaptive}, incorporate the information that $g$ maximizes $\norm{X_g^Ty}_2/w_g$.
After ruling out degenerate cases, we obtain the solution in Algorithm~\ref{algo:linfrac} by transforming to trigonometric coordinates and using calculus. In implementations, care must be taken as to the range of the inverse trigonometric functions to guarantee that the lower truncation value $v_- \geq 0$.
\begin{figure}
\begin{algorithm}[H]
 \caption{The \textsc{LinearFractional} subproblem}
 \SetAlgoLined
 \DontPrintSemicolon
 \label{algo:linfrac}
   \KwData{The $p$-vectors $a, b$ in Algorithm~\ref{algo:pval}, weights, inactive set $A^c$, a small tolerance number, we use $10^{-10}$}
   \KwResult{Solution pair $(v_-, v_+)$.}\;
   \For{$h$ in $A^c$}{
     \eIf{$\norm{b_h}_2 = 0$ or $\norm{a_h}_2/\norm{b_h}_2 < \text{tol}$}{
       $(v^-_h, v^+_h) \leftarrow (0, \infty)$\;
     }{
        $\theta_c \leftarrow a_h^Tb_h/(\norm{a_h}_2 \norm{b_h}_2)$\;
        $\theta_s \leftarrow \sqrt{1-\theta_c^2}$\;
        $\theta \leftarrow \arccos (\theta_c)$\;
        $\phi_s \leftarrow \theta_s \norm{b_h}_2 / w_h$\;
       \eIf{$\phi_s > 1$}{
          $(v^-_h, v^+_h) \leftarrow (0, \infty)$\;
       }{
          $\phi \leftarrow \arcsin(\phi_s)$\;
          $z_\pm \leftarrow s_\pm \norm{a_h}_2 \cos(\phi) /(w_h - s_\pm \norm{b_h}_2 \cos(\theta - \phi))$ for $s_\pm = \pm 1$\;
         \eIf{$\norm{b_h}_2 < w_h$}{
            $(v_h^-, v_h^+) \leftarrow (\max \{ z_+, z_- \}, \infty)$\;
         }{
            $(v_h^-, v_h^+) \leftarrow (\min \{ z_+, z_- \}, \max \{ z_+, z_- \})$\;
         }
       }
     }
   }
    $v_- \leftarrow \max_h v_h^-$\;
    $v_+ \leftarrow \min_h v_h^+$\;
   \KwRet{$(v_-, v_+)$}
\end{algorithm}
\end{figure}

After adding the first group in forward stepwise and computing the $T\chi$ $p$-value, we orthogonalize the residual and all remaining groups with respect to the group just added. This imposes further constraints that in principle should be used when computing $p$-values at subsequent steps. For now we ignore these constraints and iterate the global null test, but work on incorporating all known constraints is ongoing. The effect of not tracking all constraints is what causes the nominal $p$-value to be increasingly stochastically larger than uniform further down the forward stepwise path, as can be seen in Figures~\ref{fig:binary} and \ref{fig:eyedata}.

\section{Simulations}
\label{sec:simulations}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

To study the behavior of the $T\chi$ test after taking steps in the forward stepwise path and to understand its power to detect various departures from the global null we performed simulations with several classes of random design matrices
including Gaussian and categorical, and a real design matrix from the eye data \citep{eyedata}.
Simulated categorical matrices
were generated by first choosing a vector of probabilities for a given
variable from a Dirichlet prior, and then sampling category levels with
that vector of probabilities. The eye data consists of gene expression data from microarray experiments on mammalian eye tissue.

Since we are interested in variable selection we generate signals with
nonzero coefficients on the scale of $\gamma := \sqrt{2 \log(G)/n}$,
where recall $G$ is the number of groups. This scale is small enough to
be challenging for variable selection procedures like lasso and forward
stepwise. Coefficients within a
nonzero group have roughly the same magnitude and are normalized
so that $\| \beta_g \|_2$ has scale $\gamma$ independently of $p_g$.
The magnitudes for nonzero groups all lie in an interval, for
example $[1.5\gamma, 2\gamma]$.

Plots show observed quantiles of test statistics against the uniform quantiles plotted at various steps of forward stepwise. Since forward stepwise does not always capture the true variables first, there are some realizations where the null hypothesis does not hold at step $k+1$. To account for this, we report the false discovery rate at each step as well. If the false discovery rate is less than 5\% at step $k$, then we know that in more than 95\% of the realizations the null hypothesis holds at step $k+1$ and the test statistics should be close to uniform. The performance of forward stepwise can be judged by how low the false discovery rate remains until step $k$, with larger false discovery rates occurring in difficult problems where the design matrices likely have high coherence.

Aside from the $T\chi$ and $\chi^2$ test statistics, we also compute a Monte-Carlo estimate, motivated by \cite{buja_discussion}, based on comparing the norm achieved by the variable being added to the model, $\| X_{g^*}^T r \|_2$, to quantiles of the Monte Carlo sample
\[
\max_{g \in A^c} \| U_g^T z_m \|_2, \quad z_m \sim N(0, I_n) \quad \text{for } m = 1,\ldots,M
\]
with $A^c$ being the set of candidate variables, including $g^*$. Here $U_g$ is a modified version of $X_g$ to account for variables already regressed out of the residual.

\begin{remark}
In all the plots we see that the $T\chi$ $p$-value is smaller than uniform while forward stepwise is still adding true variables, usually up to step $k$, then uniform at the step after the last true variable has been added, usually step $k+1$, and increasingly larger than uniform at later steps.
\end{remark}

When the false discovery rate is high, it means that noise variables are being added before some true variables, and in these cases the $p$-values remain a bit smaller until later steps when the true variables have been fully recovered.
A reasonable stopping rule based
on this behavior is to pick the model including all variables up to and
including the last one with a significant $p$-value. We call this the
\last stopping rule, and it selects the first $\hat k_\textit{last}$
variables in the forward stepwise path where, if
$t_j$ denotes the $T\chi$ $p$-value calculated at the $j$-th step, then
\begin{equation}
  \label{eq:klast}
    \hat k_\textit{last} = \max \{ j \leq steps : t_j < \alpha \}. \\
\end{equation}
The behavior of this stopping rule is shown in Figure~\ref{fig:selection_sparsity} where it performs on par with \ric.

It is a strong assumption that the errors of the linear model are normal with known covariance $\Sigma$. Work is ongoing to relax this assumption, but in Figure~\ref{fig:noisedist} we see that even when the assumption does not hold the $T\chi$ test appears somewhat robust. The Monte-Carlo estimated statistic is more sensitive perturbations in the error distribution.

\section{Discussion}

When adding the next variable in forward stepwise, our hypothesis test
roughly depends on the improvement gained by this variable compared to
the next best variable.
Thus, the $T\chi$ $p$-value is small when there is a large
gap between the variable to be included and all the remaining variables.
If multiple predictors have truly nonzero coefficients that are close
in magnitude, the $p$-value may be large until forward stepwise has
included all of these.
This motivated us to consider the \last stopping
rule \eqref{eq:klast} for model selection. In simulations we
found this stopping rule to have good performance in terms of power, 
but it does not control the false discovery rate
unless the truly nonzero coefficients are large enough to guarantee
forward stepwise picks the corresponding variables before picking too
many noise variables.

Forward stepwise and the $T\chi$ test 
are well-suited to parallel computation since operations can be performed
separately on each group of variables.
The authors will release an R
package implementing the methods in this paper and future improvements.

\section{Software}
\label{sec5}

Software in the form of R code and complete documentation is available on request from the corresponding author (joftius@stanford.edu).

% \section{Supplementary Material}
% \label{sec6}

% Supplementary material is available online at
% \href{http://biostatistics.oxfordjournals.org}%
% {http://biostatistics.oxfordjournals.org}.

\section*{Acknowledgments}
The authors would like to thank Robert Tibshirani and Trevor Hastie for many helpful comments and suggestions.

{\it Conflict of Interest}: None declared.

\bibliographystyle{biorefs}
\bibliography{refs}

\begin{figure}
\begin{center}
\includegraphics[width=.9\textwidth]{nonnull_G100_high4_binary.eps}
\caption{Quantile-quantile plots of test statistics computed at various steps in forward stepwise. Design matrices are random with $n = 100$ rows and $p = 200$ columns encoding $G = 100$ binary predictors generated from a Beta-Bernoulli model. Nonzero entries of $\beta$ are multiples of $\sqrt{2 \log (p)/n}$ with multipliers between 4 and 2. The outcome is $y_i = x_i\beta + \epsilon_i$ with $\epsilon_i \sim N(0,1)$. Lines are solid for $T\chi$, dashed for the Monte-Carlo estimate, and dotted for $\chi^2$.}
\label{fig:binary}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=.9\textwidth]{nonnull_G200_high15_data.eps}
\caption{Quantile-quantile plots of test statistics computed at various steps in forward stepwise. The design matrix has $n = 120$ rows and $p = 200$ columns consisting of gene expression data from microarray experiments. Nonzero entries of $\beta$ are multiples of $\sqrt{2 \log (p)/n}$ with multipliers between 15 and 12. The outcome is $y_i = x_i\beta + \epsilon_i$ with $\epsilon_i \sim N(0,1)$. Lines are solid for $T\chi$, dashed for the Monte-Carlo estimate, and dotted for $\chi^2$.}
\label{fig:eyedata}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=\textwidth]{selection_G200_high2.eps}
\caption{\em For various sparsity levels $k$, the true positive rate
  in the left panel and 1 minus the false discovery rate in the right
  panel, for models selected by forward stepwise using various
  stopping criteria. Design matrices have standard Gaussian entries
  with $n = 100$ rows and $p = 200$ columns. Nonzero entries of
  $\beta$ are multiples of $\sqrt{2 \log (p)/n}$ with multipliers
  between 2 and 1.5. Forward stepwise ran with 50 steps and from the
  resulting sequence models were selected by \bic, dotted, by \ric,
  dashed, by \forward in \cite{sequential:fdr}, dot-dash, and by
  \last, described in equation \eqref{eq:klast}, solid.}
\label{fig:selection_sparsity}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
\includegraphics[width=.9\textwidth]{global_null_G200.eps}
\caption{Test statistic behavior when the noise distribution is not correctly specified. For overestimated and underestimated plots, the true $\sigma$ is 20\% smaller or larger, respectively, than the value used in computing each statistic. Heavy-tailed errors are from a $t$-distribution with standard error 50\% larger, using 18/5 degrees of freedom. Skewed errors are a standard normal plus a lognormal, with standard error 50\% larger than assumed. Lines are solid for $T\chi$, dashed for the Monte-Carlo test, and dotted for $\chi^2$. All plots are at the first step and under the global null.}
\label{fig:noisedist}
\end{center}
\end{figure}

\end{document}

### 2014-10-08 biostatistics/paper.tex
Added diabetes example

@@
 that it performs well.
+As a simple example, and for comparison to other methods using the same data, we apply our method on the diabetes dataset described in \cite{lars}.
+The outcome is a measurement of disease progression, and we regress this on baseline measurements of various health indicators.
+Table~\ref{tab:diabetes} shows a comparison between the $T\chi$ $p$-values and standard $\chi^2$ $p$-values.
 
@@
 
+% latex table generated in R 3.1.0 by xtable 1.7-3 package
+% Tue Oct  7 22:25:31 2014
+\begin{table}[ht]
+\centering
+\begin{tabular}{r|rrrrrrrrr}
+  \hline
+Variable & bmi & ltg & map & hdl & sex & ldl & glu & tc & tch \\ 
+  \hline
+T$\chi$ $p$-value & 0.000 & 0.000 & 0.008 & 0.729 & 0.002 & 0.622 & 0.353 & 0.652 & 0.420 \\ 
+$\chi^2$ $p$-value & 0.000 & 0.000 & 0.000 & 0.001 & 0.000 & 0.073 & 0.277 & 0.101 & 0.266 \\ 
+   \hline
+\end{tabular}
+\caption{Forward stepwise results on diabetes dataset.
+Chi-squared $p$-values are computed using the drop in residual sum of squares.
+Variables are listed in the order of inclusion by forward stepwise.}
+\label{tab:diabetes}
+\end{table}
+
 \begin{figure}

### 2014-10-12 diabetes.R
Added diabetes example and funding information

@@ -1,0 +1,32 @@
+library(MASS)
+library(Matrix)
+library(devtools)
+library(parallel)
+load_all("package/fstep/")
+source("plots.R") # for qq_plot
+
+# Load data
+library(lars)
+data(diabetes)
+x <- diabetes$x
+labels <- 1:ncol(x)
+index <- labels
+X <- scale_groups(x, index, scale = FALSE)
+Xnorm <- scale_groups(X, index)
+Y <- diabetes$y
+Y <- Y - mean(Y)
+
+# Scale Y
+bigmodel <- lm(Y ~ Xnorm)
+Y <-  Y/sqrt(mean(resid(bigmodel)^2))
+
+# Run forward stepwise
+fit <- fstep(x=Xnorm, y=Y, Sigma=1, steps = 9)
+
+# Create table of output
+pvals <- data.frame(round(rbind(fit$p.value, fit$log$chisq), 3))
+colnames(pvals) <- colnames(X)[fit$variable]
+rownames(pvals) <- c("Tchi", "chisq")
+
+xtable(pvals, caption = "Forward stepwise results on diabetes dataset. Chi-squared $p$-values are computed using the drop in residual sum of squares. Variables are listed in the order of inclusion by forward stepwise.", digits = 3)
+
